### 2021-06-20 draftReport.Rmd
default files

---
title: "CYO Project Report - Brazilian Ecommerce"
author: "Asham Vohra"
date: "6/21/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Clustering:

Some customers didn't write a review. But why are they happy or mad?

Sales Prediction:

With purchase date information you'll be able to predict future sales.

Delivery Performance:

You will also be able to work through delivery performance and find ways to optimize delivery times.

Product Quality: (Let's see if we can find this)

Enjoy yourself discovering the products categories that are more prone to customer insatisfaction.

Attention

- An order might have multiple items.
- Each item might be fulfilled by a distinct seller.
- All text identifying stores and partners where replaced by the names of Game of Thrones great houses.

olist_customers_dataset.csv
This dataset has information about the customer and its location. Use it to identify unique customers in the orders dataset and to find the orders delivery location.
At our system each order is assigned to a unique customerid. This means that the same customer will get different ids for different orders. The purpose of having a customerunique_id on the dataset is to allow you to identify customers that made repurchases at the store. Otherwise you would find that each order had a different customer associated with.

olist_order_reviews_dataset.csv
This dataset includes data about the reviews made by the customers.

After a customer purchases the product from Olist Store a seller gets notified to fulfill that order. Once the customer receives the product, or the estimated delivery date is due, the customer gets a satisfaction survey by email where he can give a note for the purchase experience and write down some comments.

olist_orders_dataset.csv
This is the core dataset. From each order you might find all other information.

olist_products_dataset.csv
Products Dataset
This dataset includes data about the products sold by Olist.

# Overview
Data set: https://www.kaggle.com/olistbr/brazilian-ecommerce

# Analysis
# Results
# Conclusion

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2021-06-20 analysis.R

@@ -1,0 +1,3 @@
+# TODO re
+library(dplyr)
+products_dataset <- read.csv(filte.path('data','olist_products_dataset.csv'),header=TRUE)

### 2021-06-27 draftReport.Rmd
added transformations for geo data

@@
 # Analysis
+
+Product Quality: (Let's see if we can find this)
+
+Trying to identify the products categories that are more prone to customer insatisfaction. Some things come to mind
+- categories with more poorly rated products? data set--> order_reviews_dataset
+- categories which tend to be delivered late? data set-->
+- expensive products maybe? do not get value for money?
+- bulky products maybe?
+- some categories maybe limited to few sellers and they may not be selling good products?
+
+Likely classification problem that you select category and given paramters should it come undder customer satisfaction or not.
+Anything less than 3 is unsatisfied
+3 or more is satisfied
+
+but for same product you may have some people satisfied and some as not. 
+( I think that's where knn and related algorithms were introduced, to find mean of the points)
+
+Would be good clustering project but may not be ideal as it based on unsupervised learning and its  coverage was limited.
+
+Maybe it would be wiser to look at sales or delivery times problems.
+
+Clustering:
+
+So given order of an item, delivery time, product etc  can we try to predict customer rating. But will be so miliar to movielens project can become boring
+
+Delivery Performance:
+
+You will also be able to work through delivery performance and find ways to optimize delivery times.
+
+Let's predict delivery time/date
+-- delivery distance
+-- location of customer as some can be remotely connected and need special delivery merchanism. This is independent of delivery distance.
+-- location of seller as shipping from some locations maybe difficult. This is independent of delivery distance.
+-- size of product? 
+-- price of product (expensive nature of product as it could mean extra care or it maybe fragile?)
+-- purchase time....if in night, delivery time maybe longer
+-- order approved at time-> if payment approval came multiple hours after purchase time, this maybe longer
+-- product category -- products of same category may show similar behavior.
+-- or specific product - some products may have higher delivery time.
+-- deliveries maybe impacted due to sales, festival season.
+
+we should consider if we need time stamp or we can take purchase time as 0 and calculated approved at from it.
+though we may lose granularity in terms of time as truck may come everyday at 4.
+so after 4 you can ship , before 4 you cant
+
+estimated delivery date is present in ---> order_estimated_delivery_date
+We can calculate RMSE using that and it will be our benchmark
+
+We can see if we improved than that or not. It will give us indication of how close are we to the production model RMSE.
+
+```{r Install or load packages}
+# TODO use conditional loading or installation of library
+library(dplyr)
+library(lubridate)
+library(tidyverse)
+install.packages("geosphere")
+library(geosphere)
+```
+
+```{r Basic analysis}
+
+products_dataset <- read.csv(file.path('data','olist_products_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) %>% mutate(product_category_name = as.factor(product_category_name))
+
+products_dataset %>% head()
+products_dataset %>% str()
+
+order_reviews_dataset <- read.csv(file.path('data','olist_order_reviews_dataset.csv'),header=TRUE,stringsAsFactors = FALSE)
+
+order_reviews_dataset %>% head()
+order_reviews_dataset %>% str()
+
+# customers data set  with zip code prefix and can be used to uniquely identify a customer
+# customer_id ----> key to the orders dataset. Each order has a unique customer_id.
+# customer_zip_code_prefix --> first five digits of customer zip code
+customers_dataset <- read.csv(file.path('data','olist_customers_dataset.csv'),header=TRUE) %>% mutate(customer_zip_code_prefix=as.factor(customer_zip_code_prefix),customer_id=as.character(customer_id ))
+
+customers_dataset %>% head()
+customers_dataset %>% str()
+
+# geo location details
+# in brazil while 5 digits zip code is good enough to identify state, it may not be correct to identify city
+# it can have geolocation_zip_code_prefix shared across multiple places
+# refer: https://codigo-postal.org/en-us/brazil/
+# The Postal Code Address (CEP) is an eight-digit number set that has by main objective to guide and accelerate the distribution of objects sent through the national mail in Brazil.
+The current structure of the Brazilian postal code is 8 (eight) digits, divided into two parts:
+
+# first part composed of 5 digits where each digit represents the 1) region, 2) sub-region, 3) sector, 4) subsector, 5) subsector divider;
+# the second part is composed of 3 digits, separated by a hyphen from the first part and represents the Distribution Identifiers.
+# You can navigate through states, then cities and finally neighborhoods and streets until you find the corresponding postal code.
+
+geolocation_dataset <- read.csv(file.path('data','olist_geolocation_dataset.csv'),header=TRUE,stringsAsFactors = FALSE)
+
+geolocation_dataset %>% head()
+geolocation_dataset %>% str()
+
+# Important: many geolocation_city have same name but different ascent component. Example: niteroi and niterói
+# intially we can see 
+geolocation_dataset %>% dplyr::count(geolocation_zip_code_prefix,geolocation_city,geolocation_state) %>% nrow()
+
+# post transformation to resolve any Accent or special characters from the city name
+geolocation_dataset <- geolocation_dataset %>% mutate(geolocation_city=stri_trans_general(geolocation_city,"Any-ASCII"))
+
+geolocation_dataset %>% dplyr::count(geolocation_zip_code_prefix,geolocation_city,geolocation_state) %>% nrow()
+
+# as each geolocation_zip_code_prefix can have many data points, we will limit these to one by taking mean for lat, long location
+geolocation_dataset <- geolocation_dataset %>% group_by(geolocation_zip_code_prefix,geolocation_city,geolocation_state) %>% summarise(geolocation_lat=mean(geolocation_lat),geolocation_lng=mean(geolocation_lng)) %>% ungroup()
+
+# TODO review it
+#%>% mutate(geolocation_state=as.factor(geolocation_state)) 
+# TODO review if geolocation_zip_code_prefix and geolocation_city can have multiple states with them.
+
+# remove whitespace before and after reduces one city
+# still issues are there and can be seen by
+zip_codes_of_interest <- geolocation_dataset %>% mutate(geolocation_city=str_trim(geolocation_city)) %>% dplyr::count(geolocation_zip_code_prefix) %>% filter(n>1)  %>% pull(geolocation_zip_code_prefix)
+
+geolocation_dataset %>% filter(geolocation_zip_code_prefix %in% zip_codes_of_interest) %>% view()
+# to filter out duplicates
+# we need not worry about city or state, we can pick any as long as lat, lng values are similar
+
+# seller data set with zip code prefix
+# seller_zip_code_prefix --> first 5 digits of seller zip code
+sellers_dataset <- read.csv(file.path('data','olist_sellers_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) 
+# TODO review if we need it
+# %>% mutate(seller_zip_code_prefix=as.factor(seller_zip_code_prefix),seller_id=as.character(seller_id ))
+
+sellers_dataset %>% head()
+sellers_dataset %>% str()
+
+# Important: many geolocation_city have same name but different ascent component. Example: niteroi and niterói
+# These all need to be handled in data wrangling
+
+sellers_dataset %>% inner_join(geolocation_dataset,by=c('seller_zip_code_prefix'='geolocation_zip_code_prefix')) %>% str()
+
+# TODO there seems to be an issue
+sellers_dataset %>% inner_join(geolocation_dataset,by=c('seller_zip_code_prefix'='geolocation_zip_code_prefix')) %>% dplyr::count(seller_id) %>% slice_max(order_by=n,n=5)
+
+# geolocation_dataset seems to have 4 entries for same zip code . example 13454
+
+# we tried another approach by joining over all zip codee, city and state but it led to no match
+sellers_dataset %>% mutate(seller_city=stri_trans_general(seller_city,"Any-ASCII")) %>% inner_join(geolocation_dataset,by=c('seller_zip_code_prefix'='geolocation_zip_code_prefix','seller_city'='geolocation_city','seller_state'='geolocation_state')) %>% str()
+# TODO -- if we analyse we can see city and state for seller data differ in both datasets. so best to stick to zip code. city and state are prone to errors.
+# we used anti join to identify ones not matching.
+
+sellers_dataset %>% anti_join(geolocation_dataset,by=c('seller_zip_code_prefix'='geolocation_zip_code_prefix'))
+
+# order data set with various delivery related times
+# note each customer_id is unique per order
+# order_purchase_timestamp -> Shows the purchase timestamp.
+# order_approved_at/payment_approved_at -> Shows the payment approval timestamp.
+# order_delivered_carrier_date/handed_over_to_carrier_date -> Shows the order posting timestamp. When it was handled to the logistic partner.
+# order_delivered_customer_date --> Shows the actual order delivery date to the customer.
+# order_estimated_delivery_date --> Shows the estimated delivery date that was informed to customer at the purchase moment.
+orders_dataset <- read.csv(file.path('data','olist_orders_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) %>% rename(payment_approved_at=order_approved_at,handed_over_to_carrier_date=order_delivered_carrier_date)  %>% mutate(order_status=as.factor(order_status),order_purchase_timestamp=as_datetime(order_purchase_timestamp),payment_approved_at=as_datetime(payment_approved_at),handed_over_to_carrier_date=as_datetime(handed_over_to_carrier_date),order_delivered_customer_date=as_datetime(order_delivered_customer_date),order_estimated_delivery_date=as_datetime(order_estimated_delivery_date))
+
+orders_dataset <- orders_dataset %>% filter(order_status=='delivered') %>% filter( !is.na(order_estimated_delivery_date ) & !is.na(order_delivered_customer_date ) )
+
+orders_dataset %>% head()
+orders_dataset %>% str()
+
+# current RMSE for predicted delivery time given by order_estimated_delivery_date vs actual outcome give by order_delivered_customer_date
+# However, note that order_estimated_delivery_date has no timestamp, so we will remove timestamp component, round date using floor_date, which means any delivery done during the day say at 2017-10-10 21:25:13 is treated to be done on 2017-10-10
+
+# first we convert it to desired date
+# then we covert date to numeric i.e. to convert it to seconds
+orders_dataset %>% mutate(order_delivered_customer_date=floor_date(order_delivered_customer_date)) %>% mutate(order_delivered_customer_date=as.numeric(order_delivered_customer_date),order_estimated_delivery_date=as.numeric(order_estimated_delivery_date))  %>% head()
+
+# then we calculate time_taken_from_order_to_delivery, estimated_time_from_order_to_delivery in seconds
+
+orders_dataset %>% head() %>% mutate(order_delivered_customer_date=floor_date(order_delivered_customer_date),order_purchase_timestamp=as.numeric(order_purchase_timestamp)) %>%  mutate(time_taken_from_order_to_delivery=as.numeric(order_delivered_customer_date)-order_purchase_timestamp, estimated_time_from_order_to_delivery = as.numeric(order_estimated_delivery_date) -order_purchase_timestamp) %>% str()
+
+orders_dataset %>% mutate(order_delivered_customer_date=floor_date(order_delivered_customer_date,unit="day"),order_purchase_timestamp=as.numeric(order_purchase_timestamp)) %>%  mutate(time_taken_from_order_to_delivery=as.numeric(order_delivered_customer_date)-order_purchase_timestamp, estimated_time_from_order_to_delivery = as.numeric(order_estimated_delivery_date) -order_purchase_timestamp,diff=estimated_time_from_order_to_delivery-time_taken_from_order_to_delivery)
+
+# if we look at graph we can  see there is huge difference between estimate data and actual delivery date
+# are these different readings or mode of transport is different? 
+p1 <- orders_dataset %>% mutate(order_delivered_customer_date=floor_date(order_delivered_customer_date,unit="day"),order_purchase_timestamp=as.numeric(order_purchase_timestamp)) %>%  mutate(time_taken_from_order_to_delivery=as.numeric(order_delivered_customer_date)-order_purchase_timestamp, estimated_time_from_order_to_delivery = as.numeric(order_estimated_delivery_date) -order_purchase_timestamp,diff=estimated_time_from_order_to_delivery-time_taken_from_order_to_delivery) %>% ggplot(aes(diff)) + geom_histogram(binwidth=86400*15)
+
+library(plotly)
+ggplotly(p1)
+
+# orders data with item related seller info
+# order_item_id --->sequential number identifying number of items included in the same order.
+# so if there are three items they will be nuber 1 , 2, 3 for the given order
+# seller_limit_date ---> Shows the seller shipping limit date for handling the order over to the logistic partner.
+# freight_value --> item freight value item (if an order has more than one item the freight value is splitted between items)
+order_items_dataset <- read.csv(file.path('data','olist_order_items_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) %>% mutate(order_id=as.factor(order_id),product_id=as.factor(product_id),seller_id=as.factor(seller_id))
+
+order_items_dataset %>% str()
+order_items_dataset %>% head()
+
+# we can combine orders delivery information (orders_dataset) with orders items and seller related info (order_items_dataset)
+orders_dataset %>% inner_join(order_items_dataset,by='order_id') %>% head()
+
+# let's combine with seller details
+orders_dataset %>% inner_join(order_items_dataset,by='order_id') %>% inner_join(sellers_dataset,by='seller_id') %>% head()
+
+# let's combine with unique customer and its location dataset
+orders_dataset %>% inner_join(order_items_dataset,by='order_id') %>% inner_join(sellers_dataset,by='seller_id') %>% inner_join(customers_dataset,by='customer_id') %>% select(-customer_id) %>% head()
+
+orders_dataset %>% inner_join(order_items_dataset,by='order_id') %>% inner_join(sellers_dataset,by='seller_id') %>% inner_join(customers_dataset,by='customer_id') %>% select(-customer_id) %>% head() %>% mutate(distance_between_two_points=round(distGeo(c(-80,40),c(-81,41))/1000,2))
+
+# we will use distGeo from geosphere package to compute distance in km between seller and customer.
+considered to be highly accurate.
+if it takes more time, we can consider using distVincentyEllipsoid or distHaversine as we are fine with accuracy of a km.
+orders_dataset %>% inner_join(order_items_dataset,by='order_id') %>% inner_join(sellers_dataset,by='seller_id') %>% inner_join(customers_dataset,by='customer_id') %>% select(-customer_id) %>% head(2) %>% mutate(seller_zip_code_prefix=as.numeric(seller_zip_code_prefix)) %>% select('seller_zip_code_prefix') %>% inner_join(geolocation_dataset,by=c('seller_zip_code_prefix'='geolocation_zip_code_prefix')) %>% rename(geolocation_lat_seller=geolocation_lat,geolocation_lng_seller=geolocation_lng)
+
+orders_dataset %>% inner_join(order_items_dataset,by='order_id') %>% inner_join(sellers_dataset,by='seller_id') %>% inner_join(customers_dataset,by='customer_id') %>% select(-customer_id) %>% head(5) %>% mutate(seller_zip_code_prefix=as.numeric(seller_zip_code_prefix)) %>% mutate(customer_zip_code_prefix=as.numeric(customer_zip_code_prefix)) %>% inner_join(geolocation_dataset,by=c('seller_zip_code_prefix'='geolocation_zip_code_prefix')) %>% rename(geolocation_lat_seller=geolocation_lat,geolocation_lng_seller=geolocation_lng) %>% select(customer_zip_code_prefix) %>% inner_join(geolocation_dataset,by=c('customer_zip_code_prefix'='geolocation_zip_code_prefix')) %>% rename(geolocation_lat_customer=geolocation_lat,geolocation_lng_customer=geolocation_lng)
+
+```
+
 # Results

### 2021-06-27 draftReport.Rmd
further data cleaning

--- 2021-06-20 draftReport.Rmd
@@
 # Analysis
+
+Product Quality: (Let's see if we can find this)
+
+Trying to identify the products categories that are more prone to customer insatisfaction. Some things come to mind
+- categories with more poorly rated products? data set--> order_reviews_dataset
+- categories which tend to be delivered late? data set-->
+- expensive products maybe? do not get value for money?
+- bulky products maybe?
+- some categories maybe limited to few sellers and they may not be selling good products?
+
+Likely classification problem that you select category and given paramters should it come undder customer satisfaction or not.
+Anything less than 3 is unsatisfied
+3 or more is satisfied
+
+but for same product you may have some people satisfied and some as not. 
+( I think that's where knn and related algorithms were introduced, to find mean of the points)
+
+Would be good clustering project but may not be ideal as it based on unsupervised learning and its  coverage was limited.
+
+Maybe it would be wiser to look at sales or delivery times problems.
+
+Clustering:
+
+So given order of an item, delivery time, product etc  can we try to predict customer rating. But will be so miliar to movielens project can become boring
+
+Delivery Performance:
+
+You will also be able to work through delivery performance and find ways to optimize delivery times.
+
+Let's predict delivery time/date
+-- delivery distance
+-- location of customer as some can be remotely connected and need special delivery merchanism. This is independent of delivery distance.
+-- location of seller as shipping from some locations maybe difficult. This is independent of delivery distance.
+-- size of product? 
+-- price of product (expensive nature of product as it could mean extra care or it maybe fragile?)
+-- purchase time....if in night, delivery time maybe longer
+-- order approved at time-> if payment approval came multiple hours after purchase time, this maybe longer
+-- product category -- products of same category may show similar behavior.
+-- or specific product - some products may have higher delivery time.
+-- deliveries maybe impacted due to sales, festival season.
+
+we should consider if we need time stamp or we can take purchase time as 0 and calculated approved at from it.
+though we may lose granularity in terms of time as truck may come everyday at 4.
+so after 4 you can ship , before 4 you cant
+
+estimated delivery date is present in ---> order_estimated_delivery_date
+We can calculate RMSE using that and it will be our benchmark
+
+We can see if we improved than that or not. It will give us indication of how close are we to the production model RMSE.
+
+```{r Install or load packages}
+# TODO use conditional loading or installation of library
+library(dplyr)
+library(lubridate)
+library(tidyverse)
+install.packages("geosphere")
+library(geosphere)
+library(caret)
+```
+
+```{r Basic analysis}
+
+products_dataset <- read.csv(file.path('data','olist_products_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) %>% select(-product_name_lenght,-product_description_lenght,-product_photos_qty)
+
+products_dataset %>% head()
+products_dataset %>% str()
+
+product_category_name_translation <- read.csv(file.path('data','product_category_name_translation.csv'),header=TRUE,stringsAsFactors = FALSE)
+
+product_category_name_translation %>% head()
+product_category_name_translation %>% str()
+
+# convert product_name to english form wherever possible
+products_dataset <- products_dataset %>% left_join(product_category_name_translation,by='product_category_name') %>% mutate(product_category_name=if_else(is.na(product_category_name_english),product_category_name,product_category_name_english)) %>% select(-product_category_name_english)
+
+# further remove products for which data is not available i.e is NA
+products_dataset <- products_dataset %>% filter(!is.na(product_weight_g) & !is.na(product_length_cm) & !is.na(product_height_cm) & !is.na(product_width_cm)  )
+
+# assign category name for blank category
+products_dataset <- products_dataset %>% mutate(product_category_name=if_else(product_category_name=='','UNCATEGORIZED',product_category_name))
+
+# customers data set  with zip code prefix and can be used to uniquely identify a customer
+# customer_id ----> key to the orders dataset. Each order has a unique customer_id.
+# customer_zip_code_prefix --> first five digits of customer zip code
+customers_dataset <- read.csv(file.path('data','olist_customers_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) 
+# %>% mutate(customer_zip_code_prefix=as.factor(customer_zip_code_prefix),customer_id=as.character(customer_id ))
+
+customers_dataset %>% head()
+customers_dataset %>% str()
+
+# geo location details
+# in brazil while 5 digits zip code is good enough to identify state, it may not be correct to identify city
+# it can have geolocation_zip_code_prefix shared across multiple places
+# refer: https://codigo-postal.org/en-us/brazil/
+# The Postal Code Address (CEP) is an eight-digit number set that has by main objective to guide and accelerate the distribution of objects sent through the national mail in Brazil.
+The current structure of the Brazilian postal code is 8 (eight) digits, divided into two parts:
+
+# first part composed of 5 digits where each digit represents the 1) region, 2) sub-region, 3) sector, 4) subsector, 5) subsector divider;
+# the second part is composed of 3 digits, separated by a hyphen from the first part and represents the Distribution Identifiers.
+# You can navigate through states, then cities and finally neighborhoods and streets until you find the corresponding postal code.
+
+geolocation_dataset <- read.csv(file.path('data','olist_geolocation_dataset.csv'),header=TRUE,stringsAsFactors = FALSE)
+
+geolocation_dataset %>% head()
+geolocation_dataset %>% str()
+
+# Important: many geolocation_city have same name but different ascent component. Example: niteroi and niterói
+# intially we can see 
+# geolocation_dataset %>% dplyr::count(geolocation_zip_code_prefix,geolocation_city,geolocation_state) %>% nrow()
+
+# post transformation to resolve any Accent or special characters from the city name
+# geolocation_dataset <- geolocation_dataset %>% mutate(geolocation_city=stri_trans_general(geolocation_city,"Any-ASCII"))
+
+# above is not done as the way zip codes are in brazil, first five digits can correspond to different cities and so we will remove the city.
+We also removed state as there seems to be incorrect entries in the sellers data.
+This either needs to be fixed or we can limit ourselves to lat/long value
+
+# geolocation_dataset %>% dplyr::count(geolocation_zip_code_prefix,geolocation_city,geolocation_state) %>% nrow()
+
+# as each geolocation_zip_code_prefix can have many data points, we will limit these to one by taking mean for lat, long location
+geolocation_dataset <- geolocation_dataset %>% group_by(geolocation_zip_code_prefix) %>% summarise(geolocation_lat=mean(geolocation_lat),geolocation_lng=mean(geolocation_lng)) %>% ungroup()
+#geolocation_dataset <- geolocation_dataset %>% group_by(geolocation_zip_code_prefix,geolocation_city,geolocation_state) %>% summarise(geolocation_lat=mean(geolocation_lat),geolocation_lng=mean(geolocation_lng)) %>% ungroup()
+
+# TODO review it
+#%>% mutate(geolocation_state=as.factor(geolocation_state)) 
+# TODO review if geolocation_zip_code_prefix and geolocation_city can have multiple states with them.
+
+# remove whitespace before and after reduces one city
+# still issues are there and can be seen by
+# zip_codes_of_interest <- geolocation_dataset %>% mutate(geolocation_city=str_trim(geolocation_city)) %>% dplyr::count(geolocation_zip_code_prefix) %>% filter(n>1)  %>% pull(geolocation_zip_code_prefix)
+
+# geolocation_dataset %>% filter(geolocation_zip_code_prefix %in% zip_codes_of_interest) %>% view()
+# to filter out duplicates
+# we need not worry about city or state, we can pick any as long as lat, lng values are similar
+
+# seller data set with zip code prefix
+# seller_zip_code_prefix --> first 5 digits of seller zip code
+sellers_dataset <- read.csv(file.path('data','olist_sellers_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) 
+# TODO review if we need it
+# %>% mutate(seller_zip_code_prefix=as.factor(seller_zip_code_prefix),seller_id=as.character(seller_id ))
+
+sellers_dataset %>% head()
+sellers_dataset %>% str()
+
+# order data set with various delivery related times
+# note each customer_id is unique per order
+# order_purchase_timestamp -> Shows the purchase timestamp.
+# order_approved_at/payment_approved_at -> Shows the payment approval timestamp.
+# order_delivered_carrier_date/handed_over_to_carrier_date -> Shows the order posting timestamp. When it was handled to the logistic partner.
+# order_delivered_customer_date --> Shows the actual order delivery date to the customer.
+# order_estimated_delivery_date --> Shows the estimated delivery date that was informed to customer at the purchase moment.
+orders_dataset <- read.csv(file.path('data','olist_orders_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) %>% rename(payment_approved_at=order_approved_at,handed_over_to_carrier_date=order_delivered_carrier_date)  %>% mutate(order_status=as.factor(order_status),order_purchase_timestamp=as_datetime(order_purchase_timestamp),payment_approved_at=as_datetime(payment_approved_at),handed_over_to_carrier_date=as_datetime(handed_over_to_carrier_date),order_delivered_customer_date=as_datetime(order_delivered_customer_date),order_estimated_delivery_date=as_datetime(order_estimated_delivery_date))
+
+# TODO if we are not using payment_approved_at we can remove the filter logic based on that.
+orders_dataset <- orders_dataset %>% filter(order_status=='delivered') %>% filter( !is.na(order_estimated_delivery_date ) & !is.na(order_delivered_customer_date ) & !is.na(handed_over_to_carrier_date) & !is.na(payment_approved_at) & !is.na(order_purchase_timestamp))
+
+orders_dataset %>% head()
+orders_dataset %>% str()
+
+# orders data with item related seller info
+# order_item_id --->sequential number identifying number of items included in the same order.
+# so if there are three items they will be nuber 1 , 2, 3 for the given order
+# shipping_limit_date ---> Shows the seller shipping limit date for handling the order over to the logistic partner.
+# freight_value --> item freight value item (if an order has more than one item the freight value is splitted between items)
+order_items_dataset <- read.csv(file.path('data','olist_order_items_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) %>% mutate(order_id=as.factor(order_id),product_id=as.factor(product_id),seller_id=as.factor(seller_id), shipping_limit_date=as_datetime(shipping_limit_date))
+
+order_items_dataset %>% str()
+order_items_dataset %>% head()
+```
+
+```{r Current RMSE}
+# current RMSE for predicted delivery time given by order_estimated_delivery_date vs actual outcome give by order_delivered_customer_date
+# However, note that order_estimated_delivery_date has no timestamp, so we will remove timestamp component, round date using floor_date, which means any delivery done during the day say at 2017-10-10 21:25:13 is treated to be done on 2017-10-10
+
+# first we convert it to desired date
+# then we covert date to numeric i.e. to convert it to seconds
+orders_dataset %>% mutate(order_delivered_customer_date=floor_date(order_delivered_customer_date,unit="day")) %>% mutate(order_delivered_customer_date=as.numeric(order_delivered_customer_date),order_estimated_delivery_date=as.numeric(order_estimated_delivery_date))  %>% head()
+
+# then we calculate time_taken_from_order_to_delivery, estimated_time_from_order_to_delivery in seconds
+orders_dataset %>% mutate(order_delivered_customer_date=floor_date(order_delivered_customer_date,unit="day"),order_purchase_timestamp=as.numeric(order_purchase_timestamp)) %>%  mutate(time_taken_from_order_to_delivery=as.numeric(order_delivered_customer_date)-order_purchase_timestamp, estimated_time_from_order_to_delivery = as.numeric(order_estimated_delivery_date) -order_purchase_timestamp,diff=estimated_time_from_order_to_delivery-time_taken_from_order_to_delivery)
+
+# if we look at graph we can  see there is huge difference between estimate data and actual delivery date
+# it is possible some products are bulky or require differen mode of transport like ship 
+p1 <- orders_dataset %>% mutate(order_delivered_customer_date=floor_date(order_delivered_customer_date,unit="day"),order_purchase_timestamp=as.numeric(order_purchase_timestamp)) %>%  mutate(time_taken_from_order_to_delivery=as.numeric(order_delivered_customer_date)-order_purchase_timestamp, estimated_time_from_order_to_delivery = as.numeric(order_estimated_delivery_date) -order_purchase_timestamp,diff=estimated_time_from_order_to_delivery-time_taken_from_order_to_delivery) %>% ggplot(aes(diff)) + geom_histogram(binwidth=86400*15) + xlab('Estimation error where each bar represents 15 days')
+
+library(plotly)
+ggplotly(p1)
+```
+
+```{r Refine datasets}
+#
+# Important: many geolocation_city have same name but different ascent component. Example: niteroi and niterói
+# These all need to be handled in data wrangling
+
+# we noticed some seller_zip_code_prefix are not present in geolocation dataset. So we will ignore them
+sellers_to_ignore <- sellers_dataset %>% anti_join(geolocation_dataset,by=c('seller_zip_code_prefix'='geolocation_zip_code_prefix'))
+
+order_items_dataset %>% inner_join(sellers_to_ignore,by='seller_id') %>% nrow()
+# while there are decent number of order items for the concerned sellers, we will ignore them as we will still have sufficient data for analysis.
+
+# let's use the one of interest
+sellers_dataset <- sellers_dataset %>% inner_join(geolocation_dataset,by=c('seller_zip_code_prefix'='geolocation_zip_code_prefix')) %>% rename(geolocation_lat_seller=geolocation_lat,geolocation_lng_seller=geolocation_lng)
+
+# TODO there seems to be an issue (Should be resolved now)
+# geolocation_dataset seems to have 4 entries for same zip code . example 13454
+# sellers_dataset %>% inner_join(geolocation_dataset,by=c('seller_zip_code_prefix'='geolocation_zip_code_prefix')) %>% dplyr::count(seller_id) %>% slice_max(order_by=n,n=5)
+
+# we tried another approach by joining over all zip codee, city and state but it led to no match
+# sellers_dataset %>% mutate(seller_city=stri_trans_general(seller_city,"Any-ASCII")) %>% inner_join(geolocation_dataset,by=c('seller_zip_code_prefix'='geolocation_zip_code_prefix','seller_city'='geolocation_city','seller_state'='geolocation_state')) %>% str()
+# TODO -- if we analyse we can see city and state for seller data differ in both datasets. so best to stick to zip code. city and state are prone to errors.
+# we decided against using city and state.
+
+# similarly let see customers for whom we don't have data in geolocation_dataset
+customers_to_ignore <- customers_dataset %>% anti_join(geolocation_dataset,by=c('customer_zip_code_prefix'='geolocation_zip_code_prefix'))
+
+orders_dataset %>% inner_join(customers_to_ignore,by='customer_id') %>% nrow()
+# while there are decent number of orders for these customers, we will ignore them as we will still have sufficient data for analysis.
+
+# let's use customers of interest
+customers_dataset <- customers_dataset %>% inner_join(geolocation_dataset,by=c('customer_zip_code_prefix'='geolocation_zip_code_prefix')) %>% rename(geolocation_lat_customer=geolocation_lat,geolocation_lng_customer=geolocation_lng)
+```
+
+```{r Combine data sets}
+# we can combine orders delivery information (orders_dataset) with orders items and seller related info (order_items_dataset)
+orders_dataset %>% inner_join(order_items_dataset,by='order_id') %>% head()
+
+# let's combine with seller details
+orders_dataset %>% inner_join(order_items_dataset,by='order_id') %>% inner_join(sellers_dataset,by='seller_id') %>% head()
+
+# let's combine with unique customer and its location dataset
+orders_dataset %>% inner_join(order_items_dataset,by='order_id') %>% inner_join(sellers_dataset,by='seller_id') %>% inner_join(customers_dataset,by='customer_id') %>% select(-customer_id) %>% head()
+
+# let's combine with products data as well
+orders_dataset %>% inner_join(order_items_dataset,by='order_id') %>% inner_join(sellers_dataset,by='seller_id') %>% inner_join(customers_dataset,by='customer_id') %>% inner_join(products_dataset,by='product_id') %>% select(-customer_id) %>% head()
+
+# seller to customer distance should be right predictor to consider, as the text says "Those merchants are able to sell their products through the Olist Store and ship them directly to the customers using Olist logistics partners."
+# source: https://www.kaggle.com/olistbr/brazilian-ecommerce
+# we will use distGeo from geosphere package to compute distance in km between seller and customer.
+considered to be highly accurate.
+if it takes more time, we can consider using distVincentyEllipsoid or distHaversine as we are fine with accuracy of a km.
+orders_dataset %>% inner_join(order_items_dataset,by='order_id') %>% inner_join(sellers_dataset,by='seller_id') %>% inner_join(customers_dataset,by='customer_id') %>%  inner_join(products_dataset,by='product_id') %>% select(-customer_id) %>% head() %>% mutate(distance_between_two_points=round(distGeo(matrix(c(geolocation_lng_seller,geolocation_lat_seller),ncol=2),matrix(c(geolocation_lng_customer,geolocation_lat_seller),ncol=2))/1000,2))
+
+# so raw data
+raw_data <- orders_dataset %>% inner_join(order_items_dataset,by='order_id') %>% inner_join(sellers_dataset,by='seller_id') %>% inner_join(customers_dataset,by='customer_id') %>% inner_join(products_dataset,by='product_id') %>%  mutate(distance_between_two_points=round(distGeo(matrix(c(geolocation_lng_seller,geolocation_lat_seller),ncol=2),matrix(c(geolocation_lng_customer,geolocation_lat_seller),ncol=2))/1000,2)) %>% select(-geolocation_lat_customer,-geolocation_lng_customer,-geolocation_lat_seller,-geolocation_lng_seller,-customer_id,-order_status,-seller_city,-seller_state,-customer_city,-customer_state)
+
+p1 <- raw_data %>% ggplot(aes(distance_between_two_points)) + geom_histogram(binwidth=50)
+ggplotly(p1)
+
+# identify if there was delay in delivery to carrier
+raw_data <- raw_data %>% select(shipping_limit_date,handed_over_to_carrier_date) %>% mutate(delayed_to_carrier=if_else((as.numeric(handed_over_to_carrier_date)-as.numeric(shipping_limit_date))>0,TRUE,FALSE))
+
+```
+
 # Results
 # Conclusion
+We can make fine predictions if we can predict time for approval, time for handling of goods to carrier and time for delivery separately as final prediction would be based on that.
+But when a new order is made, we only know purchase time
 

### 2021-07-03 draftReport.Rmd
basic analysis using graphs

@@
 # TODO use conditional loading or installation of library
-library(dplyr)
-library(lubridate)
-library(tidyverse)
-install.packages("geosphere")
-library(geosphere)
-library(caret)
+if (!require('dplyr')) install.packages('dplyr'); library('dplyr')
+if (!require('lubridate')) install.packages('lubridate'); library('lubridate')
+if (!require('tidyverse')) install.packages('tidyverse'); library('tidyverse')
+if (!require('geosphere')) install.packages('geosphere'); library('geosphere')
+if (!require('caret')) install.packages('caret'); library('caret')
+if (!require('plotly')) install.packages('plotly'); library('plotly')
 ```
@@
 # freight_value --> item freight value item (if an order has more than one item the freight value is splitted between items)
-order_items_dataset <- read.csv(file.path('data','olist_order_items_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) %>% mutate(order_id=as.factor(order_id),product_id=as.factor(product_id),seller_id=as.factor(seller_id), shipping_limit_date=as_datetime(shipping_limit_date))
+order_items_dataset <- read.csv(file.path('data','olist_order_items_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) %>% mutate(shipping_limit_date=as_datetime(shipping_limit_date))
 
@@
 
-```{r Current RMSE}
-# current RMSE for predicted delivery time given by order_estimated_delivery_date vs actual outcome give by order_delivered_customer_date
-# However, note that order_estimated_delivery_date has no timestamp, so we will remove timestamp component, round date using floor_date, which means any delivery done during the day say at 2017-10-10 21:25:13 is treated to be done on 2017-10-10
-
-# first we convert it to desired date
-# then we covert date to numeric i.e. to convert it to seconds
-orders_dataset %>% mutate(order_delivered_customer_date=floor_date(order_delivered_customer_date,unit="day")) %>% mutate(order_delivered_customer_date=as.numeric(order_delivered_customer_date),order_estimated_delivery_date=as.numeric(order_estimated_delivery_date))  %>% head()
-
-# then we calculate time_taken_from_order_to_delivery, estimated_time_from_order_to_delivery in seconds
-orders_dataset %>% mutate(order_delivered_customer_date=floor_date(order_delivered_customer_date,unit="day"),order_purchase_timestamp=as.numeric(order_purchase_timestamp)) %>%  mutate(time_taken_from_order_to_delivery=as.numeric(order_delivered_customer_date)-order_purchase_timestamp, estimated_time_from_order_to_delivery = as.numeric(order_estimated_delivery_date) -order_purchase_timestamp,diff=estimated_time_from_order_to_delivery-time_taken_from_order_to_delivery)
-
-# if we look at graph we can  see there is huge difference between estimate data and actual delivery date
-# it is possible some products are bulky or require differen mode of transport like ship 
-p1 <- orders_dataset %>% mutate(order_delivered_customer_date=floor_date(order_delivered_customer_date,unit="day"),order_purchase_timestamp=as.numeric(order_purchase_timestamp)) %>%  mutate(time_taken_from_order_to_delivery=as.numeric(order_delivered_customer_date)-order_purchase_timestamp, estimated_time_from_order_to_delivery = as.numeric(order_estimated_delivery_date) -order_purchase_timestamp,diff=estimated_time_from_order_to_delivery-time_taken_from_order_to_delivery) %>% ggplot(aes(diff)) + geom_histogram(binwidth=86400*15) + xlab('Estimation error where each bar represents 15 days')
-
-library(plotly)
-ggplotly(p1)
+```{r RMSE}
+RMSE <- function(true_ratings, predicted_ratings) {
+  sqrt(mean((true_ratings - predicted_ratings)^2))
+}
 ```
@@
 # identify if there was delay in delivery to carrier
-raw_data <- raw_data %>% select(shipping_limit_date,handed_over_to_carrier_date) %>% mutate(delayed_to_carrier=if_else((as.numeric(handed_over_to_carrier_date)-as.numeric(shipping_limit_date))>0,TRUE,FALSE))
+raw_data <- raw_data %>% mutate(delayed_to_carrier=if_else((as.numeric(handed_over_to_carrier_date)-as.numeric(shipping_limit_date))>0,TRUE,FALSE))
 
+wrangled_data <- raw_data %>% select(-shipping_limit_date,-handed_over_to_carrier_date,-payment_approved_at,-order_id,-order_estimated_delivery_date,-order_item_id,-delayed_to_carrier) %>% mutate(order_purchase_timestamp=as.numeric(order_purchase_timestamp),order_delivered_customer_date=as.numeric(order_delivered_customer_date),product_id=as.factor(product_id),seller_id=as.factor(seller_id),seller_zip_code_prefix=as.factor(seller_zip_code_prefix),customer_unique_id=as.factor(customer_unique_id),customer_zip_code_prefix=as.factor(customer_zip_code_prefix),product_category_name=as.factor(product_category_name))
+
+# we will predict number of hours to deliver based on various predictors.
+This seems more appropriate.
+Though as a result of this, we won't be able to use night, day time data of purchase which may influence delivery time (say to carrier it is informed only next day)
+wrangled_data <- wrangled_data %>% mutate(number_of_hours=(order_delivered_customer_date-order_purchase_timestamp)/3600) %>% select(-order_delivered_customer_date,-order_purchase_timestamp) 
 ```
+
+```{r Explore product_category_name relation with number_of_hours}
+wrangled_data %>% group_by(product_category_name) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% mutate(product_category_name=reorder(product_category_name,mean)) %>% ggplot(aes(x=product_category_name,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))
+```
+We can see that number_of_hours of delivery varies as number product category changes
+
+```{r Explore price relation with number_of_hours}
+# first let see how are the items associated with orders priced. Most of the order items a
+p1 <- wrangled_data %>% mutate(price_group=round(price/100)+1) %>% ggplot(aes(price_group)) + geom_histogram(binwidth=1) + xlab('Each price_group is of 100 units each')
+ggplotly(p1)
+
+# let's plot errorbar for number of hours vs price_group.
+# we can see while for price groups number of hours for delivery vary but the relationship is not clear enough. Higher priced items can also be delivered early
+wrangled_data %>% mutate(price_group=round(price/50)+1) %>% mutate(price_group=as.factor(price_group)) %>% group_by(price_group) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% ggplot(aes(x=price_group,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))
+
+cor(wrangled_data$price,wrangled_data$number_of_hours)
+```
+Price Group for a product does not give any clear indication. It does not seem to be impacting delivery time of product.
+
+```{r Explore distance relationship between two points with number_of_hours}
+# let's plot errorbar for number of hours vs distance group
+# we can see a pattern that as distance increases delivery time increases. 
+wrangled_data %>% mutate(distance_group=round(distance_between_two_points/50)+1) %>% mutate(distance_group=as.factor(distance_group)) %>% group_by(distance_group) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% ggplot(aes(x=distance_group,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))
+
+cor(wrangled_data$distance_between_two_points,wrangled_data$number_of_hours)
+```
+
+Distance Group can be seen to influence delivery time.
+
+```{r Explore freight value group with number_of_hours}
+# let's plot errorbar for number of hours vs freight value group
+wrangled_data %>% mutate(freight_value_group=round(freight_value/5)+1) %>% mutate(freight_value_group=as.factor(freight_value_group)) %>% group_by(freight_value_group) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% ggplot(aes(x=freight_value_group,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))
+
+cor(wrangled_data$freight_value,wrangled_data$number_of_hours)
+```
+
+Above graphs indicates that for freight value group where each group represents 5 units, as freight value increases, number of hours for delivery increases.
+
+```{r Explore seller_zip_code_prefix with number_of_hours}
+# TODO anyway to add regularization or remove the low impact seller_zip_code_prefix from the model
+# let's plot errorbar for number of hours vs seller_zip_code_prefix
+sellers_zip_of_interest <- wrangled_data %>% dplyr::count(seller_zip_code_prefix) %>% filter(n>10) %>% pull(seller_zip_code_prefix)
+
+wrangled_data %>% filter(seller_zip_code_prefix %in% sellers_zip_of_interest) %>% group_by(seller_zip_code_prefix) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% mutate(seller_zip_code_prefix=reorder(seller_zip_code_prefix,mean)) %>%  ggplot(aes(x=seller_zip_code_prefix,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90)) + scale_y_sqrt()
+```
+Indicates role between seller_zip_code and number of hours for delivery
+
+```{r Explore customer_zip_code_prefix with number_of_hours}
+# TODO anyway to add regularization or remove the low impact customer_zip_code_prefix from the model
+# let's plot errorbar for number of hours vs customer_zip_code_prefix
+customers_zip_of_interest <- wrangled_data %>% dplyr::count(customer_zip_code_prefix) %>% filter(n>=5) %>% pull(customer_zip_code_prefix)
+
+wrangled_data %>% filter(customer_zip_code_prefix %in% customers_zip_of_interest) %>% group_by(customer_zip_code_prefix) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% mutate(customer_zip_code_prefix=reorder(customer_zip_code_prefix,mean)) %>%  ggplot(aes(x=customer_zip_code_prefix,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90)) + scale_y_sqrt()
+```
+customer_zip_code_prefix can  be seen to be  influence number of hours for delivery
+
+```{r Explore product_id with number_of_hours}
+# TODO anyway to add regularization or remove the low impact products from the model
+# let's plot errorbar for number of hours vs product_id
+products_of_interest <- wrangled_data %>% dplyr::count(product_id)  %>% filter(n>=10) %>% pull(product_id)
+
+wrangled_data %>% filter(product_id %in% products_of_interest) %>% group_by(product_id) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% mutate(product_id=reorder(product_id,mean)) %>%  ggplot(aes(x=product_id,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))
+```
+It influences the number of hours
+
+```{r Explore volume of product group with number_of_hours}
+# let's plot errorbar for number of hours vs volume of product group
+wrangled_data %>% mutate(volume_group=round((product_length_cm*product_width_cm*product_height_cm)/1000)+1) %>% mutate(volume_group=as.factor(volume_group)) %>% group_by(volume_group) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% ggplot(aes(x=volume_group,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))
+```
+It does not influencee much the number of hours
+
+```{r Explore weight of product with number_of_hours}
+# let's plot errorbar for number of hours vs weight of product
+wrangled_data %>% mutate(weight_group=round((product_weight_g)/1000)+1) %>% mutate(weight_group=as.factor(weight_group)) %>% mutate(weight_group=as.factor(weight_group)) %>% group_by(weight_group) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% ggplot(aes(x=weight_group,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))
+```
+Some clues but not strongly correlated.
+
+```{r Current RMSE}
+# current RMSE for predicted delivery time given by order_estimated_delivery_date vs actual outcome give by order_delivered_customer_date
+# However, note that order_estimated_delivery_date has no timestamp, so we will remove timestamp component, round date using floor_date, which means any delivery done during the day say at 2017-10-10 21:25:13 is treated to be done on 2017-10-10
+
+# first we convert it to desired date
+# then we covert date to numeric i.e. to convert it to seconds
+orders_dataset %>% mutate(order_delivered_customer_date=floor_date(order_delivered_customer_date,unit="day")) %>% mutate(order_delivered_customer_date=as.numeric(order_delivered_customer_date),order_estimated_delivery_date=as.numeric(order_estimated_delivery_date))  %>% head()
+
+# then we calculate time_taken_from_order_to_delivery, estimated_time_from_order_to_delivery in seconds
+orders_dataset %>% mutate(order_delivered_customer_date=floor_date(order_delivered_customer_date,unit="day"),order_purchase_timestamp=as.numeric(order_purchase_timestamp)) %>%  mutate(time_taken_from_order_to_delivery=as.numeric(order_delivered_customer_date)-order_purchase_timestamp, estimated_time_from_order_to_delivery = as.numeric(order_estimated_delivery_date) -order_purchase_timestamp,diff=estimated_time_from_order_to_delivery-time_taken_from_order_to_delivery)
+
+# if we look at graph we can  see there is huge difference between estimate data and actual delivery date
+# it is possible some products are bulky or require differen mode of transport like ship 
+p1 <- orders_dataset %>% mutate(order_delivered_customer_date=floor_date(order_delivered_customer_date,unit="day"),order_purchase_timestamp=as.numeric(order_purchase_timestamp)) %>%  mutate(time_taken_from_order_to_delivery=as.numeric(order_delivered_customer_date)-order_purchase_timestamp, estimated_time_from_order_to_delivery = as.numeric(order_estimated_delivery_date) -order_purchase_timestamp,diff=estimated_time_from_order_to_delivery-time_taken_from_order_to_delivery) %>% ggplot(aes(diff)) + geom_histogram(binwidth=86400*15) + xlab('Estimation error where each bar represents 15 days')
+
+library(plotly)
+ggplotly(p1)
+```
+
+Split data into training and validation set, where validation set will be used later
+```{r Split data into training and validation set}
+set.seed(1,sample.kind = 'Rounding')
+test_indices <- createDataPartition(wrangled_data$number_of_hours,times=1,p=0.2,list=FALSE)
+
+training_dataset <- wrangled_data[-test_indices,]
+validation_dataset <- wrangled_data[test_indices,]
+#TODO ensure validation_dataset has products only for concerned product categories which are in training_dataset 
+```
+
+We use training_dataset and split into into train and test set. So that training_dataset is sufficient to fine tune the model
+```{r Split data for training}
+# Let's split training_dataset into two parts
+set.seed(1,sample.kind = 'Rounding')
+test_indices <- createDataPartition(training_dataset$number_of_hours,times=1,p=0.2,list=FALSE)
+
+train_subset <- training_dataset[-test_indices,]
+test_subset <- training_dataset[test_indices,]
+#TODO ensure test_subset has products only for concerned product categories which are in train_subset 
+```
+
+```{r Linear model}
+# trying to reduce coefficients.
+# in dimensions we can use volume to decrase predictors
+
+########### only price predictor#############
+## only price complete data
+lm_fit <- train(number_of_hours~., train_subset %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix) %>% select(price,number_of_hours), method='lm')
+
+predictions <- predict(lm_fit,test_subset  %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix) %>% select(price))
+
+RMSE(predictions,(test_subset  %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix) %>% pull(number_of_hours) ))
+# 220.41
+
+## only price,distance_between_two_points complete data
+lm_fit <- train(number_of_hours~., train_subset %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix) %>% select(price,distance_between_two_points,number_of_hours), method='lm')
+
+predictions <- predict(lm_fit,test_subset  %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix) %>% select(price,distance_between_two_points))
+
+RMSE(predictions,(test_subset  %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix) %>% pull(number_of_hours) ))
+# 206.98
+
+## only price,distance_between_two_points,product_weight_g complete data
+lm_fit <- train(number_of_hours~., train_subset %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix) %>% select(price,distance_between_two_points,product_weight_g,number_of_hours), method='lm')
+
+predictions <- predict(lm_fit,test_subset  %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix) %>% select(price,distance_between_two_points,product_weight_g))
+
+RMSE(predictions,(test_subset  %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix) %>% pull(number_of_hours) ))
+# 206.24
+
+## only price,distance_between_two_points,product_weight_g,product_length_cm complete data
+lm_fit <- train(number_of_hours~., train_subset %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix) %>% select(price,distance_between_two_points,product_weight_g,product_length_cm,number_of_hours), method='lm')
+
+predictions <- predict(lm_fit,test_subset  %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix)  %>% select(price,distance_between_two_points,product_weight_g,product_length_cm))
+
+RMSE(predictions,(test_subset  %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix) %>% pull(number_of_hours) ))
+# 206.22
+
+## only price,distance_between_two_points,product_weight_g,product_length_cm,freight_value complete data
+lm_fit <- train(number_of_hours~., train_subset %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix)  %>% select(price,distance_between_two_points,product_weight_g,product_length_cm,freight_value,number_of_hours), method='lm')
+
+predictions <- predict(lm_fit,test_subset  %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix) %>% select(price,distance_between_two_points,product_weight_g,product_length_cm,freight_value))
+
+RMSE(predictions,(test_subset  %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix) %>% pull(number_of_hours) ))
+# 205.44
+
+## only price,distance_between_two_points,product_weight_g,product_length_cm,product_height_cm,freight_value complete data
+lm_fit <- train(number_of_hours~., train_subset %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix) %>% select(price,distance_between_two_points,product_weight_g,product_length_cm,freight_value,product_height_cm,number_of_hours), method='lm')
+
+predictions <- predict(lm_fit,test_subset  %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix) %>% select(price,distance_between_two_points,product_weight_g,product_length_cm,freight_value,product_height_cm))
+
+RMSE(predictions,(test_subset  %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix) %>% pull(number_of_hours) ))
+# 205.43
+
+## only price,distance_between_two_points,product_weight_g,product_length_cm,product_height_cm,freight_value,product_width_cm complete data
+lm_fit <- train(number_of_hours~., train_subset %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix)  %>% select(price,distance_between_two_points,product_weight_g,product_length_cm,freight_value,product_height_cm,product_width_cm,number_of_hours), method='lm')
+
+predictions <- predict(lm_fit,test_subset  %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix)  %>% select(price,distance_between_two_points,product_weight_g,product_length_cm,freight_value,product_height_cm,product_width_cm))
+
+RMSE(predictions,(test_subset  %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix) %>% pull(number_of_hours) ))
+# 205.41
+
+## only product_category_name complete data
+lm_fit <- train(number_of_hours~., train_subset %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix) %>% select(product_category_name,number_of_hours), method='lm')
+
+predictions <- predict(lm_fit,test_subset  %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix) %>% select(product_category_name))
+
+RMSE(predictions,(test_subset  %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix) %>% pull(number_of_hours) ))
+# 218.89
+
+# if we combine this with other predictors
+## only price,distance_between_two_points,product_category_name complete data
+lm_fit <- train(number_of_hours~., train_subset %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix) %>% select(price,distance_between_two_points,product_category_name,number_of_hours), method='lm')
+
+predictions <- predict(lm_fit,test_subset  %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix) %>% select(price,distance_between_two_points,product_category_name))
+
+RMSE(predictions,(test_subset  %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix) %>% pull(number_of_hours) ))
+# 204.87
+```
+
+To improve model performance consider--> changing character categorical variables to numerical categorical variables like
+
+  product_category_name=as.factor(as.numeric(product_category_name))
+
+use levels from original data for completeness
+
+look for variables which have limited to no results and we can combine them as one categorical variable
+Refer: https://www.analyticsvidhya.com/blog/2015/11/easy-methods-deal-categorical-variables-predictive-modeling/
+Also read : feature hashing
+
+Existing model algorithms should handle regularization. So we should not be required to handle ourselves.
+
+We can use conditional model i.e. for some use different model and for some use different model just like ensemble.
+So basically we could say for these pin codes, delivery time= say mean of delivery time and for others use standard algorthm like random forest etc.
 

### 2021-07-04 draftReport.Rmd
updated models with penalty for delivering late to carrier

--- 2021-06-27 draftReport.Rmd
@@
 date: "6/21/2021"
-output: pdf_document
+output:
+  html_document:
+    df_print: paged
 ---
@@
 
-Clustering:
-
-Some customers didn't write a review. But why are they happy or mad?
-
-Sales Prediction:
-
-With purchase date information you'll be able to predict future sales.
-
 Delivery Performance:
@@
 You will also be able to work through delivery performance and find ways to optimize delivery times.
-
-Product Quality: (Let's see if we can find this)
-
-Enjoy yourself discovering the products categories that are more prone to customer insatisfaction.
 
@@
 
-Product Quality: (Let's see if we can find this)
-
-Trying to identify the products categories that are more prone to customer insatisfaction. Some things come to mind
-- categories with more poorly rated products? data set--> order_reviews_dataset
-- categories which tend to be delivered late? data set-->
-- expensive products maybe? do not get value for money?
-- bulky products maybe?
-- some categories maybe limited to few sellers and they may not be selling good products?
-
-Likely classification problem that you select category and given paramters should it come undder customer satisfaction or not.
-Anything less than 3 is unsatisfied
-3 or more is satisfied
-
-but for same product you may have some people satisfied and some as not. 
-( I think that's where knn and related algorithms were introduced, to find mean of the points)
-
-Would be good clustering project but may not be ideal as it based on unsupervised learning and its  coverage was limited.
-
-Maybe it would be wiser to look at sales or delivery times problems.
-
-Clustering:
-
-So given order of an item, delivery time, product etc  can we try to predict customer rating. But will be so miliar to movielens project can become boring
-
 Delivery Performance:
@@
 -- size of product? 
+-- weight of the product. Heavier products may take time to deliver
 -- price of product (expensive nature of product as it could mean extra care or it maybe fragile?)
@@
 so after 4 you can ship , before 4 you cant
+We can be agnostic to purchase time and just predict number of hours for delivery for given purchase.
 
@@
 
-```{r Install or load packages}
+TODO -> if there is delayed to carrier precendence, we could use that probability and average delay to carrier to fine tune our results.
+Similarly we could use other metrics like payment approval time etc.
+
+```{r Install or load packages,include=FALSE}
 # TODO use conditional loading or installation of library
-library(dplyr)
-library(lubridate)
-library(tidyverse)
-install.packages("geosphere")
-library(geosphere)
-library(caret)
+# Remove plotly related code for pdfs
+if (!require('dplyr')) install.packages('dplyr'); library('dplyr')
+if (!require('lubridate')) install.packages('lubridate'); library('lubridate')
+if (!require('tidyverse')) install.packages('tidyverse'); library('tidyverse')
+if (!require('geosphere')) install.packages('geosphere'); library('geosphere')
+if (!require('caret')) install.packages('caret'); library('caret')
+if (!require('plotly')) install.packages('plotly'); library('plotly')
+if (!require('gam')) install.packages('gam'); library('gam')
 ```
 
-```{r Basic analysis}
+## Data Wrangling
+Reading from raw data source and converting data to usable form
 
+For products data set,
+- we read data, remove unnecessary fields, join with english names whereever available
+- remove products where dimension and related attributes are not available
+- assign default category name of UNCATEGORIZED for products where same was not available
+```{r Data wrangling-products data set, include=FALSE}
 products_dataset <- read.csv(file.path('data','olist_products_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) %>% select(-product_name_lenght,-product_description_lenght,-product_photos_qty)
@@
 product_category_name_translation %>% head()
-product_category_name_translation %>% str()
 
@@
 products_dataset <- products_dataset %>% mutate(product_category_name=if_else(product_category_name=='','UNCATEGORIZED',product_category_name))
+```
+The wrangled data looks like below:
+```{r Data wrangling result-products data set, echo=FALSE}
+products_dataset %>% head()
+```
 
-# customers data set  with zip code prefix and can be used to uniquely identify a customer
-# customer_id ----> key to the orders dataset. Each order has a unique customer_id.
-# customer_zip_code_prefix --> first five digits of customer zip code
-customers_dataset <- read.csv(file.path('data','olist_customers_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) 
-# %>% mutate(customer_zip_code_prefix=as.factor(customer_zip_code_prefix),customer_id=as.character(customer_id ))
-
-customers_dataset %>% head()
-customers_dataset %>% str()
-
+For geolocation data set
+```{r Data wrangling- geolocation data set, echo=FALSE}
 # geo location details
@@
 # we need not worry about city or state, we can pick any as long as lat, lng values are similar
+```
 
+```{r Data wrangling-customers data set, echo=FALSE}
+# customers data set  with zip code prefix and can be used to uniquely identify a customer
+# customer_id ----> key to the orders dataset. Each order has a unique customer_id.
+# customer_zip_code_prefix --> first five digits of customer zip code
+customers_dataset <- read.csv(file.path('data','olist_customers_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) 
+# %>% mutate(customer_zip_code_prefix=as.factor(customer_zip_code_prefix),customer_id=as.character(customer_id ))
+
+customers_dataset %>% head()
+customers_dataset %>% str()
+```
+
+```{r Data wrangling-sellers data set, echo=FALSE}
 # seller data set with zip code prefix
@@
 sellers_dataset %>% str()
+```
 
+Wrangled customers dataset
+- removed data where delivery was not made
+- removed data which was delivered but all the delivery details were not available
+```{r Data wrangling-orders data set, echo=FALSE}
 # order data set with various delivery related times
@@
 
-# TODO if we are not using payment_approved_at we can remove the filter logic based on that.
+# TODO if we are not using payment_approved_at or few fields we can remove the filter logic based on that.
 orders_dataset <- orders_dataset %>% filter(order_status=='delivered') %>% filter( !is.na(order_estimated_delivery_date ) & !is.na(order_delivered_customer_date ) & !is.na(handed_over_to_carrier_date) & !is.na(payment_approved_at) & !is.na(order_purchase_timestamp))
@@
 orders_dataset %>% str()
+```
 
+```{r Data wrangling-order items data set, echo=FALSE}
 # orders data with item related seller info
@@
 
-```{r Current RMSE}
-# current RMSE for predicted delivery time given by order_estimated_delivery_date vs actual outcome give by order_delivered_customer_date
-# However, note that order_estimated_delivery_date has no timestamp, so we will remove timestamp component, round date using floor_date, which means any delivery done during the day say at 2017-10-10 21:25:13 is treated to be done on 2017-10-10
-
-# first we convert it to desired date
-# then we covert date to numeric i.e. to convert it to seconds
-orders_dataset %>% mutate(order_delivered_customer_date=floor_date(order_delivered_customer_date,unit="day")) %>% mutate(order_delivered_customer_date=as.numeric(order_delivered_customer_date),order_estimated_delivery_date=as.numeric(order_estimated_delivery_date))  %>% head()
-
-# then we calculate time_taken_from_order_to_delivery, estimated_time_from_order_to_delivery in seconds
-orders_dataset %>% mutate(order_delivered_customer_date=floor_date(order_delivered_customer_date,unit="day"),order_purchase_timestamp=as.numeric(order_purchase_timestamp)) %>%  mutate(time_taken_from_order_to_delivery=as.numeric(order_delivered_customer_date)-order_purchase_timestamp, estimated_time_from_order_to_delivery = as.numeric(order_estimated_delivery_date) -order_purchase_timestamp,diff=estimated_time_from_order_to_delivery-time_taken_from_order_to_delivery)
-
-# if we look at graph we can  see there is huge difference between estimate data and actual delivery date
-# it is possible some products are bulky or require differen mode of transport like ship 
-p1 <- orders_dataset %>% mutate(order_delivered_customer_date=floor_date(order_delivered_customer_date,unit="day"),order_purchase_timestamp=as.numeric(order_purchase_timestamp)) %>%  mutate(time_taken_from_order_to_delivery=as.numeric(order_delivered_customer_date)-order_purchase_timestamp, estimated_time_from_order_to_delivery = as.numeric(order_estimated_delivery_date) -order_purchase_timestamp,diff=estimated_time_from_order_to_delivery-time_taken_from_order_to_delivery) %>% ggplot(aes(diff)) + geom_histogram(binwidth=86400*15) + xlab('Estimation error where each bar represents 15 days')
-
-library(plotly)
-ggplotly(p1)
+Performance metric
+```{r RMSE, echo=FALSE}
+RMSE <- function(true_outcomes, predicted_outcomes) {
+  sqrt(mean((true_outcomes - predicted_outcomes)^2))
+}
 ```
 
-```{r Refine datasets}
+Filter out data where sellers location details are not available
+```{r Refine sellers dataset, include=FALSE}
 #
@@
 # we decided against using city and state.
+```
 
+Wrangled sellers dataset ,looks like below:
+```{r Refined sellers dataset result, echo=FALSE}
+sellers_dataset %>% head()
+```
+
+```{r Refine customers data set, include=FALSE}
 # similarly let see customers for whom we don't have data in geolocation_dataset
@@
 
-```{r Combine data sets}
+Wrangled customers data set looks like below:
+
+```{r Refined customers dataset result, echo=FALSE}
+customers_dataset %>% head()
+```
+
+Combine segregated data sets
+- create derived info i.e. distance between two points
+
+```{r Combine data sets, include=FALSE}
 # we can combine orders delivery information (orders_dataset) with orders items and seller related info (order_items_dataset)
@@
 if it takes more time, we can consider using distVincentyEllipsoid or distHaversine as we are fine with accuracy of a km.
-orders_dataset %>% inner_join(order_items_dataset,by='order_id') %>% inner_join(sellers_dataset,by='seller_id') %>% inner_join(customers_dataset,by='customer_id') %>%  inner_join(products_dataset,by='product_id') %>% select(-customer_id) %>% head() %>% mutate(distance_between_two_points=round(distGeo(matrix(c(geolocation_lng_seller,geolocation_lat_seller),ncol=2),matrix(c(geolocation_lng_customer,geolocation_lat_seller),ncol=2))/1000,2))
+orders_dataset %>% inner_join(order_items_dataset,by='order_id') %>% inner_join(sellers_dataset,by='seller_id') %>% inner_join(customers_dataset,by='customer_id') %>%  inner_join(products_dataset,by='product_id') %>% select(-customer_id) %>% head() %>% mutate(distance_between_two_points=round(distGeo(matrix(c(geolocation_lng_seller,geolocation_lat_seller),ncol=2),matrix(c(geolocation_lng_customer,geolocation_lat_customer),ncol=2))/1000,2))
 
 # so raw data
-raw_data <- orders_dataset %>% inner_join(order_items_dataset,by='order_id') %>% inner_join(sellers_dataset,by='seller_id') %>% inner_join(customers_dataset,by='customer_id') %>% inner_join(products_dataset,by='product_id') %>%  mutate(distance_between_two_points=round(distGeo(matrix(c(geolocation_lng_seller,geolocation_lat_seller),ncol=2),matrix(c(geolocation_lng_customer,geolocation_lat_seller),ncol=2))/1000,2)) %>% select(-geolocation_lat_customer,-geolocation_lng_customer,-geolocation_lat_seller,-geolocation_lng_seller,-customer_id,-order_status,-seller_city,-seller_state,-customer_city,-customer_state)
+raw_data <- orders_dataset %>% inner_join(order_items_dataset,by='order_id') %>% inner_join(sellers_dataset,by='seller_id') %>% inner_join(customers_dataset,by='customer_id') %>% inner_join(products_dataset,by='product_id') %>%  mutate(distance_between_two_points=round(distGeo(matrix(c(geolocation_lng_seller,geolocation_lat_seller),ncol=2),matrix(c(geolocation_lng_customer,geolocation_lat_customer),ncol=2))/1000,2)) %>% select(-geolocation_lat_customer,-geolocation_lng_customer,-geolocation_lat_seller,-geolocation_lng_seller,-customer_id,-order_status,-seller_city,-seller_state,-customer_city,-customer_state)
 
@@
 
+# TODO we maybe able to remove this as it is not used, because this is the info which is generally not available
 # identify if there was delay in delivery to carrier
-raw_data <- raw_data %>% select(shipping_limit_date,handed_over_to_carrier_date) %>% mutate(delayed_to_carrier=if_else((as.numeric(handed_over_to_carrier_date)-as.numeric(shipping_limit_date))>0,TRUE,FALSE))
+raw_data <- raw_data %>% mutate(delayed_to_carrier=if_else((as.numeric(handed_over_to_carrier_date)-as.numeric(shipping_limit_date))>0,TRUE,FALSE))
 
+wrangled_data <- raw_data %>% select(-shipping_limit_date,-handed_over_to_carrier_date,-payment_approved_at,-order_id,-order_estimated_delivery_date,-order_item_id,-delayed_to_carrier) %>% mutate(order_purchase_timestamp=as.numeric(order_purchase_timestamp),order_delivered_customer_date=as.numeric(order_delivered_customer_date),product_id=as.factor(product_id),seller_id=as.factor(seller_id),seller_zip_code_prefix=as.factor(seller_zip_code_prefix),customer_unique_id=as.factor(customer_unique_id),customer_zip_code_prefix=as.factor(customer_zip_code_prefix),product_category_name=as.factor(product_category_name))
+
+# we will predict number of hours to deliver based on various predictors.
+This seems more appropriate.
+Though as a result of this, we won't be able to use night, day time data of purchase which may influence delivery time (say to carrier it is informed only next day)
+wrangled_data <- wrangled_data %>% mutate(number_of_hours=(order_delivered_customer_date-order_purchase_timestamp)/3600) %>% select(-order_delivered_customer_date,-order_purchase_timestamp) 
+```
+
+```{r Final wrangled data,echo=FALSE}
+wrangled_data %>% head()
+```
+
+```{r Cleanup and save required dataset}
+save(wrangled_data,file='rda/wrangled_data.rda')
+rm(geolocation_dataset, order_items_dataset, orders_dataset, p1, product_category_name_translation, products_dataset, sellers_dataset, sellers_to_ignore, customers_dataset, customers_to_ignore)
+```
+
+## Explore product_category_name relation with number_of_hours
+
+```{r Explore product_category_name relation with number_of_hours, echo=FALSE}
+wrangled_data %>% group_by(product_category_name) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% mutate(product_category_name=reorder(product_category_name,mean)) %>% ggplot(aes(x=product_category_name,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))
+```
+
+We can see that number_of_hours of delivery varies as number product category changes
+
+##Explore price relation with number_of_hours
+
+```{r Explore price relation with number_of_hours, echo = FALSE}
+# first let see how are the items associated with orders priced. Most of the order items a
+wrangled_data %>% mutate(price_group=round(price/100)+1) %>% ggplot(aes(price_group)) + geom_histogram(binwidth=1) + xlab('Each price_group is of 100 units each')
+
+# let's plot errorbar for number of hours vs price_group.
+# we can see while for price groups number of hours for delivery vary but the relationship is not clear enough. Higher priced items can also be delivered early
+wrangled_data %>% mutate(price_group=round(price/50)+1) %>% mutate(price_group=as.factor(price_group)) %>% group_by(price_group) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% ggplot(aes(x=price_group,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))
+
+cor(wrangled_data$price,wrangled_data$number_of_hours)
+```
+
+Price Group for a product does not give any clear indication. It does not seem to be impacting delivery time of product.
+
+## Explore distance relationship between two points with number_of_hours
+
+```{r Explore distance relationship between two points with number_of_hours, echo = FALSE}
+# let's plot errorbar for number of hours vs distance group
+# we can see a pattern that as distance increases delivery time increases. 
+wrangled_data %>% mutate(distance_group=round(distance_between_two_points/50)+1) %>% mutate(distance_group=as.factor(distance_group)) %>% group_by(distance_group) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% ggplot(aes(x=distance_group,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))
+
+cor(wrangled_data$distance_between_two_points,wrangled_data$number_of_hours)
+```
+
+Distance Group can be seen to influence delivery time.
+
+```{r Explore freight value group with number_of_hours}
+# let's plot errorbar for number of hours vs freight value group
+wrangled_data %>% mutate(freight_value_group=round(freight_value/5)+1) %>% mutate(freight_value_group=as.factor(freight_value_group)) %>% group_by(freight_value_group) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% ggplot(aes(x=freight_value_group,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))
+
+cor(wrangled_data$freight_value,wrangled_data$number_of_hours)
+```
+
+Above graphs indicates that for freight value group where each group represents 5 units, as freight value increases, number of hours for delivery increases.
+
+However, it dips near the end
+```{r smooth fit for freight value, echo=FALSE}
+wrangled_data %>% ggplot(aes(x=freight_value,y=number_of_hours)) + geom_smooth() 
+```
+
+## Explore seller_zip_code_prefix with number_of_hours
+
+```{r Explore seller_zip_code_prefix with number_of_hours, echo = FALSE}
+# TODO anyway to add regularization or remove the low impact seller_zip_code_prefix from the model
+# let's plot errorbar for number of hours vs seller_zip_code_prefix
+sellers_zip_of_interest <- wrangled_data %>% dplyr::count(seller_zip_code_prefix) %>% filter(n>10) %>% pull(seller_zip_code_prefix)
+
+wrangled_data %>% filter(seller_zip_code_prefix %in% sellers_zip_of_interest) %>% group_by(seller_zip_code_prefix) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% mutate(seller_zip_code_prefix=reorder(seller_zip_code_prefix,mean)) %>%  ggplot(aes(x=seller_zip_code_prefix,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90)) + scale_y_sqrt()
+```
+Indicates role between seller_zip_code and number of hours for delivery
+
+## Explore customer_zip_code_prefix with number_of_hours
+
+```{r Explore customer_zip_code_prefix with number_of_hours, echo=FALSE}
+# TODO anyway to add regularization or remove the low impact customer_zip_code_prefix from the model
+# let's plot errorbar for number of hours vs customer_zip_code_prefix
+customers_zip_of_interest <- wrangled_data %>% dplyr::count(customer_zip_code_prefix) %>% filter(n>=5) %>% pull(customer_zip_code_prefix)
+
+wrangled_data %>% filter(customer_zip_code_prefix %in% customers_zip_of_interest) %>% group_by(customer_zip_code_prefix) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% mutate(customer_zip_code_prefix=reorder(customer_zip_code_prefix,mean)) %>%  ggplot(aes(x=customer_zip_code_prefix,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90)) + scale_y_sqrt()
+```
+customer_zip_code_prefix can  be seen to be  influence number of hours for delivery
+
+## Explore product_id with number_of_hours
+
+```{r Explore product_id with number_of_hours}
+# TODO anyway to add regularization or remove the low impact products from the model
+# let's plot errorbar for number of hours vs product_id
+products_of_interest <- wrangled_data %>% dplyr::count(product_id)  %>% filter(n>=10) %>% pull(product_id)
+
+wrangled_data %>% filter(product_id %in% products_of_interest) %>% group_by(product_id) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% mutate(product_id=reorder(product_id,mean)) %>%  ggplot(aes(x=product_id,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_blank())
+```
+It influences the number of hours
+
+## Explore volume of product group with number_of_hours
+
+```{r Explore volume of product group with number_of_hours, echo=FALSE}
+# let's plot errorbar for number of hours vs volume of product group
+wrangled_data %>% mutate(volume_group=round((product_length_cm*product_width_cm*product_height_cm)/1000)+1) %>% mutate(volume_group=as.factor(volume_group)) %>% group_by(volume_group) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% ggplot(aes(x=volume_group,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))
+```
+It does not influencee much the number of hours
+
+However a smooth fit shows some pattern 
+```{r smooth fit for volume, echo=FALSE}
+wrangled_data %>% mutate(volume=product_length_cm*product_width_cm*product_height_cm) %>% ggplot(aes(volume,number_of_hours)) + geom_smooth()
+```
+
+## Explore weight of product with number_of_hours
+
+```{r Explore weight of product with number_of_hours}
+# let's plot errorbar for number of hours vs weight of product
+wrangled_data %>% mutate(weight_group=round((product_weight_g)/1000)+1) %>% mutate(weight_group=as.factor(weight_group)) %>% mutate(weight_group=as.factor(weight_group)) %>% group_by(weight_group) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% ggplot(aes(x=weight_group,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))
+```
+Some clues but not strongly correlated.
+
+```{r Smooth fit for product, echo= FALSE}
+wrangled_data  %>% ggplot(aes(product_weight_g,number_of_hours)) + geom_smooth()
+```
+
+Calculating current RMSE range
+```{r Current RMSE}
+# current RMSE can be calculated by finding number of hours for predicted delivery time given by order_estimated_delivery_date vs actual outcome/number of hours give by order_delivered_customer_date
+# However, note that order_estimated_delivery_date has no timestamp, so we will remove timestamp component, round date using floor_date, which means any delivery done during the day say at 2017-10-10 21:25:13 is treated to be done on 2017-10-10
+
+# first we convert it to desired date
+# then we covert date to numeric i.e. to convert it to seconds
+orders_dataset %>% mutate(order_delivered_customer_date_low=floor_date(order_delivered_customer_date,unit="day"),order_delivered_customer_date_high=ceiling_date(order_delivered_customer_date,unit="day")) %>% mutate(order_delivered_customer_date_low=as.numeric(order_delivered_customer_date_low),order_delivered_customer_date_hugh=as.numeric(order_delivered_customer_date_high), order_estimated_delivery_date=as.numeric(order_estimated_delivery_date))  %>% head()
+
+# then we calculate time_taken_from_order_to_delivery, estimated_time_from_order_to_delivery in seconds
+data_for_rmse <- orders_dataset %>% mutate(order_delivered_customer_date_low=floor_date(order_delivered_customer_date,unit="day"),order_delivered_customer_date_high=ceiling_date(order_delivered_customer_date,unit="day"),order_purchase_timestamp=as.numeric(order_purchase_timestamp)) %>%  mutate(time_taken_from_order_to_delivery_rangemax=as.numeric(order_delivered_customer_date_high)-order_purchase_timestamp, time_taken_from_order_to_delivery_rangemin=as.numeric(order_delivered_customer_date_low)-order_purchase_timestamp, estimated_time_from_order_to_delivery = as.numeric(order_estimated_delivery_date) -order_purchase_timestamp) %>% mutate(estimated_time_from_order_to_delivery=estimated_time_from_order_to_delivery/3600,time_taken_from_order_to_delivery_rangemax = time_taken_from_order_to_delivery_rangemax/3600,time_taken_from_order_to_delivery_rangemin = time_taken_from_order_to_delivery_rangemin/3600) %>%  select(estimated_time_from_order_to_delivery,time_taken_from_order_to_delivery_rangemax,time_taken_from_order_to_delivery_rangemin)
+
+RMSE_value_based_on_max <- RMSE(data_for_rmse$time_taken_from_order_to_delivery_rangemax,data_for_rmse$estimated_time_from_order_to_delivery)
+# 357.53
+
+RMSE_value_based_on_min <- RMSE(data_for_rmse$time_taken_from_order_to_delivery_rangemin,data_for_rmse$estimated_time_from_order_to_delivery)
+# 375.408
+
+# if we look at graph we can  see there is huge difference between estimate data and actual delivery date
+# it is possible some products are bulky or require differen mode of transport like ship 
+p1 <- data_for_rmse %>% mutate(diff=estimated_time_from_order_to_delivery-1/2*(time_taken_from_order_to_delivery_rangemax+time_taken_from_order_to_delivery_rangemin)) %>% ggplot(aes(diff)) + geom_histogram(binwidth = 24*2) + xlab('Estimation error where each bar represents 2 days')
+
+library(plotly)
+ggplotly(p1)
+```
+
+Split data into training and validation set, where validation set will be used later
+```{r Split data into training and validation set}
+set.seed(1,sample.kind = 'Rounding')
+test_indices <- createDataPartition(wrangled_data$number_of_hours,times=1,p=0.2,list=FALSE)
+
+training_dataset <- wrangled_data[-test_indices,]
+validation_dataset <- wrangled_data[test_indices,]
+
+# ensure validation_dataset has products only for concerned product categories which are in training_dataset 
+validation_dataset <- validation_dataset %>% semi_join(training_dataset,by='product_category_name')
+
+rm(test_indices)
+```
+
+We use training_dataset and split into into train and test set. So that training_dataset is sufficient to fine tune the model
+```{r Split data for training}
+# Let's split training_dataset into two parts
+set.seed(1,sample.kind = 'Rounding')
+training_dataset <- training_dataset %>% mutate(volume=product_length_cm*product_height_cm*product_width_cm) %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix,-product_length_cm,-product_height_cm,-product_width_cm)
+
+test_indices <- createDataPartition(training_dataset$number_of_hours,times=1,p=0.2,list=FALSE)
+
+train_subset <- training_dataset[-test_indices,]
+test_subset <- training_dataset[test_indices,]
+# ensure test_subset has products only for concerned product categories which are in train_subset
+test_subset <- test_subset %>% semi_join(train_subset,by='product_category_name')
+
+rm(test_indices)
+```
+
+### Linear model trial and analysis
+
+```{r Linear model, eval=FALSE, echo = FALSE}
+# trying to reduce coefficients.
+# in dimensions we can use volume to decrase predictors
+
+########### only price predictor#############
+## only price complete data
+lm_fit <- train(number_of_hours~., train_subset %>% select(price,number_of_hours), method='lm')
+
+lm_fit
+
+predictions <- predict(lm_fit,test_subset %>% select(price))
+
+RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
+# 220.4085
+
+## only price,distance_between_two_points complete data
+lm_fit <- train(number_of_hours~., train_subset %>% select(price,distance_between_two_points,number_of_hours), method='lm')
+
+lm_fit
+
+predictions <- predict(lm_fit,test_subset %>% select(price,distance_between_two_points))
+
+RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
+# 202.0175
+
+## only price,distance_between_two_points,product_weight_g complete data
+lm_fit <- train(number_of_hours~., train_subset %>% select(price,distance_between_two_points,product_weight_g,number_of_hours), method='lm')
+
+lm_fit
+
+predictions <- predict(lm_fit,test_subset %>% select(price,distance_between_two_points,product_weight_g))
+
+RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
+# 201.1412
+
+## only price,distance_between_two_points,product_weight_g,volume complete data
+lm_fit <- train(number_of_hours~., train_subset %>% select(price,distance_between_two_points,product_weight_g,volume,number_of_hours), method='lm')
+
+lm_fit
+
+predictions <- predict(lm_fit,test_subset %>% select(price,distance_between_two_points,product_weight_g,volume))
+
+RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
+# 201.1086
+
+## only price,distance_between_two_points,product_weight_g,volume,freight_value complete data
+lm_fit <- train(number_of_hours~., train_subset %>% select(price,distance_between_two_points,product_weight_g,volume,freight_value,number_of_hours), method='lm')
+
+lm_fit
+
+predictions <- predict(lm_fit,test_subset %>% select(price,distance_between_two_points,product_weight_g,volume,freight_value))
+
+RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
+# 201.1047
+
+## only product_category_name complete data
+lm_fit <- train(number_of_hours~., train_subset %>% select(product_category_name,number_of_hours), method='lm')
+
+lm_fit
+
+predictions <- predict(lm_fit,test_subset %>% select(product_category_name))
+
+RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
+# 218.8943
+
+# if we combine this with other predictors
+## only price,distance_between_two_points,product_category_name complete data
+lm_fit <- train(number_of_hours~., train_subset %>% select(price,distance_between_two_points,product_category_name,number_of_hours), method='lm')
+
+lm_fit
+
+predictions <- predict(lm_fit,test_subset %>% select(price,distance_between_two_points,product_category_name))
+
+RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
+# 199.8844
+```
+
+```{r Linear model with all predictors, echo=FALSE}
+# above is a fine model
+# with all predictors
+lm_fit <- train(number_of_hours~., train_subset, method='lm')
+
+lm_fit
+
+predictions <- predict(lm_fit,test_subset %>% select(-number_of_hours))
+
+RMSE((test_subset %>% pull(number_of_hours) ), predictions)
+# 199.4791
+
+save(lm_fit,file='rda/lm_model.rda')
+```
+Warnings received:
+prediction from a rank-deficient fit may be misleading
+
+To improve model performance consider--> changing character categorical variables to numerical categorical variables like
+
+  product_category_name=as.factor(as.numeric(product_category_name))
+
+use levels from original data for completeness
+
+look for variables which have limited to no results and we can combine them as one categorical variable
+Refer: https://www.analyticsvidhya.com/blog/2015/11/easy-methods-deal-categorical-variables-predictive-modeling/
+Also read : feature hashing
+
+Existing model algorithms should handle regularization. So we should not be required to handle ourselves.
+
+We can use conditional model i.e. for some use different model and for some use different model just like ensemble.
+So basically we could say for these pin codes, delivery time= say mean of delivery time and for others use standard algorthm like random forest etc.
+
+```{r knn model, echo = FALSE}
+set.seed(seed=1,sample.kind = 'Rounding')
+#tuningControl=trainControl(method = "cv", number = 2, p = .8)
+#knn_fit <- train(number_of_hours~., train_subset, method='knn',tuneGrid=data.frame(k=c(15,21,30)),trControl=tuningControl)
+tuningControl=trainControl(method = "cv", number = 3, p = .8)
+knn_fit <- train(number_of_hours~., train_subset, method='knn',tuneGrid=data.frame(k=c(15,23,31)),trControl=tuningControl)
+
+knn_fit
+
+predictions <- predict(knn_fit,test_subset %>% select(-number_of_hours))
+
+RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
+#199.8518
+
+save(knn_fit,file='rda/knn_model.rda')
+```
+
+gam?
+gamLoess??
+
+```{r GamLoess model, echo = FALSE}
+set.seed(seed=1,sample.kind = 'Rounding')
+tuningControl=trainControl(method = "cv", number = 3, p = .8)
+#grid <- expand.grid(span = seq(0.15, 0.65, len = 10), degree = 1)
+grid <- expand.grid(span = seq(0.15, 0.65, len = 3), degree = 1)
+gamLoess_fit <- train(number_of_hours~., train_subset, method='gamLoess',tuneGrid=grid,trControl=tuningControl)
+
+gamLoess_fit
+
+predictions <- predict(gamLoess_fit,test_subset %>% select(-number_of_hours))
+
+RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
+#197.0024
+
+save(gamLoess_fit,file='rda/gamLoess_model.rda')
+```
+
+Warnings received:
+extrapolation not allowed with blending
+ (Discovered by lowesd)liv too small.    
+ (Discovered by lowesd)lv too small
+ extrapolation not allowed with blending
+ non-list contrasts argument ignored
+
+```{r Gam model, echo = FALSE}
+set.seed(seed=1,sample.kind = 'Rounding')
+tuningControl=trainControl(method = "cv", number =2, p = .8)
+gam_fit <- train(number_of_hours~., train_subset, method='gam',trControl=tuningControl) # check for tuning parameters
+
+gam_fit
+
+predictions <- predict(gam_fit,test_subset %>% select(-number_of_hours))
+
+RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
+# 197.5325
+
+save(gam_fit,file='rda/gam_model.rda')
+```
+
+Random forest
+However, it is important to know your data and keep in mind that a Random Forest can’t extrapolate.
+It can only make a prediction that is an average of previously observed labels.
+In this sense it is very similar to KNN.
+In other words, in a regression problem, the range of predictions a Random Forest can make is bound by the highest and lowest labels in the training data.
+This behavior becomes problematic in situations where the training and prediction inputs differ in their range and/or distributions.
+This is called covariate shift and it is difficult for most models to handle but especially for Random Forest, because it can’t extrapolate.
+From : https://towardsdatascience.com/a-limitation-of-random-forest-regression-db8ed7419e9f
+
+```{r rf model, echo = FALSE}
+set.seed(seed=1,sample.kind = 'Rounding')
+tuningControl=trainControl(method = "cv", number = 2, p = .8)
+rf_fit <- train(number_of_hours~., train_subset, method='rf',tuneGrid = data.frame(mtry = seq(2, 6, 2)),trControl=tuningControl) # nodesize = 50,maxnodes=2000,ntree=200
+
+rf_fit
+
+predictions <- predict(rf_fit,test_subset %>% select(-number_of_hours))
+
+RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
+# 192.9998
+
+save(rf_fit,file='rda/rf_model.rda')
+```
+
+```{r Glm model}
+# Let's try with glm model
+set.seed(seed=1,sample.kind = 'Rounding')
+tuningControl=trainControl(method = "cv", number = 2, p = .8)
+glm_fit <- train(number_of_hours~., train_subset, method='glm',trControl=tuningControl)
+
+glm_fit
+
+predictions <- predict(glm_fit, test_subset %>% select(-number_of_hours))
+
+RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
+# 199.4791 (improves if we use seller_id as predictor as well)
+```
+Warning:
+prediction from a rank-deficient fit may be misleading
+
+## Experimets which failed
+
+Let's play with seller_id. It deriorates rf model and for gam model there was no improvement
+```{r play with seller_id, eval=FALSE, echo = FALSE}
+set.seed(1,sample.kind = 'Rounding')
+test_indices <- createDataPartition(wrangled_data$number_of_hours,times=1,p=0.2,list=FALSE)
+dummy_data <- wrangled_data[-test_indices,]
+dummy_data <- dummy_data %>% mutate(volume=product_length_cm*product_height_cm*product_width_cm) %>% mutate(seller_id=factor(as.numeric(seller_id))) %>% select(-customer_unique_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix,-product_length_cm,-product_height_cm,-product_width_cm)
+
+set.seed(1,sample.kind = 'Rounding')
+test_indices <- createDataPartition(dummy_data$number_of_hours,times=1,p=0.2,list=FALSE)
+dummy_train_subset <- dummy_data[-test_indices,]
+dummy_test_subset <- dummy_data[test_indices,]
+
+# only keeping those rows which have seller_id in training set
+dummy_test_subset <- dummy_test_subset %>% semi_join(dummy_train_subset,by='seller_id')
+
+# used gam model but it did not use seller id at all 
+# Let's try with rf model
+set.seed(seed=1,sample.kind = 'Rounding')
+tuningControl=trainControl(method = "cv", number = 2, p = .8)
+rf_fit_dummy <- train(number_of_hours~., dummy_train_subset, method='rf',tuneGrid = data.frame(mtry = seq(5, 7, 1)),trControl=tuningControl) # nodesize = 50,maxnodes=2000,ntree=200
+
+rf_fit_dummy
+
+predictions <- predict(rf_fit_dummy,dummy_test_subset %>% select(-number_of_hours))
+
+RMSE(predictions,(dummy_test_subset %>% pull(number_of_hours) ))
+# 203.6908
+
+# Let's try with glm model
+set.seed(seed=1,sample.kind = 'Rounding')
+tuningControl=trainControl(method = "cv", number = 2, p = .8)
+glm_fit_dummy <- train(number_of_hours~., dummy_train_subset, method='glm',trControl=tuningControl)
+
+glm_fit_dummy
+
+predictions <- predict(glm_fit_dummy,dummy_test_subset %>% select(-number_of_hours))
+
+RMSE(predictions,(dummy_test_subset %>% pull(number_of_hours) ))
+# 195.4884
+```
+
+```{r ensemble, echo = FALSE}
+load('rda/gamLoess_model.rda')
+load('rda/gam_model.rda')
+load('rda/knn_model.rda')
+load('rda/lm_model.rda')
+load('rda/rf_model.rda')
+
+predictions_gamLoess <- predict(gamLoess_fit,test_subset %>% select(-number_of_hours))
+predictions_knn <- predict(knn_fit,test_subset %>% select(-number_of_hours))
+predictions_rf <- predict(rf_fit,test_subset %>% select(-number_of_hours))
+predictions_gam <- predict(gam_fit,test_subset %>% select(-number_of_hours))
+predictions_glm <- predict(glm_fit,test_subset %>% select(-number_of_hours))
+
+predictions_df <- data.frame(predictions_gamLoess,predictions_knn,predictions_rf,predictions_gam,predictions_glm)
+
+RMSE(rowMeans(predictions_df),(test_subset %>% pull(number_of_hours) ))
+# 194.8623
+
+RMSE(apply(predictions_df %>% select(predictions_rf,predictions_gam,predictions_knn) %>% as.matrix(),1,'median'),(test_subset %>% pull(number_of_hours) ))
+# 194.25
+```
+
+TODO-- what if we remove freight value as it depends on number of items in order too.as it is divided by number of items
+
+```{r delayed to carrier experiment, eval=FALSE}
+# let's see how delay in carrier influences overall delivery time
+raw_data  %>% mutate(order_purchase_timestamp=as.numeric(order_purchase_timestamp),order_delivered_customer_date=as.numeric(order_delivered_customer_date),product_id=as.factor(product_id),seller_id=as.factor(seller_id),seller_zip_code_prefix=as.factor(seller_zip_code_prefix),customer_unique_id=as.factor(customer_unique_id),customer_zip_code_prefix=as.factor(customer_zip_code_prefix),product_category_name=as.factor(product_category_name)) %>% mutate(number_of_hours=(order_delivered_customer_date-order_purchase_timestamp)/3600) %>% select(-order_delivered_customer_date,-order_purchase_timestamp)  %>% group_by(delayed_to_carrier) %>% summarise(mean(number_of_hours))
+
+# let's see time contribution due to delay or early drop to carrier 
+raw_data  %>% mutate(order_purchase_timestamp=as.numeric(order_purchase_timestamp),order_delivered_customer_date=as.numeric(order_delivered_customer_date),product_id=as.factor(product_id),seller_id=as.factor(seller_id),seller_zip_code_prefix=as.factor(seller_zip_code_prefix),customer_unique_id=as.factor(customer_unique_id),customer_zip_code_prefix=as.factor(customer_zip_code_prefix),product_category_name=as.factor(product_category_name)) %>% mutate(number_of_hours=(order_delivered_customer_date-order_purchase_timestamp)/3600) %>% select(-order_delivered_customer_date,-order_purchase_timestamp)  %>% mutate(delay= (as.numeric(handed_over_to_carrier_date) -as.numeric(shipping_limit_date))/3600) %>% select(delay,delayed_to_carrier) %>% group_by(delayed_to_carrier) %>% summarise(mean(delay))
+
+# sellers of interest - we have sellers to be penalized(Add delivery time) and sellers to be corrected(reduce delivery time)
+seller_ids_of_interest <- raw_data  %>% mutate(order_purchase_timestamp=as.numeric(order_purchase_timestamp),order_delivered_customer_date=as.numeric(order_delivered_customer_date),product_id=as.factor(product_id),seller_id=as.factor(seller_id),seller_zip_code_prefix=as.factor(seller_zip_code_prefix),customer_unique_id=as.factor(customer_unique_id),customer_zip_code_prefix=as.factor(customer_zip_code_prefix),product_category_name=as.factor(product_category_name)) %>% mutate(number_of_hours=(order_delivered_customer_date-order_purchase_timestamp)/3600) %>% select(-order_delivered_customer_date,-order_purchase_timestamp)  %>% mutate(delay= (as.numeric(handed_over_to_carrier_date) -as.numeric(shipping_limit_date))/3600) %>% select(delayed_to_carrier, seller_id) %>% dplyr::count(seller_id,delayed_to_carrier) %>% spread(delayed_to_carrier,n,fill=0) %>% mutate(delay_percentage=`TRUE`/(`TRUE`+`FALSE`),total=`TRUE`+`FALSE`) %>% filter(total>=10 & (delay_percentage >=0.75 | delay_percentage <=0.25)) %>% pull(seller_id)
+
+# dummy data points
+dummy_wrangled_data <- raw_data  %>% mutate(order_purchase_timestamp=as.numeric(order_purchase_timestamp),order_delivered_customer_date=as.numeric(order_delivered_customer_date),product_id=as.factor(product_id),seller_id=as.factor(seller_id),seller_zip_code_prefix=as.factor(seller_zip_code_prefix),customer_unique_id=as.factor(customer_unique_id),customer_zip_code_prefix=as.factor(customer_zip_code_prefix),product_category_name=as.factor(product_category_name)) %>% mutate(number_of_hours=(order_delivered_customer_date-order_purchase_timestamp)/3600) %>% select(-order_delivered_customer_date,-order_purchase_timestamp)  %>% filter(seller_id %in% seller_ids_of_interest) %>% mutate(delay= (as.numeric(handed_over_to_carrier_date) -as.numeric(shipping_limit_date))/3600) %>% select(-shipping_limit_date,-handed_over_to_carrier_date,-payment_approved_at,-order_id,-order_estimated_delivery_date,-order_item_id,-delayed_to_carrier)
+
+penalty_data <- raw_data  %>% mutate(order_purchase_timestamp=as.numeric(order_purchase_timestamp),order_delivered_customer_date=as.numeric(order_delivered_customer_date),product_id=as.factor(product_id),seller_id=as.factor(seller_id),seller_zip_code_prefix=as.factor(seller_zip_code_prefix),customer_unique_id=as.factor(customer_unique_id),customer_zip_code_prefix=as.factor(customer_zip_code_prefix),product_category_name=as.factor(product_category_name)) %>% mutate(number_of_hours=(order_delivered_customer_date-order_purchase_timestamp)/3600) %>% select(-order_delivered_customer_date,-order_purchase_timestamp)  %>% filter(seller_id %in% seller_ids_of_interest) %>% mutate(delay= (as.numeric(handed_over_to_carrier_date) -as.numeric(shipping_limit_date))/3600) %>% group_by(seller_id) %>% summarise(mean_penalty_to_consider=mean(delay))
+
+set.seed(1,sample.kind = 'Rounding')
+test_indices <- createDataPartition(dummy_wrangled_data$number_of_hours,times=1,p=0.2,list=FALSE)
+dummy_data <- dummy_wrangled_data[-test_indices,]
+dummy_data <- dummy_data %>% mutate(volume=product_length_cm*product_height_cm*product_width_cm) %>%  select(-customer_unique_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix,-product_length_cm,-product_height_cm,-product_width_cm)
+
+set.seed(1,sample.kind = 'Rounding')
+test_indices <- createDataPartition(dummy_data$number_of_hours,times=1,p=0.2,list=FALSE)
+dummy_train_subset <- dummy_data[-test_indices,]
+dummy_test_subset <- dummy_data[test_indices,]
+
+# this penalty data can later be reused with test data as well
+penalty_data <- dummy_train_subset %>% group_by(seller_id) %>% summarise(mean_penalty_to_consider=mean(delay))
+
+dummy_train_subset <- dummy_train_subset %>% left_join(penalty_data,by='seller_id')
+
+# only keeping those rows which have seller_id in training set
+dummy_test_subset <- dummy_test_subset %>% semi_join(dummy_train_subset,by='seller_id')
+
+dummy_train_subset <- dummy_train_subset %>% select(-seller_id,-delay)
+# Let's try with rf model
+set.seed(seed=1,sample.kind = 'Rounding')
+tuningControl=trainControl(method = "cv", number = 2, p = .8)
+rf_fit_dummy <- train(number_of_hours~., dummy_train_subset, method='rf',tuneGrid = data.frame(mtry = seq(5, 7, 1)),trControl=tuningControl) # nodesize = 50,maxnodes=2000,ntree=200
+
+rf_fit_dummy
+
+dummy_test_subset <- dummy_test_subset %>% left_join(penalty_data,by='seller_id') %>% select(-seller_id,-delay)
+predictions <- predict(rf_fit_dummy,dummy_test_subset %>% select(-number_of_hours))
+
+RMSE(predictions,(dummy_test_subset %>% pull(number_of_hours) ))
+#188.2743
 ```

### 2021-07-06 draftReport.Rmd
updated for scaling and one hot encoding

@@
 if (!require('caret')) install.packages('caret'); library('caret')
+if (!require('randomForest')) install.packages('randomForest'); library('randomForest')
 if (!require('plotly')) install.packages('plotly'); library('plotly')
@@
 # let's combine with products data as well
-orders_dataset %>% inner_join(order_items_dataset,by='order_id') %>% inner_join(sellers_dataset,by='seller_id') %>% inner_join(customers_dataset,by='customer_id') %>% inner_join(products_dataset,by='product_id') %>% select(-customer_id) %>% head()
-
-# seller to customer distance should be right predictor to consider, as the text says "Those merchants are able to sell their products through the Olist Store and ship them directly to the customers using Olist logistics partners."
-# source: https://www.kaggle.com/olistbr/brazilian-ecommerce
-# we will use distGeo from geosphere package to compute distance in km between seller and customer.
-considered to be highly accurate.
-if it takes more time, we can consider using distVincentyEllipsoid or distHaversine as we are fine with accuracy of a km.
-orders_dataset %>% inner_join(order_items_dataset,by='order_id') %>% inner_join(sellers_dataset,by='seller_id') %>% inner_join(customers_dataset,by='customer_id') %>%  inner_join(products_dataset,by='product_id') %>% select(-customer_id) %>% head() %>% mutate(distance_between_two_points=round(distGeo(matrix(c(geolocation_lng_seller,geolocation_lat_seller),ncol=2),matrix(c(geolocation_lng_customer,geolocation_lat_customer),ncol=2))/1000,2))
+orders_dataset %>% inner_join(order_items_dataset,by='order_id') %>% inner_join(sellers_dataset,by='seller_id') %>% inner_join(customers_dataset,by='customer_id') %>% inner_join(products_dataset,by='product_id') %>% select(-customer_id,-order_status,-seller_city,-seller_state,-customer_city,-customer_state) %>% head()
 
 # so raw data
-raw_data <- orders_dataset %>% inner_join(order_items_dataset,by='order_id') %>% inner_join(sellers_dataset,by='seller_id') %>% inner_join(customers_dataset,by='customer_id') %>% inner_join(products_dataset,by='product_id') %>%  mutate(distance_between_two_points=round(distGeo(matrix(c(geolocation_lng_seller,geolocation_lat_seller),ncol=2),matrix(c(geolocation_lng_customer,geolocation_lat_customer),ncol=2))/1000,2)) %>% select(-geolocation_lat_customer,-geolocation_lng_customer,-geolocation_lat_seller,-geolocation_lng_seller,-customer_id,-order_status,-seller_city,-seller_state,-customer_city,-customer_state)
+raw_data <- orders_dataset %>% 
+  inner_join(order_items_dataset,by='order_id') %>%
+  inner_join(sellers_dataset,by='seller_id') %>% 
+  inner_join(customers_dataset,by='customer_id') %>%  
+  inner_join(products_dataset,by='product_id') %>%
+  mutate(order_purchase_timestamp=as.numeric(order_purchase_timestamp),
+         product_id=as.factor(product_id), 
+         seller_id=as.factor(seller_id),
+         seller_zip_code_prefix=as.factor(seller_zip_code_prefix),
+         customer_unique_id=as.factor(customer_unique_id),
+         customer_zip_code_prefix=as.factor(customer_zip_code_prefix),
+         product_category_name=as.factor(product_category_name)) %>%
+  select(-customer_id,-order_status,-seller_city,-seller_state,-customer_city,-customer_state, -payment_approved_at,-order_id, -order_item_id, -product_id,-seller_zip_code_prefix,-customer_unique_id,-customer_zip_code_prefix)
+```
 
-p1 <- raw_data %>% ggplot(aes(distance_between_two_points)) + geom_histogram(binwidth=50)
+Calculating current RMSE range
+```{r Current RMSE, echo = FALSE}
+# current RMSE can be calculated by finding number of hours for predicted delivery time given by order_estimated_delivery_date vs actual outcome/number of hours give by order_delivered_customer_date
+# However, note that order_estimated_delivery_date has no timestamp, so we will remove timestamp component, round date using floor_date, which means any delivery done during the day say at 2017-10-10 21:25:13 is treated to be done on 2017-10-10
+
+# first we convert it to desired date
+# then we covert date to numeric i.e. to convert it to seconds
+# then we calculate time_taken_from_order_to_delivery, estimated_time_from_order_to_delivery in seconds
+data_for_rmse <- raw_data %>%
+  select(order_purchase_timestamp,order_delivered_customer_date,order_estimated_delivery_date) %>%
+  mutate(order_delivered_customer_date_low=floor_date(order_delivered_customer_date,unit="day"),
+         order_delivered_customer_date_high=ceiling_date(order_delivered_customer_date,unit="day")) %>% 
+  mutate(order_delivered_customer_date_low=as.numeric(order_delivered_customer_date_low),
+         order_delivered_customer_date_high=as.numeric(order_delivered_customer_date_high),
+         order_estimated_delivery_date=as.numeric(order_estimated_delivery_date))  %>%
+  mutate(time_taken_from_order_to_delivery_rangemax=order_delivered_customer_date_high-order_purchase_timestamp,
+         time_taken_from_order_to_delivery_rangemin=order_delivered_customer_date_low-order_purchase_timestamp,
+         estimated_time_from_order_to_delivery=order_estimated_delivery_date-order_purchase_timestamp) %>% 
+  mutate(estimated_time_from_order_to_delivery=estimated_time_from_order_to_delivery/3600,
+         time_taken_from_order_to_delivery_rangemax = time_taken_from_order_to_delivery_rangemax/3600,
+         time_taken_from_order_to_delivery_rangemin = time_taken_from_order_to_delivery_rangemin/3600) %>%
+  select(estimated_time_from_order_to_delivery,time_taken_from_order_to_delivery_rangemax,time_taken_from_order_to_delivery_rangemin)
+
+RMSE_value_based_on_max <- RMSE(data_for_rmse$time_taken_from_order_to_delivery_rangemax,data_for_rmse$estimated_time_from_order_to_delivery)
+# 359.9466
+
+RMSE_value_based_on_min <- RMSE(data_for_rmse$time_taken_from_order_to_delivery_rangemin,data_for_rmse$estimated_time_from_order_to_delivery)
+# 377.9515
+
+# if we look at graph we can  see there is huge difference between estimate data and actual delivery date
+# it is possible some products are bulky or require differen mode of transport like ship 
+p1 <- data_for_rmse %>% mutate(diff=estimated_time_from_order_to_delivery-1/2*(time_taken_from_order_to_delivery_rangemax+time_taken_from_order_to_delivery_rangemin)) %>% ggplot(aes(diff)) + geom_histogram(binwidth = 24*2) + xlab('Estimation error where each bar represents 2 days')
+
+library(plotly)
 ggplotly(p1)
 
-# TODO we maybe able to remove this as it is not used, because this is the info which is generally not available
-# identify if there was delay in delivery to carrier
-raw_data <- raw_data %>% mutate(delayed_to_carrier=if_else((as.numeric(handed_over_to_carrier_date)-as.numeric(shipping_limit_date))>0,TRUE,FALSE))
+rm(data_for_rmse)
+```
 
-wrangled_data <- raw_data %>% select(-shipping_limit_date,-handed_over_to_carrier_date,-payment_approved_at,-order_id,-order_estimated_delivery_date,-order_item_id,-delayed_to_carrier) %>% mutate(order_purchase_timestamp=as.numeric(order_purchase_timestamp),order_delivered_customer_date=as.numeric(order_delivered_customer_date),product_id=as.factor(product_id),seller_id=as.factor(seller_id),seller_zip_code_prefix=as.factor(seller_zip_code_prefix),customer_unique_id=as.factor(customer_unique_id),customer_zip_code_prefix=as.factor(customer_zip_code_prefix),product_category_name=as.factor(product_category_name))
-
+```{r Prepare final wrangled data set, echo=FALSE}
 # we will predict number of hours to deliver based on various predictors.
@@
 Though as a result of this, we won't be able to use night, day time data of purchase which may influence delivery time (say to carrier it is informed only next day)
-wrangled_data <- wrangled_data %>% mutate(number_of_hours=(order_delivered_customer_date-order_purchase_timestamp)/3600) %>% select(-order_delivered_customer_date,-order_purchase_timestamp) 
+wrangled_data <- raw_data %>%
+  mutate(order_delivered_customer_date=as.numeric(order_delivered_customer_date)) %>%
+  mutate(number_of_hours=(order_delivered_customer_date-order_purchase_timestamp)/3600) %>%
+  select(-order_purchase_timestamp,-order_delivered_customer_date,-order_estimated_delivery_date)
+```
+
+```{r Apply one hot encoding to product_category_name, include = FALSE}
+set.seed(1,sample.kind = 'Rounding')
+# fullRank parameter to make sure there is no perfect collinearity
+product_category_dummyvars <- data.frame(predict(dummyVars(~product_category_name, data=wrangled_data, fullRank = TRUE), newdata = wrangled_data))
+
+wrangled_data <- wrangled_data %>% cbind.data.frame(product_category_dummyvars) %>% select(-product_category_name)
+
+wrangled_data %>% str()
+rm(product_category_dummyvars)
 ```
@@
 save(wrangled_data,file='rda/wrangled_data.rda')
-rm(geolocation_dataset, order_items_dataset, orders_dataset, p1, product_category_name_translation, products_dataset, sellers_dataset, sellers_to_ignore, customers_dataset, customers_to_ignore)
+rm(geolocation_dataset, order_items_dataset, orders_dataset, p1, product_category_name_translation, products_dataset, sellers_dataset, sellers_to_ignore, customers_dataset, customers_to_ignore, raw_data)
+```
+
+Derived data set should be generated later on observations
+- geolocation usage to capture distance
+- delay and mean delay in handling to carrier
+- number of hours for delivery should be generated later
+```{r Method to derive distance data, include = FALSE}
+# we will use distGeo from geosphere package to compute distance in km between seller and customer.
+considered to be highly accurate.
+if it takes more time, we can consider using distVincentyEllipsoid or distHaversine as we are fine with accuracy of a km.
+extract_distance_between_seller_customer <- function(df) { 
+  df %>% mutate(distance_between_two_points=round(distGeo(
+    matrix(c(geolocation_lng_seller,geolocation_lat_seller),ncol=2),
+    matrix(c(geolocation_lng_customer,geolocation_lat_customer),ncol=2)
+    )/1000,2
+    )) %>% select(-geolocation_lat_customer,-geolocation_lng_customer,-geolocation_lat_seller,-geolocation_lng_seller)
+}
+```
+
+```{r Method to derive delay in handover to carrier, include = FALSE}
+# calculated in hours
+get_delay_in_handover_to_carrier_per_seller_id <- function(df) {
+  df %>% mutate(delay= (as.numeric(handed_over_to_carrier_date) -as.numeric(shipping_limit_date))/3600)  %>% 
+    group_by(seller_id) %>% 
+    summarise(mean_penalty_to_consider=mean(delay))
+}
+```
+
+```{r Method to compute volume from available fields, include = FALSE}
+# compute volume
+compute_volume_using_product_info <- function(df) {
+  df %>% mutate(volume=product_length_cm*product_height_cm*product_width_cm) %>%
+    select(-product_length_cm,-product_height_cm,-product_width_cm)
+}
+```
+
+Split data into training and validation set, where validation set will be used later
+```{r Split data into training and validation set, include = FALSE}
+set.seed(1,sample.kind = 'Rounding')
+test_indices <- createDataPartition(wrangled_data$number_of_hours,times=1,p=0.2,list=FALSE)
+
+training_dataset <- wrangled_data[-test_indices,]
+validation_dataset <- wrangled_data[test_indices,]
+
+# ensure validation_dataset has products only for concerned product categories and seller_ids which are in training_dataset 
+# validation_dataset <- validation_dataset %>% semi_join(training_dataset,by='product_category_name') %>% semi_join(training_dataset,by='seller_id')
+validation_dataset <- validation_dataset %>% semi_join(training_dataset,by='seller_id')
+
+rm(test_indices)
+```
+
+We use training_dataset and split into into train and test set. So that training_dataset is sufficient to fine tune the model
+```{r Split data for training, include=FALSE}
+# Let's split training_dataset into two parts
+set.seed(1,sample.kind = 'Rounding')
+training_dataset <- training_dataset %>% compute_volume_using_product_info() %>% extract_distance_between_seller_customer()
+
+test_indices <- createDataPartition(training_dataset$number_of_hours,times=1,p=0.2,list=FALSE)
+
+train_subset <- training_dataset[-test_indices,]
+test_subset <- training_dataset[test_indices,]
+# ensure test_subset has products only for concerned product categories and seller_ids which are in train_subset
+# test_subset <- test_subset %>% semi_join(train_subset,by='product_category_name') %>% semi_join(train_subset,by='seller_id')
+test_subset <- test_subset %>% semi_join(train_subset,by='seller_id')
+
+penalty_data <- train_subset %>% get_delay_in_handover_to_carrier_per_seller_id()
+
+rm(test_indices)
 ```
@@
 ```{r Explore product_category_name relation with number_of_hours, echo=FALSE}
-wrangled_data %>% group_by(product_category_name) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% mutate(product_category_name=reorder(product_category_name,mean)) %>% ggplot(aes(x=product_category_name,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))
+# TODO review it, fix it as required
+training_dataset %>% group_by(product_category_name) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% mutate(product_category_name=reorder(product_category_name,mean)) %>% ggplot(aes(x=product_category_name,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))
 ```
@@
 # first let see how are the items associated with orders priced. Most of the order items a
-wrangled_data %>% mutate(price_group=round(price/100)+1) %>% ggplot(aes(price_group)) + geom_histogram(binwidth=1) + xlab('Each price_group is of 100 units each')
+training_dataset %>% mutate(price_group=round(price/100)+1) %>% ggplot(aes(price_group)) + geom_histogram(binwidth=1) + xlab('Each price_group is of 100 units each')
 
@@
 # we can see while for price groups number of hours for delivery vary but the relationship is not clear enough. Higher priced items can also be delivered early
-wrangled_data %>% mutate(price_group=round(price/50)+1) %>% mutate(price_group=as.factor(price_group)) %>% group_by(price_group) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% ggplot(aes(x=price_group,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))
+training_dataset %>% mutate(price_group=round(price/50)+1) %>% mutate(price_group=as.factor(price_group)) %>% group_by(price_group) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% ggplot(aes(x=price_group,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))
 
-cor(wrangled_data$price,wrangled_data$number_of_hours)
+cor(training_dataset$price,training_dataset$number_of_hours)
 ```
@@
 # we can see a pattern that as distance increases delivery time increases. 
-wrangled_data %>% mutate(distance_group=round(distance_between_two_points/50)+1) %>% mutate(distance_group=as.factor(distance_group)) %>% group_by(distance_group) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% ggplot(aes(x=distance_group,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))
+training_dataset %>% mutate(distance_group=round(distance_between_two_points/50)+1) %>% mutate(distance_group=as.factor(distance_group)) %>% group_by(distance_group) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% ggplot(aes(x=distance_group,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))
 
-cor(wrangled_data$distance_between_two_points,wrangled_data$number_of_hours)
+cor(training_dataset$distance_between_two_points,training_dataset$number_of_hours)
 ```
@@
 # let's plot errorbar for number of hours vs freight value group
-wrangled_data %>% mutate(freight_value_group=round(freight_value/5)+1) %>% mutate(freight_value_group=as.factor(freight_value_group)) %>% group_by(freight_value_group) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% ggplot(aes(x=freight_value_group,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))
+training_dataset %>% mutate(freight_value_group=round(freight_value/5)+1) %>% mutate(freight_value_group=as.factor(freight_value_group)) %>% group_by(freight_value_group) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% ggplot(aes(x=freight_value_group,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))
 
-cor(wrangled_data$freight_value,wrangled_data$number_of_hours)
+cor(training_dataset$freight_value,training_dataset$number_of_hours)
 ```
@@
 ```{r smooth fit for freight value, echo=FALSE}
-wrangled_data %>% ggplot(aes(x=freight_value,y=number_of_hours)) + geom_smooth() 
+training_dataset %>% ggplot(aes(x=freight_value,y=number_of_hours)) + geom_smooth() 
+```
+
+## Explore volume of product group with number_of_hours
+
+```{r Explore volume of product group with number_of_hours, echo=FALSE}
+# let's plot errorbar for number of hours vs volume of product group
+training_dataset %>% mutate(volume_group=round(volume/1000)+1) %>% mutate(volume_group=as.factor(volume_group)) %>% group_by(volume_group) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% ggplot(aes(x=volume_group,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))
+```
+It does not influencee much the number of hours
+
+However a smooth fit shows some pattern 
+```{r smooth fit for volume, echo=FALSE}
+training_dataset %>% ggplot(aes(volume,number_of_hours)) + geom_smooth()
+```
+
+## Explore weight of product with number_of_hours
+
+```{r Explore weight of product with number_of_hours}
+# let's plot errorbar for number of hours vs weight of product
+training_dataset %>% mutate(weight_group=round((product_weight_g)/1000)+1) %>% mutate(weight_group=as.factor(weight_group)) %>% mutate(weight_group=as.factor(weight_group)) %>% group_by(weight_group) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% ggplot(aes(x=weight_group,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))
+```
+Some clues but not strongly correlated.
+
+```{r Smooth fit for product, echo= FALSE}
+training_dataset  %>% ggplot(aes(product_weight_g,number_of_hours)) + geom_smooth()
+```
+
+```{r Prepare training and test data for ,include=FALSE}
+# penalty_data is computed on train_subset and is used with test_subset as historical learning of efficiency of each seller in delivering to carrier
+train_subset <- train_subset %>% left_join(penalty_data,by='seller_id') %>% select(-handed_over_to_carrier_date, -shipping_limit_date, -seller_id)
+
+test_subset <- test_subset %>% left_join(penalty_data,by='seller_id') %>% select(-handed_over_to_carrier_date, -shipping_limit_date, -seller_id)
+```
+
+```{r knn model, echo = FALSE}
+set.seed(seed=1,sample.kind = 'Rounding')
+
+tuningControl=trainControl(method = "cv", number = 3, p = .8)
+knn_fit <- train(number_of_hours~., train_subset, method='knn',tuneGrid=data.frame(k=c(27,34,41)),trControl=tuningControl)
+
+knn_fit
+
+predictions <- predict(knn_fit,test_subset %>% select(-number_of_hours))
+
+RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
+#199.448 for k=34
+
+save(knn_fit,file='rda/knn_model.rda')
+```
+
+```{r knn model data scaline under test, echo=FALSE}
+# TODO review and improve or remove this as required
+set.seed(seed=1,sample.kind = 'Rounding')
+scaled_train_subset <- train_subset %>% mutate(across(c(mean_penalty_to_consider,distance_between_two_points,volume,product_weight_g,freight_value,price),scale)) 
+scaled_test_subset <- test_subset %>% mutate(across(c(mean_penalty_to_consider,distance_between_two_points,volume,product_weight_g,freight_value,price),scale))
+
+tuningControl=trainControl(method = "cv", number = 3, p = .8)
+knn_fit <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(27,34,41)),trControl=tuningControl)
+
+knn_fit
+
+predictions <- predict(knn_fit,scaled_test_subset %>% select(-number_of_hours))
+
+RMSE(predictions,(scaled_test_subset %>% pull(number_of_hours) ))
+#194.8233 for k=34
+
+save(knn_fit,file='rda/knn_model.rda')
+```
+
+Random forest
+However, it is important to know your data and keep in mind that a Random Forest can’t extrapolate.
+It can only make a prediction that is an average of previously observed labels.
+In this sense it is very similar to KNN.
+In other words, in a regression problem, the range of predictions a Random Forest can make is bound by the highest and lowest labels in the training data.
+This behavior becomes problematic in situations where the training and prediction inputs differ in their range and/or distributions.
+This is called covariate shift and it is difficult for most models to handle but especially for Random Forest, because it can’t extrapolate.
+From : https://towardsdatascience.com/a-limitation-of-random-forest-regression-db8ed7419e9f
+
+```{r rf model, echo = FALSE}
+set.seed(seed=1,sample.kind = 'Rounding')
+tuningControl=trainControl(method = "cv", number = 2, p = .8)
+rf_fit <- train(number_of_hours~., train_subset, method='rf',tuneGrid = data.frame(mtry = seq(5, 50, 10)),trControl=tuningControl,ntree=250) # nodesize = 50,maxnodes=2000,ntree=200
+
+rf_fit
+
+predictions <- predict(rf_fit,test_subset %>% select(-number_of_hours))
+
+RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
+# 184.1709 for mtry 15. Technically there are total of 79 variables due to 73 product categories and 6 other predictors
+
+save(rf_fit,file='rda/rf_model.rda')
+```
+
+```{r Gam model, echo = FALSE}
+set.seed(seed=1,sample.kind = 'Rounding')
+tuningControl=trainControl(method = "cv", number =2, p = .8)
+gam_fit <- train(number_of_hours~., train_subset, method='gam',trControl=tuningControl) # check for tuning parameters
+
+gam_fit
+
+predictions <- predict(gam_fit,test_subset %>% select(-number_of_hours))
+
+RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
+# 195.3235
+
+save(gam_fit,file='rda/gam_model.rda')
+```
+
+```{r Glm model}
+# Let's try with glm model
+set.seed(seed=1,sample.kind = 'Rounding')
+tuningControl=trainControl(method = "cv", number = 2, p = .8)
+glm_fit <- train(number_of_hours~., train_subset, method='glm',trControl=tuningControl)
+
+glm_fit
+
+predictions <- predict(glm_fit, test_subset %>% select(-number_of_hours))
+
+RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
+# 198.8575 (improves if we use seller_id as predictor as well)
+```
+Warning:
+prediction from a rank-deficient fit may be misleading
+
+```{r GamLoess model, echo = FALSE}
+set.seed(seed=1,sample.kind = 'Rounding')
+tuningControl=trainControl(method = "cv", number = 3, p = .8)
+#grid <- expand.grid(span = seq(0.15, 0.65, len = 10), degree = 1)
+grid <- expand.grid(span = seq(0.15, 0.65, len = 3), degree = 1)
+gamLoess_fit <- train(number_of_hours~., train_subset, method='gamLoess',tuneGrid=grid,trControl=tuningControl)
+
+gamLoess_fit
+
+predictions <- predict(gamLoess_fit,test_subset %>% select(-number_of_hours))
+
+RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
+#194.6215
+
+save(gamLoess_fit,file='rda/gamLoess_model.rda')
+```
+
+Warnings received:
+extrapolation not allowed with blending
+ (Discovered by lowesd)liv too small.    
+ (Discovered by lowesd)lv too small
+ extrapolation not allowed with blending
+ non-list contrasts argument ignored
+
+```{r GamLoess model with scaling, echo = FALSE}
+# TODO review-keep or remove it
+set.seed(seed=1,sample.kind = 'Rounding')
+scaled_train_subset <- train_subset %>% mutate(across(c(mean_penalty_to_consider,distance_between_two_points,volume,product_weight_g,freight_value,price),scale)) 
+scaled_test_subset <- test_subset %>% mutate(across(c(mean_penalty_to_consider,distance_between_two_points,volume,product_weight_g,freight_value,price),scale))
+
+tuningControl=trainControl(method = "cv", number = 3, p = .8)
+#grid <- expand.grid(span = seq(0.15, 0.65, len = 10), degree = 1)
+grid <- expand.grid(span = seq(0.15, 0.65, len = 3), degree = 1)
+gamLoess_fit <- train(number_of_hours~., scaled_train_subset, method='gamLoess',tuneGrid=grid,trControl=tuningControl)
+
+gamLoess_fit
+
+predictions <- predict(gamLoess_fit,scaled_test_subset %>% select(-number_of_hours))
+
+RMSE(predictions,(scaled_test_subset %>% pull(number_of_hours) ))
+#194.7017
+
+save(gamLoess_fit,file='rda/gamLoess_model.rda')
+```
+Warnings received:
+extrapolation not allowed with blending
+ (Discovered by lowesd)liv too small.    
+ (Discovered by lowesd)lv too small
+ extrapolation not allowed with blending
+ non-list contrasts argument ignored
+
+```{r Final Model Training, echo = FALSE}
+set.seed(seed=1,sample.kind = 'Rounding')
+
+penalty_data <- training_dataset %>% get_delay_in_handover_to_carrier_per_seller_id()
+
+training_dataset <- training_dataset %>% left_join(penalty_data,by='seller_id') %>% select(-seller_id,-handed_over_to_carrier_date,-shipping_limit_date)
+
+mtry_to_use <- rf_fit$finalModel$mtry
+ntree_to_use <- 500 #rf_fit$finalModel$ntree
+tuningControl <- trainControl(method = "cv", number = 2, p = .8)
+final_model <- train(number_of_hours~., training_dataset, method='rf',tuneGrid = data.frame(mtry = c(mtry_to_use)),trControl=tuningControl,ntree=ntree_to_use)
+
+# final_model <- randomForest(number_of_hours~., training_dataset,mtry= mtry_to_use, ntree=ntree_to_use)
+
+final_model
+
+save(final_model,file='rda/final_model.rda')
+```
+
+```{r Final movel validation, echo = FALSE}
+validation_dataset <- validation_dataset %>% compute_volume_using_product_info() %>% extract_distance_between_seller_customer() %>% left_join(penalty_data,by='seller_id') %>% select(-seller_id,-handed_over_to_carrier_date,-shipping_limit_date)
+
+predictions <- predict(final_model,validation_dataset %>% select(-number_of_hours))
+
+RMSE((validation_dataset %>% pull(number_of_hours) ),predictions)
+# 194.6748 for mtry=25, ntree=500
+```
+
+===== TODO review from here
+
+Data used maybe incorrect
+Review joins and if multiple orders of same type are being delivered at the same time
+refer
+> training_dataset %>% filter(number_of_hours>=354.73 & number_of_hours<354.74)
+   price freight_value product_category_name product_weight_g number_of_hours volume
+1   59.0         13.43          garden_tools             1550        354.7386  19800
+2   59.0         13.43          garden_tools             1550        354.7386  19800
+3   59.0         13.43          garden_tools             1550        354.7386  19800
+4   59.0         13.43          garden_tools             1550        354.7386  19800
+5   59.0         13.43          garden_tools             1550        354.7386  19800
+6   59.0         13.43          garden_tools             1550        354.7386  19800
+7   59.0         13.43          garden_tools             1550        354.7386  19800
+8   59.0         13.43          garden_tools             1550        354.7386  19800
+9   59.0         13.43          garden_tools             1550        354.7386  19800
+10  59.0         13.43          garden_tools             1550        354.7386  19800
+11  59.0         13.43          garden_tools             1550        354.7386  19800
+12  66.0         14.21       furniture_decor             1150        354.7319  28175
+13  46.9         17.63        sports_leisure              300        354.7378   2304
+   distance_between_two_points mean_penalty_to_consider
+1                       469.74                -51.60514
+2                       469.74                -51.60514
+3                       469.74                -51.60514
+4                       469.74                -51.60514
+5                       469.74                -51.60514
+6                       469.74                -51.60514
+7                       469.74                -51.60514
+8                       469.74                -51.60514
+9                       469.74                -51.60514
+10                      469.74                -51.60514
+11                      469.74                -51.60514
+12                      807.32                -56.57523
+13                     1083.46               -147.48749
+
+```{r Derived data}
+# seller to customer distance should be right predictor to consider, as the text says "Those merchants are able to sell their products through the Olist Store and ship them directly to the customers using Olist logistics partners."
+# source: https://www.kaggle.com/olistbr/brazilian-ecommerce
+wrangled_data %>% head() %>% extract_distance_between_seller_customer()
+
+# plot of how 
+p1 <- wrangled_data %>% extract_distance_between_seller_customer() %>% ggplot(aes(distance_between_two_points)) + geom_histogram(binwidth=50)
+ggplotly(p1)
+```
+
+```{r Explore delayed handover to carrier, echo=FALSE }
+# TODO we maybe able to remove this as it is not used, because this is the info which is generally not available
+# identify if there was delay in delivery to carrier
+wrangled_data <- wrangled_data %>% mutate(delayed_to_carrier=if_else((as.numeric(handed_over_to_carrier_date)-as.numeric(shipping_limit_date))>0,TRUE,FALSE))
+
+wrangled_data <- wrangled_data %>% select(-shipping_limit_date,-handed_over_to_carrier_date,-order_estimated_delivery_date)
 ```
@@
 # let's plot errorbar for number of hours vs seller_zip_code_prefix
-sellers_zip_of_interest <- wrangled_data %>% dplyr::count(seller_zip_code_prefix) %>% filter(n>10) %>% pull(seller_zip_code_prefix)
+sellers_zip_of_interest <- training_dataset %>% dplyr::count(seller_zip_code_prefix) %>% filter(n>10) %>% pull(seller_zip_code_prefix)
 
-wrangled_data %>% filter(seller_zip_code_prefix %in% sellers_zip_of_interest) %>% group_by(seller_zip_code_prefix) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% mutate(seller_zip_code_prefix=reorder(seller_zip_code_prefix,mean)) %>%  ggplot(aes(x=seller_zip_code_prefix,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90)) + scale_y_sqrt()
+training_dataset %>% filter(seller_zip_code_prefix %in% sellers_zip_of_interest) %>% group_by(seller_zip_code_prefix) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% mutate(seller_zip_code_prefix=reorder(seller_zip_code_prefix,mean)) %>%  ggplot(aes(x=seller_zip_code_prefix,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90)) + scale_y_sqrt()
 ```
@@
 # let's plot errorbar for number of hours vs customer_zip_code_prefix
-customers_zip_of_interest <- wrangled_data %>% dplyr::count(customer_zip_code_prefix) %>% filter(n>=5) %>% pull(customer_zip_code_prefix)
+customers_zip_of_interest <- training_dataset %>% dplyr::count(customer_zip_code_prefix) %>% filter(n>=5) %>% pull(customer_zip_code_prefix)
 
-wrangled_data %>% filter(customer_zip_code_prefix %in% customers_zip_of_interest) %>% group_by(customer_zip_code_prefix) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% mutate(customer_zip_code_prefix=reorder(customer_zip_code_prefix,mean)) %>%  ggplot(aes(x=customer_zip_code_prefix,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90)) + scale_y_sqrt()
+training_dataset %>% filter(customer_zip_code_prefix %in% customers_zip_of_interest) %>% group_by(customer_zip_code_prefix) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% mutate(customer_zip_code_prefix=reorder(customer_zip_code_prefix,mean)) %>%  ggplot(aes(x=customer_zip_code_prefix,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90)) + scale_y_sqrt()
 ```
@@
 # let's plot errorbar for number of hours vs product_id
-products_of_interest <- wrangled_data %>% dplyr::count(product_id)  %>% filter(n>=10) %>% pull(product_id)
+products_of_interest <- training_dataset %>% dplyr::count(product_id)  %>% filter(n>=10) %>% pull(product_id)
 
-wrangled_data %>% filter(product_id %in% products_of_interest) %>% group_by(product_id) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% mutate(product_id=reorder(product_id,mean)) %>%  ggplot(aes(x=product_id,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_blank())
+training_dataset %>% filter(product_id %in% products_of_interest) %>% group_by(product_id) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% mutate(product_id=reorder(product_id,mean)) %>%  ggplot(aes(x=product_id,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_blank())
 ```
@@
 
-## Explore volume of product group with number_of_hours
-
-```{r Explore volume of product group with number_of_hours, echo=FALSE}
-# let's plot errorbar for number of hours vs volume of product group
-wrangled_data %>% mutate(volume_group=round((product_length_cm*product_width_cm*product_height_cm)/1000)+1) %>% mutate(volume_group=as.factor(volume_group)) %>% group_by(volume_group) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% ggplot(aes(x=volume_group,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))
-```
-It does not influencee much the number of hours
-
-However a smooth fit shows some pattern 
-```{r smooth fit for volume, echo=FALSE}
-wrangled_data %>% mutate(volume=product_length_cm*product_width_cm*product_height_cm) %>% ggplot(aes(volume,number_of_hours)) + geom_smooth()
-```
-
-## Explore weight of product with number_of_hours
-
-```{r Explore weight of product with number_of_hours}
-# let's plot errorbar for number of hours vs weight of product
-wrangled_data %>% mutate(weight_group=round((product_weight_g)/1000)+1) %>% mutate(weight_group=as.factor(weight_group)) %>% mutate(weight_group=as.factor(weight_group)) %>% group_by(weight_group) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% ggplot(aes(x=weight_group,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))
-```
-Some clues but not strongly correlated.
-
-```{r Smooth fit for product, echo= FALSE}
-wrangled_data  %>% ggplot(aes(product_weight_g,number_of_hours)) + geom_smooth()
-```
-
-Calculating current RMSE range
-```{r Current RMSE}
-# current RMSE can be calculated by finding number of hours for predicted delivery time given by order_estimated_delivery_date vs actual outcome/number of hours give by order_delivered_customer_date
-# However, note that order_estimated_delivery_date has no timestamp, so we will remove timestamp component, round date using floor_date, which means any delivery done during the day say at 2017-10-10 21:25:13 is treated to be done on 2017-10-10
-
-# first we convert it to desired date
-# then we covert date to numeric i.e. to convert it to seconds
-orders_dataset %>% mutate(order_delivered_customer_date_low=floor_date(order_delivered_customer_date,unit="day"),order_delivered_customer_date_high=ceiling_date(order_delivered_customer_date,unit="day")) %>% mutate(order_delivered_customer_date_low=as.numeric(order_delivered_customer_date_low),order_delivered_customer_date_hugh=as.numeric(order_delivered_customer_date_high), order_estimated_delivery_date=as.numeric(order_estimated_delivery_date))  %>% head()
-
-# then we calculate time_taken_from_order_to_delivery, estimated_time_from_order_to_delivery in seconds
-data_for_rmse <- orders_dataset %>% mutate(order_delivered_customer_date_low=floor_date(order_delivered_customer_date,unit="day"),order_delivered_customer_date_high=ceiling_date(order_delivered_customer_date,unit="day"),order_purchase_timestamp=as.numeric(order_purchase_timestamp)) %>%  mutate(time_taken_from_order_to_delivery_rangemax=as.numeric(order_delivered_customer_date_high)-order_purchase_timestamp, time_taken_from_order_to_delivery_rangemin=as.numeric(order_delivered_customer_date_low)-order_purchase_timestamp, estimated_time_from_order_to_delivery = as.numeric(order_estimated_delivery_date) -order_purchase_timestamp) %>% mutate(estimated_time_from_order_to_delivery=estimated_time_from_order_to_delivery/3600,time_taken_from_order_to_delivery_rangemax = time_taken_from_order_to_delivery_rangemax/3600,time_taken_from_order_to_delivery_rangemin = time_taken_from_order_to_delivery_rangemin/3600) %>%  select(estimated_time_from_order_to_delivery,time_taken_from_order_to_delivery_rangemax,time_taken_from_order_to_delivery_rangemin)
-
-RMSE_value_based_on_max <- RMSE(data_for_rmse$time_taken_from_order_to_delivery_rangemax,data_for_rmse$estimated_time_from_order_to_delivery)
-# 357.53
-
-RMSE_value_based_on_min <- RMSE(data_for_rmse$time_taken_from_order_to_delivery_rangemin,data_for_rmse$estimated_time_from_order_to_delivery)
-# 375.408
-
-# if we look at graph we can  see there is huge difference between estimate data and actual delivery date
-# it is possible some products are bulky or require differen mode of transport like ship 
-p1 <- data_for_rmse %>% mutate(diff=estimated_time_from_order_to_delivery-1/2*(time_taken_from_order_to_delivery_rangemax+time_taken_from_order_to_delivery_rangemin)) %>% ggplot(aes(diff)) + geom_histogram(binwidth = 24*2) + xlab('Estimation error where each bar represents 2 days')
-
-library(plotly)
-ggplotly(p1)
-```
-
-Split data into training and validation set, where validation set will be used later
-```{r Split data into training and validation set}
-set.seed(1,sample.kind = 'Rounding')
-test_indices <- createDataPartition(wrangled_data$number_of_hours,times=1,p=0.2,list=FALSE)
-
-training_dataset <- wrangled_data[-test_indices,]
-validation_dataset <- wrangled_data[test_indices,]
-
-# ensure validation_dataset has products only for concerned product categories which are in training_dataset 
-validation_dataset <- validation_dataset %>% semi_join(training_dataset,by='product_category_name')
-
-rm(test_indices)
-```
-
-We use training_dataset and split into into train and test set. So that training_dataset is sufficient to fine tune the model
-```{r Split data for training}
-# Let's split training_dataset into two parts
-set.seed(1,sample.kind = 'Rounding')
-training_dataset <- training_dataset %>% mutate(volume=product_length_cm*product_height_cm*product_width_cm) %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix,-product_length_cm,-product_height_cm,-product_width_cm)
-
-test_indices <- createDataPartition(training_dataset$number_of_hours,times=1,p=0.2,list=FALSE)
-
-train_subset <- training_dataset[-test_indices,]
-test_subset <- training_dataset[test_indices,]
-# ensure test_subset has products only for concerned product categories which are in train_subset
-test_subset <- test_subset %>% semi_join(train_subset,by='product_category_name')
-
-rm(test_indices)
-```
+====TODO review above
 
@@
 
-```{r knn model, echo = FALSE}
-set.seed(seed=1,sample.kind = 'Rounding')
-#tuningControl=trainControl(method = "cv", number = 2, p = .8)
-#knn_fit <- train(number_of_hours~., train_subset, method='knn',tuneGrid=data.frame(k=c(15,21,30)),trControl=tuningControl)
-tuningControl=trainControl(method = "cv", number = 3, p = .8)
-knn_fit <- train(number_of_hours~., train_subset, method='knn',tuneGrid=data.frame(k=c(15,23,31)),trControl=tuningControl)
-
-knn_fit
-
-predictions <- predict(knn_fit,test_subset %>% select(-number_of_hours))
-
-RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
-#199.8518
-
-save(knn_fit,file='rda/knn_model.rda')
-```
-
 gam?
 gamLoess??
-
-```{r GamLoess model, echo = FALSE}
-set.seed(seed=1,sample.kind = 'Rounding')
-tuningControl=trainControl(method = "cv", number = 3, p = .8)
-#grid <- expand.grid(span = seq(0.15, 0.65, len = 10), degree = 1)
-grid <- expand.grid(span = seq(0.15, 0.65, len = 3), degree = 1)
-gamLoess_fit <- train(number_of_hours~., train_subset, method='gamLoess',tuneGrid=grid,trControl=tuningControl)
-
-gamLoess_fit
-
-predictions <- predict(gamLoess_fit,test_subset %>% select(-number_of_hours))
-
-RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
-#197.0024
-
-save(gamLoess_fit,file='rda/gamLoess_model.rda')
-```
-
-Warnings received:
-extrapolation not allowed with blending
- (Discovered by lowesd)liv too small.    
- (Discovered by lowesd)lv too small
- extrapolation not allowed with blending
- non-list contrasts argument ignored
-
-```{r Gam model, echo = FALSE}
-set.seed(seed=1,sample.kind = 'Rounding')
-tuningControl=trainControl(method = "cv", number =2, p = .8)
-gam_fit <- train(number_of_hours~., train_subset, method='gam',trControl=tuningControl) # check for tuning parameters
-
-gam_fit
-
-predictions <- predict(gam_fit,test_subset %>% select(-number_of_hours))
-
-RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
-# 197.5325
-
-save(gam_fit,file='rda/gam_model.rda')
-```
-
-Random forest
-However, it is important to know your data and keep in mind that a Random Forest can’t extrapolate.
-It can only make a prediction that is an average of previously observed labels.
-In this sense it is very similar to KNN.
-In other words, in a regression problem, the range of predictions a Random Forest can make is bound by the highest and lowest labels in the training data.
-This behavior becomes problematic in situations where the training and prediction inputs differ in their range and/or distributions.
-This is called covariate shift and it is difficult for most models to handle but especially for Random Forest, because it can’t extrapolate.
-From : https://towardsdatascience.com/a-limitation-of-random-forest-regression-db8ed7419e9f
-
-```{r rf model, echo = FALSE}
-set.seed(seed=1,sample.kind = 'Rounding')
-tuningControl=trainControl(method = "cv", number = 2, p = .8)
-rf_fit <- train(number_of_hours~., train_subset, method='rf',tuneGrid = data.frame(mtry = seq(2, 6, 2)),trControl=tuningControl) # nodesize = 50,maxnodes=2000,ntree=200
-
-rf_fit
-
-predictions <- predict(rf_fit,test_subset %>% select(-number_of_hours))
-
-RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
-# 192.9998
-
-save(rf_fit,file='rda/rf_model.rda')
-```
-
-```{r Glm model}
-# Let's try with glm model
-set.seed(seed=1,sample.kind = 'Rounding')
-tuningControl=trainControl(method = "cv", number = 2, p = .8)
-glm_fit <- train(number_of_hours~., train_subset, method='glm',trControl=tuningControl)
-
-glm_fit
-
-predictions <- predict(glm_fit, test_subset %>% select(-number_of_hours))
-
-RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
-# 199.4791 (improves if we use seller_id as predictor as well)
-```
-Warning:
-prediction from a rank-deficient fit may be misleading
 
@@
 
-penalty_data <- raw_data  %>% mutate(order_purchase_timestamp=as.numeric(order_purchase_timestamp),order_delivered_customer_date=as.numeric(order_delivered_customer_date),product_id=as.factor(product_id),seller_id=as.factor(seller_id),seller_zip_code_prefix=as.factor(seller_zip_code_prefix),customer_unique_id=as.factor(customer_unique_id),customer_zip_code_prefix=as.factor(customer_zip_code_prefix),product_category_name=as.factor(product_category_name)) %>% mutate(number_of_hours=(order_delivered_customer_date-order_purchase_timestamp)/3600) %>% select(-order_delivered_customer_date,-order_purchase_timestamp)  %>% filter(seller_id %in% seller_ids_of_interest) %>% mutate(delay= (as.numeric(handed_over_to_carrier_date) -as.numeric(shipping_limit_date))/3600) %>% group_by(seller_id) %>% summarise(mean_penalty_to_consider=mean(delay))
+penalty_data <- wrangled_data %>% get_delay_in_handover_to_carrier_per_seller_id()
 
@@
 tuningControl=trainControl(method = "cv", number = 2, p = .8)
-rf_fit_dummy <- train(number_of_hours~., dummy_train_subset, method='rf',tuneGrid = data.frame(mtry = seq(5, 7, 1)),trControl=tuningControl) # nodesize = 50,maxnodes=2000,ntree=200
+# rf_fit_dummy <- train(number_of_hours~., dummy_train_subset, method='rf',tuneGrid = data.frame(mtry = seq(5, 15, 4)),trControl=tuningControl) # nodesize = 50,maxnodes=2000,ntree=200
+rf_fit_dummy <- train(number_of_hours~., dummy_train_subset, method='rf',tuneGrid = data.frame(mtry = seq(20, 40, 8)),trControl=tuningControl) # nodesize = 50,maxnodes=2000,ntree=200
 
@@
 RMSE(predictions,(dummy_test_subset %>% pull(number_of_hours) ))
-#188.2743
+#184.0301 for mtry 13. Technically there are total of ~79 variables due to 72-74 product categories and 6 other predictors
 ```

### 2021-07-11 finalReport.Rmd
added target encoding for seller_id and feature scaling for knn

--- 2021-07-06 draftReport.Rmd
@@
 
-```{r Install or load packages,include=FALSE}
+```{r Install or load packages,include=FALSE, echo = FALSE}
 # TODO use conditional loading or installation of library
@@
 if (!require('randomForest')) install.packages('randomForest'); library('randomForest')
-if (!require('plotly')) install.packages('plotly'); library('plotly')
-if (!require('gam')) install.packages('gam'); library('gam')
+if (!require('knitr')) install.packages('knitr'); library('knitr')
 ```
@@
 - assign default category name of UNCATEGORIZED for products where same was not available
-```{r Data wrangling-products data set, include=FALSE}
+```{r Data wrangling-products data set, include=FALSE, echo=FALSE}
 products_dataset <- read.csv(file.path('data','olist_products_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) %>% select(-product_name_lenght,-product_description_lenght,-product_photos_qty)
@@
 ```{r Data wrangling result-products data set, echo=FALSE}
-products_dataset %>% head()
+products_dataset %>% head() %>% knitr::kable()
 ```
@@
 For geolocation data set
-```{r Data wrangling- geolocation data set, echo=FALSE}
+```{r Data wrangling- geolocation data set, include=FALSE, echo=FALSE}
 # geo location details
@@
 
-```{r Data wrangling-customers data set, echo=FALSE}
+```{r Data wrangling-customers data set, include=FALSE, echo=FALSE}
 # customers data set  with zip code prefix and can be used to uniquely identify a customer
@@
 
-```{r Data wrangling-sellers data set, echo=FALSE}
+```{r Data wrangling-sellers data set, echo=FALSE, include=FALSE}
 # seller data set with zip code prefix
@@
 - removed data which was delivered but all the delivery details were not available
-```{r Data wrangling-orders data set, echo=FALSE}
+```{r Data wrangling-orders data set, include=FALSE, echo=FALSE}
 # order data set with various delivery related times
@@
 # order_estimated_delivery_date --> Shows the estimated delivery date that was informed to customer at the purchase moment.
-orders_dataset <- read.csv(file.path('data','olist_orders_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) %>% rename(payment_approved_at=order_approved_at,handed_over_to_carrier_date=order_delivered_carrier_date)  %>% mutate(order_status=as.factor(order_status),order_purchase_timestamp=as_datetime(order_purchase_timestamp),payment_approved_at=as_datetime(payment_approved_at),handed_over_to_carrier_date=as_datetime(handed_over_to_carrier_date),order_delivered_customer_date=as_datetime(order_delivered_customer_date),order_estimated_delivery_date=as_datetime(order_estimated_delivery_date))
+orders_dataset <- read.csv(file.path('data','olist_orders_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) %>% 
+  rename(payment_approved_at=order_approved_at,
+         handed_over_to_carrier_date=order_delivered_carrier_date)  %>% 
+  mutate(order_status=as.factor(order_status),
+         order_purchase_timestamp=as_datetime(order_purchase_timestamp),
+         payment_approved_at=as_datetime(payment_approved_at),
+         handed_over_to_carrier_date=as_datetime(handed_over_to_carrier_date),
+         order_delivered_customer_date=as_datetime(order_delivered_customer_date),
+         order_estimated_delivery_date=as_datetime(order_estimated_delivery_date))
 
 # TODO if we are not using payment_approved_at or few fields we can remove the filter logic based on that.
-orders_dataset <- orders_dataset %>% filter(order_status=='delivered') %>% filter( !is.na(order_estimated_delivery_date ) & !is.na(order_delivered_customer_date ) & !is.na(handed_over_to_carrier_date) & !is.na(payment_approved_at) & !is.na(order_purchase_timestamp))
+orders_dataset <- orders_dataset %>% 
+  filter(order_status=='delivered') %>% 
+  filter( !is.na(order_estimated_delivery_date ) & !is.na(order_delivered_customer_date ) & !is.na(handed_over_to_carrier_date) & !is.na(payment_approved_at) & !is.na(order_purchase_timestamp))
 
@@
 
-```{r Data wrangling-order items data set, echo=FALSE}
+```{r Data wrangling-order items data set, include=FALSE, echo=FALSE}
 # orders data with item related seller info
@@
 Filter out data where sellers location details are not available
-```{r Refine sellers dataset, include=FALSE}
+```{r Refine sellers dataset, include=FALSE, echo=FALSE}
 #
@@
 ```{r Refined sellers dataset result, echo=FALSE}
-sellers_dataset %>% head()
-```
-
-```{r Refine customers data set, include=FALSE}
+sellers_dataset %>% head() %>% knitr::kable()
+```
+
+```{r Refine customers data set, include=FALSE, echo=FALSE}
 # similarly let see customers for whom we don't have data in geolocation_dataset
@@
 ```{r Refined customers dataset result, echo=FALSE}
-customers_dataset %>% head()
+customers_dataset %>% head() %>% knitr::kable()
 ```
@@
 
-```{r Combine data sets, include=FALSE}
+```{r Combine data sets, include=FALSE, echo=FALSE}
 # we can combine orders delivery information (orders_dataset) with orders items and seller related info (order_items_dataset)
-orders_dataset %>% inner_join(order_items_dataset,by='order_id') %>% head()
+# orders_dataset %>% inner_join(order_items_dataset,by='order_id') %>% head()
 
 # let's combine with seller details
-orders_dataset %>% inner_join(order_items_dataset,by='order_id') %>% inner_join(sellers_dataset,by='seller_id') %>% head()
+# orders_dataset %>% inner_join(order_items_dataset,by='order_id') %>% inner_join(sellers_dataset,by='seller_id') %>% head()
 
 # let's combine with unique customer and its location dataset
-orders_dataset %>% inner_join(order_items_dataset,by='order_id') %>% inner_join(sellers_dataset,by='seller_id') %>% inner_join(customers_dataset,by='customer_id') %>% select(-customer_id) %>% head()
+# orders_dataset %>% inner_join(order_items_dataset,by='order_id') %>% inner_join(sellers_dataset,by='seller_id') %>% inner_join(customers_dataset,by='customer_id') %>% select(-customer_id) %>% head()
 
 # let's combine with products data as well
-orders_dataset %>% inner_join(order_items_dataset,by='order_id') %>% inner_join(sellers_dataset,by='seller_id') %>% inner_join(customers_dataset,by='customer_id') %>% inner_join(products_dataset,by='product_id') %>% select(-customer_id,-order_status,-seller_city,-seller_state,-customer_city,-customer_state) %>% head()
+# orders_dataset %>% inner_join(order_items_dataset,by='order_id') %>% inner_join(sellers_dataset,by='seller_id') %>% inner_join(customers_dataset,by='customer_id') %>% inner_join(products_dataset,by='product_id') %>% select(-customer_id,-order_status,-seller_city,-seller_state,-customer_city,-customer_state) %>% head()
 
@@
          product_category_name=as.factor(product_category_name)) %>%
-  select(-customer_id,-order_status,-seller_city,-seller_state,-customer_city,-customer_state, -payment_approved_at,-order_id, -order_item_id, -product_id,-seller_zip_code_prefix,-customer_unique_id,-customer_zip_code_prefix)
+  select(-customer_id,-order_status,-seller_city,-seller_state,-customer_city,-customer_state, -payment_approved_at,-order_id,-order_item_id,-product_id,-seller_zip_code_prefix,-customer_unique_id,-customer_zip_code_prefix)
+#  select(-customer_id,-order_status,-seller_city,-seller_state,-customer_city,-customer_state, -payment_approved_at,-order_id, -order_item_id, -product_id,-seller_zip_code_prefix,-customer_unique_id,-customer_zip_code_prefix)
 ```
@@
 Calculating current RMSE range
-```{r Current RMSE, echo = FALSE}
+```{r Current RMSE, include=FALSE, echo = FALSE}
 # current RMSE can be calculated by finding number of hours for predicted delivery time given by order_estimated_delivery_date vs actual outcome/number of hours give by order_delivered_customer_date
@@
 
-# if we look at graph we can  see there is huge difference between estimate data and actual delivery date
+# if we look at graph we can  see there is huge difference between estimated data and actual delivery date
 # it is possible some products are bulky or require differen mode of transport like ship 
-p1 <- data_for_rmse %>% mutate(diff=estimated_time_from_order_to_delivery-1/2*(time_taken_from_order_to_delivery_rangemax+time_taken_from_order_to_delivery_rangemin)) %>% ggplot(aes(diff)) + geom_histogram(binwidth = 24*2) + xlab('Estimation error where each bar represents 2 days')
-
-library(plotly)
-ggplotly(p1)
+data_for_rmse %>% mutate(diff=estimated_time_from_order_to_delivery-1/2*(time_taken_from_order_to_delivery_rangemax+time_taken_from_order_to_delivery_rangemin)) %>% ggplot(aes(diff)) + geom_histogram(binwidth = 24*2) + xlab('Estimation error where each bar represents 2 days')
 
@@
 ```
-
-```{r Prepare final wrangled data set, echo=FALSE}
+```{r Range of current rmse, echo=FALSE}
+data.frame(type=character(),value=numeric()) %>% 
+  add_row(type='RMSE Max',value=RMSE_value_based_on_max) %>% 
+  add_row(type='RMSE Min',value=RMSE_value_based_on_min) %>% 
+  knitr::kable(caption='Current RMSE')
+```
+
+Final wrangled data looks like below:
+```{r Prepare final wrangled data set, include=FALSE}
 # we will predict number of hours to deliver based on various predictors.
@@
 
-wrangled_data <- wrangled_data %>% cbind.data.frame(product_category_dummyvars) %>% select(-product_category_name)
+wrangled_data <- wrangled_data %>% cbind.data.frame(product_category_dummyvars)
 
@@
 ```{r Final wrangled data,echo=FALSE}
-wrangled_data %>% head()
-```
-
-```{r Cleanup and save required dataset}
+wrangled_data %>% head() %>% knitr::kable()
+```
+
+```{r Cleanup and save required dataset, include=FALSE, echo=FALSE}
 save(wrangled_data,file='rda/wrangled_data.rda')
-rm(geolocation_dataset, order_items_dataset, orders_dataset, p1, product_category_name_translation, products_dataset, sellers_dataset, sellers_to_ignore, customers_dataset, customers_to_ignore, raw_data)
-```
+rm(geolocation_dataset, order_items_dataset, orders_dataset, product_category_name_translation, products_dataset, sellers_dataset, sellers_to_ignore, customers_dataset, customers_to_ignore, raw_data)
+```
+
+TODO
+Use target encoding for categorical variables
+Use preprocess for scaling data against training data set. Then use mean, sd from this trained data for test data.
 
@@
     group_by(seller_id) %>% 
-    summarise(mean_penalty_to_consider=mean(delay))
+    summarise(mean_delay_in_handover=mean(delay))
 }
@@
 Split data into training and validation set, where validation set will be used later
-```{r Split data into training and validation set, include = FALSE}
+```{r Split data into training and validation set, include = FALSE, echo = FALSE}
 set.seed(1,sample.kind = 'Rounding')
@@
 # ensure validation_dataset has products only for concerned product categories and seller_ids which are in training_dataset 
-# validation_dataset <- validation_dataset %>% semi_join(training_dataset,by='product_category_name') %>% semi_join(training_dataset,by='seller_id')
-validation_dataset <- validation_dataset %>% semi_join(training_dataset,by='seller_id')
+validation_dataset <- validation_dataset %>%
+  semi_join(training_dataset,by='product_category_name') %>%
+  semi_join(training_dataset,by='seller_id')
+# validation_dataset <- validation_dataset %>% semi_join(training_dataset,by='seller_id')
 
@@
 # ensure test_subset has products only for concerned product categories and seller_ids which are in train_subset
-# test_subset <- test_subset %>% semi_join(train_subset,by='product_category_name') %>% semi_join(train_subset,by='seller_id')
-test_subset <- test_subset %>% semi_join(train_subset,by='seller_id')
+test_subset <- test_subset %>% semi_join(train_subset,by='product_category_name') %>% semi_join(train_subset,by='seller_id')
+#test_subset <- test_subset %>% semi_join(train_subset,by='seller_id')
 
@@
 ```{r Explore product_category_name relation with number_of_hours, echo=FALSE}
-# TODO review it, fix it as required
 training_dataset %>% group_by(product_category_name) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% mutate(product_category_name=reorder(product_category_name,mean)) %>% ggplot(aes(x=product_category_name,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))
@@
 # penalty_data is computed on train_subset and is used with test_subset as historical learning of efficiency of each seller in delivering to carrier
-train_subset <- train_subset %>% left_join(penalty_data,by='seller_id') %>% select(-handed_over_to_carrier_date, -shipping_limit_date, -seller_id)
-
-test_subset <- test_subset %>% left_join(penalty_data,by='seller_id') %>% select(-handed_over_to_carrier_date, -shipping_limit_date, -seller_id)
+train_subset <- train_subset %>% left_join(penalty_data,by='seller_id') %>% select(-handed_over_to_carrier_date, -shipping_limit_date)
+
+test_subset <- test_subset %>% left_join(penalty_data,by='seller_id') %>% select(-handed_over_to_carrier_date, -shipping_limit_date)
 ```
@@
 ```{r knn model, echo = FALSE}
+#TODO can be removed
 set.seed(seed=1,sample.kind = 'Rounding')
@@
 
-```{r knn model data scaline under test, echo=FALSE}
+```{r Apply Target encoding, echo = FALSE}
+  train_subset_for_encoding <- train_subset[1:30000,]
+  global_mean <- mean(train_subset_for_encoding$number_of_hours)
+
+  set.seed(1,sample.kind='Rounding')
+  test_weight_indices <- createDataPartition(train_subset_for_encoding$number_of_hours,p=0.2,times=3,list=FALSE)
+
+  weights <- seq(0,6,3)
+
+  weight_to_use_per_partition <- sapply(1:3, function(index) {
+  #lambda_data_set <- prepare_lambda_data_set(index)
+  #train_lambda_set <- lambda_data_set[[1]]
+  #test_lambda_set <- lambda_data_set[[2]]  
+  train_weight_set <- train_subset_for_encoding[-test_weight_indices[,index],]
+  test_weight_set <- train_subset_for_encoding[test_weight_indices[,index],]
+
+  test_weight_set <- test_weight_set %>%
+    semi_join(train_weight_set,by='product_category_name') %>%
+    semi_join(train_weight_set,by='seller_id')
+
+  weight_rmses <- sapply(weights, function(weight) {
+      seller_id_effects <- train_weight_set %>% 
+        group_by(seller_id) %>% 
+        summarise(mean=mean(number_of_hours),n=n()) %>%
+        mutate(mean=(n*mean+weight*global_mean)/(n+weight)) %>% 
+        select(-n) %>% rename()
+
+      train_weight_set <- train_weight_set %>% left_join(seller_id_effects,by='seller_id') %>% rename(seller_id_effect=mean) %>% select(-product_category_name,-seller_id)
+
+      test_weight_set <- test_weight_set %>% left_join(seller_id_effects,by='seller_id') %>% rename(seller_id_effect=mean) %>% select(-product_category_name,-seller_id)
+
+      set.seed(seed=1,sample.kind = 'Rounding')
+      preProcessScalingValues <- preProcess(train_weight_set %>% select(-number_of_hours), method = c("center", "scale"))
+
+      scaled_train_subset <- predict(preProcessScalingValues, train_weight_set)
+      scaled_test_subset <- predict(preProcessScalingValues, test_weight_set)
+
+      # TODO remove below scaling was erroneous. train scaling needs to be used for test data
+      # below is just demo of how to use scale
+      # scaled_train_subset <- train_subset %>% mutate(across(c(mean_delay_in_handover,distance_between_two_points,volume,product_weight_g,freight_value,price),scale)) 
+      # scaled_test_subset <- test_subset %>% mutate(across(c(mean_delay_in_handover,distance_between_two_points,volume,product_weight_g,freight_value,price),scale))
+
+      # TODO can we skip modelling while evaluating weight to use or for target encoding?
+
+      tuningControl=trainControl(method = "cv", number = 2, p = .8)
+      knn_fit_for_weight <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(27,34,41)),trControl=tuningControl)
+
+      knn_fit_for_weight
+
+      predictions <- predict(knn_fit_for_weight,scaled_test_subset %>% select(-number_of_hours))
+
+      RMSE(scaled_test_subset$number_of_hours,predictions) 
+  })
+
+  # selecting weight leading to lowest RMSE
+  weight_to_use <- weights[which.min(weight_rmses)]
+  paste('Weight to use: ', weight_to_use, ', corresponding RMSE: ', weight_rmses[which.min(weight_rmses)])
+
+  print(weight_rmses)
+  weight_to_use
+})
+
+weight_to_use_per_partition
+
+weight_to_use <- mean(weight_to_use_per_partition)  
+```
+
+```{r Post regularization, echo=FALSE}
+
+# prepare seller_if_effect based on regularisation weight we have identified
+  global_mean <- mean(train_subset$number_of_hours)
+
+  set.seed(1,sample.kind='Rounding')
+  test_weight_indices <- createDataPartition(train_subset$number_of_hours,p=0.2,times=5,list=FALSE)
+
+seller_id_effects_list <- lapply(1:5, function(index) {
+    train_weight_set <- train_subset[-test_weight_indices[,index],]
+
+    seller_id_effects <- train_weight_set %>% 
+      group_by(seller_id) %>%
+      summarise(mean=mean(number_of_hours),n=n()) %>%
+      mutate(mean=(n*mean+weight_to_use*global_mean)/(n+weight_to_use)) %>% 
+      select(-n) %>% rename()
+
+    seller_id_effects
+  }
+)
+
+sie <- seller_id_effects_list[[1]]
+for (index in 2:5) {
+   sie <- sie %>% full_join(seller_id_effects_list[[index]],by='seller_id', suffix=c("1",index))
+}
+
+# whereever value is NA, we have used global mean
+sie <- sie %>% mutate(across(-seller_id,~if_else(is.na(.x),global_mean,.x)))
+
+# as different partitions may not have exactly same data set combined the results
+seller_id_effects <- tibble(sie[,1], seller_id_effect=rowMeans(sie[,-1],na.rm = TRUE))
+
+train_subset <- train_subset %>% left_join(seller_id_effects,by='seller_id') %>% select(-product_category_name,-seller_id) %>% mutate(seller_id_effect=if_else(is.na(seller_id_effect),global_mean,seller_id_effect))
+
+test_subset <- test_subset %>% left_join(seller_id_effects,by='seller_id') %>% select(-product_category_name,-seller_id)
+```
+
+```{r knn model data scaled under test, echo=FALSE}
 # TODO review and improve or remove this as required
 set.seed(seed=1,sample.kind = 'Rounding')
-scaled_train_subset <- train_subset %>% mutate(across(c(mean_penalty_to_consider,distance_between_two_points,volume,product_weight_g,freight_value,price),scale)) 
-scaled_test_subset <- test_subset %>% mutate(across(c(mean_penalty_to_consider,distance_between_two_points,volume,product_weight_g,freight_value,price),scale))
+preProcessScalingValues <- preProcess(train_subset %>% select(-number_of_hours), method = c("center", "scale"))
+
+scaled_train_subset <- predict(preProcessScalingValues, train_subset)
+scaled_test_subset <- predict(preProcessScalingValues, test_subset)
+
+# TODO remove below scaling was erroneous. train scaling needs to be used for test data
+# below is just demo of how to use scale
+# scaled_train_subset <- train_subset %>% mutate(across(c(mean_delay_in_handover,distance_between_two_points,volume,product_weight_g,freight_value,price),scale)) 
+# scaled_test_subset <- test_subset %>% mutate(across(c(mean_delay_in_handover,distance_between_two_points,volume,product_weight_g,freight_value,price),scale))
 
 tuningControl=trainControl(method = "cv", number = 3, p = .8)
-knn_fit <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(27,34,41)),trControl=tuningControl)
+knn_fit <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(20,27,34)),trControl=tuningControl)
 
@@
 RMSE(predictions,(scaled_test_subset %>% pull(number_of_hours) ))
-#194.8233 for k=34
+#192.8755 for k=27
 
@@
 set.seed(seed=1,sample.kind = 'Rounding')
-scaled_train_subset <- train_subset %>% mutate(across(c(mean_penalty_to_consider,distance_between_two_points,volume,product_weight_g,freight_value,price),scale)) 
-scaled_test_subset <- test_subset %>% mutate(across(c(mean_penalty_to_consider,distance_between_two_points,volume,product_weight_g,freight_value,price),scale))
+scaled_train_subset <- train_subset %>% mutate(across(c(mean_delay_in_handover,distance_between_two_points,volume,product_weight_g,freight_value,price),scale)) 
+scaled_test_subset <- test_subset %>% mutate(across(c(mean_delay_in_handover,distance_between_two_points,volume,product_weight_g,freight_value,price),scale))
 
@@
 13  46.9         17.63        sports_leisure              300        354.7378   2304
-   distance_between_two_points mean_penalty_to_consider
+   distance_between_two_points mean_delay_in_handover
 1                       469.74                -51.60514
@@
 # this penalty data can later be reused with test data as well
-penalty_data <- dummy_train_subset %>% group_by(seller_id) %>% summarise(mean_penalty_to_consider=mean(delay))
+penalty_data <- dummy_train_subset %>% group_by(seller_id) %>% summarise(mean_delay_in_handover=mean(delay))
 

### 2021-07-11 finalReport.Rmd
updated code and tested for random forest with seller_id changes

@@
 
-```{r knn model, echo = FALSE}
-#TODO can be removed
-set.seed(seed=1,sample.kind = 'Rounding')
-
-tuningControl=trainControl(method = "cv", number = 3, p = .8)
-knn_fit <- train(number_of_hours~., train_subset, method='knn',tuneGrid=data.frame(k=c(27,34,41)),trControl=tuningControl)
-
-knn_fit
-
-predictions <- predict(knn_fit,test_subset %>% select(-number_of_hours))
-
-RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
-#199.448 for k=34
-
-save(knn_fit,file='rda/knn_model.rda')
-```
-
 ```{r Apply Target encoding, echo = FALSE}
@@
 RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
-# 184.1709 for mtry 15. Technically there are total of 79 variables due to 73 product categories and 6 other predictors
+# 184.5549 for mtry 15. Technically there are total of 79 variables due to 73 product categories and 6 other predictors
 

### 2021-07-12 finalReport.Rmd
updated with target encoding for customer starte, seller state, seller id.

@@
          product_category_name=as.factor(product_category_name)) %>%
-  select(-customer_id,-order_status,-seller_city,-seller_state,-customer_city,-customer_state, -payment_approved_at,-order_id,-order_item_id,-product_id,-seller_zip_code_prefix,-customer_unique_id,-customer_zip_code_prefix)
+  select(-customer_id,-order_status,-seller_city,-customer_city, -payment_approved_at,-order_id,-order_item_id,-product_id,-seller_zip_code_prefix,-customer_unique_id,-customer_zip_code_prefix)
 #  select(-customer_id,-order_status,-seller_city,-seller_state,-customer_city,-customer_state, -payment_approved_at,-order_id, -order_item_id, -product_id,-seller_zip_code_prefix,-customer_unique_id,-customer_zip_code_prefix)
@@
 
-```{r Apply Target encoding, echo = FALSE}
+==TODO note penalty data should be calculated again if train_subset is changed. Only trained data should be used for test, so that there is no data leak.
+
+```{r Apply Target encoding for seller_id, echo = FALSE}
+#TODO you may need to remove customer_state and seller_state from dataset train_subset_for_encoding before proceeding further.
+
   train_subset_for_encoding <- train_subset[1:30000,]
@@
 weight_to_use <- mean(weight_to_use_per_partition)  
-```
-
-```{r Post regularization, echo=FALSE}
+# 3
+```
+
+```{r Post regularization for seller_id effect, echo=FALSE}
 
@@
 
+```{r Apply Target encoding for customer_state, echo = FALSE}
+  train_subset_for_encoding <- train_subset[1:40000,] %>% select(-seller_state)
+  global_mean <- mean(train_subset_for_encoding$number_of_hours)
+
+  set.seed(1,sample.kind='Rounding')
+  test_weight_indices <- createDataPartition(train_subset_for_encoding$number_of_hours,p=0.2,times=3,list=FALSE)
+
+  weights <- seq(0,30,15)
+
+  weight_to_use_per_partition <- sapply(1:3, function(index) {
+  #lambda_data_set <- prepare_lambda_data_set(index)
+  #train_lambda_set <- lambda_data_set[[1]]
+  #test_lambda_set <- lambda_data_set[[2]]  
+  train_weight_set <- train_subset_for_encoding[-test_weight_indices[,index],]
+  test_weight_set <- train_subset_for_encoding[test_weight_indices[,index],]
+
+  test_weight_set <- test_weight_set %>%
+    #semi_join(train_weight_set,by='product_category_name') %>%
+    semi_join(train_weight_set,by='customer_state')
+
+  weight_rmses <- sapply(weights, function(weight) {
+      customer_state_effects <- train_weight_set %>% 
+        group_by(customer_state) %>% 
+        summarise(mean=mean(number_of_hours),n=n()) %>%
+        mutate(mean=(n*mean+weight*global_mean)/(n+weight)) %>% 
+        select(-n) %>% rename()
+
+      train_weight_set <- train_weight_set %>% left_join(customer_state_effects,by='customer_state') %>% rename(customer_state_effect=mean) %>% select(-customer_state) #-product_category_name
+
+      test_weight_set <- test_weight_set %>% left_join(customer_state_effects,by='customer_state') %>% rename(customer_state_effect=mean) %>% select(-customer_state) # -product_category_name,
+
+      set.seed(seed=1,sample.kind = 'Rounding')
+      preProcessScalingValues <- preProcess(train_weight_set %>% select(-number_of_hours), method = c("center", "scale"))
+
+      scaled_train_subset <- predict(preProcessScalingValues, train_weight_set)
+      scaled_test_subset <- predict(preProcessScalingValues, test_weight_set)
+
+      # TODO remove below scaling was erroneous. train scaling needs to be used for test data
+      # below is just demo of how to use scale
+      # scaled_train_subset <- train_subset %>% mutate(across(c(mean_delay_in_handover,distance_between_two_points,volume,product_weight_g,freight_value,price),scale)) 
+      # scaled_test_subset <- test_subset %>% mutate(across(c(mean_delay_in_handover,distance_between_two_points,volume,product_weight_g,freight_value,price),scale))
+
+      # TODO can we skip modelling while evaluating weight to use or for target encoding?
+
+      tuningControl=trainControl(method = "cv", number = 2, p = .8)
+      knn_fit_for_weight <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(20,27,34)),trControl=tuningControl)
+
+      knn_fit_for_weight
+
+      print(knn_fit_for_weight)
+
+      predictions <- predict(knn_fit_for_weight,scaled_test_subset %>% select(-number_of_hours))
+
+      RMSE(scaled_test_subset$number_of_hours,predictions) 
+  })
+
+  # selecting weight leading to lowest RMSE
+  weight_to_use <- weights[which.min(weight_rmses)]
+  paste('Weight to use: ', weight_to_use, ', corresponding RMSE: ', weight_rmses[which.min(weight_rmses)])
+
+  print(weight_rmses)
+  weight_to_use
+})
+
+weight_to_use_per_partition
+
+weight_to_use <- mean(weight_to_use_per_partition) 
+# 20
+```
+
+```{r Post regularization with consumer state effect, echo=FALSE}
+
+# prepare consumer_state_effect based on regularisation weight we have identified
+  global_mean <- mean(train_subset$number_of_hours)
+
+  set.seed(1,sample.kind='Rounding')
+  test_weight_indices <- createDataPartition(train_subset$number_of_hours,p=0.2,times=5,list=FALSE)
+
+customer_state_effects_list <- lapply(1:5, function(index) {
+    train_weight_set <- train_subset[-test_weight_indices[,index],]
+
+    customer_state_effects <- train_weight_set %>% 
+      group_by(customer_state) %>%
+      summarise(mean=mean(number_of_hours),n=n()) %>%
+      mutate(mean=(n*mean+weight_to_use*global_mean)/(n+weight_to_use)) %>% 
+      select(-n) %>% rename()
+
+    customer_state_effects
+  }
+)
+
+cse <- customer_state_effects_list[[1]]
+for (index in 2:5) {
+   cse <- cse %>% full_join(customer_state_effects_list[[index]],by='customer_state', suffix=c("1",index))
+}
+
+# whereever value is NA, we have used global mean
+cse <- cse %>% mutate(across(-customer_state,~if_else(is.na(.x),global_mean,.x)))
+
+# as different partitions may not have exactly same data set combined the results
+customer_state_effects <- tibble(cse[,1], customer_state_effect=rowMeans(cse[,-1],na.rm = TRUE))
+
+train_subset <- train_subset %>% left_join(customer_state_effects,by='customer_state') %>% select(-customer_state) %>% mutate(customer_state_effect=if_else(is.na(customer_state_effect),global_mean,customer_state_effect))
+
+test_subset <- test_subset %>% left_join(customer_state_effects,by='customer_state') %>% select(-customer_state)
+```
+
+```{r Apply Target encoding for seller_state, echo = FALSE}
+  train_subset_for_encoding <- train_subset[1:40000,] 
+  global_mean <- mean(train_subset_for_encoding$number_of_hours)
+
+  set.seed(1,sample.kind='Rounding')
+  test_weight_indices <- createDataPartition(train_subset_for_encoding$number_of_hours,p=0.2,times=3,list=FALSE)
+
+  weights <- seq(0,30,15)
+
+  weight_to_use_per_partition <- sapply(1:3, function(index) {
+  #lambda_data_set <- prepare_lambda_data_set(index)
+  #train_lambda_set <- lambda_data_set[[1]]
+  #test_lambda_set <- lambda_data_set[[2]]  
+  train_weight_set <- train_subset_for_encoding[-test_weight_indices[,index],]
+  test_weight_set <- train_subset_for_encoding[test_weight_indices[,index],]
+
+  test_weight_set <- test_weight_set %>%
+    #semi_join(train_weight_set,by='product_category_name') %>%
+    semi_join(train_weight_set,by='seller_state')
+
+  weight_rmses <- sapply(weights, function(weight) {
+      seller_state_effects <- train_weight_set %>% 
+        group_by(seller_state) %>% 
+        summarise(mean=mean(number_of_hours),n=n()) %>%
+        mutate(mean=(n*mean+weight*global_mean)/(n+weight)) %>% 
+        select(-n) %>% rename()
+
+      train_weight_set <- train_weight_set %>% left_join(seller_state_effects,by='seller_state') %>% rename(seller_state_effect=mean) %>% select(-seller_state) #-product_category_name
+
+      test_weight_set <- test_weight_set %>% left_join(seller_state_effects,by='seller_state') %>% rename(seller_state_effect=mean) %>% select(-seller_state) # -product_category_name,
+
+      set.seed(seed=1,sample.kind = 'Rounding')
+      preProcessScalingValues <- preProcess(train_weight_set %>% select(-number_of_hours), method = c("center", "scale"))
+
+      scaled_train_subset <- predict(preProcessScalingValues, train_weight_set)
+      scaled_test_subset <- predict(preProcessScalingValues, test_weight_set)
+
+      # TODO remove below scaling was erroneous. train scaling needs to be used for test data
+      # below is just demo of how to use scale
+      # scaled_train_subset <- train_subset %>% mutate(across(c(mean_delay_in_handover,distance_between_two_points,volume,product_weight_g,freight_value,price),scale)) 
+      # scaled_test_subset <- test_subset %>% mutate(across(c(mean_delay_in_handover,distance_between_two_points,volume,product_weight_g,freight_value,price),scale))
+
+      # TODO can we skip modelling while evaluating weight to use or for target encoding?
+
+      tuningControl=trainControl(method = "cv", number = 2, p = .8)
+      knn_fit_for_weight <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(20,27,34)),trControl=tuningControl)
+
+      knn_fit_for_weight
+
+      print(knn_fit_for_weight)
+
+      predictions <- predict(knn_fit_for_weight,scaled_test_subset %>% select(-number_of_hours))
+
+      RMSE(scaled_test_subset$number_of_hours,predictions) 
+  })
+
+  # selecting weight leading to lowest RMSE
+  weight_to_use <- weights[which.min(weight_rmses)]
+  paste('Weight to use: ', weight_to_use, ', corresponding RMSE: ', weight_rmses[which.min(weight_rmses)])
+
+  print(weight_rmses)
+  weight_to_use
+})
+
+weight_to_use_per_partition
+
+weight_to_use <- mean(weight_to_use_per_partition)  
+# 15
+```
+
+```{r Post regularization with seller state effect, echo=FALSE}
+
+# prepare consumer_state_effect based on regularisation weight we have identified
+  global_mean <- mean(train_subset$number_of_hours)
+
+  set.seed(1,sample.kind='Rounding')
+  test_weight_indices <- createDataPartition(train_subset$number_of_hours,p=0.2,times=10,list=FALSE)
+
+seller_state_effects_list <- lapply(1:10, function(index) {
+    train_weight_set <- train_subset[-test_weight_indices[,index],]
+
+    seller_state_effects <- train_weight_set %>% 
+      group_by(seller_state) %>%
+      summarise(mean=mean(number_of_hours),n=n()) %>%
+      mutate(mean=(n*mean+weight_to_use*global_mean)/(n+weight_to_use)) %>% 
+      select(-n) %>% rename()
+
+    seller_state_effects
+  }
+)
+
+sse <- seller_state_effects_list[[1]]
+for (index in 2:10) {
+   sse <- sse %>% full_join(seller_state_effects_list[[index]],by='seller_state', suffix=c("1",index))
+}
+
+# whereever value is NA, we have used global mean
+sse <- sse %>% mutate(across(-seller_state,~if_else(is.na(.x),global_mean,.x)))
+
+# as different partitions may not have exactly same data set combined the results
+seller_state_effects <- tibble(sse[,1], seller_state_effect=rowMeans(sse[,-1],na.rm = TRUE))
+
+train_subset <- train_subset %>% left_join(seller_state_effects,by='seller_state') %>% select(-seller_state) %>% mutate(seller_state_effect=if_else(is.na(seller_state_effect),global_mean,seller_state_effect))
+
+test_subset <- test_subset %>% left_join(seller_state_effects,by='seller_state') %>% select(-seller_state)
+```
+
 ```{r knn model data scaled under test, echo=FALSE}
@@
 tuningControl=trainControl(method = "cv", number = 3, p = .8)
-knn_fit <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(20,27,34)),trControl=tuningControl)
+knn_fit <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(27,34,41)),trControl=tuningControl)
 
@@
 RMSE(predictions,(scaled_test_subset %>% pull(number_of_hours) ))
-#192.8755 for k=27
+#190.243 for k=41. The seller_state is not of much help. it decrease RMSE by .14
 
@@
 tuningControl=trainControl(method = "cv", number = 2, p = .8)
-rf_fit <- train(number_of_hours~., train_subset, method='rf',tuneGrid = data.frame(mtry = seq(5, 50, 10)),trControl=tuningControl,ntree=250) # nodesize = 50,maxnodes=2000,ntree=200
+rf_fit <- train(number_of_hours~., train_subset, method='rf',tuneGrid = data.frame(mtry = seq(5, 50, 10)),trControl=tuningControl,ntree=250)
+#rf_fit <- train(number_of_hours~., train_subset %>% select(-seller_state), method='rf',tuneGrid = data.frame(mtry = seq(5, 50, 10)),trControl=tuningControl,ntree=250) # nodesize = 50,maxnodes=2000,ntree=200
 
@@
 RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
-# 184.5549 for mtry 15. Technically there are total of 79 variables due to 73 product categories and 6 other predictors
+# 181.8166 for mtry 15.
+Technically there are total of 82 variables due to 73 product categories and 9 other predictors.
+The seller_state is not of much help.
+it increases RMSE by .6.
+So we should consider ignoring it.
 

### 2021-07-13 finalReport.Rmd
tested model with customer loc effect aand there is improvement in the model knn, gam and rf

@@
          product_category_name=as.factor(product_category_name)) %>%
-  select(-customer_id,-order_status,-seller_city,-customer_city, -payment_approved_at,-order_id,-order_item_id,-product_id,-seller_zip_code_prefix,-customer_unique_id,-customer_zip_code_prefix)
-#  select(-customer_id,-order_status,-seller_city,-seller_state,-customer_city,-customer_state, -payment_approved_at,-order_id, -order_item_id, -product_id,-seller_zip_code_prefix,-customer_unique_id,-customer_zip_code_prefix)
+  select(-customer_id,-order_status,-seller_city,-seller_state,-customer_city, -payment_approved_at,-order_id,-order_item_id,-product_id,-seller_zip_code_prefix,-customer_unique_id)
 ```
@@
 save(wrangled_data,file='rda/wrangled_data.rda')
-rm(geolocation_dataset, order_items_dataset, orders_dataset, product_category_name_translation, products_dataset, sellers_dataset, sellers_to_ignore, customers_dataset, customers_to_ignore, raw_data)
+rm(geolocation_dataset, order_items_dataset, orders_dataset, product_category_name_translation, products_dataset, sellers_dataset, sellers_to_ignore, customers_dataset, customers_to_ignore)
+# TODO delete raw_data as well.
 ```
@@
 
+```{r Method to reduce levels in customer_zip_code_prefix, include = FALSE}
+# trimming zip code to three digits to be able to extract zip code component which represents region, sub-region and sector data.
+compute_rough_customer_location <- function(df) {
+  df %>% mutate(customer_zip_code_prefix=str_pad(customer_zip_code_prefix, c(3),pad=c('0'),side='right')) %>% 
+    mutate(customer_loc = as.numeric(str_extract(customer_zip_code_prefix,"\\d\\d\\d"))) 
+  #%>% select(-customer_zip_code_prefix)
+}
+```
+
 ```{r Method to derive delay in handover to carrier, include = FALSE}
@@
   semi_join(training_dataset,by='product_category_name') %>%
-  semi_join(training_dataset,by='seller_id')
+  semi_join(training_dataset,by='seller_id') %>% 
+  semi_join(training_dataset,by="customer_zip_code_prefix")
+## TODO review if we want to change this join
 # validation_dataset <- validation_dataset %>% semi_join(training_dataset,by='seller_id')
@@
 set.seed(1,sample.kind = 'Rounding')
-training_dataset <- training_dataset %>% compute_volume_using_product_info() %>% extract_distance_between_seller_customer()
+training_dataset <- training_dataset %>% compute_volume_using_product_info() %>% extract_distance_between_seller_customer() %>% compute_rough_customer_location()
 
@@
 # ensure test_subset has products only for concerned product categories and seller_ids which are in train_subset
-test_subset <- test_subset %>% semi_join(train_subset,by='product_category_name') %>% semi_join(train_subset,by='seller_id')
+test_subset <- test_subset %>% semi_join(train_subset,by='product_category_name') %>% semi_join(train_subset,by='seller_id') %>% 
+  semi_join(train_subset,by="customer_zip_code_prefix")
 #test_subset <- test_subset %>% semi_join(train_subset,by='seller_id')
@@
 
+## Explore truncated customer_zip_code_prefix with number_of_hours
+
+```{r Explore truncated customer_zip_code_prefix with number_of_hours}
+# let's plot errorbar for number of hours vs truncated customer_zip_code_prefix
+training_dataset %>% mutate(czc = as.numeric(str_extract(customer_zip_code_prefix,"\\d\\d\\d"))) %>% group_by(czc) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% mutate(czc=reorder(czc,mean)) %>% ggplot(aes(x=czc,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))
+```
+
 ```{r Prepare training and test data for ,include=FALSE}
@@
 
-  train_subset_for_encoding <- train_subset[1:30000,]
+  train_subset_for_encoding <- train_subset[1:30000,] %>% select(-customer_loc, -customer_zip_code_prefix, -customer_state)
+
   global_mean <- mean(train_subset_for_encoding$number_of_hours)
@@
 ```{r Apply Target encoding for customer_state, echo = FALSE}
-  train_subset_for_encoding <- train_subset[1:40000,] %>% select(-seller_state)
+  train_subset_for_encoding <- train_subset[1:40000,] %>% select(-customer_loc,-customer_zip_code_prefix)
   global_mean <- mean(train_subset_for_encoding$number_of_hours)
@@
 
-```{r Apply Target encoding for seller_state, echo = FALSE}
-  train_subset_for_encoding <- train_subset[1:40000,] 
+```{r Apply Target encoding for customer_loc, echo = FALSE}
+  train_subset_for_encoding <- train_subset[1:40000,] %>% select(-customer_zip_code_prefix)
   global_mean <- mean(train_subset_for_encoding$number_of_hours)
@@
 
-  weights <- seq(0,30,15)
+  weights <- seq(0,10,5)
 
@@
     #semi_join(train_weight_set,by='product_category_name') %>%
-    semi_join(train_weight_set,by='seller_state')
+    semi_join(train_weight_set,by='customer_loc')
 
   weight_rmses <- sapply(weights, function(weight) {
-      seller_state_effects <- train_weight_set %>% 
-        group_by(seller_state) %>% 
+      customer_loc_effects <- train_weight_set %>% 
+        group_by(customer_loc) %>% 
         summarise(mean=mean(number_of_hours),n=n()) %>%
@@
 
-      train_weight_set <- train_weight_set %>% left_join(seller_state_effects,by='seller_state') %>% rename(seller_state_effect=mean) %>% select(-seller_state) #-product_category_name
-
-      test_weight_set <- test_weight_set %>% left_join(seller_state_effects,by='seller_state') %>% rename(seller_state_effect=mean) %>% select(-seller_state) # -product_category_name,
+      train_weight_set <- train_weight_set %>% left_join(customer_loc_effects,by='customer_loc') %>% rename(customer_loc_effect=mean) %>% select(-customer_loc) #-product_category_name
+
+      test_weight_set <- test_weight_set %>% left_join(customer_loc_effects,by='customer_loc') %>% rename(customer_loc_effect=mean) %>% select(-customer_loc) # -product_category_name,
 
@@
 weight_to_use <- mean(weight_to_use_per_partition) 
-# 15
-```
-
-```{r Post regularization with seller state effect, echo=FALSE}
-
-# prepare consumer_state_effect based on regularisation weight we have identified
+# 5/3 = 1.66667
+```
+
+```{r Post regularization with customer loc effect, echo=FALSE}
+
+# prepare customer_loc_effect based on regularisation weight we have identified
   global_mean <- mean(train_subset$number_of_hours)
@@
 
-seller_state_effects_list <- lapply(1:10, function(index) {
+customer_loc_effects_list <- lapply(1:10, function(index) {
     train_weight_set <- train_subset[-test_weight_indices[,index],]
 
-    seller_state_effects <- train_weight_set %>% 
-      group_by(seller_state) %>%
+    customer_loc_effects <- train_weight_set %>% 
+      group_by(customer_loc) %>%
       summarise(mean=mean(number_of_hours),n=n()) %>%
@@
 
-    seller_state_effects
+    customer_loc_effects
   }
@@
 
-sse <- seller_state_effects_list[[1]]
+cle <- customer_loc_effects_list[[1]]
 for (index in 2:10) {
-   sse <- sse %>% full_join(seller_state_effects_list[[index]],by='seller_state', suffix=c("1",index))
+   cle <- cle %>% full_join(customer_loc_effects_list[[index]],by='customer_loc', suffix=c("1",index))
 }
@@
 # whereever value is NA, we have used global mean
-sse <- sse %>% mutate(across(-seller_state,~if_else(is.na(.x),global_mean,.x)))
+cle <- cle %>% mutate(across(-customer_loc,~if_else(is.na(.x),global_mean,.x)))
 
 # as different partitions may not have exactly same data set combined the results
-seller_state_effects <- tibble(sse[,1], seller_state_effect=rowMeans(sse[,-1],na.rm = TRUE))
-
-train_subset <- train_subset %>% left_join(seller_state_effects,by='seller_state') %>% select(-seller_state) %>% mutate(seller_state_effect=if_else(is.na(seller_state_effect),global_mean,seller_state_effect))
-
-test_subset <- test_subset %>% left_join(seller_state_effects,by='seller_state') %>% select(-seller_state)
+customer_loc_effects <- tibble(cle[,1], customer_loc_effect=rowMeans(cle[,-1],na.rm = TRUE))
+
+train_subset <- train_subset %>% left_join(customer_loc_effects,by='customer_loc') %>% select(-customer_loc,-customer_zip_code_prefix) %>% mutate(customer_loc_effect=if_else(is.na(customer_loc_effect),global_mean,customer_loc_effect))
+
+test_subset <- test_subset %>% left_join(customer_loc_effects,by='customer_loc') %>% select(-customer_loc,-customer_zip_code_prefix)
 ```
@@
 RMSE(predictions,(scaled_test_subset %>% pull(number_of_hours) ))
-#190.243 for k=41. The seller_state is not of much help. it decrease RMSE by .14
+#184.7906 for k=34.
 
 save(knn_fit,file='rda/knn_model.rda')
+```
+
+```{r correlation, eval=FALSE}
+cor(train_subset %>% select(price,freight_value,customer_loc_effect,customer_state_effect,seller_id_effect,distance_between_two_points,mean_delay_in_handover,volume,number_of_hours,product_weight_g),use="pairwise.complete.obs") %>% knitr::kable()
 ```
@@
 tuningControl=trainControl(method = "cv", number = 2, p = .8)
-rf_fit <- train(number_of_hours~., train_subset, method='rf',tuneGrid = data.frame(mtry = seq(5, 50, 10)),trControl=tuningControl,ntree=250)
+rf_fit <- train(number_of_hours~., train_subset, method='rf',tuneGrid = data.frame(mtry = seq(5, 40, 10)),trControl=tuningControl,ntree=250)
 #rf_fit <- train(number_of_hours~., train_subset %>% select(-seller_state), method='rf',tuneGrid = data.frame(mtry = seq(5, 50, 10)),trControl=tuningControl,ntree=250) # nodesize = 50,maxnodes=2000,ntree=200
@@
 RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
-# 181.8166 for mtry 15.
+# 175.0988 for mtry 15.
 Technically there are total of 82 variables due to 73 product categories and 9 other predictors.
-The seller_state is not of much help.
-it increases RMSE by .6.
-So we should consider ignoring it.
+The seller_state was not of much help.
+it was actually increasing RMSE by .6.
+So have ignored it.
 
@@
 set.seed(seed=1,sample.kind = 'Rounding')
-tuningControl=trainControl(method = "cv", number =2, p = .8)
-gam_fit <- train(number_of_hours~., train_subset, method='gam',trControl=tuningControl) # check for tuning parameters
+preProcessScalingValues <- preProcess(train_subset %>% select(-number_of_hours), method = c("center", "scale"))
+
+scaled_train_subset <- predict(preProcessScalingValues, train_subset) 
+scaled_test_subset <- predict(preProcessScalingValues, test_subset)
+
+tuningControl=trainControl(method = "cv", number =3, p = .8)
+gam_fit <- train(number_of_hours~., scaled_train_subset, method='gam',trControl=tuningControl) # check for tuning parameters
 
@@
 
-predictions <- predict(gam_fit,test_subset %>% select(-number_of_hours))
-
-RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
-# 195.3235
+predictions <- predict(gam_fit,scaled_test_subset %>% select(-number_of_hours))
+
+RMSE(predictions,(scaled_test_subset %>% pull(number_of_hours) ))
+# 185.1074 uses GCV.Cp method and select=FALSE
 

### 2021-07-13 finalReport.Rmd
fine tuned code for final model and predictions

@@
 
+```{r Dataframe to store target encoding weights, include=FALSE}
+target_encoding_weights <- data.frame(effect=character(0),weight=double(0)) 
+```
+
 ```{r Apply Target encoding for seller_id, echo = FALSE}
@@
 # 3
+
+target_encoding_weights <- target_encoding_weights %>% add_row(effect='seller_id',weight=weight_to_use)
 ```
@@
 # 20
+
+target_encoding_weights <- target_encoding_weights %>% add_row(effect='customer_state',weight=weight_to_use)
 ```
@@
 # 5/3 = 1.66667
+
+target_encoding_weights <- target_encoding_weights %>% add_row(effect='customer_loc',weight=weight_to_use)
 ```
@@
 
-```{r Glm model}
-# Let's try with glm model
+```{r Prepare seller_id effect based on target encoding, include = FALSE}
+# prepare seller_id_effect based on regularisation weight we have identified
+prepare_seller_id_effect <- function(train_df, weight_to_use) {
+    global_mean <- mean(train_df$number_of_hours)
+
+    set.seed(1,sample.kind='Rounding')
+    test_weight_indices <- createDataPartition(train_df$number_of_hours, p=0.2,times=10,list=FALSE)
+
+    seller_id_effects_list <- lapply(1:10, function(index) {
+      train_weight_set <- train_df[-test_weight_indices[,index],]
+
+      seller_id_effects <- train_weight_set %>% 
+        group_by(seller_id) %>%
+        summarise(mean=mean(number_of_hours),n=n()) %>%
+        mutate(mean=(n*mean+weight_to_use*global_mean)/(n+weight_to_use)) %>% 
+        select(-n) %>% rename()
+
+      seller_id_effects
+    })
+
+    sie <- seller_id_effects_list[[1]]
+    for (index in 2:10) {
+       sie <- sie %>% 
+         full_join(seller_id_effects_list[[index]],by='seller_id', suffix=c("1",index))
+    }
+
+    # whereever value is NA, we have used global mean
+    sie <- sie %>% mutate(across(-seller_id,~if_else(is.na(.x),global_mean,.x)))
+
+    # as different partitions may not have exactly same data set combined the results
+    seller_id_effects <- tibble(sie[,1], seller_id_effect=rowMeans(sie[,-1],na.rm = TRUE))
+
+    #seller id effects are ready
+    seller_id_effects
+}
+```
+
+```{r Prepare customer_state effect based on target encoding, include = FALSE}
+# prepare consumer_state_effect based on regularisation weight we have identified
+prepare_customer_state_effect <- function(train_df, weight_to_use) {
+  global_mean <- mean(train_df$number_of_hours)
+
+  set.seed(1,sample.kind='Rounding')
+  test_weight_indices <- createDataPartition(train_df$number_of_hours,p=0.2,times=10,list=FALSE)
+
+  customer_state_effects_list <- lapply(1:10, function(index) {
+      train_weight_set <- train_df[-test_weight_indices[,index],]
+
+      customer_state_effects <- train_weight_set %>% 
+        group_by(customer_state) %>%
+        summarise(mean=mean(number_of_hours),n=n()) %>%
+        mutate(mean=(n*mean+weight_to_use*global_mean)/(n+weight_to_use)) %>% 
+        select(-n) %>% rename()
+
+      customer_state_effects
+    }
+  )
+
+  cse <- customer_state_effects_list[[1]]
+  for (index in 2:10) {
+     cse <- cse %>% 
+       full_join(customer_state_effects_list[[index]],by='customer_state', suffix=c("1",index))
+  }
+
+  # whereever value is NA, we have used global mean
+  cse <- cse %>% mutate(across(-customer_state,~if_else(is.na(.x),global_mean,.x)))
+
+  # as different partitions may not have exactly same data set combined the results
+  customer_state_effects <- tibble(cse[,1], customer_state_effect=rowMeans(cse[,-1],na.rm = TRUE))
+
+  #customer state effects are ready
+  customer_state_effects
+}
+```
+
+```{r Prepare customer_loc effect based on target encoding, include = FALSE}
+# prepare customer_loc_effect based on regularisation weight we have identified
+prepare_customer_loc_effect <- function(train_df, weight_to_use) {
+  global_mean <- mean(train_df$number_of_hours)
+
+  set.seed(1,sample.kind='Rounding')
+  test_weight_indices <- createDataPartition(train_df$number_of_hours,p=0.2,times=10,list=FALSE)
+
+  customer_loc_effects_list <- lapply(1:10, function(index) {
+      train_weight_set <- train_df[-test_weight_indices[,index],]
+
+      customer_loc_effects <- train_weight_set %>% 
+        group_by(customer_loc) %>%
+        summarise(mean=mean(number_of_hours),n=n()) %>%
+        mutate(mean=(n*mean+weight_to_use*global_mean)/(n+weight_to_use)) %>% 
+        select(-n) %>% rename()
+
+      customer_loc_effects
+    }
+  )
+
+  cle <- customer_loc_effects_list[[1]]
+  for (index in 2:10) {
+     cle <- cle %>% full_join(customer_loc_effects_list[[index]],by='customer_loc', suffix=c("1",index))
+  }
+
+  # where-ever value is NA, we have used global mean
+  cle <- cle %>% mutate(across(-customer_loc,~if_else(is.na(.x),global_mean,.x)))
+
+  # as different partitions may not have exactly same data set combined the results
+  customer_loc_effects <- tibble(cle[,1], customer_loc_effect=rowMeans(cle[,-1],na.rm = TRUE))
+
+  #customer loc effects are ready
+  customer_loc_effects
+}
+```
+
+```{r Enhance training and test dataset for seller_id effect based on target encoding, include = FALSE}
+enhance_datasets_using_seller_id_effect <- function(seller_id_effects,train_df,test_df,cleanup) {
+  #we will use seller id effects computed using training data in test data.
+  train_df <- train_df %>% left_join(seller_id_effects,by='seller_id') %>% mutate(seller_id_effect=if_else(is.na(seller_id_effect),global_mean,seller_id_effect))
+
+  test_df <- test_df %>% left_join(seller_id_effects,by='seller_id')
+
+  # cleanup
+  if (cleanup) {
+    train_df <- train_df %>% select(-product_category_name,-seller_id)
+    test_df <- test_df %>% select(-product_category_name,-seller_id)
+  }
+
+  list(train_df,test_df)
+}
+```
+
+```{r Enhance training and test dataset for customer_state effect based on target encoding, include = FALSE}
+enhance_datasets_using_customer_state_effect <- function(customer_state_effects,train_df,test_df,cleanup) {
+  #we will use customer state effects computed using training data in test data.
+  train_df <- train_df %>% left_join(customer_state_effects,by='customer_state') %>%
+    mutate(customer_state_effect=if_else(is.na(customer_state_effect),
+                                         global_mean,
+                                         customer_state_effect))
+
+  test_df <- test_df %>% left_join(customer_state_effects,by='customer_state')
+
+  # cleanup
+  if (cleanup) {
+    train_df <- train_df %>% select(-customer_state)
+    test_df <- test_df %>% select(-customer_state)
+  }
+
+  list(train_df,test_df)
+}
+```
+
+```{r Enhance training and test dataset for customer_loc effect based on target encoding, include = FALSE}
+enhance_datasets_using_customer_loc_effect <- function(customer_loc_effects,train_df,test_df,cleanup) {
+  #we will use customer loc effects computed using training data in test data.
+  train_df <- train_df %>% 
+    left_join(customer_loc_effects,by='customer_loc') %>%
+    mutate(customer_loc_effect=if_else(is.na(customer_loc_effect),global_mean,customer_loc_effect))
+
+  test_df <- test_df %>% left_join(customer_loc_effects,by='customer_loc')
+
+  # cleanup
+  if (cleanup) {
+    train_df <- train_df %>% select(-customer_loc,-customer_zip_code_prefix)
+    test_df <- test_df %>% select(-customer_loc,-customer_zip_code_prefix)
+  }
+
+  list(train_df,test_df)
+}
+```
+
+```{r Prepare data set for training, include = FALSE}
+# prepares data for training and testing
+# returns list of data frames. First list has training dataframe. Second list has test data frame
+prepare_data_for_training_testing <- function(train_ds, test_ds, cleanup) {
+  print('Adding column for delay in handover')
+  penalty_data <- train_ds %>% get_delay_in_handover_to_carrier_per_seller_id()
+
+  #we will use penalty_data computed using training data in test data.
+  train_ds <- train_ds %>% left_join(penalty_data,by='seller_id') 
+  test_ds <- test_ds %>% left_join(penalty_data,by='seller_id')
+  if (cleanup) {
+    train_ds <- train_ds %>% select(-handed_over_to_carrier_date,-shipping_limit_date)
+    test_ds <- test_ds %>% select(-handed_over_to_carrier_date,-shipping_limit_date)
+  }
+
+  print('Enhancing data for seller id effects')
+  weight_to_use <- target_encoding_weights %>% filter(effect=='seller_id') %>% pull(weight)
+  seller_id_effects <- prepare_seller_id_effect(train_ds,weight_to_use)
+  dataset_list <- enhance_datasets_using_seller_id_effect(seller_id_effects,train_ds,test_ds,cleanup)
+  train_ds <- dataset_list[[1]]
+  test_ds <- dataset_list[[2]]
+
+  print('Enhancing data for customer state effects')
+  weight_to_use <- target_encoding_weights %>% filter(effect=='customer_state') %>% pull(weight)
+  customer_state_effects <- prepare_customer_state_effect(train_ds,weight_to_use)
+  dataset_list <- enhance_datasets_using_customer_state_effect(customer_state_effects,train_ds,test_ds,cleanup)
+  train_ds <- dataset_list[[1]]
+  test_ds <- dataset_list[[2]]
+
+  print('Enhancing data for customer loc effects')
+  weight_to_use <- target_encoding_weights %>% filter(effect=='customer_loc') %>% pull(weight)
+  customer_loc_effects <- prepare_customer_loc_effect(train_ds,weight_to_use)
+  dataset_list <- enhance_datasets_using_customer_loc_effect(customer_loc_effects,train_ds,test_ds,cleanup)
+  dataset_list
+}
+```
+
+```{r Prepare dataset for final model training, echo = FALSE}
 set.seed(seed=1,sample.kind = 'Rounding')
-tuningControl=trainControl(method = "cv", number = 2, p = .8)
-glm_fit <- train(number_of_hours~., train_subset, method='glm',trControl=tuningControl)
-
-glm_fit
-
-predictions <- predict(glm_fit, test_subset %>% select(-number_of_hours))
-
-RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
-# 198.8575 (improves if we use seller_id as predictor as well)
-```
-Warning:
-prediction from a rank-deficient fit may be misleading
-
-```{r GamLoess model, echo = FALSE}
-set.seed(seed=1,sample.kind = 'Rounding')
-tuningControl=trainControl(method = "cv", number = 3, p = .8)
-#grid <- expand.grid(span = seq(0.15, 0.65, len = 10), degree = 1)
-grid <- expand.grid(span = seq(0.15, 0.65, len = 3), degree = 1)
-gamLoess_fit <- train(number_of_hours~., train_subset, method='gamLoess',tuneGrid=grid,trControl=tuningControl)
-
-gamLoess_fit
-
-predictions <- predict(gamLoess_fit,test_subset %>% select(-number_of_hours))
-
-RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
-#194.6215
-
-save(gamLoess_fit,file='rda/gamLoess_model.rda')
-```
-
-Warnings received:
-extrapolation not allowed with blending
- (Discovered by lowesd)liv too small.    
- (Discovered by lowesd)lv too small
- extrapolation not allowed with blending
- non-list contrasts argument ignored
-
-```{r GamLoess model with scaling, echo = FALSE}
-# TODO review-keep or remove it
-set.seed(seed=1,sample.kind = 'Rounding')
-scaled_train_subset <- train_subset %>% mutate(across(c(mean_delay_in_handover,distance_between_two_points,volume,product_weight_g,freight_value,price),scale)) 
-scaled_test_subset <- test_subset %>% mutate(across(c(mean_delay_in_handover,distance_between_two_points,volume,product_weight_g,freight_value,price),scale))
-
-tuningControl=trainControl(method = "cv", number = 3, p = .8)
-#grid <- expand.grid(span = seq(0.15, 0.65, len = 10), degree = 1)
-grid <- expand.grid(span = seq(0.15, 0.65, len = 3), degree = 1)
-gamLoess_fit <- train(number_of_hours~., scaled_train_subset, method='gamLoess',tuneGrid=grid,trControl=tuningControl)
-
-gamLoess_fit
-
-predictions <- predict(gamLoess_fit,scaled_test_subset %>% select(-number_of_hours))
-
-RMSE(predictions,(scaled_test_subset %>% pull(number_of_hours) ))
-#194.7017
-
-save(gamLoess_fit,file='rda/gamLoess_model.rda')
-```
-Warnings received:
-extrapolation not allowed with blending
- (Discovered by lowesd)liv too small.    
- (Discovered by lowesd)lv too small
- extrapolation not allowed with blending
- non-list contrasts argument ignored
+
+validation_dataset <- validation_dataset %>% 
+  compute_volume_using_product_info() %>% 
+  extract_distance_between_seller_customer() %>% 
+  compute_rough_customer_location()
+
+dataset_list <- prepare_data_for_training_testing(training_dataset,validation_dataset,FALSE)
+training_dataset <- dataset_list[[1]]
+validation_dataset <- dataset_list[[2]]
+
+validation_dataset <- validation_dataset %>%
+  semi_join(training_dataset,by='product_category_name') %>% 
+  semi_join(training_dataset,by='seller_id') %>% 
+  semi_join(training_dataset,by='customer_state') %>%
+  semi_join(training_dataset,by='customer_loc')
+
+# cleanup
+training_dataset <- training_dataset %>%
+  select(-handed_over_to_carrier_date,-shipping_limit_date,-product_category_name,-seller_id,-customer_state,-customer_loc,-customer_zip_code_prefix)
+validation_dataset <- validation_dataset %>%
+  select(-handed_over_to_carrier_date,-shipping_limit_date,-product_category_name,-seller_id,-customer_state,-customer_loc,-customer_zip_code_prefix)
+```
 
@@
 
-penalty_data <- training_dataset %>% get_delay_in_handover_to_carrier_per_seller_id()
-
-training_dataset <- training_dataset %>% left_join(penalty_data,by='seller_id') %>% select(-seller_id,-handed_over_to_carrier_date,-shipping_limit_date)
-
 mtry_to_use <- rf_fit$finalModel$mtry
 ntree_to_use <- 500 #rf_fit$finalModel$ntree
-tuningControl <- trainControl(method = "cv", number = 2, p = .8)
-final_model <- train(number_of_hours~., training_dataset, method='rf',tuneGrid = data.frame(mtry = c(mtry_to_use)),trControl=tuningControl,ntree=ntree_to_use)
-
-# final_model <- randomForest(number_of_hours~., training_dataset,mtry= mtry_to_use, ntree=ntree_to_use)
+#tuningControl <- trainControl(method = "cv", number = 2, p = .8)
+#final_model <- train(number_of_hours~., training_dataset, method='rf',tuneGrid = data.frame(mtry = c(mtry_to_use)),trControl=tuningControl,ntree=ntree_to_use)
+
+final_model <- randomForest(number_of_hours~., training_dataset,mtry= mtry_to_use, ntree=ntree_to_use)
 
@@
 ```{r Final movel validation, echo = FALSE}
-validation_dataset <- validation_dataset %>% compute_volume_using_product_info() %>% extract_distance_between_seller_customer() %>% left_join(penalty_data,by='seller_id') %>% select(-seller_id,-handed_over_to_carrier_date,-shipping_limit_date)
+validation_dataset <- validation_dataset %>% select(-seller_id,-handed_over_to_carrier_date,-shipping_limit_date)
 
@@
 
-## Explore seller_zip_code_prefix with number_of_hours
-
-```{r Explore seller_zip_code_prefix with number_of_hours, echo = FALSE}
-# TODO anyway to add regularization or remove the low impact seller_zip_code_prefix from the model
-# let's plot errorbar for number of hours vs seller_zip_code_prefix
-sellers_zip_of_interest <- training_dataset %>% dplyr::count(seller_zip_code_prefix) %>% filter(n>10) %>% pull(seller_zip_code_prefix)
-
-training_dataset %>% filter(seller_zip_code_prefix %in% sellers_zip_of_interest) %>% group_by(seller_zip_code_prefix) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% mutate(seller_zip_code_prefix=reorder(seller_zip_code_prefix,mean)) %>%  ggplot(aes(x=seller_zip_code_prefix,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90)) + scale_y_sqrt()
-```
-Indicates role between seller_zip_code and number of hours for delivery
-
-## Explore customer_zip_code_prefix with number_of_hours
-
-```{r Explore customer_zip_code_prefix with number_of_hours, echo=FALSE}
-# TODO anyway to add regularization or remove the low impact customer_zip_code_prefix from the model
-# let's plot errorbar for number of hours vs customer_zip_code_prefix
-customers_zip_of_interest <- training_dataset %>% dplyr::count(customer_zip_code_prefix) %>% filter(n>=5) %>% pull(customer_zip_code_prefix)
-
-training_dataset %>% filter(customer_zip_code_prefix %in% customers_zip_of_interest) %>% group_by(customer_zip_code_prefix) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% mutate(customer_zip_code_prefix=reorder(customer_zip_code_prefix,mean)) %>%  ggplot(aes(x=customer_zip_code_prefix,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90)) + scale_y_sqrt()
-```
-customer_zip_code_prefix can  be seen to be  influence number of hours for delivery
-
-## Explore product_id with number_of_hours
-
-```{r Explore product_id with number_of_hours}
-# TODO anyway to add regularization or remove the low impact products from the model
-# let's plot errorbar for number of hours vs product_id
-products_of_interest <- training_dataset %>% dplyr::count(product_id)  %>% filter(n>=10) %>% pull(product_id)
-
-training_dataset %>% filter(product_id %in% products_of_interest) %>% group_by(product_id) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% mutate(product_id=reorder(product_id,mean)) %>%  ggplot(aes(x=product_id,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_blank())
-```
-It influences the number of hours
-
-====TODO review above
-
-### Linear model trial and analysis
-
-```{r Linear model, eval=FALSE, echo = FALSE}
-# trying to reduce coefficients.
-# in dimensions we can use volume to decrase predictors
-
-########### only price predictor#############
-## only price complete data
-lm_fit <- train(number_of_hours~., train_subset %>% select(price,number_of_hours), method='lm')
-
-lm_fit
-
-predictions <- predict(lm_fit,test_subset %>% select(price))
-
-RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
-# 220.4085
-
-## only price,distance_between_two_points complete data
-lm_fit <- train(number_of_hours~., train_subset %>% select(price,distance_between_two_points,number_of_hours), method='lm')
-
-lm_fit
-
-predictions <- predict(lm_fit,test_subset %>% select(price,distance_between_two_points))
-
-RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
-# 202.0175
-
-## only price,distance_between_two_points,product_weight_g complete data
-lm_fit <- train(number_of_hours~., train_subset %>% select(price,distance_between_two_points,product_weight_g,number_of_hours), method='lm')
-
-lm_fit
-
-predictions <- predict(lm_fit,test_subset %>% select(price,distance_between_two_points,product_weight_g))
-
-RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
-# 201.1412
-
-## only price,distance_between_two_points,product_weight_g,volume complete data
-lm_fit <- train(number_of_hours~., train_subset %>% select(price,distance_between_two_points,product_weight_g,volume,number_of_hours), method='lm')
-
-lm_fit
-
-predictions <- predict(lm_fit,test_subset %>% select(price,distance_between_two_points,product_weight_g,volume))
-
-RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
-# 201.1086
-
-## only price,distance_between_two_points,product_weight_g,volume,freight_value complete data
-lm_fit <- train(number_of_hours~., train_subset %>% select(price,distance_between_two_points,product_weight_g,volume,freight_value,number_of_hours), method='lm')
-
-lm_fit
-
-predictions <- predict(lm_fit,test_subset %>% select(price,distance_between_two_points,product_weight_g,volume,freight_value))
-
-RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
-# 201.1047
-
-## only product_category_name complete data
-lm_fit <- train(number_of_hours~., train_subset %>% select(product_category_name,number_of_hours), method='lm')
-
-lm_fit
-
-predictions <- predict(lm_fit,test_subset %>% select(product_category_name))
-
-RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
-# 218.8943
-
-# if we combine this with other predictors
-## only price,distance_between_two_points,product_category_name complete data
-lm_fit <- train(number_of_hours~., train_subset %>% select(price,distance_between_two_points,product_category_name,number_of_hours), method='lm')
-
-lm_fit
-
-predictions <- predict(lm_fit,test_subset %>% select(price,distance_between_two_points,product_category_name))
-
-RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
-# 199.8844
-```
-
-```{r Linear model with all predictors, echo=FALSE}
-# above is a fine model
-# with all predictors
-lm_fit <- train(number_of_hours~., train_subset, method='lm')
-
-lm_fit
-
-predictions <- predict(lm_fit,test_subset %>% select(-number_of_hours))
-
-RMSE((test_subset %>% pull(number_of_hours) ), predictions)
-# 199.4791
-
-save(lm_fit,file='rda/lm_model.rda')
-```
-Warnings received:
-prediction from a rank-deficient fit may be misleading
-
-To improve model performance consider--> changing character categorical variables to numerical categorical variables like
-
-  product_category_name=as.factor(as.numeric(product_category_name))
-
-use levels from original data for completeness
+===TODO review
 
@@
 gamLoess??
-
-## Experimets which failed
-
-Let's play with seller_id. It deriorates rf model and for gam model there was no improvement
-```{r play with seller_id, eval=FALSE, echo = FALSE}
-set.seed(1,sample.kind = 'Rounding')
-test_indices <- createDataPartition(wrangled_data$number_of_hours,times=1,p=0.2,list=FALSE)
-dummy_data <- wrangled_data[-test_indices,]
-dummy_data <- dummy_data %>% mutate(volume=product_length_cm*product_height_cm*product_width_cm) %>% mutate(seller_id=factor(as.numeric(seller_id))) %>% select(-customer_unique_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix,-product_length_cm,-product_height_cm,-product_width_cm)
-
-set.seed(1,sample.kind = 'Rounding')
-test_indices <- createDataPartition(dummy_data$number_of_hours,times=1,p=0.2,list=FALSE)
-dummy_train_subset <- dummy_data[-test_indices,]
-dummy_test_subset <- dummy_data[test_indices,]
-
-# only keeping those rows which have seller_id in training set
-dummy_test_subset <- dummy_test_subset %>% semi_join(dummy_train_subset,by='seller_id')
-
-# used gam model but it did not use seller id at all 
-# Let's try with rf model
-set.seed(seed=1,sample.kind = 'Rounding')
-tuningControl=trainControl(method = "cv", number = 2, p = .8)
-rf_fit_dummy <- train(number_of_hours~., dummy_train_subset, method='rf',tuneGrid = data.frame(mtry = seq(5, 7, 1)),trControl=tuningControl) # nodesize = 50,maxnodes=2000,ntree=200
-
-rf_fit_dummy
-
-predictions <- predict(rf_fit_dummy,dummy_test_subset %>% select(-number_of_hours))
-
-RMSE(predictions,(dummy_test_subset %>% pull(number_of_hours) ))
-# 203.6908
-
-# Let's try with glm model
-set.seed(seed=1,sample.kind = 'Rounding')
-tuningControl=trainControl(method = "cv", number = 2, p = .8)
-glm_fit_dummy <- train(number_of_hours~., dummy_train_subset, method='glm',trControl=tuningControl)
-
-glm_fit_dummy
-
-predictions <- predict(glm_fit_dummy,dummy_test_subset %>% select(-number_of_hours))
-
-RMSE(predictions,(dummy_test_subset %>% pull(number_of_hours) ))
-# 195.4884
-```
 

### 2021-07-13 finalReport.Rmd
created final model

@@
 
-```{r Final movel validation, echo = FALSE}
-validation_dataset <- validation_dataset %>% select(-seller_id,-handed_over_to_carrier_date,-shipping_limit_date)
-
+```{r Final model validation, echo = FALSE}
 predictions <- predict(final_model,validation_dataset %>% select(-number_of_hours))
@@
 RMSE((validation_dataset %>% pull(number_of_hours) ),predictions)
-# 194.6748 for mtry=25, ntree=500
+# 190.184 for mtry=15, ntree=500
 ```

### 2021-07-16 finalReportPrep.Rmd
added payment_approval code

--- 2021-07-13 finalReport.Rmd
@@
 ---
-title: "CYO Project Report - Brazilian Ecommerce"
+title: "CYO Project Report - Delivery Estimation System"
 author: "Asham Vohra"
@@
 output:
+  pdf_document: default
   html_document:
@@
 # Overview
-Data set: https://www.kaggle.com/olistbr/brazilian-ecommerce
+Prediction systems play a cruical role in the modern world.
+With advent of technology and availability of data, these systems are being leveraged by retail chains, banks, colleges among other businesses.
+Depending on the use case, these systems are used to estimate delivery time, sales figures, carry out inventory management, predict occurrence of a disease, among others.
+
+Inspired by what all problems these systems can solve, we have attempted to build a delivery prediction system based on publically available [dataset of orders made at Olist Store](https://www.kaggle.com/olistbr/brazilian-ecommerce), a Brazilian ecommerce.
+
+The dataset used here is hosted at [Kaggle website](https://www.kaggle.com).
+The dataset is provided by Olist itself, which is the largest department store in Brazilian marketplaces.
+Olist connects sellers to the customers.
+It is the sellers, who are responsible to fulfill any orders placed on the platform using Olist's logistic partners.
+
+The dataset consists of **100,000+** orders and for each order the dataset contains related details like product bought and from which category, customer who bought and his/her location, seller involved and many more attributes.
+
+The delivery prediction system developed as part of the project, leverages the past orders, product category, delivery distance, customer location, actual delivery time and related attributes to predict delivery estimate for the new orders.
+
+The goal of the project was to build a delivery prediction system which can estimate number of hours required for delivery keeping RMSE(Root mean square error) minimal.
+
+In order to achieve this goal, we started by cleaning data, identifying right predictors, deriving insights from data already available to create new predictors.
+Once we had the dataset with us, we incrementally took a predictor, analysed its relationship with the number of hours of delivery, incorporated the predictor in our model after encoding, scaling it as required and tested our updated model against the test data.
+To ensure that the model performance was not due to random split of dataset, the model was trained and tuned using k fold cross validation and then only tested against the test dataset.
+Only if the predictor helped improve our metric i.e.
+RMSE, the predictor under analysis and evaluation was added to the model and the steps were repeated with new predictor.
+Otherwise the predictor was ignored
+
+This report walks through the approach, analysis and evaluation carried out to achieve our delivery prediction system
 
@@
 
-Delivery Performance:
-
-You will also be able to work through delivery performance and find ways to optimize delivery times.
-
-Let's predict delivery time/date
--- delivery distance
--- location of customer as some can be remotely connected and need special delivery merchanism. This is independent of delivery distance.
--- location of seller as shipping from some locations maybe difficult. This is independent of delivery distance.
--- size of product? 
--- weight of the product. Heavier products may take time to deliver
--- price of product (expensive nature of product as it could mean extra care or it maybe fragile?)
--- purchase time....if in night, delivery time maybe longer
--- order approved at time-> if payment approval came multiple hours after purchase time, this maybe longer
--- product category -- products of same category may show similar behavior.
--- or specific product - some products may have higher delivery time.
--- deliveries maybe impacted due to sales, festival season.
-
-we should consider if we need time stamp or we can take purchase time as 0 and calculated approved at from it.
-though we may lose granularity in terms of time as truck may come everyday at 4.
-so after 4 you can ship , before 4 you cant
-We can be agnostic to purchase time and just predict number of hours for delivery for given purchase.
-
-estimated delivery date is present in ---> order_estimated_delivery_date
-We can calculate RMSE using that and it will be our benchmark
-
-We can see if we improved than that or not. It will give us indication of how close are we to the production model RMSE.
-
-TODO -> if there is delayed to carrier precendence, we could use that probability and average delay to carrier to fine tune our results.
-Similarly we could use other metrics like payment approval time etc.
-
+## Data Wrangling
 ```{r Install or load packages,include=FALSE, echo = FALSE}
@@
 
-## Data Wrangling
-Reading from raw data source and converting data to usable form
+An important step in using any public data set is converting it to usable form.
+In our case, the data set was available in multiple csv files, each file dedicated to a specific kind of information.
+Example:
 
-For products data set,
-- we read data, remove unnecessary fields, join with english names whereever available
-- remove products where dimension and related attributes are not available
-- assign default category name of UNCATEGORIZED for products where same was not available
+* There was a file for orders which had information about product bought, purchase time, delivery timestamp and other attributes. 
+* Then there was a file with product details which had information about product category, product weight and other attributes.
+
+Each of these files due to the varied dataset required different data wrangling steps to achieve a usable form which could then be unified for further analysis.
+
+### Data wrangling for products data set
+
+The data set here had product details like category name, dimensions and other attributes.
+Here the category name was in Portuguese.
+Some important attributes like product dimensions and weight were not populated for few of the products.
+To analyse this data properly, we carried out below steps:
+
+* Read the concerned csv file and removed irrelevant columns for the analysis.
+* Removed products where dimensions and related attributes were not available.
+* Assigned default category name of UNCATEGORIZED for products where product_category_name was not available.
+* Replaced product category names with English names where-ever mapping was available. The product categories for which equivalent English name mapping was not available were left untouched.
+
 ```{r Data wrangling-products data set, include=FALSE, echo=FALSE}
-products_dataset <- read.csv(file.path('data','olist_products_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) %>% select(-product_name_lenght,-product_description_lenght,-product_photos_qty)
+# read the data from source and remove irrelevant columns
+products_dataset <- read.csv(file.path('data','olist_products_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) %>%
+  select(-product_name_lenght,-product_description_lenght,-product_photos_qty)
 
+# check how they are structured
 products_dataset %>% head()
@@
 
+# load the product category name translations
 product_category_name_translation <- read.csv(file.path('data','product_category_name_translation.csv'),header=TRUE,stringsAsFactors = FALSE)
@@
 
-# convert product_name to english form wherever possible
-products_dataset <- products_dataset %>% left_join(product_category_name_translation,by='product_category_name') %>% mutate(product_category_name=if_else(is.na(product_category_name_english),product_category_name,product_category_name_english)) %>% select(-product_category_name_english)
+# convert product_name to english form wherever possible. If not available, existing product category name is retained
+products_dataset <- products_dataset %>% 
+  left_join(product_category_name_translation,by='product_category_name') %>% 
+  mutate(product_category_name=if_else(is.na(product_category_name_english),
+                                       product_category_name,
+                                       product_category_name_english)) %>% 
+  select(-product_category_name_english)
 
 # further remove products for which data is not available i.e is NA
-products_dataset <- products_dataset %>% filter(!is.na(product_weight_g) & !is.na(product_length_cm) & !is.na(product_height_cm) & !is.na(product_width_cm)  )
+products_dataset <- products_dataset %>% 
+  filter(!is.na(product_weight_g) & !is.na(product_length_cm) & !is.na(product_height_cm) & !is.na(product_width_cm))
 
-# assign category name for blank category
-products_dataset <- products_dataset %>% mutate(product_category_name=if_else(product_category_name=='','UNCATEGORIZED',product_category_name))
-```
-The wrangled data looks like below:
-```{r Data wrangling result-products data set, echo=FALSE}
-products_dataset %>% head() %>% knitr::kable()
+# assign default category name of UNCATEGORIZED where-ever product_category_name is blank.
+products_dataset <- products_dataset %>%
+  mutate(product_category_name=if_else(product_category_name=='','UNCATEGORIZED',product_category_name))
 ```
 
-For geolocation data set
+The wrangled data looked like below:
+```{r Data wrangling result-products data set, echo=FALSE}
+products_dataset %>% head(2) %>% as.matrix()  %>% t() %>% 
+  knitr::kable(col.names=c('Example Value1','Example Value2'), caption='Product data set')
+```
+
+### Data wrangling for orders data set
+
+The data set here had orders details like order id, associated customer, order status, purchase time, when was the order handed over to carrier, when was it delivered to customer and estimated delivery date and few other attributes.
+
+This was a crucial dataset not only for us to build our delivery estimation system but to even identify the current performance metric i.e. RMSE for the approach Olist was using.
+
+In order to leverage this dataset properly, we carried out below steps:
+
+* Read the concerned csv file and removed irrelevant columns for the analysis.
+* Parsed columns with date as characters and converted them to desired date type.
+* Filtered out any orders where delivery was not made or where delivery was made but required delivery time columns were not available.
+
+```{r Data wrangling-orders data set, include=FALSE, echo=FALSE}
+# order data set with various delivery related times
+# note each customer_id is unique per order
+# order_purchase_timestamp -> Shows the purchase timestamp.
+# order_approved_at/payment_approved_at -> Shows the payment approval timestamp.
+# order_delivered_carrier_date/handed_over_to_carrier_date -> Shows the order posting timestamp. When it was handled to the logistic partner.
+# order_delivered_customer_date --> Shows the actual order delivery date to the customer.
+# order_estimated_delivery_date --> Shows the estimated delivery date that was informed to customer at the purchase moment.
+orders_dataset <- read.csv(file.path('data','olist_orders_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) %>% 
+  rename(payment_approved_at=order_approved_at,
+         handed_over_to_carrier_date=order_delivered_carrier_date)  %>% 
+  mutate(order_status=as.factor(order_status),
+         order_purchase_timestamp=as_datetime(order_purchase_timestamp),
+         payment_approved_at=as_datetime(payment_approved_at),
+         handed_over_to_carrier_date=as_datetime(handed_over_to_carrier_date),
+         order_delivered_customer_date=as_datetime(order_delivered_customer_date),
+         order_estimated_delivery_date=as_datetime(order_estimated_delivery_date))
+
+# TODO if we are not using payment_approved_at or few fields we can remove the filter logic based on that.
+orders_dataset <- orders_dataset %>% 
+  filter(order_status=='delivered') %>% 
+  filter( !is.na(order_estimated_delivery_date ) & !is.na(order_delivered_customer_date ) & !is.na(handed_over_to_carrier_date) & !is.na(payment_approved_at) & !is.na(order_purchase_timestamp))
+
+orders_dataset %>% head()
+orders_dataset %>% str()
+```
+
+The wrangled data looks like below:
+```{r Data wrangling result-orders dataset, echo=FALSE}
+orders_dataset %>% head(2) %>% as.matrix() %>% t() %>% 
+  knitr::kable(col.names=c('Example Value1','Example Value2'), caption='Orders data set')
+```
+
+### Data wrangling for orders items data set
+
+The data set here had order item details like which product(s) were bought for a given order, who the seller was, price and freight value and shipping limit date.
+While most of the attributes were used to connect with other data sets, it was the shipping limit date which could be used along with product handed over to carrier date from orders dataset to identify any delays in the shipping process.
+
+In order to leverage this dataset properly, we carried out below steps:
+
+* Read the concerned csv file and removed irrelevant columns for the analysis.
+* Parsed columns with date as characters and converted them to desired date type.
+
+```{r Data wrangling-order items data set, include=FALSE, echo=FALSE}
+# orders data with item related seller info
+# order_item_id --->sequential number identifying number of items included in the same order.
+# so if there are three items they will be nuber 1 , 2, 3 for the given order
+# shipping_limit_date ---> Shows the seller shipping limit date for handling the order over to the logistic partner.
+# freight_value --> item freight value item (if an order has more than one item the freight value is splitted between items)
+order_items_dataset <- read.csv(file.path('data','olist_order_items_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) %>% mutate(shipping_limit_date=as_datetime(shipping_limit_date))  %>% select(-order_item_id)
+
+order_items_dataset %>% str()
+order_items_dataset %>% head()
+```
+
+The wrangled data looks like below:
+```{r Data wrangling result-orders items dataset, echo=FALSE}
+order_items_dataset %>% head(2) %>% as.matrix() %>% t() %>% 
+  knitr::kable(col.names=c('Example Value1','Example Value2'), caption='Order Items\' data set')
+```
+### Data wrangling for geolocation data set
+
+The data set here had geolocation data. It had zip code prefix, lat, lang, city and state details. However, there was no unique identifier in the dataset.
+
+It is important to note that  Brazil uses [Postal Code Address(CEP)](https://codigo-postal.org/en-us/brazil/) which is 8-digit number created mainly from mail distribution objective.
+It is split into two parts of 5 and 3 digits resp.
+separated by a hyphen('-').
+The first part helps identify the rough delivery area, while it is the second part which provides precision and helps one reach the desired street, neighbourhood.
+
+So in the geolocation data set, zip code prefix refered to the first part of the CEP and it varied from 1 to 5 digits in the dataset available to us.
+Because the zip code prefix does not represent the precise value, it was not unique and had many duplicate values with varying value of lat, lng for same zip code prefix.
+
+In order to be able to use this data with other data sets available like customer location and seller location datasets, which are discussed later, we carried out below steps:
+
+* Read the concerned csv file and removed irrelevant columns for the analysis.
+* Calculated the mean of lat, lng for given zip code prefix.
+
 ```{r Data wrangling- geolocation data set, include=FALSE, echo=FALSE}
@@
 
-geolocation_dataset <- read.csv(file.path('data','olist_geolocation_dataset.csv'),header=TRUE,stringsAsFactors = FALSE)
+# geolocation_dataset <- read.csv(file.path('data','olist_geolocation_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) 
+geolocation_dataset <- read.csv(file.path('data','olist_geolocation_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) %>% select(-geolocation_city,-geolocation_state)
 
@@
 
+The wrangled data looks like below:
+```{r Data wrangling result-geolocation dataset, echo=FALSE}
+geolocation_dataset %>% head(2) %>% summarise_all(as.character) %>% t() %>% 
+  knitr::kable(col.names=c('Example Value1','Example Value2'), caption='Geolocation data set')
+```
+
+### Data wrangling for customers and sellers data set
+
+There were two data sets one for customers and one for sellers.
+Apart from identifier ids like customer_unique_id, seller_id, both the datasets had similar columns i.e.
+zip code prefix, for which details were in geolocation data set and city and state.
+
+To be able to leverage either of these datasets for analysis later, we carried out below steps:
+
+* Read the concerned csv file and removed irrelevant columns for the analysis.
+* Combined sellers dataset data individually with geolocation dataset in order to capture seller lat,lng. Similar operation was carried out for customers data set.
+* Remove sellers or customers for which geolocation details are not available as delivery or even its estimation is not possible without the critical fields.
+
 ```{r Data wrangling-customers data set, include=FALSE, echo=FALSE}
@@
 # customer_zip_code_prefix --> first five digits of customer zip code
-customers_dataset <- read.csv(file.path('data','olist_customers_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) 
+customers_dataset <- read.csv(file.path('data','olist_customers_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) %>% mutate(customer_zip_code_prefix=str_pad(customer_zip_code_prefix, c(5),pad=c('0'),side='right'))
 # %>% mutate(customer_zip_code_prefix=as.factor(customer_zip_code_prefix),customer_id=as.character(customer_id ))
@@
 # seller_zip_code_prefix --> first 5 digits of seller zip code
-sellers_dataset <- read.csv(file.path('data','olist_sellers_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) 
+sellers_dataset <- read.csv(file.path('data','olist_sellers_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) %>% mutate(seller_zip_code_prefix=str_pad(seller_zip_code_prefix, c(5),pad=c('0'),side='right'))
 # TODO review if we need it
@@
 
-Wrangled customers dataset
-- removed data where delivery was not made
-- removed data which was delivered but all the delivery details were not available
-```{r Data wrangling-orders data set, include=FALSE, echo=FALSE}
-# order data set with various delivery related times
-# note each customer_id is unique per order
-# order_purchase_timestamp -> Shows the purchase timestamp.
-# order_approved_at/payment_approved_at -> Shows the payment approval timestamp.
-# order_delivered_carrier_date/handed_over_to_carrier_date -> Shows the order posting timestamp. When it was handled to the logistic partner.
-# order_delivered_customer_date --> Shows the actual order delivery date to the customer.
-# order_estimated_delivery_date --> Shows the estimated delivery date that was informed to customer at the purchase moment.
-orders_dataset <- read.csv(file.path('data','olist_orders_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) %>% 
-  rename(payment_approved_at=order_approved_at,
-         handed_over_to_carrier_date=order_delivered_carrier_date)  %>% 
-  mutate(order_status=as.factor(order_status),
-         order_purchase_timestamp=as_datetime(order_purchase_timestamp),
-         payment_approved_at=as_datetime(payment_approved_at),
-         handed_over_to_carrier_date=as_datetime(handed_over_to_carrier_date),
-         order_delivered_customer_date=as_datetime(order_delivered_customer_date),
-         order_estimated_delivery_date=as_datetime(order_estimated_delivery_date))
-
-# TODO if we are not using payment_approved_at or few fields we can remove the filter logic based on that.
-orders_dataset <- orders_dataset %>% 
-  filter(order_status=='delivered') %>% 
-  filter( !is.na(order_estimated_delivery_date ) & !is.na(order_delivered_customer_date ) & !is.na(handed_over_to_carrier_date) & !is.na(payment_approved_at) & !is.na(order_purchase_timestamp))
-
-orders_dataset %>% head()
-orders_dataset %>% str()
-```
-
-```{r Data wrangling-order items data set, include=FALSE, echo=FALSE}
-# orders data with item related seller info
-# order_item_id --->sequential number identifying number of items included in the same order.
-# so if there are three items they will be nuber 1 , 2, 3 for the given order
-# shipping_limit_date ---> Shows the seller shipping limit date for handling the order over to the logistic partner.
-# freight_value --> item freight value item (if an order has more than one item the freight value is splitted between items)
-order_items_dataset <- read.csv(file.path('data','olist_order_items_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) %>% mutate(shipping_limit_date=as_datetime(shipping_limit_date))
-
-order_items_dataset %>% str()
-order_items_dataset %>% head()
-```
-
-Performance metric
-```{r RMSE, echo=FALSE}
-RMSE <- function(true_outcomes, predicted_outcomes) {
-  sqrt(mean((true_outcomes - predicted_outcomes)^2))
-}
-```
-
-Filter out data where sellers location details are not available
 ```{r Refine sellers dataset, include=FALSE, echo=FALSE}
+# Filter out data where sellers location details are not available
 #
@@
 
-Wrangled sellers dataset ,looks like below:
+The wrangled sellers dataset looks like below:
 ```{r Refined sellers dataset result, echo=FALSE}
-sellers_dataset %>% head() %>% knitr::kable()
+sellers_dataset %>% head(2) %>% as.matrix()  %>% t() %>% 
+  knitr::kable(col.names=c('Example Value1','Example Value2'), caption='Sellers data set')
 ```
@@
 
-Wrangled customers data set looks like below:
-
+The wrangled customers dataset looks like below:
 ```{r Refined customers dataset result, echo=FALSE}
-customers_dataset %>% head() %>% knitr::kable()
+customers_dataset %>% head(2) %>% as.matrix()  %>% t() %>% 
+  knitr::kable(col.names=c('Example Value1','Example Value2'), caption='Customers data set')
 ```
-
-Combine segregated data sets
-- create derived info i.e. distance between two points
 
@@
          product_category_name=as.factor(product_category_name)) %>%
-  select(-customer_id,-order_status,-seller_city,-seller_state,-customer_city, -payment_approved_at,-order_id,-order_item_id,-product_id,-seller_zip_code_prefix,-customer_unique_id)
+  select(-customer_id,-order_status,-seller_city,-seller_state,-customer_city, -payment_approved_at,-order_id,-product_id,-seller_zip_code_prefix,-customer_unique_id)
 ```
 
-Calculating current RMSE range
+Performance metric
+```{r RMSE, echo=FALSE}
+RMSE <- function(true_outcomes, predicted_outcomes) {
+  sqrt(mean((true_outcomes - predicted_outcomes)^2))
+}
+```
+
 ```{r Current RMSE, include=FALSE, echo = FALSE}
@@
 ```
-```{r Range of current rmse, echo=FALSE}
-data.frame(type=character(),value=numeric()) %>% 
-  add_row(type='RMSE Max',value=RMSE_value_based_on_max) %>% 
-  add_row(type='RMSE Min',value=RMSE_value_based_on_min) %>% 
-  knitr::kable(caption='Current RMSE')
-```
 
@@
 
-```{r Apply one hot encoding to product_category_name, include = FALSE}
-set.seed(1,sample.kind = 'Rounding')
-# fullRank parameter to make sure there is no perfect collinearity
-product_category_dummyvars <- data.frame(predict(dummyVars(~product_category_name, data=wrangled_data, fullRank = TRUE), newdata = wrangled_data))
+Having wrangled the datasets individually we combined them into a single dataset.
+While the dataset had many attributes, based on their frequency and analysis, we zeroed on the below fields.
+In addition, it is important to note that for model development we created outcome column number_of_hours based on existing fields i.e.
+order_delivered_customer_date and order_purchase_timestamp as this is what we would want to predict whenever we get a new order i.e.
+given this order, how many number of hours are required to deliver the product to the customer.
 
-wrangled_data <- wrangled_data %>% cbind.data.frame(product_category_dummyvars)
-
-wrangled_data %>% str()
-rm(product_category_dummyvars)
+```{r Final wrangled data,echo=FALSE}
+wrangled_data %>% head(2) %>% as.matrix()  %>% t() %>% 
+  knitr::kable(col.names=c('Example Value1','Example Value2'), caption='Final wrangled data set')
 ```
 
-```{r Final wrangled data,echo=FALSE}
-wrangled_data %>% head() %>% knitr::kable()
+Calculating current performance metric i.e. RMSE.
+```{r Range of current rmse, echo=FALSE}
+data.frame(type=character(),value=numeric()) %>% 
+  add_row(type='RMSE Max',value=RMSE_value_based_on_max) %>% 
+  add_row(type='RMSE Min',value=RMSE_value_based_on_min) %>% 
+  knitr::kable(caption='Current RMSE')
 ```
@@
 ```
+
+Delivery Performance:
+
+You will also be able to work through delivery performance and find ways to optimize delivery times.
+
+Let's predict delivery time/date
+-- delivery distance
+-- location of customer as some can be remotely connected and need special delivery merchanism. This is independent of delivery distance.
+-- location of seller as shipping from some locations maybe difficult. This is independent of delivery distance.
+-- size of product? 
+-- weight of the product. Heavier products may take time to deliver
+-- price of product (expensive nature of product as it could mean extra care or it maybe fragile?)
+-- purchase time....if in night, delivery time maybe longer
+-- order approved at time-> if payment approval came multiple hours after purchase time, this maybe longer
+-- product category -- products of same category may show similar behavior.
+-- or specific product - some products may have higher delivery time.
+-- deliveries maybe impacted due to sales, festival season.
+
+we should consider if we need time stamp or we can take purchase time as 0 and calculated approved at from it.
+though we may lose granularity in terms of time as truck may come everyday at 4.
+so after 4 you can ship , before 4 you cant
+We can be agnostic to purchase time and just predict number of hours for delivery for given purchase.
+
+estimated delivery date is present in ---> order_estimated_delivery_date
+We can calculate RMSE using that and it will be our benchmark
+
+We can see if we improved than that or not. It will give us indication of how close are we to the production model RMSE.
+
+TODO -> if there is delayed to carrier precendence, we could use that probability and average delay to carrier to fine tune our results.
+Similarly we could use other metrics like payment approval time etc.
 
@@
 
+```{r Get delay in payment approval based on customer zip code prefix, include = FALSE}
+# calculated in hours
+# can be due to fraud check etc.
+get_delay_in_payment_approval_per_customer_zip_code <- function(df) {
+  df %>% mutate(payment_approved_at=as.numeric(payment_approved_at)) %>%
+    mutate(time_to_approve_payment_hours=(payment_approved_at-order_purchase_timestamp)/3600) %>% 
+    group_by(customer_zip_code_prefix) %>% 
+    summarise(mean_time_to_approve_payment_hours=mean(time_to_approve_payment_hours))
+}
+
+```
+
 ```{r Method to compute volume from available fields, include = FALSE}
@@
 ```
+
+## Paritioning dataset for training and validation
+
+Before using the wrangled data for analysis and data modeling, we partitioned the data into training and validation data sets.
+The idea was to keep validation data set aside and use only training data for any analysis, model development and tuning.
+
+A split of 80/20 was made here between training and validation set to ensure we have enough data points to train and validate the results.
+Secondly, the split was done such that distribution of outcome
+
+The data sets were named as below for furture reference:
+
+* training data set as **edx** 
+* validation data set as **validation**
 
@@
 
-## Explore product_category_name relation with number_of_hours
-
-```{r Explore product_category_name relation with number_of_hours, echo=FALSE}
-training_dataset %>% group_by(product_category_name) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% mutate(product_category_name=reorder(product_category_name,mean)) %>% ggplot(aes(x=product_category_name,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))
-```
-
-We can see that number_of_hours of delivery varies as number product category changes
-
-##Explore price relation with number_of_hours
-
-```{r Explore price relation with number_of_hours, echo = FALSE}
-# first let see how are the items associated with orders priced. Most of the order items a
-training_dataset %>% mutate(price_group=round(price/100)+1) %>% ggplot(aes(price_group)) + geom_histogram(binwidth=1) + xlab('Each price_group is of 100 units each')
-
-# let's plot errorbar for number of hours vs price_group.
-# we can see while for price groups number of hours for delivery vary but the relationship is not clear enough. Higher priced items can also be delivered early
-training_dataset %>% mutate(price_group=round(price/50)+1) %>% mutate(price_group=as.factor(price_group)) %>% group_by(price_group) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% ggplot(aes(x=price_group,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))
-
-cor(training_dataset$price,training_dataset$number_of_hours)
-```
-
-Price Group for a product does not give any clear indication. It does not seem to be impacting delivery time of product.
-
-## Explore distance relationship between two points with number_of_hours
-
-```{r Explore distance relationship between two points with number_of_hours, echo = FALSE}
-# let's plot errorbar for number of hours vs distance group
-# we can see a pattern that as distance increases delivery time increases. 
-training_dataset %>% mutate(distance_group=round(distance_between_two_points/50)+1) %>% mutate(distance_group=as.factor(distance_group)) %>% group_by(distance_group) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% ggplot(aes(x=distance_group,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))
-
-cor(training_dataset$distance_between_two_points,training_dataset$number_of_hours)
-```
-
-Distance Group can be seen to influence delivery time.
-
-```{r Explore freight value group with number_of_hours}
-# let's plot errorbar for number of hours vs freight value group
-training_dataset %>% mutate(freight_value_group=round(freight_value/5)+1) %>% mutate(freight_value_group=as.factor(freight_value_group)) %>% group_by(freight_value_group) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% ggplot(aes(x=freight_value_group,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))
-
-cor(training_dataset$freight_value,training_dataset$number_of_hours)
-```
-
-Above graphs indicates that for freight value group where each group represents 5 units, as freight value increases, number of hours for delivery increases.
-
-However, it dips near the end
-```{r smooth fit for freight value, echo=FALSE}
-training_dataset %>% ggplot(aes(x=freight_value,y=number_of_hours)) + geom_smooth() 
-```
-
-## Explore volume of product group with number_of_hours
-
-```{r Explore volume of product group with number_of_hours, echo=FALSE}
-# let's plot errorbar for number of hours vs volume of product group
-training_dataset %>% mutate(volume_group=round(volume/1000)+1) %>% mutate(volume_group=as.factor(volume_group)) %>% group_by(volume_group) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% ggplot(aes(x=volume_group,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))
-```
-It does not influencee much the number of hours
-
-However a smooth fit shows some pattern 
-```{r smooth fit for volume, echo=FALSE}
-training_dataset %>% ggplot(aes(volume,number_of_hours)) + geom_smooth()
-```
-
-## Explore weight of product with number_of_hours
-
-```{r Explore weight of product with number_of_hours}
-# let's plot errorbar for number of hours vs weight of product
-training_dataset %>% mutate(weight_group=round((product_weight_g)/1000)+1) %>% mutate(weight_group=as.factor(weight_group)) %>% mutate(weight_group=as.factor(weight_group)) %>% group_by(weight_group) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% ggplot(aes(x=weight_group,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))
-```
-Some clues but not strongly correlated.
-
-```{r Smooth fit for product, echo= FALSE}
-training_dataset  %>% ggplot(aes(product_weight_g,number_of_hours)) + geom_smooth()
-```
-
-## Explore truncated customer_zip_code_prefix with number_of_hours
-
-```{r Explore truncated customer_zip_code_prefix with number_of_hours}
-# let's plot errorbar for number of hours vs truncated customer_zip_code_prefix
-training_dataset %>% mutate(czc = as.numeric(str_extract(customer_zip_code_prefix,"\\d\\d\\d"))) %>% group_by(czc) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% mutate(czc=reorder(czc,mean)) %>% ggplot(aes(x=czc,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))
-```
-
-```{r Prepare training and test data for ,include=FALSE}
-# penalty_data is computed on train_subset and is used with test_subset as historical learning of efficiency of each seller in delivering to carrier
-train_subset <- train_subset %>% left_join(penalty_data,by='seller_id') %>% select(-handed_over_to_carrier_date, -shipping_limit_date)
-
-test_subset <- test_subset %>% left_join(penalty_data,by='seller_id') %>% select(-handed_over_to_carrier_date, -shipping_limit_date)
-```
-
-==TODO note penalty data should be calculated again if train_subset is changed. Only trained data should be used for test, so that there is no data leak.
-
-```{r Dataframe to store target encoding weights, include=FALSE}
-target_encoding_weights <- data.frame(effect=character(0),weight=double(0)) 
-```
-
-```{r Apply Target encoding for seller_id, echo = FALSE}
-#TODO you may need to remove customer_state and seller_state from dataset train_subset_for_encoding before proceeding further.
-
-  train_subset_for_encoding <- train_subset[1:30000,] %>% select(-customer_loc, -customer_zip_code_prefix, -customer_state)
-
-  global_mean <- mean(train_subset_for_encoding$number_of_hours)
-
-  set.seed(1,sample.kind='Rounding')
-  test_weight_indices <- createDataPartition(train_subset_for_encoding$number_of_hours,p=0.2,times=3,list=FALSE)
-
-  weights <- seq(0,6,3)
-
-  weight_to_use_per_partition <- sapply(1:3, function(index) {
-  #lambda_data_set <- prepare_lambda_data_set(index)
-  #train_lambda_set <- lambda_data_set[[1]]
-  #test_lambda_set <- lambda_data_set[[2]]  
-  train_weight_set <- train_subset_for_encoding[-test_weight_indices[,index],]
-  test_weight_set <- train_subset_for_encoding[test_weight_indices[,index],]
-
-  test_weight_set <- test_weight_set %>%
-    semi_join(train_weight_set,by='product_category_name') %>%
-    semi_join(train_weight_set,by='seller_id')
-
-  weight_rmses <- sapply(weights, function(weight) {
-      seller_id_effects <- train_weight_set %>% 
-        group_by(seller_id) %>% 
-        summarise(mean=mean(number_of_hours),n=n()) %>%
-        mutate(mean=(n*mean+weight*global_mean)/(n+weight)) %>% 
-        select(-n) %>% rename()
-
-      train_weight_set <- train_weight_set %>% left_join(seller_id_effects,by='seller_id') %>% rename(seller_id_effect=mean) %>% select(-product_category_name,-seller_id)
-
-      test_weight_set <- test_weight_set %>% left_join(seller_id_effects,by='seller_id') %>% rename(seller_id_effect=mean) %>% select(-product_category_name,-seller_id)
-
-      set.seed(seed=1,sample.kind = 'Rounding')
-      preProcessScalingValues <- preProcess(train_weight_set %>% select(-number_of_hours), method = c("center", "scale"))
-
-      scaled_train_subset <- predict(preProcessScalingValues, train_weight_set)
-      scaled_test_subset <- predict(preProcessScalingValues, test_weight_set)
-
-      # TODO remove below scaling was erroneous. train scaling needs to be used for test data
-      # below is just demo of how to use scale
-      # scaled_train_subset <- train_subset %>% mutate(across(c(mean_delay_in_handover,distance_between_two_points,volume,product_weight_g,freight_value,price),scale)) 
-      # scaled_test_subset <- test_subset %>% mutate(across(c(mean_delay_in_handover,distance_between_two_points,volume,product_weight_g,freight_value,price),scale))
-
-      # TODO can we skip modelling while evaluating weight to use or for target encoding?
-
-      tuningControl=trainControl(method = "cv", number = 2, p = .8)
-      knn_fit_for_weight <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(27,34,41)),trControl=tuningControl)
-
-      knn_fit_for_weight
-
-      predictions <- predict(knn_fit_for_weight,scaled_test_subset %>% select(-number_of_hours))
-
-      RMSE(scaled_test_subset$number_of_hours,predictions) 
-  })
-
-  # selecting weight leading to lowest RMSE
-  weight_to_use <- weights[which.min(weight_rmses)]
-  paste('Weight to use: ', weight_to_use, ', corresponding RMSE: ', weight_rmses[which.min(weight_rmses)])
-
-  print(weight_rmses)
-  weight_to_use
-})
-
-weight_to_use_per_partition
-
-weight_to_use <- mean(weight_to_use_per_partition)  
-# 3
-
-target_encoding_weights <- target_encoding_weights %>% add_row(effect='seller_id',weight=weight_to_use)
-```
-
-```{r Post regularization for seller_id effect, echo=FALSE}
-
-# prepare seller_if_effect based on regularisation weight we have identified
-  global_mean <- mean(train_subset$number_of_hours)
-
-  set.seed(1,sample.kind='Rounding')
-  test_weight_indices <- createDataPartition(train_subset$number_of_hours,p=0.2,times=5,list=FALSE)
-
-seller_id_effects_list <- lapply(1:5, function(index) {
-    train_weight_set <- train_subset[-test_weight_indices[,index],]
-
-    seller_id_effects <- train_weight_set %>% 
-      group_by(seller_id) %>%
-      summarise(mean=mean(number_of_hours),n=n()) %>%
-      mutate(mean=(n*mean+weight_to_use*global_mean)/(n+weight_to_use)) %>% 
-      select(-n) %>% rename()
-
-    seller_id_effects
-  }
-)
-
-sie <- seller_id_effects_list[[1]]
-for (index in 2:5) {
-   sie <- sie %>% full_join(seller_id_effects_list[[index]],by='seller_id', suffix=c("1",index))
-}
-
-# whereever value is NA, we have used global mean
-sie <- sie %>% mutate(across(-seller_id,~if_else(is.na(.x),global_mean,.x)))
-
-# as different partitions may not have exactly same data set combined the results
-seller_id_effects <- tibble(sie[,1], seller_id_effect=rowMeans(sie[,-1],na.rm = TRUE))
-
-train_subset <- train_subset %>% left_join(seller_id_effects,by='seller_id') %>% select(-product_category_name,-seller_id) %>% mutate(seller_id_effect=if_else(is.na(seller_id_effect),global_mean,seller_id_effect))
-
-test_subset <- test_subset %>% left_join(seller_id_effects,by='seller_id') %>% select(-product_category_name,-seller_id)
-```
-
-```{r Apply Target encoding for customer_state, echo = FALSE}
-  train_subset_for_encoding <- train_subset[1:40000,] %>% select(-customer_loc,-customer_zip_code_prefix)
-  global_mean <- mean(train_subset_for_encoding$number_of_hours)
-
-  set.seed(1,sample.kind='Rounding')
-  test_weight_indices <- createDataPartition(train_subset_for_encoding$number_of_hours,p=0.2,times=3,list=FALSE)
-
-  weights <- seq(0,30,15)
-
-  weight_to_use_per_partition <- sapply(1:3, function(index) {
-  #lambda_data_set <- prepare_lambda_data_set(index)
-  #train_lambda_set <- lambda_data_set[[1]]
-  #test_lambda_set <- lambda_data_set[[2]]  
-  train_weight_set <- train_subset_for_encoding[-test_weight_indices[,index],]
-  test_weight_set <- train_subset_for_encoding[test_weight_indices[,index],]
-
-  test_weight_set <- test_weight_set %>%
-    #semi_join(train_weight_set,by='product_category_name') %>%
-    semi_join(train_weight_set,by='customer_state')
-
-  weight_rmses <- sapply(weights, function(weight) {
-      customer_state_effects <- train_weight_set %>% 
-        group_by(customer_state) %>% 
-        summarise(mean=mean(number_of_hours),n=n()) %>%
-        mutate(mean=(n*mean+weight*global_mean)/(n+weight)) %>% 
-        select(-n) %>% rename()
-
-      train_weight_set <- train_weight_set %>% left_join(customer_state_effects,by='customer_state') %>% rename(customer_state_effect=mean) %>% select(-customer_state) #-product_category_name
-
-      test_weight_set <- test_weight_set %>% left_join(customer_state_effects,by='customer_state') %>% rename(customer_state_effect=mean) %>% select(-customer_state) # -product_category_name,
-
-      set.seed(seed=1,sample.kind = 'Rounding')
-      preProcessScalingValues <- preProcess(train_weight_set %>% select(-number_of_hours), method = c("center", "scale"))
-
-      scaled_train_subset <- predict(preProcessScalingValues, train_weight_set)
-      scaled_test_subset <- predict(preProcessScalingValues, test_weight_set)
-
-      # TODO remove below scaling was erroneous. train scaling needs to be used for test data
-      # below is just demo of how to use scale
-      # scaled_train_subset <- train_subset %>% mutate(across(c(mean_delay_in_handover,distance_between_two_points,volume,product_weight_g,freight_value,price),scale)) 
-      # scaled_test_subset <- test_subset %>% mutate(across(c(mean_delay_in_handover,distance_between_two_points,volume,product_weight_g,freight_value,price),scale))
-
-      # TODO can we skip modelling while evaluating weight to use or for target encoding?
-
-      tuningControl=trainControl(method = "cv", number = 2, p = .8)
-      knn_fit_for_weight <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(20,27,34)),trControl=tuningControl)
-
-      knn_fit_for_weight
-
-      print(knn_fit_for_weight)
-
-      predictions <- predict(knn_fit_for_weight,scaled_test_subset %>% select(-number_of_hours))
-
-      RMSE(scaled_test_subset$number_of_hours,predictions) 
-  })
-
-  # selecting weight leading to lowest RMSE
-  weight_to_use <- weights[which.min(weight_rmses)]
-  paste('Weight to use: ', weight_to_use, ', corresponding RMSE: ', weight_rmses[which.min(weight_rmses)])
-
-  print(weight_rmses)
-  weight_to_use
-})
-
-weight_to_use_per_partition
-
-weight_to_use <- mean(weight_to_use_per_partition) 
-# 20
-
-target_encoding_weights <- target_encoding_weights %>% add_row(effect='customer_state',weight=weight_to_use)
-```
-
-```{r Post regularization with consumer state effect, echo=FALSE}
-
-# prepare consumer_state_effect based on regularisation weight we have identified
-  global_mean <- mean(train_subset$number_of_hours)
-
-  set.seed(1,sample.kind='Rounding')
-  test_weight_indices <- createDataPartition(train_subset$number_of_hours,p=0.2,times=5,list=FALSE)
-
-customer_state_effects_list <- lapply(1:5, function(index) {
-    train_weight_set <- train_subset[-test_weight_indices[,index],]
-
-    customer_state_effects <- train_weight_set %>% 
-      group_by(customer_state) %>%
-      summarise(mean=mean(number_of_hours),n=n()) %>%
-      mutate(mean=(n*mean+weight_to_use*global_mean)/(n+weight_to_use)) %>% 
-      select(-n) %>% rename()
-
-    customer_state_effects
-  }
-)
-
-cse <- customer_state_effects_list[[1]]
-for (index in 2:5) {
-   cse <- cse %>% full_join(customer_state_effects_list[[index]],by='customer_state', suffix=c("1",index))
-}
-
-# whereever value is NA, we have used global mean
-cse <- cse %>% mutate(across(-customer_state,~if_else(is.na(.x),global_mean,.x)))
-
-# as different partitions may not have exactly same data set combined the results
-customer_state_effects <- tibble(cse[,1], customer_state_effect=rowMeans(cse[,-1],na.rm = TRUE))
-
-train_subset <- train_subset %>% left_join(customer_state_effects,by='customer_state') %>% select(-customer_state) %>% mutate(customer_state_effect=if_else(is.na(customer_state_effect),global_mean,customer_state_effect))
-
-test_subset <- test_subset %>% left_join(customer_state_effects,by='customer_state') %>% select(-customer_state)
-```
-
-```{r Apply Target encoding for customer_loc, echo = FALSE}
-  train_subset_for_encoding <- train_subset[1:40000,] %>% select(-customer_zip_code_prefix)
-  global_mean <- mean(train_subset_for_encoding$number_of_hours)
-
-  set.seed(1,sample.kind='Rounding')
-  test_weight_indices <- createDataPartition(train_subset_for_encoding$number_of_hours,p=0.2,times=3,list=FALSE)
-
-  weights <- seq(0,10,5)
-
-  weight_to_use_per_partition <- sapply(1:3, function(index) {
-  #lambda_data_set <- prepare_lambda_data_set(index)
-  #train_lambda_set <- lambda_data_set[[1]]
-  #test_lambda_set <- lambda_data_set[[2]]  
-  train_weight_set <- train_subset_for_encoding[-test_weight_indices[,index],]
-  test_weight_set <- train_subset_for_encoding[test_weight_indices[,index],]
-
-  test_weight_set <- test_weight_set %>%
-    #semi_join(train_weight_set,by='product_category_name') %>%
-    semi_join(train_weight_set,by='customer_loc')
-
-  weight_rmses <- sapply(weights, function(weight) {
-      customer_loc_effects <- train_weight_set %>% 
-        group_by(customer_loc) %>% 
-        summarise(mean=mean(number_of_hours),n=n()) %>%
-        mutate(mean=(n*mean+weight*global_mean)/(n+weight)) %>% 
-        select(-n) %>% rename()
-
-      train_weight_set <- train_weight_set %>% left_join(customer_loc_effects,by='customer_loc') %>% rename(customer_loc_effect=mean) %>% select(-customer_loc) #-product_category_name
-
-      test_weight_set <- test_weight_set %>% left_join(customer_loc_effects,by='customer_loc') %>% rename(customer_loc_effect=mean) %>% select(-customer_loc) # -product_category_name,
-
-      set.seed(seed=1,sample.kind = 'Rounding')
-      preProcessScalingValues <- preProcess(train_weight_set %>% select(-number_of_hours), method = c("center", "scale"))
-
-      scaled_train_subset <- predict(preProcessScalingValues, train_weight_set)
-      scaled_test_subset <- predict(preProcessScalingValues, test_weight_set)
-
-      # TODO remove below scaling was erroneous. train scaling needs to be used for test data
-      # below is just demo of how to use scale
-      # scaled_train_subset <- train_subset %>% mutate(across(c(mean_delay_in_handover,distance_between_two_points,volume,product_weight_g,freight_value,price),scale)) 
-      # scaled_test_subset <- test_subset %>% mutate(across(c(mean_delay_in_handover,distance_between_two_points,volume,product_weight_g,freight_value,price),scale))
-
-      # TODO can we skip modelling while evaluating weight to use or for target encoding?
-
-      tuningControl=trainControl(method = "cv", number = 2, p = .8)
-      knn_fit_for_weight <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(20,27,34)),trControl=tuningControl)
-
-      knn_fit_for_weight
-
-      print(knn_fit_for_weight)
-
-      predictions <- predict(knn_fit_for_weight,scaled_test_subset %>% select(-number_of_hours))
-
-      RMSE(scaled_test_subset$number_of_hours,predictions) 
-  })
-
-  # selecting weight leading to lowest RMSE
-  weight_to_use <- weights[which.min(weight_rmses)]
-  paste('Weight to use: ', weight_to_use, ', corresponding RMSE: ', weight_rmses[which.min(weight_rmses)])
-
-  print(weight_rmses)
-  weight_to_use
-})
-
-weight_to_use_per_partition
-
-weight_to_use <- mean(weight_to_use_per_partition) 
-# 5/3 = 1.66667
-
-target_encoding_weights <- target_encoding_weights %>% add_row(effect='customer_loc',weight=weight_to_use)
-```
-
-```{r Post regularization with customer loc effect, echo=FALSE}
-
-# prepare customer_loc_effect based on regularisation weight we have identified
-  global_mean <- mean(train_subset$number_of_hours)
-
-  set.seed(1,sample.kind='Rounding')
-  test_weight_indices <- createDataPartition(train_subset$number_of_hours,p=0.2,times=10,list=FALSE)
-
-customer_loc_effects_list <- lapply(1:10, function(index) {
-    train_weight_set <- train_subset[-test_weight_indices[,index],]
-
-    customer_loc_effects <- train_weight_set %>% 
-      group_by(customer_loc) %>%
-      summarise(mean=mean(number_of_hours),n=n()) %>%
-      mutate(mean=(n*mean+weight_to_use*global_mean)/(n+weight_to_use)) %>% 
-      select(-n) %>% rename()
-
-    customer_loc_effects
-  }
-)
-
-cle <- customer_loc_effects_list[[1]]
-for (index in 2:10) {
-   cle <- cle %>% full_join(customer_loc_effects_list[[index]],by='customer_loc', suffix=c("1",index))
-}
-
-# whereever value is NA, we have used global mean
-cle <- cle %>% mutate(across(-customer_loc,~if_else(is.na(.x),global_mean,.x)))
-
-# as different partitions may not have exactly same data set combined the results
-customer_loc_effects <- tibble(cle[,1], customer_loc_effect=rowMeans(cle[,-1],na.rm = TRUE))
-
-train_subset <- train_subset %>% left_join(customer_loc_effects,by='customer_loc') %>% select(-customer_loc,-customer_zip_code_prefix) %>% mutate(customer_loc_effect=if_else(is.na(customer_loc_effect),global_mean,customer_loc_effect))
-
-test_subset <- test_subset %>% left_join(customer_loc_effects,by='customer_loc') %>% select(-customer_loc,-customer_zip_code_prefix)
-```
-
-```{r knn model data scaled under test, echo=FALSE}
-# TODO review and improve or remove this as required
-set.seed(seed=1,sample.kind = 'Rounding')
-preProcessScalingValues <- preProcess(train_subset %>% select(-number_of_hours), method = c("center", "scale"))
-
-scaled_train_subset <- predict(preProcessScalingValues, train_subset) 
-scaled_test_subset <- predict(preProcessScalingValues, test_subset)
-
-# TODO remove below scaling was erroneous. train scaling needs to be used for test data
-# below is just demo of how to use scale
-# scaled_train_subset <- train_subset %>% mutate(across(c(mean_delay_in_handover,distance_between_two_points,volume,product_weight_g,freight_value,price),scale)) 
-# scaled_test_subset <- test_subset %>% mutate(across(c(mean_delay_in_handover,distance_between_two_points,volume,product_weight_g,freight_value,price),scale))
-
-tuningControl=trainControl(method = "cv", number = 3, p = .8)
-knn_fit <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(27,34,41)),trControl=tuningControl)
-
-knn_fit
-
-predictions <- predict(knn_fit,scaled_test_subset %>% select(-number_of_hours))
-
-RMSE(predictions,(scaled_test_subset %>% pull(number_of_hours) ))
-#184.7906 for k=34.
-
-save(knn_fit,file='rda/knn_model.rda')
-```
-
-```{r correlation, eval=FALSE}
-cor(train_subset %>% select(price,freight_value,customer_loc_effect,customer_state_effect,seller_id_effect,distance_between_two_points,mean_delay_in_handover,volume,number_of_hours,product_weight_g),use="pairwise.complete.obs") %>% knitr::kable()
-```
-
-Random forest
-However, it is important to know your data and keep in mind that a Random Forest can’t extrapolate.
-It can only make a prediction that is an average of previously observed labels.
-In this sense it is very similar to KNN.
-In other words, in a regression problem, the range of predictions a Random Forest can make is bound by the highest and lowest labels in the training data.
-This behavior becomes problematic in situations where the training and prediction inputs differ in their range and/or distributions.
-This is called covariate shift and it is difficult for most models to handle but especially for Random Forest, because it can’t extrapolate.
-From : https://towardsdatascience.com/a-limitation-of-random-forest-regression-db8ed7419e9f
-
-```{r rf model, echo = FALSE}
-set.seed(seed=1,sample.kind = 'Rounding')
-tuningControl=trainControl(method = "cv", number = 2, p = .8)
-rf_fit <- train(number_of_hours~., train_subset, method='rf',tuneGrid = data.frame(mtry = seq(5, 40, 10)),trControl=tuningControl,ntree=250)
-#rf_fit <- train(number_of_hours~., train_subset %>% select(-seller_state), method='rf',tuneGrid = data.frame(mtry = seq(5, 50, 10)),trControl=tuningControl,ntree=250) # nodesize = 50,maxnodes=2000,ntree=200
-
-rf_fit
-
-predictions <- predict(rf_fit,test_subset %>% select(-number_of_hours))
-
-RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
-# 175.0988 for mtry 15.
-Technically there are total of 82 variables due to 73 product categories and 9 other predictors.
-The seller_state was not of much help.
-it was actually increasing RMSE by .6.
-So have ignored it.
-
-save(rf_fit,file='rda/rf_model.rda')
-```
-
-```{r Gam model, echo = FALSE}
-set.seed(seed=1,sample.kind = 'Rounding')
-preProcessScalingValues <- preProcess(train_subset %>% select(-number_of_hours), method = c("center", "scale"))
-
-scaled_train_subset <- predict(preProcessScalingValues, train_subset) 
-scaled_test_subset <- predict(preProcessScalingValues, test_subset)
-
-tuningControl=trainControl(method = "cv", number =3, p = .8)
-gam_fit <- train(number_of_hours~., scaled_train_subset, method='gam',trControl=tuningControl) # check for tuning parameters
-
-gam_fit
-
-predictions <- predict(gam_fit,scaled_test_subset %>% select(-number_of_hours))
-
-RMSE(predictions,(scaled_test_subset %>% pull(number_of_hours) ))
-# 185.1074 uses GCV.Cp method and select=FALSE
-
-save(gam_fit,file='rda/gam_model.rda')
-```
-
-```{r Prepare seller_id effect based on target encoding, include = FALSE}
-# prepare seller_id_effect based on regularisation weight we have identified
-prepare_seller_id_effect <- function(train_df, weight_to_use) {
-    global_mean <- mean(train_df$number_of_hours)
-
-    set.seed(1,sample.kind='Rounding')
-    test_weight_indices <- createDataPartition(train_df$number_of_hours, p=0.2,times=10,list=FALSE)
-
-    seller_id_effects_list <- lapply(1:10, function(index) {
-      train_weight_set <- train_df[-test_weight_indices[,index],]
-
-      seller_id_effects <- train_weight_set %>% 
-        group_by(seller_id) %>%
-        summarise(mean=mean(number_of_hours),n=n()) %>%
-        mutate(mean=(n*mean+weight_to_use*global_mean)/(n+weight_to_use)) %>% 
-        select(-n) %>% rename()
-
-      seller_id_effects
-    })
-
-    sie <- seller_id_effects_list[[1]]
-    for (index in 2:10) {
-       sie <- sie %>% 
-         full_join(seller_id_effects_list[[index]],by='seller_id', suffix=c("1",index))
-    }
-
-    # whereever value is NA, we have used global mean
-    sie <- sie %>% mutate(across(-seller_id,~if_else(is.na(.x),global_mean,.x)))
-
-    # as different partitions may not have exactly same data set combined the results
-    seller_id_effects <- tibble(sie[,1], seller_id_effect=rowMeans(sie[,-1],na.rm = TRUE))
-
-    #seller id effects are ready
-    seller_id_effects
-}
-```
-
-```{r Prepare customer_state effect based on target encoding, include = FALSE}
-# prepare consumer_state_effect based on regularisation weight we have identified
-prepare_customer_state_effect <- function(train_df, weight_to_use) {
-  global_mean <- mean(train_df$number_of_hours)
-
-  set.seed(1,sample.kind='Rounding')
-  test_weight_indices <- createDataPartition(train_df$number_of_hours,p=0.2,times=10,list=FALSE)
-
-  customer_state_effects_list <- lapply(1:10, function(index) {
-      train_weight_set <- train_df[-test_weight_indices[,index],]
-
-      customer_state_effects <- train_weight_set %>% 
-        group_by(customer_state) %>%
-        summarise(mean=mean(number_of_hours),n=n()) %>%
-        mutate(mean=(n*mean+weight_to_use*global_mean)/(n+weight_to_use)) %>% 
-        select(-n) %>% rename()
-
-      customer_state_effects
-    }
-  )
-
-  cse <- customer_state_effects_list[[1]]
-  for (index in 2:10) {
-     cse <- cse %>% 
-       full_join(customer_state_effects_list[[index]],by='customer_state', suffix=c("1",index))
-  }
-
-  # whereever value is NA, we have used global mean
-  cse <- cse %>% mutate(across(-customer_state,~if_else(is.na(.x),global_mean,.x)))
-
-  # as different partitions may not have exactly same data set combined the results
-  customer_state_effects <- tibble(cse[,1], customer_state_effect=rowMeans(cse[,-1],na.rm = TRUE))
-
-  #customer state effects are ready
-  customer_state_effects
-}
-```
-
-```{r Prepare customer_loc effect based on target encoding, include = FALSE}
-# prepare customer_loc_effect based on regularisation weight we have identified
-prepare_customer_loc_effect <- function(train_df, weight_to_use) {
-  global_mean <- mean(train_df$number_of_hours)
-
-  set.seed(1,sample.kind='Rounding')
-  test_weight_indices <- createDataPartition(train_df$number_of_hours,p=0.2,times=10,list=FALSE)
-
-  customer_loc_effects_list <- lapply(1:10, function(index) {
-      train_weight_set <- train_df[-test_weight_indices[,index],]
-
-      customer_loc_effects <- train_weight_set %>% 
-        group_by(customer_loc) %>%
-        summarise(mean=mean(number_of_hours),n=n()) %>%
-        mutate(mean=(n*mean+weight_to_use*global_mean)/(n+weight_to_use)) %>% 
-        select(-n) %>% rename()
-
-      customer_loc_effects
-    }
-  )
-
-  cle <- customer_loc_effects_list[[1]]
-  for (index in 2:10) {
-     cle <- cle %>% full_join(customer_loc_effects_list[[index]],by='customer_loc', suffix=c("1",index))
-  }
-
-  # where-ever value is NA, we have used global mean
-  cle <- cle %>% mutate(across(-customer_loc,~if_else(is.na(.x),global_mean,.x)))
-
-  # as different partitions may not have exactly same data set combined the results
-  customer_loc_effects <- tibble(cle[,1], customer_loc_effect=rowMeans(cle[,-1],na.rm = TRUE))
-
-  #customer loc effects are ready
-  customer_loc_effects
-}
-```
-
-```{r Enhance training and test dataset for seller_id effect based on target encoding, include = FALSE}
-enhance_datasets_using_seller_id_effect <- function(seller_id_effects,train_df,test_df,cleanup) {
-  #we will use seller id effects computed using training data in test data.
-  train_df <- train_df %>% left_join(seller_id_effects,by='seller_id') %>% mutate(seller_id_effect=if_else(is.na(seller_id_effect),global_mean,seller_id_effect))
-
-  test_df <- test_df %>% left_join(seller_id_effects,by='seller_id')
-
-  # cleanup
-  if (cleanup) {
-    train_df <- train_df %>% select(-product_category_name,-seller_id)
-    test_df <- test_df %>% select(-product_category_name,-seller_id)
-  }
-
-  list(train_df,test_df)
-}
-```
-
-```{r Enhance training and test dataset for customer_state effect based on target encoding, include = FALSE}
-enhance_datasets_using_customer_state_effect <- function(customer_state_effects,train_df,test_df,cleanup) {
-  #we will use customer state effects computed using training data in test data.
-  train_df <- train_df %>% left_join(customer_state_effects,by='customer_state') %>%
-    mutate(customer_state_effect=if_else(is.na(customer_state_effect),
-                                         global_mean,
-                                         customer_state_effect))
-
-  test_df <- test_df %>% left_join(customer_state_effects,by='customer_state')
-
-  # cleanup
-  if (cleanup) {
-    train_df <- train_df %>% select(-customer_state)
-    test_df <- test_df %>% select(-customer_state)
-  }
-
-  list(train_df,test_df)
-}
-```
-
-```{r Enhance training and test dataset for customer_loc effect based on target encoding, include = FALSE}
-enhance_datasets_using_customer_loc_effect <- function(customer_loc_effects,train_df,test_df,cleanup) {
-  #we will use customer loc effects computed using training data in test data.
-  train_df <- train_df %>% 
-    left_join(customer_loc_effects,by='customer_loc') %>%
-    mutate(customer_loc_effect=if_else(is.na(customer_loc_effect),global_mean,customer_loc_effect))
-
-  test_df <- test_df %>% left_join(customer_loc_effects,by='customer_loc')
-
-  # cleanup
-  if (cleanup) {
-    train_df <- train_df %>% select(-customer_loc,-customer_zip_code_prefix)
-    test_df <- test_df %>% select(-customer_loc,-customer_zip_code_prefix)
-  }
-
-  list(train_df,test_df)
-}
-```
-
-```{r Prepare data set for training, include = FALSE}
-# prepares data for training and testing
-# returns list of data frames. First list has training dataframe. Second list has test data frame
-prepare_data_for_training_testing <- function(train_ds, test_ds, cleanup) {
-  print('Adding column for delay in handover')
-  penalty_data <- train_ds %>% get_delay_in_handover_to_carrier_per_seller_id()
-
-  #we will use penalty_data computed using training data in test data.
-  train_ds <- train_ds %>% left_join(penalty_data,by='seller_id') 
-  test_ds <- test_ds %>% left_join(penalty_data,by='seller_id')
-  if (cleanup) {
-    train_ds <- train_ds %>% select(-handed_over_to_carrier_date,-shipping_limit_date)
-    test_ds <- test_ds %>% select(-handed_over_to_carrier_date,-shipping_limit_date)
-  }
-
-  print('Enhancing data for seller id effects')
-  weight_to_use <- target_encoding_weights %>% filter(effect=='seller_id') %>% pull(weight)
-  seller_id_effects <- prepare_seller_id_effect(train_ds,weight_to_use)
-  dataset_list <- enhance_datasets_using_seller_id_effect(seller_id_effects,train_ds,test_ds,cleanup)
-  train_ds <- dataset_list[[1]]
-  test_ds <- dataset_list[[2]]
-
-  print('Enhancing data for customer state effects')
-  weight_to_use <- target_encoding_weights %>% filter(effect=='customer_state') %>% pull(weight)
-  customer_state_effects <- prepare_customer_state_effect(train_ds,weight_to_use)
-  dataset_list <- enhance_datasets_using_customer_state_effect(customer_state_effects,train_ds,test_ds,cleanup)
-  train_ds <- dataset_list[[1]]
-  test_ds <- dataset_list[[2]]
-
-  print('Enhancing data for customer loc effects')
-  weight_to_use <- target_encoding_weights %>% filter(effect=='customer_loc') %>% pull(weight)
-  customer_loc_effects <- prepare_customer_loc_effect(train_ds,weight_to_use)
-  dataset_list <- enhance_datasets_using_customer_loc_effect(customer_loc_effects,train_ds,test_ds,cleanup)
-  dataset_list
-}
-```
-
-```{r Prepare dataset for final model training, echo = FALSE}
-set.seed(seed=1,sample.kind = 'Rounding')
-
-validation_dataset <- validation_dataset %>% 
-  compute_volume_using_product_info() %>% 
-  extract_distance_between_seller_customer() %>% 
-  compute_rough_customer_location()
-
-dataset_list <- prepare_data_for_training_testing(training_dataset,validation_dataset,FALSE)
-training_dataset <- dataset_list[[1]]
-validation_dataset <- dataset_list[[2]]
-
-validation_dataset <- validation_dataset %>%
-  semi_join(training_dataset,by='product_category_name') %>% 
-  semi_join(training_dataset,by='seller_id') %>% 
-  semi_join(training_dataset,by='customer_state') %>%
-  semi_join(training_dataset,by='customer_loc')
-
-# cleanup
-training_dataset <- training_dataset %>%
-  select(-handed_over_to_carrier_date,-shipping_limit_date,-product_category_name,-seller_id,-customer_state,-customer_loc,-customer_zip_code_prefix)
-validation_dataset <- validation_dataset %>%
-  select(-handed_over_to_carrier_date,-shipping_limit_date,-product_category_name,-seller_id,-customer_state,-customer_loc,-customer_zip_code_prefix)
-```
-
-```{r Final Model Training, echo = FALSE}
-set.seed(seed=1,sample.kind = 'Rounding')
-
-mtry_to_use <- rf_fit$finalModel$mtry
-ntree_to_use <- 500 #rf_fit$finalModel$ntree
-#tuningControl <- trainControl(method = "cv", number = 2, p = .8)
-#final_model <- train(number_of_hours~., training_dataset, method='rf',tuneGrid = data.frame(mtry = c(mtry_to_use)),trControl=tuningControl,ntree=ntree_to_use)
-
-final_model <- randomForest(number_of_hours~., training_dataset,mtry= mtry_to_use, ntree=ntree_to_use)
-
-final_model
-
-save(final_model,file='rda/final_model.rda')
-```
-
-```{r Final model validation, echo = FALSE}
-predictions <- predict(final_model,validation_dataset %>% select(-number_of_hours))
-
-RMSE((validation_dataset %>% pull(number_of_hours) ),predictions)
-# 190.184 for mtry=15, ntree=500
-```
-
-===== TODO review from here
-
-Data used maybe incorrect
-Review joins and if multiple orders of same type are being delivered at the same time
-refer
-> training_dataset %>% filter(number_of_hours>=354.73 & number_of_hours<354.74)
-   price freight_value product_category_name product_weight_g number_of_hours volume
-1   59.0         13.43          garden_tools             1550        354.7386  19800
-2   59.0         13.43          garden_tools             1550        354.7386  19800
-3   59.0         13.43          garden_tools             1550        354.7386  19800
-4   59.0         13.43          garden_tools             1550        354.7386  19800
-5   59.0         13.43          garden_tools             1550        354.7386  19800
-6   59.0         13.43          garden_tools             1550        354.7386  19800
-7   59.0         13.43          garden_tools             1550        354.7386  19800
-8   59.0         13.43          garden_tools             1550        354.7386  19800
-9   59.0         13.43          garden_tools             1550        354.7386  19800
-10  59.0         13.43          garden_tools             1550        354.7386  19800
-11  59.0         13.43          garden_tools             1550        354.7386  19800
-12  66.0         14.21       furniture_decor             1150        354.7319  28175
-13  46.9         17.63        sports_leisure              300        354.7378   2304
-   distance_between_two_points mean_delay_in_handover
-1                       469.74                -51.60514
-2                       469.74                -51.60514
-3                       469.74                -51.60514
-4                       469.74                -51.60514
-5                       469.74                -51.60514
-6                       469.74                -51.60514
-7                       469.74                -51.60514
-8                       469.74                -51.60514
-9                       469.74                -51.60514
-10                      469.74                -51.60514
-11                      469.74                -51.60514
-12                      807.32                -56.57523
-13                     1083.46               -147.48749
-
-```{r Derived data}
-# seller to customer distance should be right predictor to consider, as the text says "Those merchants are able to sell their products through the Olist Store and ship them directly to the customers using Olist logistics partners."
-# source: https://www.kaggle.com/olistbr/brazilian-ecommerce
-wrangled_data %>% head() %>% extract_distance_between_seller_customer()
-
-# plot of how 
-p1 <- wrangled_data %>% extract_distance_between_seller_customer() %>% ggplot(aes(distance_between_two_points)) + geom_histogram(binwidth=50)
-ggplotly(p1)
-```
-
-```{r Explore delayed handover to carrier, echo=FALSE }
-# TODO we maybe able to remove this as it is not used, because this is the info which is generally not available
-# identify if there was delay in delivery to carrier
-wrangled_data <- wrangled_data %>% mutate(delayed_to_carrier=if_else((as.numeric(handed_over_to_carrier_date)-as.numeric(shipping_limit_date))>0,TRUE,FALSE))
-
-wrangled_data <- wrangled_data %>% select(-shipping_limit_date,-handed_over_to_carrier_date,-order_estimated_delivery_date)
-```
-
-===TODO review
-
-look for variables which have limited to no results and we can combine them as one categorical variable
-Refer: https://www.analyticsvidhya.com/blog/2015/11/easy-methods-deal-categorical-variables-predictive-modeling/
-Also read : feature hashing
-
-Existing model algorithms should handle regularization. So we should not be required to handle ourselves.
-
-We can use conditional model i.e. for some use different model and for some use different model just like ensemble.
-So basically we could say for these pin codes, delivery time= say mean of delivery time and for others use standard algorthm like random forest etc.
-
-gam?
-gamLoess??
-
-```{r ensemble, echo = FALSE}
-load('rda/gamLoess_model.rda')
-load('rda/gam_model.rda')
-load('rda/knn_model.rda')
-load('rda/lm_model.rda')
-load('rda/rf_model.rda')
-
-predictions_gamLoess <- predict(gamLoess_fit,test_subset %>% select(-number_of_hours))
-predictions_knn <- predict(knn_fit,test_subset %>% select(-number_of_hours))
-predictions_rf <- predict(rf_fit,test_subset %>% select(-number_of_hours))
-predictions_gam <- predict(gam_fit,test_subset %>% select(-number_of_hours))
-predictions_glm <- predict(glm_fit,test_subset %>% select(-number_of_hours))
-
-predictions_df <- data.frame(predictions_gamLoess,predictions_knn,predictions_rf,predictions_gam,predictions_glm)
-
-RMSE(rowMeans(predictions_df),(test_subset %>% pull(number_of_hours) ))
-# 194.8623
-
-RMSE(apply(predictions_df %>% select(predictions_rf,predictions_gam,predictions_knn) %>% as.matrix(),1,'median'),(test_subset %>% pull(number_of_hours) ))
-# 194.25
-```
-
-TODO-- what if we remove freight value as it depends on number of items in order too.as it is divided by number of items
-
-```{r delayed to carrier experiment, eval=FALSE}
-# let's see how delay in carrier influences overall delivery time
-raw_data  %>% mutate(order_purchase_timestamp=as.numeric(order_purchase_timestamp),order_delivered_customer_date=as.numeric(order_delivered_customer_date),product_id=as.factor(product_id),seller_id=as.factor(seller_id),seller_zip_code_prefix=as.factor(seller_zip_code_prefix),customer_unique_id=as.factor(customer_unique_id),customer_zip_code_prefix=as.factor(customer_zip_code_prefix),product_category_name=as.factor(product_category_name)) %>% mutate(number_of_hours=(order_delivered_customer_date-order_purchase_timestamp)/3600) %>% select(-order_delivered_customer_date,-order_purchase_timestamp)  %>% group_by(delayed_to_carrier) %>% summarise(mean(number_of_hours))
-
-# let's see time contribution due to delay or early drop to carrier 
-raw_data  %>% mutate(order_purchase_timestamp=as.numeric(order_purchase_timestamp),order_delivered_customer_date=as.numeric(order_delivered_customer_date),product_id=as.factor(product_id),seller_id=as.factor(seller_id),seller_zip_code_prefix=as.factor(seller_zip_code_prefix),customer_unique_id=as.factor(customer_unique_id),customer_zip_code_prefix=as.factor(customer_zip_code_prefix),product_category_name=as.factor(product_category_name)) %>% mutate(number_of_hours=(order_delivered_customer_date-order_purchase_timestamp)/3600) %>% select(-order_delivered_customer_date,-order_purchase_timestamp)  %>% mutate(delay= (as.numeric(handed_over_to_carrier_date) -as.numeric(shipping_limit_date))/3600) %>% select(delay,delayed_to_carrier) %>% group_by(delayed_to_carrier) %>% summarise(mean(delay))
-
-# sellers of interest - we have sellers to be penalized(Add delivery time) and sellers to be corrected(reduce delivery time)
-seller_ids_of_interest <- raw_data  %>% mutate(order_purchase_timestamp=as.numeric(order_purchase_timestamp),order_delivered_customer_date=as.numeric(order_delivered_customer_date),product_id=as.factor(product_id),seller_id=as.factor(seller_id),seller_zip_code_prefix=as.factor(seller_zip_code_prefix),customer_unique_id=as.factor(customer_unique_id),customer_zip_code_prefix=as.factor(customer_zip_code_prefix),product_category_name=as.factor(product_category_name)) %>% mutate(number_of_hours=(order_delivered_customer_date-order_purchase_timestamp)/3600) %>% select(-order_delivered_customer_date,-order_purchase_timestamp)  %>% mutate(delay= (as.numeric(handed_over_to_carrier_date) -as.numeric(shipping_limit_date))/3600) %>% select(delayed_to_carrier, seller_id) %>% dplyr::count(seller_id,delayed_to_carrier) %>% spread(delayed_to_carrier,n,fill=0) %>% mutate(delay_percentage=`TRUE`/(`TRUE`+`FALSE`),total=`TRUE`+`FALSE`) %>% filter(total>=10 & (delay_percentage >=0.75 | delay_percentage <=0.25)) %>% pull(seller_id)
-
-# dummy data points
-dummy_wrangled_data <- raw_data  %>% mutate(order_purchase_timestamp=as.numeric(order_purchase_timestamp),order_delivered_customer_date=as.numeric(order_delivered_customer_date),product_id=as.factor(product_id),seller_id=as.factor(seller_id),seller_zip_code_prefix=as.factor(seller_zip_code_prefix),customer_unique_id=as.factor(customer_unique_id),customer_zip_code_prefix=as.factor(customer_zip_code_prefix),product_category_name=as.factor(product_category_name)) %>% mutate(number_of_hours=(order_delivered_customer_date-order_purchase_timestamp)/3600) %>% select(-order_delivered_customer_date,-order_purchase_timestamp)  %>% filter(seller_id %in% seller_ids_of_interest) %>% mutate(delay= (as.numeric(handed_over_to_carrier_date) -as.numeric(shipping_limit_date))/3600) %>% select(-shipping_limit_date,-handed_over_to_carrier_date,-payment_approved_at,-order_id,-order_estimated_delivery_date,-order_item_id,-delayed_to_carrier)
-
-penalty_data <- wrangled_data %>% get_delay_in_handover_to_carrier_per_seller_id()
-
-set.seed(1,sample.kind = 'Rounding')
-test_indices <- createDataPartition(dummy_wrangled_data$number_of_hours,times=1,p=0.2,list=FALSE)
-dummy_data <- dummy_wrangled_data[-test_indices,]
-dummy_data <- dummy_data %>% mutate(volume=product_length_cm*product_height_cm*product_width_cm) %>%  select(-customer_unique_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix,-product_length_cm,-product_height_cm,-product_width_cm)
-
-set.seed(1,sample.kind = 'Rounding')
-test_indices <- createDataPartition(dummy_data$number_of_hours,times=1,p=0.2,list=FALSE)
-dummy_train_subset <- dummy_data[-test_indices,]
-dummy_test_subset <- dummy_data[test_indices,]
-
-# this penalty data can later be reused with test data as well
-penalty_data <- dummy_train_subset %>% group_by(seller_id) %>% summarise(mean_delay_in_handover=mean(delay))
-
-dummy_train_subset <- dummy_train_subset %>% left_join(penalty_data,by='seller_id')
-
-# only keeping those rows which have seller_id in training set
-dummy_test_subset <- dummy_test_subset %>% semi_join(dummy_train_subset,by='seller_id')
-
-dummy_train_subset <- dummy_train_subset %>% select(-seller_id,-delay)
-# Let's try with rf model
-set.seed(seed=1,sample.kind = 'Rounding')
-tuningControl=trainControl(method = "cv", number = 2, p = .8)
-# rf_fit_dummy <- train(number_of_hours~., dummy_train_subset, method='rf',tuneGrid = data.frame(mtry = seq(5, 15, 4)),trControl=tuningControl) # nodesize = 50,maxnodes=2000,ntree=200
-rf_fit_dummy <- train(number_of_hours~., dummy_train_subset, method='rf',tuneGrid = data.frame(mtry = seq(20, 40, 8)),trControl=tuningControl) # nodesize = 50,maxnodes=2000,ntree=200
-
-rf_fit_dummy
-
-dummy_test_subset <- dummy_test_subset %>% left_join(penalty_data,by='seller_id') %>% select(-seller_id,-delay)
-predictions <- predict(rf_fit_dummy,dummy_test_subset %>% select(-number_of_hours))
-
-RMSE(predictions,(dummy_test_subset %>% pull(number_of_hours) ))
-#184.0301 for mtry 13. Technically there are total of ~79 variables due to 72-74 product categories and 6 other predictors
-```
-
-# Results
-# Conclusion
-We can make fine predictions if we can predict time for approval, time for handling of goods to carrier and time for delivery separately as final prediction would be based on that.
-But when a new order is made, we only know purchase time
-
-## R Markdown
-
-This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
-
-When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
-You can embed an R code chunk like this:
-
-```{r cars}
-summary(cars)
-```
-
-## Including Plots
-
-You can also embed plots, for example:
-
-```{r pressure, echo=FALSE}
-plot(pressure)
-```
-
-Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
+Regression problem.
+Algorithms considered
+- Random Forest
+- knn - k nearest neighbors 
+- gam

### 2021-07-17 finalReportPrep.Rmd
saved rf model

@@
 # customer_zip_code_prefix --> first five digits of customer zip code
-customers_dataset <- read.csv(file.path('data','olist_customers_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) %>% mutate(customer_zip_code_prefix=str_pad(customer_zip_code_prefix, c(5),pad=c('0'),side='right'))
+customers_dataset <- read.csv(file.path('data','olist_customers_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) 
 # %>% mutate(customer_zip_code_prefix=as.factor(customer_zip_code_prefix),customer_id=as.character(customer_id ))
@@
 # seller_zip_code_prefix --> first 5 digits of seller zip code
-sellers_dataset <- read.csv(file.path('data','olist_sellers_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) %>% mutate(seller_zip_code_prefix=str_pad(seller_zip_code_prefix, c(5),pad=c('0'),side='right'))
+sellers_dataset <- read.csv(file.path('data','olist_sellers_dataset.csv'),header=TRUE,stringsAsFactors = FALSE)
 # TODO review if we need it
@@
          customer_zip_code_prefix=as.factor(customer_zip_code_prefix),
-         product_category_name=as.factor(product_category_name)) %>%
-  select(-customer_id,-order_status,-seller_city,-seller_state,-customer_city, -payment_approved_at,-order_id,-product_id,-seller_zip_code_prefix,-customer_unique_id)
+         product_category_name=as.factor(product_category_name),
+         customer_zip_code_prefix=str_pad(customer_zip_code_prefix, c(5),pad=c('0'),side='right')) %>%
+  select(-customer_id,-order_status,-seller_city,-seller_state,-customer_city,-order_id,-product_id,-seller_zip_code_prefix,-customer_unique_id)
 ```
 
-Performance metric
 ```{r RMSE, echo=FALSE}
+# Performance metric
 RMSE <- function(true_outcomes, predicted_outcomes) {
@@
 
-Final wrangled data looks like below:
 ```{r Prepare final wrangled data set, include=FALSE}
@@
   mutate(number_of_hours=(order_delivered_customer_date-order_purchase_timestamp)/3600) %>%
-  select(-order_purchase_timestamp,-order_delivered_customer_date,-order_estimated_delivery_date)
+  select(-order_delivered_customer_date,-order_estimated_delivery_date)
 ```
@@
 
+The wrangled dataset after incorporating various datasets available to us looked like below:
 ```{r Final wrangled data,echo=FALSE}
@@
 
-Calculating current performance metric i.e. RMSE.
+### Performance Metric
+
+Before we dived into building our delivery prediction system, we identified our metric of choice i.e.
+RMSE for evaluating current delivery estimates being made by Olist and for comparing our delivery prediction system against that, in addition to improving our system as we built it.
+
+*  RMSE is the root mean square error and that's what was used for this project to evaluate models and to optimize them for fitment.
+* As we know, $RMSE = sqrt( 1/N * sum((\hat{y}-y)^2) )$
+    + Here, $\hat{y}$ is the prediction. In context of this project, predicted number of hours for delivering the order item,
+    + and y is the actual outcome. In context of this project, actual number of hours taken for delivering the order item.
+
+#### Current RMSE
+
+Before we dived into building our delivery prediction system, we looked into current RMSE.
+To calculate this, we leveraged order_estimated_delivery_date i.e.
+estimated delivery date and order_delivered_customer_date i.e.
+actual delivery timestamp, which was available in orders data set.
+
+While order_estimated_delivery_date did not have time details, order_delivered_customer_date did.
+Since order_delivered_customer_date could be rounded off to same day or next day, we computed range of RMSE which can be calculated using either of the approach.
+
+This RMSE i.e. our performance metric was later used as benchmark for modeling. The delivery prediction system being built attempted to improve on this metric.
+
+Below were the current RMSE calculations based on how we interpret the actual delivery timestamp.
 ```{r Range of current rmse, echo=FALSE}
-data.frame(type=character(),value=numeric()) %>% 
-  add_row(type='RMSE Max',value=RMSE_value_based_on_max) %>% 
-  add_row(type='RMSE Min',value=RMSE_value_based_on_min) %>% 
-  knitr::kable(caption='Current RMSE')
+performance_metric <- data.frame(type=character(),value=numeric()) %>% 
+  add_row(type='Current RMSE Max',value=RMSE_value_based_on_max) %>% 
+  add_row(type='Current RMSE Min',value=RMSE_value_based_on_min)
+
+performance_metric %>% knitr::kable(caption='Performance Metric(RMSE) Summary')
 ```
@@
 save(wrangled_data,file='rda/wrangled_data.rda')
-rm(geolocation_dataset, order_items_dataset, orders_dataset, product_category_name_translation, products_dataset, sellers_dataset, sellers_to_ignore, customers_dataset, customers_to_ignore)
-# TODO delete raw_data as well.
+rm(geolocation_dataset, order_items_dataset, orders_dataset, product_category_name_translation, products_dataset, sellers_dataset, sellers_to_ignore, customers_dataset, customers_to_ignore, raw_data)
 ```
 
-Delivery Performance:
-
-You will also be able to work through delivery performance and find ways to optimize delivery times.
-
-Let's predict delivery time/date
--- delivery distance
--- location of customer as some can be remotely connected and need special delivery merchanism. This is independent of delivery distance.
--- location of seller as shipping from some locations maybe difficult. This is independent of delivery distance.
--- size of product? 
--- weight of the product. Heavier products may take time to deliver
--- price of product (expensive nature of product as it could mean extra care or it maybe fragile?)
--- purchase time....if in night, delivery time maybe longer
--- order approved at time-> if payment approval came multiple hours after purchase time, this maybe longer
--- product category -- products of same category may show similar behavior.
--- or specific product - some products may have higher delivery time.
--- deliveries maybe impacted due to sales, festival season.
-
-we should consider if we need time stamp or we can take purchase time as 0 and calculated approved at from it.
-though we may lose granularity in terms of time as truck may come everyday at 4.
-so after 4 you can ship , before 4 you cant
-We can be agnostic to purchase time and just predict number of hours for delivery for given purchase.
-
-estimated delivery date is present in ---> order_estimated_delivery_date
-We can calculate RMSE using that and it will be our benchmark
-
-We can see if we improved than that or not. It will give us indication of how close are we to the production model RMSE.
-
-TODO -> if there is delayed to carrier precendence, we could use that probability and average delay to carrier to fine tune our results.
-Similarly we could use other metrics like payment approval time etc.
-
-TODO
-Use target encoding for categorical variables
-Use preprocess for scaling data against training data set. Then use mean, sd from this trained data for test data.
-
-Derived data set should be generated later on observations
-- geolocation usage to capture distance
-- delay and mean delay in handling to carrier
-- number of hours for delivery should be generated later
 ```{r Method to derive distance data, include = FALSE}
@@
     group_by(customer_zip_code_prefix) %>% 
-    summarise(mean_time_to_approve_payment_hours=mean(time_to_approve_payment_hours))
+    summarise(mean_delay_in_payment_approval=mean(time_to_approve_payment_hours))
 }
-
 ```
@@
 
-A split of 80/20 was made here between training and validation set to ensure we have enough data points to train and validate the results.
-Secondly, the split was done such that distribution of outcome
+A split of 80/20 was made here between training and validation set to ensure we have enough data points to train and to validate the results.
+Secondly, the split was done such that distribution of outcome i.e.
+number_of_hours was similar in both training and test data.
 
@@
 
-* training data set as **edx** 
-* validation data set as **validation**
+* training data set as **training_dataset** 
+* validation data set as **validation_dataset**
 
-Split data into training and validation set, where validation set will be used later
 ```{r Split data into training and validation set, include = FALSE, echo = FALSE}
+# Split data into training and validation set, where validation set will be used later
 set.seed(1,sample.kind = 'Rounding')
@@
 
+Below were the details about the elements in each of the partitioned datasets.
+```{r Basic info about training and validation dataset,  echo= FALSE}
+# The partitioned data were named training_dataset and validation_dataset, where training_dataset was our training data set and validation_dataset was our validation data set, which we kept aside as planned earlier.
+
+data.frame(data_set=c("Training data set","Validation data set"),name=c('training_dataset','validation_dataset'), count=c(nrow(training_dataset),nrow(validation_dataset))) %>% knitr::kable()
+```
+
+For analysis, we used only the training dataset.
+
+Since we needed to build delivery prediction system which could estimate number of hours required for delivery for new orders, while maintaining minimum RMSE, we used our intution and understanding of the domain to guide us in deriving insights from available predictors.
+The same are shared below:
+
+* _**Distance between seller and customer**_ could likely influence delivery estimate.
+* _**Delay in payment approval**_ could influence delivery time as it would delay shippment from the seller.
+This could be related to customer zip code as often some customer locations require fraud analysis before payment is approved.
+* _**Delay in handover of product**_ from seller to carrier would influence delivery time as any time spent there is wasted.
+This could be seller related as some sellers maybe known to take more time in processing order compared to other sellers.
+* Some _**customer locations**_ maybe difficult to reach or need special delivery mechanism due to their remote nature.
+Such locations may need more delivery time compared to other locations irrespective of the distance involved.
+We need not analyse for each customer location but a rough area could be used based on combination of zip codes.
+* _**Weight of the product**_ could influence delivery time as heavy items say bed, furniture may be transported slowly while lighter items maybe shipped more quickly.
+* _**Price of the product**_ could influence delivery time.
+Expensive items may require considerable more attention and could influence delivery like priority shipping or longer delivery time due to caution.
+* _**Volume of the product**_ could influence delivery time as smaller products can be shipped together while others may need separate truck and therefore more time for dedicated truck to be available.
+* _**Product category**_ could be used for analysis as different product categories would have different packing, assembly, processing requiremts and shelf life would could lead to impact in delivery time.
+* _**Freight value**_ could indicate complexity of shipping like fragility of the product, distance, toll booth charges on the way among others. This could be considered for analysis.
+* _**Customer state**_ could influence the delivery estimate due to state border checks, paper work before goods delivery among others.
+* Different _**sellers**_ may take different amount of time or sell specific type of products which may require different shipping time.
+So products from a seller selling apparels may be delivered quickly while products from one selling customised products or say furniture may take relatively more time.
+This could be useful information for predicting delivery time.
+
+To start building our algorithm, we further paritioned our training data set i.e.
+training_dataset into a split of 80/20 between train_subset and test_subset to ensure we use only the train_subset for training and tuning the model and test_subset to evaluate the model and measure RMSE.
+This would ensure there is no data leak.
+The split like earlier was done such that distribution of outcome i.e.
+number_of_hours was similar in both training and test data
+
+TODO
+Use target encoding for categorical variables
+Use preprocess for scaling data against training data set. Then use mean, sd from this trained data for test data.
+
+Derived data set should be generated later on observations
+- geolocation usage to capture distance
+- delay and mean delay in handling to carrier
+- number of hours for delivery should be generated later
+
 We use training_dataset and split into into train and test set. So that training_dataset is sufficient to fine tune the model
@@
 # ensure test_subset has products only for concerned product categories and seller_ids which are in train_subset
-test_subset <- test_subset %>% semi_join(train_subset,by='product_category_name') %>% semi_join(train_subset,by='seller_id') %>% 
-  semi_join(train_subset,by="customer_zip_code_prefix")
-#test_subset <- test_subset %>% semi_join(train_subset,by='seller_id')
+test_subset <- test_subset %>% semi_join(train_subset,by='product_category_name') %>%
+  semi_join(train_subset,by='seller_id') %>% 
+  semi_join(train_subset,by="customer_zip_code_prefix") %>% 
+  semi_join(train_subset,by="customer_loc")
 
@@
 
+delay_in_payment_approval_data <- train_subset %>% get_delay_in_payment_approval_per_customer_zip_code()
+
 rm(test_indices)
+```
+
+```{r Prepare training and test data for ,include=FALSE}
+# penalty_data is computed on train_subset and is used with test_subset as historical learning of efficiency of each seller in delivering to carrier
+train_subset <- train_subset %>% 
+  left_join(penalty_data,by='seller_id') %>% 
+  left_join(delay_in_payment_approval_data,by='customer_zip_code_prefix') %>% 
+  select(-handed_over_to_carrier_date, -shipping_limit_date,-order_purchase_timestamp,-payment_approved_at)
+
+test_subset <- test_subset %>% 
+  left_join(penalty_data,by='seller_id') %>% 
+  left_join(delay_in_payment_approval_data,by='customer_zip_code_prefix') %>% 
+  select(-handed_over_to_carrier_date, -shipping_limit_date,-order_purchase_timestamp,-payment_approved_at)
+```
+
+==TODO note penalty data should be calculated again if train_subset is changed. Only trained data should be used for test, so that there is no data leak.
+
+== TODO additive smoothing/Bayes estimator with target encoding
+
+```{r Dataframe to store target encoding weights, include=FALSE}
+target_encoding_weights <- data.frame(effect=character(0),weight=double(0)) 
+```
+
+```{r Prepare product_category effect based on target encoding, include = FALSE}
+# prepare product_category_effect based on regularisation weight we have identified
+prepare_product_category_effect <- function(train_df, weight_to_use) {
+  global_mean <- mean(train_df$number_of_hours)
+
+  set.seed(1,sample.kind='Rounding')
+  test_weight_indices <- createDataPartition(train_df$number_of_hours,p=0.2,times=10,list=FALSE)
+
+  product_category_effects_list <- lapply(1:10, function(index) {
+      train_weight_set <- train_df[-test_weight_indices[,index],]
+
+      product_category_effects <- train_weight_set %>% 
+        group_by(product_category_name) %>%
+        summarise(mean=mean(number_of_hours),n=n()) %>%
+        mutate(mean=(n*mean+weight_to_use*global_mean)/(n+weight_to_use)) %>% 
+        select(-n) %>% rename()
+
+      product_category_effects
+    }
+  )
+
+  pce <- product_category_effects_list[[1]]
+  for (index in 2:10) {
+     pce <- pce %>% full_join(product_category_effects_list[[index]],by='product_category_name', suffix=c("1",index))
+  }
+
+  # wherever value is NA, we have used global mean
+  pce <- pce %>% mutate(across(-product_category_name,~if_else(is.na(.x),global_mean,.x)))
+
+  # as different partitions may not have exactly same data set combined the results
+  product_category_effects <- tibble(pce[,1], product_category_effect=rowMeans(pce[,-1],na.rm = TRUE))
+
+  #product_category effects are ready
+  product_category_effects
+}
+```
+
+```{r Prepare seller_id effect based on target encoding, include = FALSE}
+# prepare seller_id_effect based on regularisation weight we have identified
+prepare_seller_id_effect <- function(train_df, weight_to_use) {
+    global_mean <- mean(train_df$number_of_hours)
+
+    set.seed(1,sample.kind='Rounding')
+    test_weight_indices <- createDataPartition(train_df$number_of_hours, p=0.2,times=10,list=FALSE)
+
+    seller_id_effects_list <- lapply(1:10, function(index) {
+      train_weight_set <- train_df[-test_weight_indices[,index],]
+
+      seller_id_effects <- train_weight_set %>% 
+        group_by(seller_id) %>%
+        summarise(mean=mean(number_of_hours),n=n()) %>%
+        mutate(mean=(n*mean+weight_to_use*global_mean)/(n+weight_to_use)) %>% 
+        select(-n) %>% rename()
+
+      seller_id_effects
+    })
+
+    sie <- seller_id_effects_list[[1]]
+    for (index in 2:10) {
+       sie <- sie %>% 
+         full_join(seller_id_effects_list[[index]],by='seller_id', suffix=c("1",index))
+    }
+
+    # wherever value is NA, we have used global mean
+    sie <- sie %>% mutate(across(-seller_id,~if_else(is.na(.x),global_mean,.x)))
+
+    # as different partitions may not have exactly same data set combined the results
+    seller_id_effects <- tibble(sie[,1], seller_id_effect=rowMeans(sie[,-1],na.rm = TRUE))
+
+    #seller id effects are ready
+    seller_id_effects
+}
+```
+
+```{r Prepare customer_state effect based on target encoding, include = FALSE}
+# prepare consumer_state_effect based on regularisation weight we have identified
+prepare_customer_state_effect <- function(train_df, weight_to_use) {
+  global_mean <- mean(train_df$number_of_hours)
+
+  set.seed(1,sample.kind='Rounding')
+  test_weight_indices <- createDataPartition(train_df$number_of_hours,p=0.2,times=10,list=FALSE)
+
+  customer_state_effects_list <- lapply(1:10, function(index) {
+      train_weight_set <- train_df[-test_weight_indices[,index],]
+
+      customer_state_effects <- train_weight_set %>% 
+        group_by(customer_state) %>%
+        summarise(mean=mean(number_of_hours),n=n()) %>%
+        mutate(mean=(n*mean+weight_to_use*global_mean)/(n+weight_to_use)) %>% 
+        select(-n) %>% rename()
+
+      customer_state_effects
+    }
+  )
+
+  cse <- customer_state_effects_list[[1]]
+  for (index in 2:10) {
+     cse <- cse %>% 
+       full_join(customer_state_effects_list[[index]],by='customer_state', suffix=c("1",index))
+  }
+
+  # wherever value is NA, we have used global mean
+  cse <- cse %>% mutate(across(-customer_state,~if_else(is.na(.x),global_mean,.x)))
+
+  # as different partitions may not have exactly same data set combined the results
+  customer_state_effects <- tibble(cse[,1], customer_state_effect=rowMeans(cse[,-1],na.rm = TRUE))
+
+  #customer state effects are ready
+  customer_state_effects
+}
+```
+
+```{r Prepare customer_loc effect based on target encoding, include = FALSE}
+# prepare customer_loc_effect based on regularisation weight we have identified
+prepare_customer_loc_effect <- function(train_df, weight_to_use) {
+  global_mean <- mean(train_df$number_of_hours)
+
+  set.seed(1,sample.kind='Rounding')
+  test_weight_indices <- createDataPartition(train_df$number_of_hours,p=0.2,times=10,list=FALSE)
+
+  customer_loc_effects_list <- lapply(1:10, function(index) {
+      train_weight_set <- train_df[-test_weight_indices[,index],]
+
+      customer_loc_effects <- train_weight_set %>% 
+        group_by(customer_loc) %>%
+        summarise(mean=mean(number_of_hours),n=n()) %>%
+        mutate(mean=(n*mean+weight_to_use*global_mean)/(n+weight_to_use)) %>% 
+        select(-n) %>% rename()
+
+      customer_loc_effects
+    }
+  )
+
+  cle <- customer_loc_effects_list[[1]]
+  for (index in 2:10) {
+     cle <- cle %>% full_join(customer_loc_effects_list[[index]],by='customer_loc', suffix=c("1",index))
+  }
+
+  # where-ever value is NA, we have used global mean
+  cle <- cle %>% mutate(across(-customer_loc,~if_else(is.na(.x),global_mean,.x)))
+
+  # as different partitions may not have exactly same data set combined the results
+  customer_loc_effects <- tibble(cle[,1], customer_loc_effect=rowMeans(cle[,-1],na.rm = TRUE))
+
+  #customer loc effects are ready
+  customer_loc_effects
+}
+```
+
+```{r Enhance training and test dataset for delay in handover to carrier, include=FALSE}
+enhance_datasets_using_delay_in_handover_to_carrier <- function(mean_delay_in_handover_to_carrier_data,train_df,test_df,cleanup) {
+  #we will use mean_delay_in_handover_to_carrier_data computed using training data in test data.
+  train_df <- train_df %>% left_join(mean_delay_in_handover_to_carrier_data,by='seller_id') 
+  test_df <- test_df %>% left_join(mean_delay_in_handover_to_carrier_data,by='seller_id')
+
+  #cleanup
+  if (cleanup) {
+    train_df <- train_df %>% select(-handed_over_to_carrier_date,-shipping_limit_date)
+    test_df <- test_df %>% select(-handed_over_to_carrier_date,-shipping_limit_date)
+  }
+
+  list(train_df,test_df)
+}
+```
+
+```{r Enhance training and test dataset for product_category effect based on target encoding, include = FALSE}
+enhance_datasets_using_product_category_effect <- function(product_category_effects,train_df,test_df,cleanup) {
+  global_mean <- mean(train_df$number_of_hours)
+
+  #we will use product category effects computed using training data in test data.
+  train_df <- train_df %>% left_join(product_category_effects,by='product_category_name') %>%
+    mutate(product_category_effect=if_else(is.na(product_category_effect),global_mean,product_category_effect))
+
+  test_df <- test_df %>% left_join(product_category_effects,by='product_category_name')
+
+  # cleanup
+  if (cleanup) {
+    train_df <- train_df %>% select(-product_category_name)
+    test_df <- test_df %>% select(-product_category_name)
+  }
+
+  list(train_df,test_df)
+}
+```
+
+```{r Enhance training and test dataset for seller_id effect based on target encoding, include = FALSE}
+enhance_datasets_using_seller_id_effect <- function(seller_id_effects,train_df,test_df,cleanup) {
+  global_mean <- mean(train_df$number_of_hours)
+
+  #we will use seller id effects computed using training data in test data.
+  train_df <- train_df %>% left_join(seller_id_effects,by='seller_id') %>%
+    mutate(seller_id_effect=if_else(is.na(seller_id_effect),global_mean,seller_id_effect))
+
+  test_df <- test_df %>% left_join(seller_id_effects,by='seller_id')
+
+  # cleanup
+  if (cleanup) {
+    train_df <- train_df %>% select(-seller_id)
+    test_df <- test_df %>% select(-seller_id)
+  }
+
+  list(train_df,test_df)
+}
+```
+
+```{r Enhance training and test dataset for customer_state effect based on target encoding, include = FALSE}
+enhance_datasets_using_customer_state_effect <- function(customer_state_effects,train_df,test_df,cleanup) {
+  global_mean <- mean(train_df$number_of_hours)
+
+  #we will use customer state effects computed using training data in test data.
+  train_df <- train_df %>% left_join(customer_state_effects,by='customer_state') %>%
+    mutate(customer_state_effect=if_else(is.na(customer_state_effect),
+                                         global_mean,
+                                         customer_state_effect))
+
+  test_df <- test_df %>% left_join(customer_state_effects,by='customer_state')
+
+  # cleanup
+  if (cleanup) {
+    train_df <- train_df %>% select(-customer_state)
+    test_df <- test_df %>% select(-customer_state)
+  }
+
+  list(train_df,test_df)
+}
+```
+
+```{r Enhance training and test dataset for customer_loc effect based on target encoding, include = FALSE}
+enhance_datasets_using_customer_loc_effect <- function(customer_loc_effects,train_df,test_df,cleanup) {
+  global_mean <- mean(train_df$number_of_hours)
+
+  #we will use customer loc effects computed using training data in test data.
+  train_df <- train_df %>% 
+    left_join(customer_loc_effects,by='customer_loc') %>%
+    mutate(customer_loc_effect=if_else(is.na(customer_loc_effect),global_mean,customer_loc_effect))
+
+  test_df <- test_df %>% left_join(customer_loc_effects,by='customer_loc')
+
+  # cleanup
+  if (cleanup) {
+    train_df <- train_df %>% select(-customer_loc, -customer_zip_code_prefix)
+    test_df <- test_df %>% select(-customer_loc, -customer_zip_code_prefix)
+  }
+
+  list(train_df,test_df)
+}
+```
+
+```{r Prepare data set for training, include = FALSE}
+# prepares data for training and testing
+# returns list of data frames. First list has training dataframe. Second list has test data frame
+prepare_data_for_training_testing <- function(train_ds, test_ds, cleanup) {
+  print('Adding column for delay in handover')
+  penalty_data <- train_ds %>% get_delay_in_handover_to_carrier_per_seller_id()
+
+  #we will use penalty_data computed using training data in test data.
+  train_ds <- train_ds %>% left_join(penalty_data,by='seller_id') 
+  test_ds <- test_ds %>% left_join(penalty_data,by='seller_id')
+  if (cleanup) {
+    train_ds <- train_ds %>% select(-handed_over_to_carrier_date,-shipping_limit_date)
+    test_ds <- test_ds %>% select(-handed_over_to_carrier_date,-shipping_limit_date)
+  }
+
+  print('Adding column for delay in payment approval')
+  delay_in_payment_approval <- train_ds %>% get_delay_in_payment_approval_per_customer_zip_code()
+
+  #we will use delay_in_payment_approval computed using training data in test data.
+  train_ds <- train_ds %>% left_join(delay_in_payment_approval_data,by='customer_zip_code_prefix')
+  test_ds <- test_ds %>% left_join(delay_in_payment_approval_data,by='customer_zip_code_prefix')
+  if (cleanup) {
+    train_df <- train_df %>% select(-customer_zip_code_prefix)
+    test_df <- test_df %>% select(-customer_zip_code_prefix)
+  }
+
+  print('Enhancing data for product category effects')
+  weight_to_use <- target_encoding_weights %>% 
+    filter(effect=='product_category') %>% 
+    pull(weight)
+  product_category_effects <- prepare_product_category_effect(train_ds,weight_to_use)
+  dataset_list <- enhance_datasets_using_product_category_effect(product_category_effects,train_ds,test_ds,cleanup)
+  train_ds <- dataset_list[[1]]
+  test_ds <- dataset_list[[2]]
+
+  print('Enhancing data for seller id effects')
+  weight_to_use <- target_encoding_weights %>% filter(effect=='seller_id') %>% pull(weight)
+  seller_id_effects <- prepare_seller_id_effect(train_ds,weight_to_use)
+  dataset_list <- enhance_datasets_using_seller_id_effect(seller_id_effects,train_ds,test_ds,cleanup)
+  train_ds <- dataset_list[[1]]
+  test_ds <- dataset_list[[2]]
+
+  print('Enhancing data for customer state effects')
+  weight_to_use <- target_encoding_weights %>% filter(effect=='customer_state') %>% pull(weight)
+  customer_state_effects <- prepare_customer_state_effect(train_ds,weight_to_use)
+  dataset_list <- enhance_datasets_using_customer_state_effect(customer_state_effects,train_ds,test_ds,cleanup)
+  train_ds <- dataset_list[[1]]
+  test_ds <- dataset_list[[2]]
+
+  print('Enhancing data for customer loc effects')
+  weight_to_use <- target_encoding_weights %>% filter(effect=='customer_loc') %>% pull(weight)
+  customer_loc_effects <- prepare_customer_loc_effect(train_ds,weight_to_use)
+  dataset_list <- enhance_datasets_using_customer_loc_effect(customer_loc_effects,train_ds,test_ds,cleanup)
+  dataset_list
+}
+```
+
+### Product category effects
+As part of this section, we looked into our intution that different product categories may have different delivery timelines and explored its influence by incoporating it in our model for analysis.
+
+Since product categories are categorical in nature, we applied target encoding on the product categories attribute.
+
+===TODO talk about weight
+===TODO talk about need of regularization.
+
+Below histogram depicts how mean rating of movie across movies is distributed.
+
+```{r Apply Target encoding for product_category_name, echo = FALSE}
+#TODO you may need to remove customer_state and seller_state from dataset train_subset_for_encoding before proceeding further.
+
+  train_subset_for_encoding <- train_subset[1:30000,] %>% select(-seller_id,-customer_loc, -customer_zip_code_prefix, -customer_state)
+
+  global_mean <- mean(train_subset_for_encoding$number_of_hours)
+
+  set.seed(1,sample.kind='Rounding')
+  test_weight_indices <- createDataPartition(train_subset_for_encoding$number_of_hours,p=0.2,times=3,list=FALSE)
+
+  weights <- seq(0,10,5)
+
+  weight_to_use_per_partition <- sapply(1:3, function(index) {
+  #lambda_data_set <- prepare_lambda_data_set(index)
+  #train_lambda_set <- lambda_data_set[[1]]
+  #test_lambda_set <- lambda_data_set[[2]]  
+  train_weight_set <- train_subset_for_encoding[-test_weight_indices[,index],]
+  test_weight_set <- train_subset_for_encoding[test_weight_indices[,index],]
+
+  test_weight_set <- test_weight_set %>%
+    semi_join(train_weight_set,by='product_category_name')
+
+  weight_rmses <- sapply(weights, function(weight) {
+      product_category_effects <- train_weight_set %>% 
+        group_by(product_category_name) %>% 
+        summarise(mean=mean(number_of_hours),n=n()) %>%
+        mutate(mean=(n*mean+weight*global_mean)/(n+weight)) %>% 
+        select(-n) %>% rename()
+
+      train_weight_set <- train_weight_set %>% left_join(product_category_effects,by='product_category_name') %>% rename(product_category_effect=mean) %>% select(-product_category_name)
+
+      test_weight_set <- test_weight_set %>% left_join(product_category_effects,by='product_category_name') %>% rename(product_category_effect=mean) %>% select(-product_category_name)
+
+      set.seed(seed=1,sample.kind = 'Rounding')
+      preProcessScalingValues <- preProcess(train_weight_set %>% select(-number_of_hours), method = c("center", "scale"))
+
+      scaled_train_subset <- predict(preProcessScalingValues, train_weight_set)
+      scaled_test_subset <- predict(preProcessScalingValues, test_weight_set)
+
+      # TODO remove below scaling was erroneous. train scaling needs to be used for test data
+      # below is just demo of how to use scale
+      # scaled_train_subset <- train_subset %>% mutate(across(c(mean_delay_in_handover,distance_between_two_points,volume,product_weight_g,freight_value,price),scale)) 
+      # scaled_test_subset <- test_subset %>% mutate(across(c(mean_delay_in_handover,distance_between_two_points,volume,product_weight_g,freight_value,price),scale))
+
+      # TODO can we skip modelling while evaluating weight to use or for target encoding?
+
+      tuningControl=trainControl(method = "cv", number = 3, p = .8)
+      knn_fit_for_weight <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(27,34,41)),trControl=tuningControl)
+
+      knn_fit_for_weight
+      print(knn_fit_for_weight)
+
+      predictions <- predict(knn_fit_for_weight,scaled_test_subset %>% select(-number_of_hours))
+
+      RMSE(scaled_test_subset$number_of_hours,predictions) 
+  })
+
+  # selecting weight leading to lowest RMSE
+  weight_to_use <- weights[which.min(weight_rmses)]
+  paste('Weight to use: ', weight_to_use, ', corresponding RMSE: ', weight_rmses[which.min(weight_rmses)])
+
+  print(weight_rmses)
+  weight_to_use
+})
+
+weight_to_use_per_partition
+
+weight_to_use <- mean(weight_to_use_per_partition)  
+# 20/3
+
+target_encoding_weights <- target_encoding_weights %>% add_row(effect='product_category',weight=weight_to_use)
+```
+
+```{r Post regularization for product_category effect, include=FALSE}
+weight_to_use <- target_encoding_weights %>% filter(effect=='product_category') %>% pull(weight)
+
+product_category_effects <- prepare_product_category_effect(train_subset,weight_to_use)
+dataset_list <- enhance_datasets_using_product_category_effect(product_category_effects,train_subset,test_subset,TRUE)
+
+train_subset <- dataset_list[[1]]
+test_subset <- dataset_list[[2]]
+```
+
+```{r Apply Target encoding for seller_id, echo = FALSE}
+#TODO you may need to remove customer_state and seller_state from dataset train_subset_for_encoding before proceeding further.
+
+  train_subset_for_encoding <- train_subset[1:30000,] %>% select(-customer_loc, -customer_zip_code_prefix, -customer_state)
+
+  global_mean <- mean(train_subset_for_encoding$number_of_hours)
+
+  set.seed(1,sample.kind='Rounding')
+  test_weight_indices <- createDataPartition(train_subset_for_encoding$number_of_hours,p=0.2,times=3,list=FALSE)
+
+  weights <- seq(0,6,3)
+
+  weight_to_use_per_partition <- sapply(1:3, function(index) {
+  #lambda_data_set <- prepare_lambda_data_set(index)
+  #train_lambda_set <- lambda_data_set[[1]]
+  #test_lambda_set <- lambda_data_set[[2]]  
+  train_weight_set <- train_subset_for_encoding[-test_weight_indices[,index],]
+  test_weight_set <- train_subset_for_encoding[test_weight_indices[,index],]
+
+  test_weight_set <- test_weight_set %>% semi_join(train_weight_set,by='seller_id')
+
+  weight_rmses <- sapply(weights, function(weight) {
+      seller_id_effects <- train_weight_set %>% 
+        group_by(seller_id) %>% 
+        summarise(mean=mean(number_of_hours),n=n()) %>%
+        mutate(mean=(n*mean+weight*global_mean)/(n+weight)) %>% 
+        select(-n) %>% rename()
+
+      train_weight_set <- train_weight_set %>% left_join(seller_id_effects,by='seller_id') %>% rename(seller_id_effect=mean) %>% select(-seller_id)
+
+      test_weight_set <- test_weight_set %>% left_join(seller_id_effects,by='seller_id') %>% rename(seller_id_effect=mean) %>% select(-seller_id)
+
+      set.seed(seed=1,sample.kind = 'Rounding')
+      preProcessScalingValues <- preProcess(train_weight_set %>% select(-number_of_hours), method = c("center", "scale"))
+
+      scaled_train_subset <- predict(preProcessScalingValues, train_weight_set)
+      scaled_test_subset <- predict(preProcessScalingValues, test_weight_set)
+
+      # TODO remove below scaling was erroneous. train scaling needs to be used for test data
+      # below is just demo of how to use scale
+      # scaled_train_subset <- train_subset %>% mutate(across(c(mean_delay_in_handover,distance_between_two_points,volume,product_weight_g,freight_value,price),scale)) 
+      # scaled_test_subset <- test_subset %>% mutate(across(c(mean_delay_in_handover,distance_between_two_points,volume,product_weight_g,freight_value,price),scale))
+
+      # TODO can we skip modelling while evaluating weight to use or for target encoding?
+
+      tuningControl=trainControl(method = "cv", number = 3, p = .8)
+      knn_fit_for_weight <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(27,34,41)),trControl=tuningControl)
+
+      knn_fit_for_weight
+
+      print(knn_fit_for_weight)
+
+      predictions <- predict(knn_fit_for_weight,scaled_test_subset %>% select(-number_of_hours))
+
+      RMSE(scaled_test_subset$number_of_hours,predictions) 
+  })
+
+  # selecting weight leading to lowest RMSE
+  weight_to_use <- weights[which.min(weight_rmses)]
+  paste('Weight to use: ', weight_to_use, ', corresponding RMSE: ', weight_rmses[which.min(weight_rmses)])
+
+  print(weight_rmses)
+  weight_to_use
+})
+
+weight_to_use_per_partition
+
+weight_to_use <- mean(weight_to_use_per_partition)  
+# 3
+
+target_encoding_weights <- target_encoding_weights %>% add_row(effect='seller_id',weight=weight_to_use)
+```
+
+```{r Post regularization for seller_id effect, include=FALSE}
+weight_to_use <- target_encoding_weights %>% filter(effect=='seller_id') %>% pull(weight)
+
+seller_id_effects <- prepare_seller_id_effect(train_subset,weight_to_use)
+dataset_list <- enhance_datasets_using_seller_id_effect(seller_id_effects,train_subset,test_subset,TRUE)
+
+train_subset <- dataset_list[[1]]
+test_subset <- dataset_list[[2]]
+```
+
+```{r Apply Target encoding for customer_state, echo = FALSE}
+  train_subset_for_encoding <- train_subset[1:40000,] %>% select(-customer_loc,-customer_zip_code_prefix)
+  global_mean <- mean(train_subset_for_encoding$number_of_hours)
+
+  set.seed(1,sample.kind='Rounding')
+  test_weight_indices <- createDataPartition(train_subset_for_encoding$number_of_hours,p=0.2,times=3,list=FALSE)
+
+  weights <- seq(0,30,15)
+
+  weight_to_use_per_partition <- sapply(1:3, function(index) {
+  #lambda_data_set <- prepare_lambda_data_set(index)
+  #train_lambda_set <- lambda_data_set[[1]]
+  #test_lambda_set <- lambda_data_set[[2]]  
+  train_weight_set <- train_subset_for_encoding[-test_weight_indices[,index],]
+  test_weight_set <- train_subset_for_encoding[test_weight_indices[,index],]
+
+  test_weight_set <- test_weight_set %>% semi_join(train_weight_set,by='customer_state')
+
+  weight_rmses <- sapply(weights, function(weight) {
+      customer_state_effects <- train_weight_set %>% 
+        group_by(customer_state) %>% 
+        summarise(mean=mean(number_of_hours),n=n()) %>%
+        mutate(mean=(n*mean+weight*global_mean)/(n+weight)) %>% 
+        select(-n) %>% rename()
+
+      train_weight_set <- train_weight_set %>% left_join(customer_state_effects,by='customer_state') %>% rename(customer_state_effect=mean) %>% select(-customer_state) #-product_category_name
+
+      test_weight_set <- test_weight_set %>% left_join(customer_state_effects,by='customer_state') %>% rename(customer_state_effect=mean) %>% select(-customer_state) # -product_category_name,
+
+      set.seed(seed=1,sample.kind = 'Rounding')
+      preProcessScalingValues <- preProcess(train_weight_set %>% select(-number_of_hours), method = c("center", "scale"))
+
+      scaled_train_subset <- predict(preProcessScalingValues, train_weight_set)
+      scaled_test_subset <- predict(preProcessScalingValues, test_weight_set)
+
+      # TODO remove below scaling was erroneous. train scaling needs to be used for test data
+      # below is just demo of how to use scale
+      # scaled_train_subset <- train_subset %>% mutate(across(c(mean_delay_in_handover,distance_between_two_points,volume,product_weight_g,freight_value,price),scale)) 
+      # scaled_test_subset <- test_subset %>% mutate(across(c(mean_delay_in_handover,distance_between_two_points,volume,product_weight_g,freight_value,price),scale))
+
+      # TODO can we skip modelling while evaluating weight to use or for target encoding?
+
+      tuningControl=trainControl(method = "cv", number = 3, p = .8)
+      knn_fit_for_weight <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(20,27,34)),trControl=tuningControl)
+
+      knn_fit_for_weight
+
+      print(knn_fit_for_weight)
+
+      predictions <- predict(knn_fit_for_weight,scaled_test_subset %>% select(-number_of_hours))
+
+      RMSE(scaled_test_subset$number_of_hours,predictions) 
+  })
+
+  # selecting weight leading to lowest RMSE
+  weight_to_use <- weights[which.min(weight_rmses)]
+  paste('Weight to use: ', weight_to_use, ', corresponding RMSE: ', weight_rmses[which.min(weight_rmses)])
+
+  print(weight_rmses)
+  weight_to_use
+})
+
+weight_to_use_per_partition
+
+weight_to_use <- mean(weight_to_use_per_partition) 
+# 5
+
+target_encoding_weights <- target_encoding_weights %>% add_row(effect='customer_state',weight=weight_to_use)
+```
+
+```{r Post regularization with consumer state effect, include=FALSE}
+weight_to_use <- target_encoding_weights %>% filter(effect=='customer_state') %>% pull(weight)
+
+customer_state_effects <- prepare_customer_state_effect(train_subset,weight_to_use)
+dataset_list <- enhance_datasets_using_customer_state_effect(customer_state_effects,train_subset,test_subset,TRUE)
+
+train_subset <- dataset_list[[1]]
+test_subset <- dataset_list[[2]]
+```
+
+```{r Apply Target encoding for customer_loc, echo = FALSE}
+  train_subset_for_encoding <- train_subset[1:40000,] %>% select(-customer_zip_code_prefix)
+  global_mean <- mean(train_subset_for_encoding$number_of_hours)
+
+  set.seed(1,sample.kind='Rounding')
+  test_weight_indices <- createDataPartition(train_subset_for_encoding$number_of_hours,p=0.2,times=3,list=FALSE)
+
+  weights <- seq(0,10,5)
+
+  weight_to_use_per_partition <- sapply(1:3, function(index) {
+  #lambda_data_set <- prepare_lambda_data_set(index)
+  #train_lambda_set <- lambda_data_set[[1]]
+  #test_lambda_set <- lambda_data_set[[2]]  
+  train_weight_set <- train_subset_for_encoding[-test_weight_indices[,index],]
+  test_weight_set <- train_subset_for_encoding[test_weight_indices[,index],]
+
+  test_weight_set <- test_weight_set %>% semi_join(train_weight_set,by='customer_loc')
+
+  weight_rmses <- sapply(weights, function(weight) {
+      customer_loc_effects <- train_weight_set %>% 
+        group_by(customer_loc) %>% 
+        summarise(mean=mean(number_of_hours),n=n()) %>%
+        mutate(mean=(n*mean+weight*global_mean)/(n+weight)) %>% 
+        select(-n) %>% rename()
+
+      train_weight_set <- train_weight_set %>% left_join(customer_loc_effects,by='customer_loc') %>% rename(customer_loc_effect=mean) %>% select(-customer_loc) #-product_category_name
+
+      test_weight_set <- test_weight_set %>% left_join(customer_loc_effects,by='customer_loc') %>% rename(customer_loc_effect=mean) %>% select(-customer_loc) # -product_category_name,
+
+      set.seed(seed=1,sample.kind = 'Rounding')
+      preProcessScalingValues <- preProcess(train_weight_set %>% select(-number_of_hours), method = c("center", "scale"))
+
+      scaled_train_subset <- predict(preProcessScalingValues, train_weight_set)
+      scaled_test_subset <- predict(preProcessScalingValues, test_weight_set)
+
+      # TODO remove below scaling was erroneous. train scaling needs to be used for test data
+      # below is just demo of how to use scale
+      # scaled_train_subset <- train_subset %>% mutate(across(c(mean_delay_in_handover,distance_between_two_points,volume,product_weight_g,freight_value,price),scale)) 
+      # scaled_test_subset <- test_subset %>% mutate(across(c(mean_delay_in_handover,distance_between_two_points,volume,product_weight_g,freight_value,price),scale))
+
+      # TODO can we skip modelling while evaluating weight to use or for target encoding?
+
+      tuningControl=trainControl(method = "cv", number = 3, p = .8)
+      knn_fit_for_weight <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(20,27,34)),trControl=tuningControl)
+
+      knn_fit_for_weight
+
+      print(knn_fit_for_weight)
+
+      predictions <- predict(knn_fit_for_weight,scaled_test_subset %>% select(-number_of_hours))
+
+      RMSE(scaled_test_subset$number_of_hours,predictions) 
+  })
+
+  # selecting weight leading to lowest RMSE
+  weight_to_use <- weights[which.min(weight_rmses)]
+  paste('Weight to use: ', weight_to_use, ', corresponding RMSE: ', weight_rmses[which.min(weight_rmses)])
+
+  print(weight_rmses)
+  weight_to_use
+})
+
+weight_to_use_per_partition
+
+weight_to_use <- mean(weight_to_use_per_partition) 
+# 5
+
+target_encoding_weights <- target_encoding_weights %>% add_row(effect='customer_loc',weight=weight_to_use)
+```
+
+```{r Post regularization with customer loc effect, include=FALSE}
+weight_to_use <- target_encoding_weights %>% filter(effect=='customer_loc') %>% pull(weight)
+
+customer_loc_effects <- prepare_customer_loc_effect(train_subset,weight_to_use)
+dataset_list <- enhance_datasets_using_customer_loc_effect(customer_loc_effects,train_subset,test_subset,TRUE)
+
+train_subset <- dataset_list[[1]]
+test_subset <- dataset_list[[2]]  
+```
+
+5 fold cross validation
+80/20 split
+TODO remove p=0.8 from knn trainControl as it is not used.
+
+```{r knn model data scaled under test, eval=FALSE, include=FALSE}
+## TODO make echo=FALSE 
+# TODO review and improve or remove this as required
+set.seed(seed=1,sample.kind = 'Rounding')
+preProcessScalingValues <- preProcess(train_subset %>% select(-number_of_hours), method = c("center", "scale"))
+
+scaled_train_subset <- predict(preProcessScalingValues, train_subset) 
+scaled_test_subset <- predict(preProcessScalingValues, test_subset)
+
+# TODO remove below scaling was erroneous. train scaling needs to be used for test data
+# below is just demo of how to use scale
+# scaled_train_subset <- train_subset %>% mutate(across(c(mean_delay_in_handover,distance_between_two_points,volume,product_weight_g,freight_value,price),scale)) 
+# scaled_test_subset <- test_subset %>% mutate(across(c(mean_delay_in_handover,distance_between_two_points,volume,product_weight_g,freight_value,price),scale))
+
+tuningControl=trainControl(method = "cv", number = 5, p = .8)
+knn_fit <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(56,63,70,77)),trControl=tuningControl)
+
+knn_fit
+
+predictions <- predict(knn_fit,scaled_test_subset %>% select(-number_of_hours))
+
+performance_metric <- performance_metric %>% 
+  add_row(type='knn model',
+          value=RMSE((scaled_test_subset %>% pull(number_of_hours)),predictions))
+
+performance_metric
+#183.8551 for k=70
+
+save(knn_fit,file='rda/knn_model.rda')
 ```
@@
 - gam
+
+Random forest
+However, it is important to know your data and keep in mind that a Random Forest can’t extrapolate.
+It can only make a prediction that is an average of previously observed labels.
+In this sense it is very similar to KNN.
+In other words, in a regression problem, the range of predictions a Random Forest can make is bound by the highest and lowest labels in the training data.
+This behavior becomes problematic in situations where the training and prediction inputs differ in their range and/or distributions.
+This is called covariate shift and it is difficult for most models to handle but especially for Random Forest, because it can’t extrapolate.
+From : https://towardsdatascience.com/a-limitation-of-random-forest-regression-db8ed7419e9f
+
+```{r rf model, eval = FALSE, include= FALSE}
+# TODO make it echo=FALSE
+set.seed(seed=1,sample.kind = 'Rounding')
+tuningControl=trainControl(method = "cv", number = 4)
+rf_fit <- train(number_of_hours~., train_subset, method='rf',tuneGrid = data.frame(mtry = c(1,2,3,4,5)),trControl=tuningControl,ntree=200)
+#rf_fit <- train(number_of_hours~., train_subset %>% select(-seller_state), method='rf',tuneGrid = data.frame(mtry = seq(5, 50, 10)),trControl=tuningControl,ntree=250) # nodesize = 50,maxnodes=2000,ntree=200
+
+rf_fit
+
+predictions <- predict(rf_fit,test_subset %>% select(-number_of_hours))
+
+performance_metric <- performance_metric %>% 
+  add_row(type='random forest model',
+          value=RMSE((test_subset %>% pull(number_of_hours)),predictions))
+
+performance_metric
+# 174.0326 for mtry 2. Technically there are total of 11 predictors. The seller_state was not of much help. it was actually increasing RMSE. So have ignored it.
+
+save(rf_fit,file='rda/rf_model.rda')
+```
+
+five sets each having a training set and testing set.
+This ensured that we used only edx dataset while we developed our model.
+In addition, the multiple partitions helped to remove any randomness from our tuning parameters and to get an appopriate estimate of performance for our incremental model.
+This helped reduced the probability of overfitting.
+
+we should consider if we need time stamp or we can take purchase time as 0 and calculated approved at from it.
+though we may lose granularity in terms of time as truck may come everyday at 4.
+so after 4 you can ship , before 4 you cant
+We can be agnostic to purchase time and just predict number of hours for delivery for given purchase.
+
+TODO -> if there is delayed to carrier precendence, we could use that probability and average delay to carrier to fine tune our results.
+Similarly we could use other metrics like payment approval time etc.

### 2021-07-17 finalReportPrep.Rmd
stabalized report with models and did cleanup

@@
 ```
-
-Delivery Performance:
-
-You will also be able to work through delivery performance and find ways to optimize delivery times.
-
-Attention
-
-- An order might have multiple items.
-- Each item might be fulfilled by a distinct seller.
-- All text identifying stores and partners where replaced by the names of Game of Thrones great houses.
-
-olist_customers_dataset.csv
-This dataset has information about the customer and its location. Use it to identify unique customers in the orders dataset and to find the orders delivery location.
-At our system each order is assigned to a unique customerid.
-This means that the same customer will get different ids for different orders.
-The purpose of having a customerunique_id on the dataset is to allow you to identify customers that made repurchases at the store.
-Otherwise you would find that each order had a different customer associated with.
-
-olist_order_reviews_dataset.csv
-This dataset includes data about the reviews made by the customers.
-
-After a customer purchases the product from Olist Store a seller gets notified to fulfill that order.
-Once the customer receives the product, or the estimated delivery date is due, the customer gets a satisfaction survey by email where he can give a note for the purchase experience and write down some comments.
-
-olist_orders_dataset.csv
-This is the core dataset. From each order you might find all other information.
-
-olist_products_dataset.csv
-Products Dataset
-This dataset includes data about the products sold by Olist.
 
@@
 
-```{r RMSE, echo=FALSE}
+```{r RMSE, include=FALSE}
 # Performance metric
@@
 test_subset <- training_dataset[test_indices,]
+
 # ensure test_subset has products only for concerned product categories and seller_ids which are in train_subset
@@
 
-penalty_data <- train_subset %>% get_delay_in_handover_to_carrier_per_seller_id()
-
-delay_in_payment_approval_data <- train_subset %>% get_delay_in_payment_approval_per_customer_zip_code()
-
 rm(test_indices)
-```
-
-```{r Prepare training and test data for ,include=FALSE}
-# penalty_data is computed on train_subset and is used with test_subset as historical learning of efficiency of each seller in delivering to carrier
-train_subset <- train_subset %>% 
-  left_join(penalty_data,by='seller_id') %>% 
-  left_join(delay_in_payment_approval_data,by='customer_zip_code_prefix') %>% 
-  select(-handed_over_to_carrier_date, -shipping_limit_date,-order_purchase_timestamp,-payment_approved_at)
-
-test_subset <- test_subset %>% 
-  left_join(penalty_data,by='seller_id') %>% 
-  left_join(delay_in_payment_approval_data,by='customer_zip_code_prefix') %>% 
-  select(-handed_over_to_carrier_date, -shipping_limit_date,-order_purchase_timestamp,-payment_approved_at)
 ```
@@
 target_encoding_weights <- data.frame(effect=character(0),weight=double(0)) 
+```
+
+```{r Prepare delay in handover to carried data, include = FALSE}
+# prepare delay_in_handover_to_carrier_data based on regularisation weight we have identified
+prepare_delay_in_handover_to_carrier <- function(train_df, weight_to_use) {
+  global_mean <- mean(train_df %>% 
+                        get_delay_in_handover_to_carrier_per_seller_id() %>% 
+                        pull(mean_delay_in_handover))
+
+  set.seed(1,sample.kind='Rounding')
+  test_weight_indices <- createDataPartition(train_df$number_of_hours,p=0.2,times=10,list=FALSE)
+
+  delay_in_handover_to_carrier_data_list <- lapply(1:10, function(index) {
+      train_weight_set <- train_df[-test_weight_indices[,index],]
+
+      delay_in_handover_to_carrier_data <- train_weight_set %>% 
+        get_delay_in_handover_to_carrier_per_seller_id()
+
+      delay_in_handover_to_carrier_data
+    }
+  )
+
+  dhc <- delay_in_handover_to_carrier_data_list[[1]]
+  for (index in 2:10) {
+     dhc <- dhc %>% full_join(delay_in_handover_to_carrier_data_list[[index]],by='seller_id', suffix=c("1",index))
+  }
+
+  # wherever value is NA, we have used global mean
+  dhc <- dhc %>% mutate(across(-seller_id,~if_else(is.na(.x),global_mean,.x)))
+
+  # as different partitions may not have exactly same data set combined the results
+  delay_in_handover_to_carrier_data <- tibble(dhc[,1],
+                                              mean_delay_in_handover=rowMeans(dhc[,-1],
+                                                                              na.rm = TRUE))
+
+  #delay_in_handover_to_carrier_data is ready
+  delay_in_handover_to_carrier_data
+}
+```
+
+```{r Prepare delay in payment approval, include = FALSE}
+# prepare delay_in_payment_approval based on regularisation weight we have identified
+prepare_delay_in_payment_approval <- function(train_df, weight_to_use) {
+  global_mean <- mean(train_df %>% 
+                        get_delay_in_payment_approval_per_customer_zip_code() %>% 
+                        pull(mean_delay_in_payment_approval))
+
+  set.seed(1,sample.kind='Rounding')
+  test_weight_indices <- createDataPartition(train_df$number_of_hours,p=0.2,times=10,list=FALSE)
+
+  delay_in_payment_approval_data_list <- lapply(1:10, function(index) {
+      train_weight_set <- train_df[-test_weight_indices[,index],]
+
+      delay_in_payment_approval_data <- train_weight_set %>% 
+        get_delay_in_payment_approval_per_customer_zip_code()
+
+      delay_in_payment_approval_data
+    }
+  )
+
+  dpa <- delay_in_payment_approval_data_list[[1]]
+  for (index in 2:10) {
+     dpa <- dpa %>% full_join(delay_in_payment_approval_data_list[[index]],by='customer_zip_code_prefix', suffix=c("1",index))
+  }
+
+  # wherever value is NA, we have used global mean
+  dpa <- dpa %>% mutate(across(-customer_zip_code_prefix,~if_else(is.na(.x),global_mean,.x)))
+
+  # as different partitions may not have exactly same data set combined the results
+  delay_in_payment_approval_data <- tibble(dpa[,1],
+                                              mean_delay_in_payment_approval=rowMeans(dpa[,-1],
+                                                                              na.rm = TRUE))
+
+  #delay_in_payment_approval_data is ready
+  delay_in_payment_approval_data
+}
 ```
@@
 
+```{r Enhance training and test dataset for delay in payment approval, include=FALSE}
+enhance_datasets_using_delay_in_payment_approval <- function(delay_in_payment_approval_data,train_df,test_df,cleanup) {
+  #we will use delay_in_payment_approval_data computed using training data in test data.  
+  train_df <- train_df %>%
+    left_join(delay_in_payment_approval_data,by='customer_zip_code_prefix')
+
+  test_df <- test_df %>% left_join(delay_in_payment_approval_data,by='customer_zip_code_prefix')
+
+  #cleanup
+  if (cleanup) {
+    train_df <- train_df %>%
+      select(-customer_zip_code_prefix,-order_purchase_timestamp,-payment_approved_at)
+    test_df <- test_df %>% 
+      select(-customer_zip_code_prefix,-order_purchase_timestamp,-payment_approved_at)
+  }
+
+  list(train_df,test_df)
+}
+```
+
 ```{r Enhance training and test dataset for product_category effect based on target encoding, include = FALSE}
@@
 
+```{r Enhance training and testing datasets with derived features, include =FALSE}
+# uses training data to derive features and then enhances training and test set with those features
+enhance_datasets_with_derived_features <- function(train_ds,test_ds,cleanup) {
+  print('Adding column for delay in handover')
+
+  delay_in_handover_to_carrier_data <- prepare_delay_in_handover_to_carrier(train_ds,0)
+
+  dataset_list <- enhance_datasets_using_delay_in_handover_to_carrier(delay_in_handover_to_carrier_data,train_ds,test_ds,cleanup)
+  train_ds <- dataset_list[[1]]
+  test_ds <- dataset_list[[2]]
+
+  print('Adding column for delay in payment approval')
+  delay_in_payment_approval_data <- prepare_delay_in_payment_approval(train_ds,0)
+
+  #we will use delay_in_payment_approval computed using training data in test data.
+  dataset_list <- enhance_datasets_using_delay_in_payment_approval(delay_in_payment_approval_data,train_ds,test_ds,cleanup)
+
+  dataset_list
+}  
+```
+
 ```{r Prepare data set for training, include = FALSE}
@@
 prepare_data_for_training_testing <- function(train_ds, test_ds, cleanup) {
-  print('Adding column for delay in handover')
-  penalty_data <- train_ds %>% get_delay_in_handover_to_carrier_per_seller_id()
-
-  #we will use penalty_data computed using training data in test data.
-  train_ds <- train_ds %>% left_join(penalty_data,by='seller_id') 
-  test_ds <- test_ds %>% left_join(penalty_data,by='seller_id')
-  if (cleanup) {
-    train_ds <- train_ds %>% select(-handed_over_to_carrier_date,-shipping_limit_date)
-    test_ds <- test_ds %>% select(-handed_over_to_carrier_date,-shipping_limit_date)
-  }
-
-  print('Adding column for delay in payment approval')
-  delay_in_payment_approval <- train_ds %>% get_delay_in_payment_approval_per_customer_zip_code()
-
-  #we will use delay_in_payment_approval computed using training data in test data.
-  train_ds <- train_ds %>% left_join(delay_in_payment_approval_data,by='customer_zip_code_prefix')
-  test_ds <- test_ds %>% left_join(delay_in_payment_approval_data,by='customer_zip_code_prefix')
-  if (cleanup) {
-    train_df <- train_df %>% select(-customer_zip_code_prefix)
-    test_df <- test_df %>% select(-customer_zip_code_prefix)
-  }
+  test_ds <- test_ds %>%
+    semi_join(train_ds,by='product_category_name') %>%
+    semi_join(train_ds,by='seller_id') %>% 
+    semi_join(train_ds,by='customer_zip_code_prefix') %>% 
+    semi_join(train_ds,by='customer_state') %>% 
+    semi_join(train_ds,by='customer_loc')
+
+  #adding derived features and target encoded categorical features
+  print('Adding derived features')
+  dataset_list <- enhance_datasets_with_derived_features(train_ds,test_ds,cleanup)
+  train_ds <- dataset_list[[1]]
+  test_ds <- dataset_list[[2]]
 
@@
 
-```{r Apply Target encoding for product_category_name, echo = FALSE}
-#TODO you may need to remove customer_state and seller_state from dataset train_subset_for_encoding before proceeding further.
-
-  train_subset_for_encoding <- train_subset[1:30000,] %>% select(-seller_id,-customer_loc, -customer_zip_code_prefix, -customer_state)
+```{r Apply Target encoding for product_category_name, include = FALSE, eval=FALSE}
+#TODO marked eval=FALSE to help generate report quickly
+  train_subset_for_encoding <- train_subset[1:40000,] %>% select(-customer_loc, -customer_state)
 
@@
   test_weight_set <- test_weight_set %>%
-    semi_join(train_weight_set,by='product_category_name')
+    semi_join(train_weight_set,by='product_category_name') %>%
+    semi_join(train_weight_set,by='seller_id') %>%
+    semi_join(train_weight_set,by='customer_zip_code_prefix')
+
+  dataset_list <- enhance_datasets_with_derived_features(train_weight_set,test_weight_set,TRUE)
+  train_weight_set <- dataset_list[[1]] %>% select(-seller_id)#,-customer_zip_code_prefix)
+  test_weight_set <- dataset_list[[2]] %>% select(-seller_id)#,-customer_zip_code_prefix)
 
@@
         mutate(mean=(n*mean+weight*global_mean)/(n+weight)) %>% 
-        select(-n) %>% rename()
+        select(-n)
 
@@
       tuningControl=trainControl(method = "cv", number = 3, p = .8)
-      knn_fit_for_weight <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(27,34,41)),trControl=tuningControl)
+      knn_fit_for_weight <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(34,41,48)),trControl=tuningControl)
 
@@
 
-```{r Post regularization for product_category effect, include=FALSE}
-weight_to_use <- target_encoding_weights %>% filter(effect=='product_category') %>% pull(weight)
-
-product_category_effects <- prepare_product_category_effect(train_subset,weight_to_use)
-dataset_list <- enhance_datasets_using_product_category_effect(product_category_effects,train_subset,test_subset,TRUE)
-
-train_subset <- dataset_list[[1]]
-test_subset <- dataset_list[[2]]
-```
-
-```{r Apply Target encoding for seller_id, echo = FALSE}
-#TODO you may need to remove customer_state and seller_state from dataset train_subset_for_encoding before proceeding further.
-
-  train_subset_for_encoding <- train_subset[1:30000,] %>% select(-customer_loc, -customer_zip_code_prefix, -customer_state)
+```{r Apply Target encoding for seller_id, include = FALSE, eval=FALSE}
+#TODO marked eval=FALSE to help generate report quickly
+  train_subset_for_encoding <- train_subset[1:40000,] %>% select(-customer_loc, -customer_state)
 
@@
 
-  test_weight_set <- test_weight_set %>% semi_join(train_weight_set,by='seller_id')
+  test_weight_set <- test_weight_set %>%
+    semi_join(train_weight_set,by='product_category_name') %>%
+    semi_join(train_weight_set,by='seller_id') %>% 
+    semi_join(train_weight_set,by='customer_zip_code_prefix')
+
+  ## adding derived features and target encoded categorical features
+  dataset_list <- enhance_datasets_with_derived_features(train_weight_set,test_weight_set,FALSE)
+  train_weight_set <- dataset_list[[1]]
+  test_weight_set <- dataset_list[[2]]
+
+  weight_to_use <- target_encoding_weights %>% 
+    filter(effect=='product_category') %>% pull(weight)
+
+  product_category_effects <- prepare_product_category_effect(train_weight_set,weight_to_use)
+  dataset_list <- enhance_datasets_using_product_category_effect(product_category_effects,train_weight_set,test_weight_set,FALSE)
+
+  train_weight_set <- dataset_list[[1]] %>% select(-order_purchase_timestamp,-handed_over_to_carrier_date,-shipping_limit_date,-customer_zip_code_prefix,-payment_approved_at,-product_category_name)
+  test_weight_set <- dataset_list[[2]] %>% select(-order_purchase_timestamp,-handed_over_to_carrier_date,-shipping_limit_date,-customer_zip_code_prefix,-payment_approved_at,-product_category_name)
+  ##
 
@@
 weight_to_use <- mean(weight_to_use_per_partition)  
-# 3
+# 5
 
@@
 
-```{r Post regularization for seller_id effect, include=FALSE}
-weight_to_use <- target_encoding_weights %>% filter(effect=='seller_id') %>% pull(weight)
-
-seller_id_effects <- prepare_seller_id_effect(train_subset,weight_to_use)
-dataset_list <- enhance_datasets_using_seller_id_effect(seller_id_effects,train_subset,test_subset,TRUE)
-
-train_subset <- dataset_list[[1]]
-test_subset <- dataset_list[[2]]
-```
-
-```{r Apply Target encoding for customer_state, echo = FALSE}
-  train_subset_for_encoding <- train_subset[1:40000,] %>% select(-customer_loc,-customer_zip_code_prefix)
+```{r Apply Target encoding for customer_state, include = FALSE,eval=FALSE}
+#TODO marked eval=FALSE to help generate report quickly
+  train_subset_for_encoding <- train_subset[1:40000,] %>% select(-customer_loc)
   global_mean <- mean(train_subset_for_encoding$number_of_hours)
@@
 
+  test_weight_set <- test_weight_set %>%
+    semi_join(train_weight_set,by='product_category_name') %>%
+    semi_join(train_weight_set,by='seller_id') %>% 
+    semi_join(train_weight_set,by='customer_zip_code_prefix') %>% 
+    semi_join(train_weight_set,by='customer_state')
+
+  ## adding derived features and target encoded categorical features
+  dataset_list <- enhance_datasets_with_derived_features(train_weight_set,test_weight_set,FALSE)
+  train_weight_set <- dataset_list[[1]]
+  test_weight_set <- dataset_list[[2]]
+
+  weight_to_use <- target_encoding_weights %>% 
+    filter(effect=='product_category') %>% pull(weight)
+
+  product_category_effects <- prepare_product_category_effect(train_weight_set,weight_to_use)
+  dataset_list <- enhance_datasets_using_product_category_effect(product_category_effects,train_weight_set,test_weight_set,FALSE)
+
+  train_weight_set <- dataset_list[[1]] %>% select(-order_purchase_timestamp,-handed_over_to_carrier_date,-shipping_limit_date,-customer_zip_code_prefix,-payment_approved_at,-product_category_name)
+  test_weight_set <- dataset_list[[2]] %>% select(-order_purchase_timestamp,-handed_over_to_carrier_date,-shipping_limit_date,-customer_zip_code_prefix,-payment_approved_at,-product_category_name)
+
+  weight_to_use <- target_encoding_weights %>% filter(effect=='seller_id') %>% pull(weight)
+
+  seller_id_effects <- prepare_seller_id_effect(train_weight_set,weight_to_use)
+  dataset_list <- enhance_datasets_using_seller_id_effect(seller_id_effects,train_weight_set,test_weight_set,FALSE)
+
+  train_weight_set <- dataset_list[[1]] %>% select(-seller_id)
+  test_weight_set <- dataset_list[[2]] %>% select(-seller_id)
+  ##
+
   weight_rmses <- sapply(weights, function(weight) {
@@
       tuningControl=trainControl(method = "cv", number = 3, p = .8)
-      knn_fit_for_weight <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(20,27,34)),trControl=tuningControl)
+      knn_fit_for_weight <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(34,41,48)),trControl=tuningControl)
 
@@
 weight_to_use <- mean(weight_to_use_per_partition) 
-# 5
-
-target_encoding_weights <- target_encoding_weights %>% add_row(effect='customer_state',weight=weight_to_use)
-```
-
-```{r Post regularization with consumer state effect, include=FALSE}
-weight_to_use <- target_encoding_weights %>% filter(effect=='customer_state') %>% pull(weight)
-
-customer_state_effects <- prepare_customer_state_effect(train_subset,weight_to_use)
-dataset_list <- enhance_datasets_using_customer_state_effect(customer_state_effects,train_subset,test_subset,TRUE)
-
-train_subset <- dataset_list[[1]]
-test_subset <- dataset_list[[2]]
-```
-
-```{r Apply Target encoding for customer_loc, echo = FALSE}
-  train_subset_for_encoding <- train_subset[1:40000,] %>% select(-customer_zip_code_prefix)
+# 10
+
+target_encoding_weights <- target_encoding_weights %>%
+  add_row(effect='customer_state',weight=weight_to_use)
+```
+
+```{r Apply Target encoding for customer_loc, include = FALSE,eval=FALSE}
+#TODO marked eval=FALSE to help generate report quickly
+  train_subset_for_encoding <- train_subset[1:40000,]
   global_mean <- mean(train_subset_for_encoding$number_of_hours)
@@
 
-  test_weight_set <- test_weight_set %>% semi_join(train_weight_set,by='customer_loc')
+  test_weight_set <- test_weight_set %>%
+    semi_join(train_weight_set,by='product_category_name') %>%
+    semi_join(train_weight_set,by='seller_id') %>% 
+    semi_join(train_weight_set,by='customer_zip_code_prefix') %>% 
+    semi_join(train_weight_set,by='customer_state') %>% 
+    semi_join(train_weight_set,by='customer_loc')
+
+  ## adding derived features and target encoded categorical features
+  dataset_list <- enhance_datasets_with_derived_features(train_weight_set,test_weight_set,FALSE)
+  train_weight_set <- dataset_list[[1]]
+  test_weight_set <- dataset_list[[2]]
+
+  weight_to_use <- target_encoding_weights %>% 
+    filter(effect=='product_category') %>% pull(weight)
+
+  product_category_effects <- prepare_product_category_effect(train_weight_set,weight_to_use)
+  dataset_list <- enhance_datasets_using_product_category_effect(product_category_effects,train_weight_set,test_weight_set,FALSE)
+
+  train_weight_set <- dataset_list[[1]] %>% select(-order_purchase_timestamp,-handed_over_to_carrier_date,-shipping_limit_date,-customer_zip_code_prefix,-payment_approved_at,-product_category_name)
+  test_weight_set <- dataset_list[[2]] %>% select(-order_purchase_timestamp,-handed_over_to_carrier_date,-shipping_limit_date,-customer_zip_code_prefix,-payment_approved_at,-product_category_name)
+
+  weight_to_use <- target_encoding_weights %>% filter(effect=='seller_id') %>% pull(weight)
+
+  seller_id_effects <- prepare_seller_id_effect(train_weight_set,weight_to_use)
+  dataset_list <- enhance_datasets_using_seller_id_effect(seller_id_effects,train_weight_set,test_weight_set,FALSE)
+
+  train_weight_set <- dataset_list[[1]] %>% select(-seller_id)
+  test_weight_set <- dataset_list[[2]] %>% select(-seller_id)
+
+  weight_to_use <- target_encoding_weights %>% filter(effect=='customer_state') %>% pull(weight)
+
+  customer_state_effects <- prepare_customer_state_effect(train_weight_set,weight_to_use)
+  dataset_list <- enhance_datasets_using_customer_state_effect(customer_state_effects,train_weight_set,test_weight_set,TRUE)
+
+  train_weight_set <- dataset_list[[1]]
+  test_weight_set <- dataset_list[[2]]
+  ##
 
@@
 
-```{r Post regularization with customer loc effect, include=FALSE}
-weight_to_use <- target_encoding_weights %>% filter(effect=='customer_loc') %>% pull(weight)
-
-customer_loc_effects <- prepare_customer_loc_effect(train_subset,weight_to_use)
-dataset_list <- enhance_datasets_using_customer_loc_effect(customer_loc_effects,train_subset,test_subset,TRUE)
-
+```{r Save target encoding weights, include=FALSE, eval=FALSE}
+save(target_encoding_weights,file='rda/target_encoding_weights.rda')
+```
+
+```{r Load target encoding weights, include=FALSE}
+# required only to generate report or resume modeling quickly
+load('rda/target_encoding_weights.rda')
+```
+
+penalty_data is computed on train_subset and is used with test_subset as historical learning of efficiency of each seller in delivering to carrier
+Similarly delay in payment approval is used to incorporate role of customer zip code in delaying this processing, which maybe due to fraud analysis etc.
+```{r Prepare training and test data for model tuning and development,include=FALSE}
+dataset_list <- prepare_data_for_training_testing(train_subset,test_subset,FALSE)
 train_subset <- dataset_list[[1]]
 test_subset <- dataset_list[[2]]
+
+# cleanup
+train_subset <- train_subset %>%
+  select(-handed_over_to_carrier_date,-shipping_limit_date,-product_category_name,-seller_id,-customer_state,-customer_loc,-customer_zip_code_prefix,-order_purchase_timestamp,-payment_approved_at)
+test_subset <- test_subset %>%
+  select(-handed_over_to_carrier_date,-shipping_limit_date,-product_category_name,-seller_id,-customer_state,-customer_loc,-customer_zip_code_prefix,-order_purchase_timestamp,-payment_approved_at)
 ```
@@
 
-```{r knn model data scaled under test, eval=FALSE, include=FALSE}
+```{r knn model data scaled under test, include=FALSE, eval=FALSE}
 ## TODO make echo=FALSE 
-# TODO review and improve or remove this as required
 set.seed(seed=1,sample.kind = 'Rounding')
@@
 
-tuningControl=trainControl(method = "cv", number = 5, p = .8)
+tuningControl=trainControl(method = "cv", number = 5)
 knn_fit <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(56,63,70,77)),trControl=tuningControl)
@@
 performance_metric <- performance_metric %>% 
-  add_row(type='knn model',
+  add_row(type='Knn model(partial dataset)',
           value=RMSE((scaled_test_subset %>% pull(number_of_hours)),predictions))
@@
 performance_metric
-#183.8551 for k=70
+#183.7767 for k=63
 
@@
 
-```{r rf model, eval = FALSE, include= FALSE}
+```{r rf model, include= FALSE, eval = FALSE}
 # TODO make it echo=FALSE
@@
 performance_metric <- performance_metric %>% 
-  add_row(type='random forest model',
+  add_row(type='Random forest model(partial dataset)',
           value=RMSE((test_subset %>% pull(number_of_hours)),predictions))
@@
 performance_metric
-# 174.0326 for mtry 2. Technically there are total of 11 predictors. The seller_state was not of much help. it was actually increasing RMSE. So have ignored it.
+# 174.2496 for mtry 2. Technically there are total of 11 predictors. The seller_state was not of much help. it was actually increasing RMSE. So have ignored it.
 
 save(rf_fit,file='rda/rf_model.rda')
+```
+
+```{r Save performance metric, include=FALSE, eval=FALSE}
+save(performance_metric,file='rda/performance_metric.rda')
+```
+
+```{r Load performance metric, include=FALSE}
+# required only to generate report
+load('rda/performance_metric.rda')
+```
+
+```{r ensemble, include=FALSE, eval= FALSE}
+load('rda/knn_model.rda')
+load('rda/rf_model.rda')
+
+predictions_knn <- predict(knn_fit,scaled_test_subset %>% select(-number_of_hours))
+predictions_rf <- predict(rf_fit,test_subset %>% select(-number_of_hours))
+
+predictions_df <- data.frame(predictions_knn,predictions_rf)
+
+RMSE(rowMeans(predictions_df),(test_subset %>% pull(number_of_hours) ))
+176.5041
+```
+
+```{r Prepare dataset for final model training, include = FALSE}
+set.seed(seed=1,sample.kind = 'Rounding')
+
+validation_dataset <- validation_dataset %>% 
+  compute_volume_using_product_info() %>% 
+  extract_distance_between_seller_customer() %>% 
+  compute_rough_customer_location()
+
+dataset_list <- prepare_data_for_training_testing(training_dataset,validation_dataset,FALSE)
+training_dataset <- dataset_list[[1]]
+validation_dataset <- dataset_list[[2]]
+
+# cleanup
+training_dataset <- training_dataset %>%
+  select(-handed_over_to_carrier_date,-shipping_limit_date,-product_category_name,-seller_id,-customer_state,-customer_loc,-customer_zip_code_prefix,-order_purchase_timestamp,-payment_approved_at)
+validation_dataset <- validation_dataset %>%
+  select(-handed_over_to_carrier_date,-shipping_limit_date,-product_category_name,-seller_id,-customer_state,-customer_loc,-customer_zip_code_prefix,-order_purchase_timestamp,-payment_approved_at)
+```
+
+```{r Final Model Training, include=FALSE,eval = FALSE}
+# TODO marked eval=FALSE to prevent it from running during report generation
+set.seed(seed=1,sample.kind = 'Rounding')
+
+mtry_to_use <- rf_fit$finalModel$mtry
+ntree_to_use <- 500 #rf_fit$finalModel$ntree
+
+final_model <- randomForest(number_of_hours~., training_dataset,mtry= mtry_to_use, ntree=ntree_to_use)
+
+final_model
+
+save(final_model,file='rda/final_model.rda')
+```
+
+```{r Final model validation, include = FALSE, eval=FALSE}
+# TODO marked eval=FALSE to prevent it from running during report generation
+predictions <- predict(final_model,validation_dataset %>% select(-number_of_hours))
+
+performance_metric <- performance_metric %>% 
+  add_row(type='Final model',
+          value=RMSE((validation_dataset %>% pull(number_of_hours) ),predictions))
+# 189.2705 for mtry=2, ntree=500
+
+performance_metric
 ```

### 2021-07-18 finalReportPrep.Rmd
stable rmd file

@@
     df_print: paged
+urlcolor: blue
 ---
@@
 # Overview
-Prediction systems play a cruical role in the modern world.
+Prediction systems play a crucial role in the modern world.
 With advent of technology and availability of data, these systems are being leveraged by retail chains, banks, colleges among other businesses.
-Depending on the use case, these systems are used to estimate delivery time, sales figures, carry out inventory management, predict occurrence of a disease, among others.
-
-Inspired by what all problems these systems can solve, we have attempted to build a delivery prediction system based on publically available [dataset of orders made at Olist Store](https://www.kaggle.com/olistbr/brazilian-ecommerce), a Brazilian ecommerce.
+Depending on the use case, these systems are used to estimate delivery time, sales figures, carry out inventory management, predict occurrence of a disease etc.
+
+Inspired by what all problems these systems can solve, we have attempted to build a delivery prediction system based on publicaly available [dataset of orders made at Olist Store](https://www.kaggle.com/olistbr/brazilian-ecommerce), a Brazilian e-commerce.
 
@@
 Olist connects sellers to the customers.
-It is the sellers, who are responsible to fulfill any orders placed on the platform using Olist's logistic partners.
+It is the sellers, who are responsible to fulfil any orders placed on the platform using Olist's logistic partners.
 
@@
 
-So in the geolocation data set, zip code prefix refered to the first part of the CEP and it varied from 1 to 5 digits in the dataset available to us.
+So in the geolocation data set, zip code prefix referred to the first part of the CEP and it varied from 1 to 5 digits in the dataset available to us.
 Because the zip code prefix does not represent the precise value, it was not unique and had many duplicate values with varying value of lat, lng for same zip code prefix.
@@
 # Important: many geolocation_city have same name but different ascent component. Example: niteroi and niterói
-# intially we can see 
+# initially we can see 
 # geolocation_dataset %>% dplyr::count(geolocation_zip_code_prefix,geolocation_city,geolocation_state) %>% nrow()
@@
          customer_unique_id=as.factor(customer_unique_id),
-         customer_zip_code_prefix=as.factor(customer_zip_code_prefix),
          product_category_name=as.factor(product_category_name),
-         customer_zip_code_prefix=str_pad(customer_zip_code_prefix, c(5),pad=c('0'),side='right')) %>%
+         customer_zip_code_prefix=str_pad(customer_zip_code_prefix, c(5),pad=c('0'),side='left')) %>% 
   select(-customer_id,-order_status,-seller_city,-seller_state,-customer_city,-order_id,-product_id,-seller_zip_code_prefix,-customer_unique_id)
@@
 Having wrangled the datasets individually we combined them into a single dataset.
-While the dataset had many attributes, based on their frequency and analysis, we zeroed on the below fields.
+While the dataset had many attributes, based on their frequency and our intuition shared earlier we zeroed on the below fields for further consideration.
 In addition, it is important to note that for model development we created outcome column number_of_hours based on existing fields i.e.
@@
 
-### Performance Metric
+## Performance Metric
 
@@
 
-#### Current RMSE
+### Current RMSE
 
@@
 
-This RMSE i.e. our performance metric was later used as benchmark for modeling. The delivery prediction system being built attempted to improve on this metric.
+This RMSE i.e. our performance metric was later used as benchmark for modelling. The delivery prediction system being built attempted to improve on this metric.
 
@@
 compute_rough_customer_location <- function(df) {
-  df %>% mutate(customer_zip_code_prefix=str_pad(customer_zip_code_prefix, c(3),pad=c('0'),side='right')) %>% 
-    mutate(customer_loc = as.numeric(str_extract(customer_zip_code_prefix,"\\d\\d\\d"))) 
-  #%>% select(-customer_zip_code_prefix)
+  df %>% mutate(customer_loc = as.numeric(str_extract(customer_zip_code_prefix,"\\d\\d\\d"))) 
 }
@@
 
-## Paritioning dataset for training and validation
-
-Before using the wrangled data for analysis and data modeling, we partitioned the data into training and validation data sets.
+## Partitioning dataset for training and validation
+
+Before using the wrangled data for analysis and data modelling, we partitioned the data into training and validation data sets.
 The idea was to keep validation data set aside and use only training data for any analysis, model development and tuning.
@@
 
-The data sets were named as below for furture reference:
+The data sets were named as below for future reference:
 
@@
 
-For analysis, we used only the training dataset.
-
-Since we needed to build delivery prediction system which could estimate number of hours required for delivery for new orders, while maintaining minimum RMSE, we used our intution and understanding of the domain to guide us in deriving insights from available predictors.
-The same are shared below:
-
-* _**Distance between seller and customer**_ could likely influence delivery estimate.
-* _**Delay in payment approval**_ could influence delivery time as it would delay shippment from the seller.
-This could be related to customer zip code as often some customer locations require fraud analysis before payment is approved.
-* _**Delay in handover of product**_ from seller to carrier would influence delivery time as any time spent there is wasted.
-This could be seller related as some sellers maybe known to take more time in processing order compared to other sellers.
-* Some _**customer locations**_ maybe difficult to reach or need special delivery mechanism due to their remote nature.
-Such locations may need more delivery time compared to other locations irrespective of the distance involved.
-We need not analyse for each customer location but a rough area could be used based on combination of zip codes.
-* _**Weight of the product**_ could influence delivery time as heavy items say bed, furniture may be transported slowly while lighter items maybe shipped more quickly.
-* _**Price of the product**_ could influence delivery time.
-Expensive items may require considerable more attention and could influence delivery like priority shipping or longer delivery time due to caution.
-* _**Volume of the product**_ could influence delivery time as smaller products can be shipped together while others may need separate truck and therefore more time for dedicated truck to be available.
-* _**Product category**_ could be used for analysis as different product categories would have different packing, assembly, processing requiremts and shelf life would could lead to impact in delivery time.
-* _**Freight value**_ could indicate complexity of shipping like fragility of the product, distance, toll booth charges on the way among others. This could be considered for analysis.
-* _**Customer state**_ could influence the delivery estimate due to state border checks, paper work before goods delivery among others.
-* Different _**sellers**_ may take different amount of time or sell specific type of products which may require different shipping time.
-So products from a seller selling apparels may be delivered quickly while products from one selling customised products or say furniture may take relatively more time.
-This could be useful information for predicting delivery time.
-
-To start building our algorithm, we further paritioned our training data set i.e.
-training_dataset into a split of 80/20 between train_subset and test_subset to ensure we use only the train_subset for training and tuning the model and test_subset to evaluate the model and measure RMSE.
-This would ensure there is no data leak.
-The split like earlier was done such that distribution of outcome i.e.
-number_of_hours was similar in both training and test data
-
-TODO
-Use target encoding for categorical variables
-Use preprocess for scaling data against training data set. Then use mean, sd from this trained data for test data.
-
-Derived data set should be generated later on observations
-- geolocation usage to capture distance
-- delay and mean delay in handling to carrier
-- number of hours for delivery should be generated later
-
-We use training_dataset and split into into train and test set. So that training_dataset is sufficient to fine tune the model
 ```{r Split data for training, include=FALSE}
+# We use training_dataset and split into into train and test set. So that training_dataset is sufficient to fine tune the model, identify weights for target encoding of categorical variables etc.
 # Let's split training_dataset into two parts
@@
 ```
-
-==TODO note penalty data should be calculated again if train_subset is changed. Only trained data should be used for test, so that there is no data leak.
-
-== TODO additive smoothing/Bayes estimator with target encoding
 
@@
 
-### Product category effects
-As part of this section, we looked into our intution that different product categories may have different delivery timelines and explored its influence by incoporating it in our model for analysis.
-
-Since product categories are categorical in nature, we applied target encoding on the product categories attribute.
-
-===TODO talk about weight
-===TODO talk about need of regularization.
-
-Below histogram depicts how mean rating of movie across movies is distributed.
+== TODO additive smoothing/Bayes estimator with target encoding
+
+## Feature Analysis and Selection
+Since we needed to build delivery prediction system which could estimate number of hours required for delivery for new orders, while maintaining minimum RMSE, we used our intuition and understanding of the domain to guide us in deriving insights from available predictors.
+The same are shared below:
+
+* _**Distance between seller and customer**_ could likely influence delivery estimate.
+* _**Delay in payment approval**_ could influence delivery time as it would delay shipment from the seller.
+This could be related to customer zip code as often some customer locations require fraud analysis before payment is approved.
+* _**Delay in handover of product from seller to carrier**_ would influence delivery time as any time spent there is wasted.
+This could be seller related as some sellers maybe known to take more time in processing order compared to other sellers.
+* Some _**customer locations**_ maybe difficult to reach or need special delivery mechanism due to their remote nature.
+Such locations may need more delivery time compared to other locations irrespective of the distance involved.
+We need not analyse for each customer location but a rough area could be used based on combination of zip codes.
+* _**Weight of the product**_ could influence delivery time as heavy items say bed, furniture may be transported slowly while lighter items maybe shipped more quickly.
+* _**Price of the product**_ could influence delivery time.
+Expensive items may require considerable more attention and could influence delivery like priority shipping or longer delivery time due to caution.
+* _**Volume of the product**_ could influence delivery time as smaller products can be shipped together while others may need separate truck and therefore more time for dedicated truck to be available.
+* _**Product category**_ could be used for analysis as different product categories would have different packing, assembly, processing requirements and shelf life would could lead to impact in delivery time.
+* _**Freight value**_ could indicate complexity of shipping like fragility of the product, distance, toll booth charges on the way among others. This could be considered for analysis.
+* _**Customer state**_ could influence the delivery estimate due to state border checks, paper work before goods delivery among others.
+* Different _**sellers**_ may take different amount of time or sell specific type of products which may require different shipping time.
+So products from a seller selling apparels may be delivered quickly while products from one selling customised products or say furniture may take relatively more time.
+This could be useful information for predicting delivery time.
+
+For analysis, we used only the training_dataset or subset of it, ensuring validation_dataset was left untouched.
+
+### Derived features
+
+#### Distance between seller and customer feature
+As part of the dataset we had customer and seller location details available in the form of latitude(lat), longitude(lng).
+We used this crucial knowledge of their location to determine shortest distance between the two on an ellipsoid.
+
+We decided to look more into this intuition.
+For this we plotted numbers of hours taken for delivery against the distance between seller and the customer.
+In order to see the trend in this over-plotted graph, gam method based smoothing curve was also plotted on the graph.
+
+```{r Explore distance_between_two_points relation with number_of_hours, echo=FALSE, message=FALSE}
+training_dataset  %>% ggplot(aes(distance_between_two_points,number_of_hours)) + geom_point() + geom_smooth() + xlab('Distance between seller and the customer') + ylab('Number of hours taken for delivery')
+```
+In above plot, we noted that number of hours taken for delivery showed stronger correlation for smaller distances.
+However, once delivery distance increased beyond a threshold, the correlation weakened between the fields.
+Though, we could see broader confidence interval for larger distances.
+
+```{r correlation analysis for distance_between_two_points with number of hours, echo=FALSE}
+# correlation analysis for distance_between_two_points with number of hours
+
+analysis_output_rows <- sapply(seq(0,2000,1000), function(offset) {  
+  list(delay_limit=str_c('>=',offset,' & <',offset+1000),
+       correlation=cor(training_dataset %>% 
+                         filter(distance_between_two_points>=offset & distance_between_two_points< offset+1000) %>% 
+                         pull(distance_between_two_points), 
+                       training_dataset %>% 
+                         filter(distance_between_two_points>=offset & distance_between_two_points< offset+1000) %>% 
+                         pull(number_of_hours))) 
+  },simplify=TRUE)
+
+analysis_output_rows %>% t() %>% rbind( c(str_c('>=',3000),
+                          round(cor(training_dataset %>% 
+                                      filter(distance_between_two_points>=3000) %>%
+                                      pull(distance_between_two_points), 
+                                    training_dataset %>% 
+                                      filter(distance_between_two_points>=3000) %>%
+                                      pull(number_of_hours)),7))
+                        ) %>% as.data.frame() %>% knitr::kable(caption='Correlation analysis between distance between seller & carrier and number of hours taken for delivery')
+
+rm(analysis_output_rows)
+```
+This behavior could be useful in estimating delivery timelines. We decided to consider this as a feature for further analysis for inclusion in the model.
+
+#### Delay in handover of product from seller to carrier feature
+As part of the dataset, there were two important attributes i.e.
+handed_over_to_carrier_date and shipping_limit_date.
+Here, handed_over_to_carrier_date is the actual timestamp when product ordered was handed over the carrier for delivery.
+While, shipping_limit_date is the date by which seller should handover the order to the carrier.
+
+Together these fields were considered useful as they could be used to understand if there was a delay or not in handing over product ordered by seller to the carrier.
+Also, this delay or early handover, could influence delivery timeline even if everything else was to take the same time.
+
+We decided to look more into this intuition.
+For this, the feature i.e mean_delay_in_handover was computed for each seller.
+The computed data was then joined with training_dataset and the below graph was used to visualize the relation between delay in handover to carrier and number of hours taken for delivery.
+In order to see the trend in this over-plotted graph, gam method based smoothing curve was also plotted on the graph.
+
+```{r Explore delay_in_handover_to_carrier relation with number_of_hours, echo=FALSE, message=FALSE}
+delay_in_handover_to_carrier_analysis_dataset <- training_dataset %>%
+  get_delay_in_handover_to_carrier_per_seller_id()
+
+analysis_joined_data <- training_dataset %>%
+  left_join(delay_in_handover_to_carrier_analysis_dataset,by='seller_id')
+
+analysis_joined_data %>% ggplot(aes(mean_delay_in_handover,number_of_hours)) + geom_point() + geom_smooth() + xlab('Delay in handover of product to carrier') + ylab('Number of hours taken for delivery')
+```
+
+In above plot, we noted that number of hours taken for delivery shows weaker correlation with  delay in handover by seller to the carrier as long as the ordered product is handed over to carried before the shipping_limit_date.
+However, post the date, we could see relatively stronger correlation between the fields.
+
+```{r correlation analysis for mean_delay_in_handover with number of hours, echo=FALSE}
+# correlation analysis for mean_delay_in_handover with number of hours
+data.frame(delay_limit=character(0),correlation=numeric(0)) %>% 
+  add_row(delay_limit='<= 0',correlation=cor(analysis_joined_data %>% 
+                                               filter(mean_delay_in_handover<= 0) %>% 
+                                               pull(mean_delay_in_handover), 
+                                             analysis_joined_data %>% 
+                                               filter(mean_delay_in_handover<= 0) %>% 
+                                               pull(number_of_hours))
+          ) %>% 
+  add_row(delay_limit='> 0',correlation=cor(analysis_joined_data %>% 
+                                              filter(mean_delay_in_handover> 0) %>% 
+                                              pull(mean_delay_in_handover), 
+                                            analysis_joined_data %>% 
+                                              filter(mean_delay_in_handover> 0) %>% 
+                                              pull(number_of_hours))) %>% 
+  knitr::kable(caption='Correlation analysis between delay in handover to carrier and number of hours taken for delivery')
+```
+
+This behavior could be useful in estimating delivery timelines.
+We decided to consider this as a feature for further analysis for inclusion in the model.
+However, as for new orders this information would not be available at time of placing the order, we were required to compute this feature using training data and apply the captured learnings later on test data.
+
+To be able to do this, as shared earlier, the feature i.e. mean_delay_in_handover was computed for each seller based on training dataset and then applied to test dataset.
+
+```{r cleanup for delay in handover to carrier feature analysis, include=FALSE}
+rm(delay_in_handover_to_carrier_analysis_dataset,analysis_joined_data)
+```
+
+#### Delay in payment approval feature
+In addition to above, as part of the dataset, there were two other important attributes i.e.
+order_purchase_timestamp and payment_approved_at.
+Here, payment_approved_at is the actual timestamp when payment is approved for a given order.
+While, order_purchase_timestamp is the timestamp at which product was ordered.
+
+Together these fields were considered useful as they could be used to understand if there was a delay or not in approving the payment.
+Also, this delay or early handover, could influence delivery timeline even if everything else was to take the same time.
+
+We decided to look more into this intuition.
+For this, the feature i.e mean_delay_in_payment_approval was computed for each customer_zip_code_prefix.
+The computed data was then joined with training_dataset and the below graph was used to visualize the relation between delay in payment approval and number of hours taken for delivery.
+In order to see the trend in this over-plotted graph, gam method based smoothing curve was also plotted on the graph.
+
+```{r Explore delay_in_payment_approval relation with number_of_hours, echo=FALSE, message=FALSE}
+delay_in_payment_approval_analysis_dataset <- training_dataset %>%
+  get_delay_in_payment_approval_per_customer_zip_code()
+
+analysis_joined_data <- training_dataset %>%
+  left_join(delay_in_payment_approval_analysis_dataset,by='customer_zip_code_prefix')
+
+analysis_joined_data %>% ggplot(aes(mean_delay_in_payment_approval,log10(number_of_hours))) + geom_point() + geom_smooth() + xlab('Delay in payment approval') + ylab('Number of hours taken for delivery [log10(n) scale]')
+```
+In above plot, we noted that number of hours taken for delivery follow a non linear pattern.
+
+```{r correlation analysis for mean_delay_in_payment_approval with number of hours, echo=FALSE}
+# correlation analysis for mean_delay_in_payment_approval with number of hours
+analysis_output_rows <- sapply(seq(0,125,25), function(offset) {  
+  list(delay_in_approval=str_c('>=',offset,' & <',offset+25),
+       correlation=cor(analysis_joined_data %>% 
+                         filter(mean_delay_in_payment_approval>=offset & mean_delay_in_payment_approval< offset+25) %>% 
+                         pull(mean_delay_in_payment_approval), 
+                       analysis_joined_data %>% 
+                         filter(mean_delay_in_payment_approval>=offset & mean_delay_in_payment_approval< offset+25) %>% 
+                         pull(number_of_hours))) 
+  },simplify=TRUE)
+
+analysis_output_rows %>% t() %>% rbind( c(str_c('>=',150),
+                          round(cor(analysis_joined_data %>% 
+                                      filter(mean_delay_in_payment_approval>=150) %>%
+                                      pull(mean_delay_in_payment_approval), 
+                                    analysis_joined_data %>% 
+                                      filter(mean_delay_in_payment_approval>=150) %>%
+                                      pull(number_of_hours)),7))
+                        ) %>% as.data.frame() %>% knitr::kable(caption='Correlation analysis between delay in payment approval and number of hours taken for delivery')
+
+rm(analysis_output_rows)
+```
+This behavior could be useful in estimating delivery timelines.
+We decided to consider this as a feature for further analysis for inclusion in the model.
+However, as for new orders this information would not be available at time of placing the order, we were required to compute this feature using training data and apply the captured learnings later on test data.
+
+To be able to do this, as shared earlier, the feature i.e. mean_delay_in_payment_approval was computed for each customer_zip_code_prefix based on training dataset and then applied to test dataset.
+
+```{r cleanup for delay in payment approval feature analysis, include=FALSE}
+rm(delay_in_payment_approval_analysis_dataset,analysis_joined_data)
+```
+
+#### Volume feature
+As part of the dataset, there were three attributes about product dimensions i.e.
+product_length_cm, product_height_cm and product_width_cm.
+Together these fields were considered useful as they could be used to derive volume of the product.
+
+We looked into the intuition that product volume could influence the delivery timelines as larger products need special delivery mechanism and even dedicated trucks at times for shipment, which may not be readily available.
+
+To understand its influence on the delivery timelines, we plotted numbers of hours taken for delivery against the volume of the product ordered in cubic centimeter($cm^3$).
+In order to see the trend in this over-plotted graph, gam method based smoothing curve was also plotted on the graph.
+
+```{r Explore volume relation with number_of_hours, echo=FALSE, message=FALSE}
+training_dataset  %>% ggplot(aes(volume,log10(number_of_hours))) + geom_point() + geom_smooth() + xlab('Volume of ordered product') + ylab('Number of hours taken for delivery [log10(n) scale]')
+```
+In above plot, we noted that number of hours taken for delivery vary as the product volume changes. There appeared to be stronger influence on delivery timelines for bulkier products beyond a limit.
+
+```{r correlation analysis for volume with number of hours, echo=FALSE}
+# correlation analysis for volume with number of hours
+
+analysis_output_rows <- sapply(seq(0,210000,70000), function(offset) {  
+  list(delay_limit=str_c('>=',offset,' & <',offset+70000),
+       correlation=cor(training_dataset %>% 
+                         filter(volume>=offset & volume< offset+70000) %>% 
+                         pull(volume), 
+                       training_dataset %>% 
+                         filter(volume>=offset & volume< offset+70000) %>% 
+                         pull(number_of_hours))) 
+  },simplify=TRUE)
+
+analysis_output_rows %>% t() %>% rbind( c(str_c('>=',280000),
+                          round(cor(training_dataset %>% 
+                                      filter(volume>=280000) %>%
+                                      pull(volume), 
+                                    training_dataset %>% 
+                                      filter(volume>=280000) %>%
+                                      pull(number_of_hours)),7))
+                        ) %>% as.data.frame() %>% knitr::kable(caption='Correlation analysis between volume and number of hours taken for delivery')
+
+rm(analysis_output_rows)
+```
+
+This behavior could be useful in estimating delivery timelines. We decided to consider this as a feature for further analysis for inclusion in the model.
+
+### Categorical features
+
+#### Product category effects
+We decided to look into our intuition that different product categories may have different delivery timelines and explore its influence on the prediction results.
+
+Below errorbar plot depicts how average number of hours taken for delivery vary across product categories while at the same time capturing the confidence interval of number of hours taken for delivery for each product category.
+
+```{r Explore product_category_name relation with number_of_hours, echo=FALSE}
+training_dataset %>% group_by(product_category_name) %>%
+  summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% 
+  mutate(product_category_name=reorder(product_category_name,mean)) %>% 
+  ggplot(aes(x=product_category_name,y=mean,ymax=mean+2*se,ymin=mean-2*se)) +
+  geom_point()+ geom_errorbar() + theme(axis.text.x = element_text(angle=90,size=8)) + xlab('Product Categories') + ylab('Avg number of hours')
+```
+
+In above plot, we noted that average number of hours taken for delivery vary considerably from one product category to another.
+So our intuition that product categories influence delivery time was confirmed.
+
+Given the influence, we decided to consider this as a feature for further analysis for inclusion in the model.
+However, product categories are categorical in nature and they are quite a few categories, incorporating them directly would not have worked for many machine learning models which require numerical values.
+Additionally in some cases, each category could end up being treated as a predictor which is also the case with one-hot encoding.
+
+However, to convert categorical variables to numeric values and still be able to limit features and make model training and evaluation less expensive, we applied **target encoding** on the product categories attribute.
+
+But to do that, first, we checked number of observations per product category and noticed that the distribution could require **regularization** to prevent overfitting.
+Secondly, target encoding could be erroneous if done against the random dataset.
+
+To solve these, we identified the regularization weight to use by carrying out target encoding for different datasets set of weights, selecting single weight per dataset based on performance metric i.e.
+RMSE and then we took a mean of weights identified across the datasets.
+
+Once weight was identified, to avoid any randomness due to choice of the dataset, we computed target encoding for product category attribute across 10 datasets and then took a mean for them to settle at the final target encoded values to use for the concerned feature.
+
+For the above analysis, knn model was used, given that it does not make assumptions about the dataset, which was useful in the early phase of analysis.
+However, to use knn model, we **scaled and normalized** the features as knn model uses Euclidean distance.
+Scaling and normalization was carried our using training data and then used for test data.
 
@@
 
-      tuningControl=trainControl(method = "cv", number = 3, p = .8)
-      knn_fit_for_weight <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(34,41,48)),trControl=tuningControl)
+      tuningControl=trainControl(method = "cv", number = 3)
+      knn_fit_for_weight <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(41,47,53)),trControl=tuningControl)
 
@@
 weight_to_use <- mean(weight_to_use_per_partition)  
-# 20/3
+# 5
 
@@
 ```
+
+#### Seller id effects
+Next we looked into the intuition that orders for products from different sellers based on what they sell, how they operate could influence delivery timelines.
+
+Below errorbar plot depicts how average number of hours taken for delivery vary across sellers while at the same time capturing the confidence interval of number of hours taken for delivery for each seller.
+Since there were **`r training_dataset %>% dplyr::distinct(seller_id) %>% nrow()`** sellers, we sampled **300** of them to represent their influence effectively in the graph.
+
+```{r Explore seller_id relation with number_of_hours, echo=FALSE, warning = FALSE}
+set.seed(1, sample.kind = 'Rounding')
+training_dataset %>% group_by(seller_id) %>%
+  summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>%
+  mutate(seller_id=reorder(seller_id,mean)) %>% 
+  filter(seller_id %in% sample(seller_id,300,replace=FALSE)) %>%
+  ggplot(aes(x=seller_id,y=mean,ymax=mean+2*se,ymin=mean-2*se)) +
+  geom_point()+ geom_errorbar() + theme(axis.text.x = element_blank()) + xlab('Seller Ids') + ylab('Avg number of hours')
+```
+In above plot, we noted that average number of hours taken for delivery vary considerably from one seller to another. So our intuition that sellers influence delivery time was confirmed.
+
+Given the influence, we decided to consider this as a feature for further analysis for inclusion in the model.
+However, seller ids similar to product categories are categorical in nature and there are **`r training_dataset %>% dplyr::distinct(seller_id) %>% nrow()`** sellers, which are a lot.
+Incorporating them as it is would not have worked for many machine learning models which require numerical values.
+As a result, for seller ids we applied target encoding and followed strategy used for product categories.
 
@@
 
-  weights <- seq(0,6,3)
+  weights <- seq(0,9,3)
 
@@
 weight_to_use <- mean(weight_to_use_per_partition)  
-# 5
+# 8
 
@@
 ```
+
+#### Customer state effects
+Next we looked into the intuition that customer states can have different regulations, paper work, approvals which could influence delivery timelines.
+
+Below errorbar plot depicts how average number of hours taken for delivery vary across customer state while at the same time capturing the confidence interval of number of hours taken for delivery for each customer state.
+
+```{r Explore customer_state relation with number_of_hours, echo=FALSE, warning = FALSE}
+set.seed(1, sample.kind = 'Rounding')
+training_dataset %>% group_by(customer_state) %>%
+  summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>%
+  mutate(customer_state=reorder(customer_state,mean)) %>% 
+  ggplot(aes(x=customer_state,y=mean,ymax=mean+2*se,ymin=mean-2*se)) +
+  geom_point()+ geom_errorbar() + theme(axis.text.x = element_text(angle=90,size=8)) + xlab('Customer States') + ylab('Avg number of hours')
+```
+In above plot, we noted that average number of hours taken for delivery vary considerably from one customer state to another.
+So our intuition that customer state influences delivery time was confirmed.
+
+Given the influence, we decided to consider this as a feature for further analysis for inclusion in the model.
+However, customer states similar to product categories are categorical in nature and there are **`r training_dataset %>% dplyr::distinct(customer_state) %>% nrow()`** customer states.
+Incorporating them as it is would not have worked for many machine learning models which require numerical values.
+As a result, for customer states we applied target encoding and followed strategy used for product categories.
 
@@
 
-      tuningControl=trainControl(method = "cv", number = 3, p = .8)
+      tuningControl=trainControl(method = "cv", number = 3)
       knn_fit_for_weight <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(34,41,48)),trControl=tuningControl)
@@
 weight_to_use <- mean(weight_to_use_per_partition) 
-# 10
+# 5
 
@@
 ```
+
+#### Customer location(loc) effects
+Next we looked into the intuition that rough information about customer location i.e.
+combination of co-located zip codes can influence the delivery timelines as customer location can be remote or require different means of transport, among others.
+
+Below errorbar plot depicts how average number of hours taken for delivery vary across customer locations while at the same time capturing the confidence interval of number of hours taken for delivery for each customer location.
+
+```{r Explore customer_loc relation with number_of_hours, echo=FALSE, warning = FALSE}
+set.seed(1, sample.kind = 'Rounding')
+training_dataset %>% group_by(customer_loc) %>%
+  summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>%
+  mutate(customer_loc=reorder(customer_loc,mean)) %>% 
+  ggplot(aes(x=customer_loc,y=mean,ymax=mean+2*se,ymin=mean-2*se)) +
+  geom_point()+ geom_errorbar() + theme(axis.text.x = element_blank()) + xlab('Customer Locations') + ylab('Avg number of hours')
+```
+In above plot, we noted that average number of hours taken for delivery vary considerably from one customer location to another.
+So our intuition that customer location influences delivery time was confirmed.
+
+Given the influence, we decided to consider this as a feature for further analysis for inclusion in the model.
+However, customer locations similar to product categories are categorical in nature and there are **`r training_dataset %>% dplyr::distinct(customer_loc) %>% nrow()`** customer locations.
+Incorporating them as it is would not have worked for many machine learning models which require numerical values.
+As a result, for customer states we applied target encoding and followed strategy used for product categories.
 
@@
 
-      tuningControl=trainControl(method = "cv", number = 3, p = .8)
-      knn_fit_for_weight <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(34,41,48)),trControl=tuningControl)
+      tuningControl=trainControl(method = "cv", number = 3)
+      knn_fit_for_weight <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(35,43,49)),trControl=tuningControl)
 
@@
 weight_to_use <- mean(weight_to_use_per_partition) 
-# 5
+# 20/3
 
@@
 
-penalty_data is computed on train_subset and is used with test_subset as historical learning of efficiency of each seller in delivering to carrier
-Similarly delay in payment approval is used to incorporate role of customer zip code in delaying this processing, which maybe due to fraud analysis etc.
+### Other features
+
+#### Price feature
+We looked into the intuition that order item's price could influence the delivery timelines as expensive items ordered may have priority shipping or be delivered with caution.
+
+To understand its influence on the delivery timelines, we plotted numbers of hours taken for delivery against the price of the ordered item.
+In order to see the trend, gam method based smoothing curve was plotted on the graph.
+
+```{r Explore price relation with number_of_hours, echo=FALSE, message=FALSE}
+training_dataset  %>% ggplot(aes(price,number_of_hours)) + geom_smooth() + xlab('Price of ordered item') + ylab('Number of hours taken for delivery')
+```
+In above plot, we noted that number of hours taken for delivery vary as the price of the ordered item changes.
+It appeared to increase significantly as the price of ordered item increased and decreased when the price reached towards higher limits.
+Though confidence interval indicated higher variability in delivery timelines at higher prices.
+
+This behavior could be useful in estimating delivery timelines. We decided to consider this as a feature for further analysis for inclusion in the model.
+
+#### Product weight feature
+Next we looked into the intuition that product weight can influence the delivery timelines as heavier items like furniture, bed may be transported slowly compared to lighter items.
+
+For this we plotted numbers of hours taken for delivery against the weight of the product ordered in grams.
+In order to see the trend in this over-plotted graph, gam method based smoothing curve was also plotted on the graph.
+
+```{r Explore product weight relation with number_of_hours, echo=FALSE, message=FALSE}
+training_dataset  %>% ggplot(aes(product_weight_g,log10(number_of_hours))) + geom_point() + geom_smooth() + xlab('Weight of ordered product in grams') + ylab('Number of hours taken for delivery [log10(n) scale]')
+```
+In above plot, we noted that number of hours taken for delivery vary as the product weight changes.
+This behavior could be useful in estimating delivery timelines.
+We decided to consider this as a feature for further analysis for inclusion in the model.
+
+#### Freight value feature
+Next we looked into the intuition that freight value can help determine the delivery timelines as higher freight value could indicate complexity in shipment, extra processing which would need time, among others.
+
+For this we plotted numbers of hours taken for delivery against the freight value of the product ordered.
+In order to see the trend in this over-plotted graph, gam method based smoothing curve was also plotted on the graph.
+
+```{r Explore freight value relation with number_of_hours, echo=FALSE, message=FALSE}
+training_dataset  %>% ggplot(aes(freight_value,log10(number_of_hours))) + geom_point() + geom_smooth() + xlab('Freight value of ordered product') + ylab('Number of hours taken for delivery [log10(n) scale]')
+```
+In above plot, we noted that number of hours taken for delivery vary as the freight value of product changes.
+It appeared to increase as freight value increased for smaller freight values, then plateaued and decreased as freight value reached higher limits The decrease could be due to priority shipping coming at a cost.
+This behavior could be useful in estimating delivery timelines.
+We decided to consider this as a feature for further analysis for inclusion in the model.
+
+## Model Selection
 ```{r Prepare training and test data for model tuning and development,include=FALSE}
@@
 
-5 fold cross validation
+Post evaluation with multiple features, the final set of features identified for model selection were as below::
+
+```{r Final set of features identified, echo=FALSE}
+train_subset %>% colnames() %>% knitr::kable(col.names=c('Feature'),caption='Features List')
+```
+
+To start building our model, training and test subset were prepared.
+For this, we further partitioned our training data set i.e.
+training_dataset into a split of 80/20 between train_subset and test_subset to ensure we use only the train_subset for training and tuning the model and test_subset to evaluate the model and measure RMSE.
+This would ensure there is no data leak.
+The split like earlier was done such that distribution of outcome i.e.
+number_of_hours was similar in both training and test data.
+
+The derived features selected earlier were prepared based on train_subset and applied over test_subset, to ensure we maintain the environmental condition where test data is not known to us.
+
+In addition, as earlier shared, categorical features were converted to numeric form with the help of target encoding with a weight term computed earlier for purpose of regularization and to avoid overfitting.
+Also, any undesired features were removed, reducing train_subset and test_subset to only the desired features.
+
+### k-nearest neighbors(knn) model
+k-nearest neighbors(knn) being based on data points of nearest neighbours was a good starting point.
+And the fact that knn model does not make assumptions about the dataset, made it a good candidate for initial performance evaluation.
+
+To use knn model, features were scaled and normalized to prevent overfitting as knn internally calculates Euclidean distance for finding the nearest data-points.
+
+In addition, for knn model training and tuning, we carried our 5-fold cross validation, which helped to remove any randomness from our tuning parameters.
+This helped reduced the probability of overfitting.
+
+5-fold cross validation was carried out.
 80/20 split
-TODO remove p=0.8 from knn trainControl as it is not used.
 
@@
 performance_metric
-#183.7767 for k=63
+#183.0028 for k=77
 
@@
 
-Random forest
+### Random forest model
 However, it is important to know your data and keep in mind that a Random Forest can’t extrapolate.
@@
 performance_metric
-# 174.2496 for mtry 2. Technically there are total of 11 predictors. The seller_state was not of much help. it was actually increasing RMSE. So have ignored it.
+# 173.3481 for mtry 2. Technically there are total of 11 predictors. The seller_state was not of much help. it was actually increasing RMSE. So have ignored it.
 
 save(rf_fit,file='rda/rf_model.rda')
+```
+
+==TODO include gam model if required.
+
+### Ensemble model 
+```{r Ensemble, include=FALSE, eval= FALSE}
+load('rda/knn_model.rda')
+load('rda/rf_model.rda')
+set.seed(1, sample.kind = 'Rounding')
+
+predictions_knn <- predict(knn_fit,scaled_test_subset %>% select(-number_of_hours))
+predictions_rf <- predict(rf_fit,test_subset %>% select(-number_of_hours))
+
+predictions_df <- data.frame(predictions_knn,predictions_rf)
+
+# mean approach
+
+performance_metric <- performance_metric %>% 
+  add_row(type='Ensemble model(partial dataset)',
+          value=RMSE(rowMeans(predictions_df),(test_subset %>% pull(number_of_hours) )))
+#175.7183
+
+# max approach
+RMSE(predictions_df %>% mutate(max=pmax(predictions_knn,predictions_rf)) %>% pull(max),(test_subset %>% pull(number_of_hours)))
+#177.7854
+
+# min approach
+RMSE(predictions_df %>% mutate(min=pmin(predictions_knn,predictions_rf)) %>% pull(min),(test_subset %>% pull(number_of_hours)))
+#178.7218
 ```
@@
 
-```{r ensemble, include=FALSE, eval= FALSE}
-load('rda/knn_model.rda')
+# Results
+Post model tuning and evaluation, random forest algorithm emerged as the right fit for our data. The tuning parameters identified for the final random forest algorithm are shared below:
+
+```{r Tuning parameters to use for final random forest, echo=FALSE}
 load('rda/rf_model.rda')
 
-predictions_knn <- predict(knn_fit,scaled_test_subset %>% select(-number_of_hours))
-predictions_rf <- predict(rf_fit,test_subset %>% select(-number_of_hours))
-
-predictions_df <- data.frame(predictions_knn,predictions_rf)
-
-RMSE(rowMeans(predictions_df),(test_subset %>% pull(number_of_hours) ))
-176.5041
-```
+data.frame(tuning_parameter='mtry', value=rf_fit$finalModel$mtry) %>% add_row(tuning_parameter='number_of_trees', value=500) %>% knitr::kable(caption='Random forest tuning parameters')
+```
+
+Till this point, we had used only training data set i.e.
+training_dataset to identify the features, evaluate various models and tune them.
+Having selected the final model, we prepared first the derived features, which as shared earlier were prepared on training_dataset.
+The features were then applied over validation_dataset, which was the dataset kept aside till now.
+
+In addition, categorical features were converted to numeric form with the help of target encoding with a weight term computed earlier for purpose of regularization and to avoid overfitting.
+Also, any undesired features were removed, reducing training_dataset and validation_dataset to only the desired features.
 
@@
           value=RMSE((validation_dataset %>% pull(number_of_hours) ),predictions))
-# 189.2705 for mtry=2, ntree=500
+# 188.5793 for mtry=2, ntree=500
 
@@
 
-five sets each having a training set and testing set.
-This ensured that we used only edx dataset while we developed our model.
-In addition, the multiple partitions helped to remove any randomness from our tuning parameters and to get an appopriate estimate of performance for our incremental model.
-This helped reduced the probability of overfitting.
-
-we should consider if we need time stamp or we can take purchase time as 0 and calculated approved at from it.
-though we may lose granularity in terms of time as truck may come everyday at 4.
-so after 4 you can ship , before 4 you cant
-We can be agnostic to purchase time and just predict number of hours for delivery for given purchase.
-
-TODO -> if there is delayed to carrier precendence, we could use that probability and average delay to carrier to fine tune our results.
-Similarly we could use other metrics like payment approval time etc.
+Below is the summary of model against the performance metric i.e. RMSE
+```{r Performance metric post final model validation, echo=FALSE}
+performance_metric %>% knitr::kable(caption='Performance Metric(RMSE) Summary')
+```
+```
+
+```{r Final knn model traing with data scaled, include=FALSE, eval= FALSE}
+set.seed(seed=1,sample.kind = 'Rounding')
+preProcessScalingValues <- preProcess(training_dataset %>% select(-number_of_hours), method = c("center", "scale"))
+
+scaled_train_subset <- predict(preProcessScalingValues, training_dataset) 
+scaled_test_subset <- predict(preProcessScalingValues, validation_dataset)
+
+final_knn_fit <- knnreg(number_of_hours~., scaled_train_subset,k = knn_fit$finalModel$k)
+
+final_knn_fit
+
+predictions <- predict(final_knn_fit,scaled_test_subset %>% select(-number_of_hours))
+
+RMSE(predictions,(scaled_test_subset %>% pull(number_of_hours) ))
+# 202.8552
+
+save(final_knn_fit,file='rda/final_knn_fit.rda')
+```
+
+```{r Final ensemble model, include=FALSE, eval= FALSE}
+load('rda/final_knn_fit.rda')
+load('rda/final_model.rda')
+
+set.seed(seed=1,sample.kind = 'Rounding')
+
+predictions_knn <- predict(final_knn_fit,scaled_test_subset %>% select(-number_of_hours))
+predictions_rf <- predict(final_model,validation_dataset %>% select(-number_of_hours))
+
+predictions_df <- data.frame(predictions_knn,predictions_rf)
+
+RMSE(rowMeans(predictions_df),(validation_dataset %>% pull(number_of_hours) ))
+# 193.3459
+```
+
+For the **final model**, the performance metric i.e.
+RMSE = **`r performance_metric %>% filter(type=='Final model') %>% pull(value)`** captured against **validation_dataset** was much better than what Olist store had as identified from the dataset.
+The newly built model can be used for predicting number of hours required for delivery of a given ordered item.
+
+# Conclusion
+As part of this project, the goal was to build a delivert prediction system which could predict the number of hours required to delivered the ordered item rating with minimal RMSE.
+
+During the course of project, we carried out data cleaning, identified features and derived features from existing information available.
+In addition, converted categorical features to numeric form and applied various techniques like target encoding, scaling and normalization as required and regularized the required features.
+
+Once the feature analysis and selection was completed, we used the features to tune models like k-nearest neighbours, random forest.
+We carried out k-fold cross validation to remove any randomness from the model training and tuning process.
+Finally we reached to a model which performs within desired range of RMSE and can be useful for predicting number of hours required for delivery of a given ordered item.
+Due care was taken to prevent overfitting from influencing our model.
+
+While the model is a significant improvement over what Olist store had, there is lot of room to improve it.
+It can be improved for incremental learning so that it can learn from new orders and their delivery details once they become available.
+In addition, it could be improved to handle new sellers or product categories.
+
+Having said that the delivery prediction system built as part of the project is performant and can be leveraged to develop further.

### 2021-07-20 finalReportPrep.Rmd
fine tuning the document

@@
 ---
-title: "CYO Project Report - Delivery Estimation System"
+title: "CYO Project Report - Delivery Prediction System"
 author: "Asham Vohra"
@@
 ```{r Install or load packages,include=FALSE, echo = FALSE}
-# TODO use conditional loading or installation of library
-# Remove plotly related code for pdfs
+#Install or load packages
 if (!require('dplyr')) install.packages('dplyr'); library('dplyr')
@@
 
-# TODO if we are not using payment_approved_at or few fields we can remove the filter logic based on that.
+# Filtered out any orders where delivery was not made 
+# Also filtered out orders where delivery was made but required delivery time columns were not available.
 orders_dataset <- orders_dataset %>% 
@@
 
+# check how they are structured
 orders_dataset %>% head()
@@
 
+# check how they are structured
 order_items_dataset %>% str()
@@
 # geo location details
-# in brazil while 5 digits zip code is good enough to identify state, it may not be correct to identify city
-# it can have geolocation_zip_code_prefix shared across multiple places
+# In brazil zip code is 8 digits. So first 5 digits zip code is good enough to identify an area but it does not uniquely identify customer location.
+# So because of this, there are multiple rows for same geolocation_zip_code_prefix
 # refer: https://codigo-postal.org/en-us/brazil/
-# The Postal Code Address (CEP) is an eight-digit number set that has by main objective to guide and accelerate the distribution of objects sent through the national mail in Brazil.
-The current structure of the Brazilian postal code is 8 (eight) digits, divided into two parts:
-
-# first part composed of 5 digits where each digit represents the 1) region, 2) sub-region, 3) sector, 4) subsector, 5) subsector divider;
-# the second part is composed of 3 digits, separated by a hyphen from the first part and represents the Distribution Identifiers.
-# You can navigate through states, then cities and finally neighborhoods and streets until you find the corresponding postal code.
-
-# geolocation_dataset <- read.csv(file.path('data','olist_geolocation_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) 
-geolocation_dataset <- read.csv(file.path('data','olist_geolocation_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) %>% select(-geolocation_city,-geolocation_state)
-
+#
+# The first five digits represent the 1) region, 2) sub-region, 3) sector, 4) subsector, 5) subsector divider;
+geolocation_dataset <- read.csv(file.path('data','olist_geolocation_dataset.csv'),
+                                header=TRUE,stringsAsFactors = FALSE) %>%
+  select(-geolocation_city,-geolocation_state)
+
+# check how they are structured
 geolocation_dataset %>% head()
@@
 
-# Important: many geolocation_city have same name but different ascent component. Example: niteroi and niterói
-# initially we can see 
-# geolocation_dataset %>% dplyr::count(geolocation_zip_code_prefix,geolocation_city,geolocation_state) %>% nrow()
-
-# post transformation to resolve any Accent or special characters from the city name
-# geolocation_dataset <- geolocation_dataset %>% mutate(geolocation_city=stri_trans_general(geolocation_city,"Any-ASCII"))
-
-# above is not done as the way zip codes are in brazil, first five digits can correspond to different cities and so we will remove the city.
-We also removed state as there seems to be incorrect entries in the sellers data.
-This either needs to be fixed or we can limit ourselves to lat/long value
-
-# geolocation_dataset %>% dplyr::count(geolocation_zip_code_prefix,geolocation_city,geolocation_state) %>% nrow()
+# check count of rows per geolocation_zip_code_prefix
+geolocation_dataset %>% dplyr::count(geolocation_zip_code_prefix) %>% nrow()
 
 # as each geolocation_zip_code_prefix can have many data points, we will limit these to one by taking mean for lat, long location
-geolocation_dataset <- geolocation_dataset %>% group_by(geolocation_zip_code_prefix) %>% summarise(geolocation_lat=mean(geolocation_lat),geolocation_lng=mean(geolocation_lng)) %>% ungroup()
-#geolocation_dataset <- geolocation_dataset %>% group_by(geolocation_zip_code_prefix,geolocation_city,geolocation_state) %>% summarise(geolocation_lat=mean(geolocation_lat),geolocation_lng=mean(geolocation_lng)) %>% ungroup()
-
-# TODO review it
-#%>% mutate(geolocation_state=as.factor(geolocation_state)) 
-# TODO review if geolocation_zip_code_prefix and geolocation_city can have multiple states with them.
-
-# remove whitespace before and after reduces one city
-# still issues are there and can be seen by
-# zip_codes_of_interest <- geolocation_dataset %>% mutate(geolocation_city=str_trim(geolocation_city)) %>% dplyr::count(geolocation_zip_code_prefix) %>% filter(n>1)  %>% pull(geolocation_zip_code_prefix)
-
-# geolocation_dataset %>% filter(geolocation_zip_code_prefix %in% zip_codes_of_interest) %>% view()
-# to filter out duplicates
-# we need not worry about city or state, we can pick any as long as lat, lng values are similar
+geolocation_dataset <- geolocation_dataset %>% group_by(geolocation_zip_code_prefix) %>% 
+  summarise(geolocation_lat=mean(geolocation_lat),geolocation_lng=mean(geolocation_lng)) %>%
+  ungroup()
 ```
@@
 customers_dataset <- read.csv(file.path('data','olist_customers_dataset.csv'),header=TRUE,stringsAsFactors = FALSE)
-# %>% mutate(customer_zip_code_prefix=as.factor(customer_zip_code_prefix),customer_id=as.character(customer_id ))
-
+
+# check how they are structured
 customers_dataset %>% head()
@@
 sellers_dataset <- read.csv(file.path('data','olist_sellers_dataset.csv'),header=TRUE,stringsAsFactors = FALSE)
-# TODO review if we need it
-# %>% mutate(seller_zip_code_prefix=as.factor(seller_zip_code_prefix),seller_id=as.character(seller_id ))
-
+
+# check how they are structured
 sellers_dataset %>% head()
@@
 # Filter out data where sellers location details are not available
-#
-# Important: many geolocation_city have same name but different ascent component. Example: niteroi and niterói
-# These all need to be handled in data wrangling
-
 # we noticed some seller_zip_code_prefix are not present in geolocation dataset. So we will ignore them
@@
 ### k-nearest neighbors(knn) model
-k-nearest neighbors(knn) being based on data points of nearest neighbours was a good starting point.
-And the fact that knn model does not make assumptions about the dataset, made it a good candidate for initial performance evaluation.
-
-To use knn model, features were scaled and normalized to prevent overfitting as knn internally calculates Euclidean distance for finding the nearest data-points.
-
-In addition, for knn model training and tuning, we carried our 5-fold cross validation, which helped to remove any randomness from our tuning parameters.
-This helped reduced the probability of overfitting.
-
-5-fold cross validation was carried out.
-80/20 split
+k-nearest neighbors(knn) relies on finding nearest neighbors and uses average of the nearest neigbbors for prediction.
+In addition, it does not make assumptions about the dataset.
+In our case limited dataset and that the alogrithm does not make any assumptions, made it a good candidate for model development and performance evaluation.
+
+To use knn model, features were scaled and normalized to prevent bias towards any features, as knn internally calculates Euclidean distance for finding the nearest data-points.
+
+In addition, for knn model training and tuning, we carried our 5-fold cross validation providing a range of values of k to use for training the model.
+Here, k means number of neighbors to be considered.
+The cross validation helped in removing any randomness from our tuning parameters.
+In addition, this helped reduce the probability of overfitting.
 
@@
 tuningControl=trainControl(method = "cv", number = 5)
-knn_fit <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(56,63,70,77)),trControl=tuningControl)
-
+knn_fit <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(45,71,77,83)),trControl=tuningControl)
+
+save(knn_fit,file='rda/knn_model.rda')
+```
+
+Details of the trained knn model can be found below:
+
+```{r trained knn model based on train_subset, echo=FALSE }
+load('rda/knn_model.rda')
 knn_fit
 
+knn_fit %>% ggplot()
+```
+
+Based on model results, k value of `r knn_fit$finalModel$k` was identified to be the one with lowest RMSE.
+
+```{r trained knn model validation, include=FALSE, eval=FALSE}
 predictions <- predict(knn_fit,scaled_test_subset %>% select(-number_of_hours))
@@
 
-save(knn_fit,file='rda/knn_model.rda')
-```
-
-Regression problem.
-Algorithms considered
-- Random Forest
-- knn - k nearest neighbors 
-- gam
+save(performance_metric,file='rda/performance_metric.rda')
+```
+
+Below is the summary of each model and knn model(which used partial dataset) against the performance metric i.e. RMSE
+
+```{r performance metric for knn model based on test data, echo =FALSE}
+load('rda/performance_metric.rda')
+performance_metric[1:3,] %>% knitr::kable(caption='Performance Metric(RMSE) Summary')
+```
+
+As we can see the performance metric i.e. RMSE is significantly better for knn model (though with partial dataset at this point of time), than the one being used by Olist store.
 
 ### Random forest model
-However, it is important to know your data and keep in mind that a Random Forest can’t extrapolate.
-It can only make a prediction that is an average of previously observed labels.
-In this sense it is very similar to KNN.
-In other words, in a regression problem, the range of predictions a Random Forest can make is bound by the highest and lowest labels in the training data.
-This behavior becomes problematic in situations where the training and prediction inputs differ in their range and/or distributions.
-This is called covariate shift and it is difficult for most models to handle but especially for Random Forest, because it can’t extrapolate.
-From : https://towardsdatascience.com/a-limitation-of-random-forest-regression-db8ed7419e9f
+Next, we used Random forest model, which is basically ensemble of multiple decision trees.
+The training dataset is split into multiple subsets.
+For each such subset, decision tree is created.
+However, randomness is introduced in creation of decision trees as at each node, the tree leverages a number of random features to determine the best split.
+Once the various models are ready, they are used to prepare predictions.
+The final prediction is reached by taking average of all the trees.
+
+The ensemble of multiple decision trees helps prevent overfitting and creates a more stable prediction.
+While, the number of random features at each node of decision tree to determine best split, helps reduce bias and variance in the model.
+It is these qualities of random forest, which prompted us to consider this for model development and evaluation.
+
+To train and tune random forest model, we carried our 4-fold cross validation providing a range of values of mtry to use for training the model.
+Here, mtry means number of random features used to determine the best split at each node of the decision tree.
+In addition, we configured the number of decision trees to create as 200.
+Here, the cross validation not only helped to reduce the probability of overfitting but also helped in identifying the tuning parameter i.e.
+mtry which would work across randomly selected datasets.
 
@@
 
+save(rf_fit,file='rda/rf_model.rda')
+```
+
+Details of the trained random forest model can be found below:
+```{r trained random forest model based on train_subset, echo=FALSE}
+load('rda/rf_model.rda')
 rf_fit
 
+rf_fit %>% ggplot()
+```
+
+```{r trained random forest model validation, include=FALSE, eval=FALSE}
 predictions <- predict(rf_fit,test_subset %>% select(-number_of_hours))
@@
 # 173.3481 for mtry 2. Technically there are total of 11 predictors. The seller_state was not of much help. it was actually increasing RMSE. So have ignored it.
-
-save(rf_fit,file='rda/rf_model.rda')
-```
-
-==TODO include gam model if required.
+```
+
+Based on model results, mtry value of `r rf_fit$finalModel$mtry` was identified to be the one with lowest RMSE.
+
+Below is the summary of each model and random forest model(which used partial dataset) against the performance metric i.e. RMSE
+
+```{r performance metric for random forest model based on test data, echo =FALSE}
+load('rda/performance_metric.rda')
+performance_metric[1:4,] %>% knitr::kable(caption='Performance Metric(RMSE) Summary')
+```
+
+As we can see the performance metric i.e.
+RMSE was significantly better for random forest model (though with partial dataset at this point of time), than the one being used by Olist store.
+In addition, we could see improvement in RMSE when compared to knn model for the same dataset.
 
 ### Ensemble model 
-```{r Ensemble, include=FALSE, eval= FALSE}
+The ensemble models use multiple algorithms and if properly developed, they can be built to take learnings from the different models and generate better predictions.
+Since we had knn and random forest models tuned and prepared, we used them to create variety of ensemble models.
+
+#### Average based ensemble model
+We took previously trained and tuned knn and random forest model and created ensemble of both.
+For this, predictions using both knn and random forest models were computed.
+Then mean of the predicitions was taken.
+This was used to calculate performance metric i.e.
+RMSE.
+
+```{r Ensemble of knn and rf model using partial data, include=FALSE, eval= FALSE}
 load('rda/knn_model.rda')
@@
 performance_metric <- performance_metric %>% 
-  add_row(type='Ensemble model(partial dataset)',
+  add_row(type='Avg based ensemble model(partial dataset)',
           value=RMSE(rowMeans(predictions_df),(test_subset %>% pull(number_of_hours) )))
@@
 
-```{r Load performance metric, include=FALSE}
+Below is the summary of each model and ensemble model(which used partial dataset) against the performance metric i.e. RMSE
+
+```{r performance metric for ensemble model based on test data, echo =FALSE}
 # required only to generate report
 load('rda/performance_metric.rda')
-```
+performance_metric[1:5,] %>% knitr::kable(caption='Performance Metric(RMSE) Summary')
+```
+
+As we can see the performance metric i.e.
+RMSE was significantly better for ensemble model (though with partial dataset at this point of time), than the one being used by Olist store.
+However, it was not better than the performance metric captured for random forest model.
+
+#### Weighted average based ensemble model
+Other alternative ensemble technique like weighted average approach was explored.
+Again, we took previously trained and tuned knn and random forest model and created ensemble of both.
+However, this time, final predictions were based on weighted average of the predictions of the constituient models.
+
+The results of performance metric i.e. RMSE for the ensemble model against increasing weights for knn model based predictions are plotted on the graph.
+
+```{r weighted average ensemble approach with partial dataset, include=FALSE, eval=FALSE}
+# weighted approach
+rmses_weighted_ensemble_model <- sapply(seq(0,1,0.15),function(a) {
+  RMSE((test_subset %>% pull(number_of_hours)),
+        predictions_df %>% mutate(value=predictions_knn*a+predictions_rf*(1-a)) %>% 
+          pull(value)
+       )
+  })
+save(rmses_weighted_ensemble_model,file='rda/rmses_weighted_ensemble_model.rda')
+```
+
+```{r Summary of performance metric for weighted ensemble approach, echo=FALSE}
+load('rda/rmses_weighted_ensemble_model.rda')
+weights_for_ensemble_model <- seq(0,1,0.15)
+
+data.frame(weight_for_knn_pred=weights_for_ensemble_model,rmse=rmses_weighted_ensemble_model) %>%
+  ggplot(aes(weight_for_knn_pred,rmse)) + geom_line() + geom_point() + 
+  xlab('Weight used(for knn prediction)') + ylab('Resultant RMSE')
+```
+
+We noted that weighted ensemble model had lowest RMSE of `r min(rmses_weighted_ensemble_model)` at weight=`r weights_for_ensemble_model[which.min(rmses_weighted_ensemble_model)]`
+
+In addition, other ensembling techniques like stacking were explored, however they could not improve the RMSE.
+
+### Selected model
+Given above results of performance metric i.e. RMSE across knn, random forest and ensemble models, we selected random forest as the final model for our delivery prediction system.
 
 # Results
-Post model tuning and evaluation, random forest algorithm emerged as the right fit for our data. The tuning parameters identified for the final random forest algorithm are shared below:
+Post model tuning and evaluation, random forest model emerged as the best fit for our data among the models considered for selection.
+The tuning parameters identified for the final random forest algorithm are shared below:
 
@@
 
-final_model
-
 save(final_model,file='rda/final_model.rda')
+```
+
+```{r trained final random forest model based on training_dataset, echo=FALSE}
+load('rda/final_model.rda')
+#final_model
+
+plot(final_model,main='Random forest model error against number of trees')
 ```
@@
 
-Below is the summary of model against the performance metric i.e. RMSE
+Below is the summary of models against the performance metric i.e. RMSE
 ```{r Performance metric post final model validation, echo=FALSE}
 performance_metric %>% knitr::kable(caption='Performance Metric(RMSE) Summary')
-```
 ```
@@
 # Conclusion
-As part of this project, the goal was to build a delivert prediction system which could predict the number of hours required to delivered the ordered item rating with minimal RMSE.
+As part of this project, the goal was to build a delivery prediction system which could predict the number of hours required to delivered the ordered item rating with minimal RMSE.
 
@@
 
-Once the feature analysis and selection was completed, we used the features to tune models like k-nearest neighbours, random forest.
+Once the feature analysis and selection was completed, we used the features to tune models like k-nearest neighbors, random forest.
 We carried out k-fold cross validation to remove any randomness from the model training and tuning process.
-Finally we reached to a model which performs within desired range of RMSE and can be useful for predicting number of hours required for delivery of a given ordered item.
+This additionally helped to reduce probability of overfitting.
+Post this, we explored multiple ensemble models using techniques like averaging, weighted average, stacking, among others.
+Finally we selected random forest based model and the model so trained performs within desired range of RMSE and can be useful for predicting number of hours required for delivery of a given ordered item.
 Due care was taken to prevent overfitting from influencing our model.

### 2021-07-20 finalReportPrep.Rmd
more or less final

@@
 
+# while there are small number of order items for the concerned sellers, we will ignore them as we will still have sufficient data for analysis.
 order_items_dataset %>% inner_join(sellers_to_ignore,by='seller_id') %>% nrow()
-# while there are decent number of order items for the concerned sellers, we will ignore them as we will still have sufficient data for analysis.
 
 # let's use the one of interest
-sellers_dataset <- sellers_dataset %>% inner_join(geolocation_dataset,by=c('seller_zip_code_prefix'='geolocation_zip_code_prefix')) %>% rename(geolocation_lat_seller=geolocation_lat,geolocation_lng_seller=geolocation_lng)
-
-# TODO there seems to be an issue (Should be resolved now)
-# geolocation_dataset seems to have 4 entries for same zip code . example 13454
-# sellers_dataset %>% inner_join(geolocation_dataset,by=c('seller_zip_code_prefix'='geolocation_zip_code_prefix')) %>% dplyr::count(seller_id) %>% slice_max(order_by=n,n=5)
-
-# we tried another approach by joining over all zip codee, city and state but it led to no match
-# sellers_dataset %>% mutate(seller_city=stri_trans_general(seller_city,"Any-ASCII")) %>% inner_join(geolocation_dataset,by=c('seller_zip_code_prefix'='geolocation_zip_code_prefix','seller_city'='geolocation_city','seller_state'='geolocation_state')) %>% str()
-# TODO -- if we analyse we can see city and state for seller data differ in both datasets. so best to stick to zip code. city and state are prone to errors.
-# we decided against using city and state.
+sellers_dataset <- sellers_dataset %>% 
+  inner_join(geolocation_dataset,by=c('seller_zip_code_prefix'='geolocation_zip_code_prefix')) %>% 
+  rename(geolocation_lat_seller=geolocation_lat,geolocation_lng_seller=geolocation_lng)
 ```
@@
 
+# while there are small number of orders for these customers, we will ignore them as we will still have sufficient data for analysis.
 orders_dataset %>% inner_join(customers_to_ignore,by='customer_id') %>% nrow()
-# while there are decent number of orders for these customers, we will ignore them as we will still have sufficient data for analysis.
 
 # let's use customers of interest
-customers_dataset <- customers_dataset %>% inner_join(geolocation_dataset,by=c('customer_zip_code_prefix'='geolocation_zip_code_prefix')) %>% rename(geolocation_lat_customer=geolocation_lat,geolocation_lng_customer=geolocation_lng)
+customers_dataset <- customers_dataset %>% 
+  inner_join(geolocation_dataset,by=c('customer_zip_code_prefix'='geolocation_zip_code_prefix')) %>% 
+  rename(geolocation_lat_customer=geolocation_lat,geolocation_lng_customer=geolocation_lng) 
 ```
@@
 ```{r Combine data sets, include=FALSE, echo=FALSE}
-# we can combine orders delivery information (orders_dataset) with orders items and seller related info (order_items_dataset)
-# orders_dataset %>% inner_join(order_items_dataset,by='order_id') %>% head()
-
-# let's combine with seller details
-# orders_dataset %>% inner_join(order_items_dataset,by='order_id') %>% inner_join(sellers_dataset,by='seller_id') %>% head()
-
-# let's combine with unique customer and its location dataset
-# orders_dataset %>% inner_join(order_items_dataset,by='order_id') %>% inner_join(sellers_dataset,by='seller_id') %>% inner_join(customers_dataset,by='customer_id') %>% select(-customer_id) %>% head()
-
-# let's combine with products data as well
-# orders_dataset %>% inner_join(order_items_dataset,by='order_id') %>% inner_join(sellers_dataset,by='seller_id') %>% inner_join(customers_dataset,by='customer_id') %>% inner_join(products_dataset,by='product_id') %>% select(-customer_id,-order_status,-seller_city,-seller_state,-customer_city,-customer_state) %>% head()
-
 # so raw data
+# in Brazil first digit of the five code zip code prefix can be 0. So we explicitly add 0 to it.
 raw_data <- orders_dataset %>% 
@@
 
+```{r Prepare final wrangled data set, include=FALSE}
+# Prepare final wrangled data set we will predict number of hours to deliver based on 
+# various predictors. This seems more appropriate. Though as a result of this, we won't be able 
+# to use night, day time data of purchase which may influence delivery time (say to carrier it 
+# is informed only next day)
+wrangled_data <- raw_data %>%
+  mutate(order_delivered_customer_date=as.numeric(order_delivered_customer_date)) %>%
+  mutate(number_of_hours=(order_delivered_customer_date-order_purchase_timestamp)/3600) %>%
+  select(-order_delivered_customer_date,-order_estimated_delivery_date)
+```
+
+Having wrangled the datasets individually we combined them into a single dataset.
+While the dataset had many attributes, based on their frequency and our intuition shared earlier we zeroed on the below fields for further consideration.
+In addition, it is important to note that for model development we created outcome column number_of_hours based on existing fields i.e.
+order_delivered_customer_date and order_purchase_timestamp as this is what we would want to predict whenever we get a new order i.e.
+given this order, how many number of hours are required to deliver the product to the customer.
+
+The wrangled dataset after incorporating various datasets available to us looked like below:
+```{r Final wrangled data,echo=FALSE}
+wrangled_data %>% head(2) %>% as.matrix()  %>% t() %>% 
+  knitr::kable(col.names=c('Example Value1','Example Value2'), caption='Final wrangled data set')
+```
+
+## Performance Metric
+
+Before we dived into building our delivery prediction system, we identified our metric of choice i.e.
+RMSE for evaluating current delivery estimates being made by Olist and for comparing our delivery prediction system against that, in addition to improving our system as we built it.
+
+*  RMSE is the root mean square error and that's what was used for this project to evaluate models and to optimize them for fitment.
+* As we know, $RMSE = sqrt( 1/N * sum((\hat{y}-y)^2) )$
+    + Here, $\hat{y}$ is the prediction. In context of this project, predicted number of hours for delivering the order item,
+    + and y is the actual outcome. In context of this project, actual number of hours taken for delivering the order item.
+
 ```{r RMSE, include=FALSE}
@@
 
+### Current RMSE
+
+Before we dived into building our delivery prediction system, we looked into current RMSE.
+To calculate this, we leveraged order_estimated_delivery_date i.e.
+estimated delivery date and order_delivered_customer_date i.e.
+actual delivery timestamp, which was available in orders data set.
+
+While order_estimated_delivery_date did not have time details, order_delivered_customer_date did.
+Since order_delivered_customer_date could be rounded off to same day or next day, we computed range of RMSE which can be calculated using either of the approach.
+
+This RMSE i.e. our performance metric was later used as benchmark for modelling. The delivery prediction system being built attempted to improve on this metric.
+
+Below were the current RMSE calculations based on how we interpret the actual delivery timestamp.
+
 ```{r Current RMSE, include=FALSE, echo = FALSE}
-# current RMSE can be calculated by finding number of hours for predicted delivery time given by order_estimated_delivery_date vs actual outcome/number of hours give by order_delivered_customer_date
-# However, note that order_estimated_delivery_date has no timestamp, so we will remove timestamp component, round date using floor_date, which means any delivery done during the day say at 2017-10-10 21:25:13 is treated to be done on 2017-10-10
-
-# first we convert it to desired date
+# current RMSE can be calculated by finding number of hours for predicted delivery time 
+# order_estimated_delivery_date vs actual outcome/number of hours give by order_delivered_customer_date
+# However, note that order_estimated_delivery_date has no timestamp, so we will calculate RMSE 
+# for both cases i.e. delivery date was taken as previous date or next date. 
+# For this, from order_delivered_customer_date we will remove timestamp component, round date 
+# using floor_date, which means any delivery done during the day say at 2017-10-10 21:25:13 is treated to be done on 2017-10-10
+# Similarly we will round upwards using ceiling_date.
+
+# For this, first we convert it to desired date
 # then we covert date to numeric i.e. to convert it to seconds
-# then we calculate time_taken_from_order_to_delivery, estimated_time_from_order_to_delivery in seconds
+# then we calculate time_taken_from_order_to_delivery (min and max as shared above), estimated_time_from_order_to_delivery in seconds
 data_for_rmse <- raw_data %>%
@@
 
-# if we look at graph we can  see there is huge difference between estimated data and actual delivery date
-# it is possible some products are bulky or require differen mode of transport like ship 
-data_for_rmse %>% mutate(diff=estimated_time_from_order_to_delivery-1/2*(time_taken_from_order_to_delivery_rangemax+time_taken_from_order_to_delivery_rangemin)) %>% ggplot(aes(diff)) + geom_histogram(binwidth = 24*2) + xlab('Estimation error where each bar represents 2 days')
-
 rm(data_for_rmse)
@@
 
-```{r Prepare final wrangled data set, include=FALSE}
-# we will predict number of hours to deliver based on various predictors.
-This seems more appropriate.
-Though as a result of this, we won't be able to use night, day time data of purchase which may influence delivery time (say to carrier it is informed only next day)
-wrangled_data <- raw_data %>%
-  mutate(order_delivered_customer_date=as.numeric(order_delivered_customer_date)) %>%
-  mutate(number_of_hours=(order_delivered_customer_date-order_purchase_timestamp)/3600) %>%
-  select(-order_delivered_customer_date,-order_estimated_delivery_date)
-```
-
-Having wrangled the datasets individually we combined them into a single dataset.
-While the dataset had many attributes, based on their frequency and our intuition shared earlier we zeroed on the below fields for further consideration.
-In addition, it is important to note that for model development we created outcome column number_of_hours based on existing fields i.e.
-order_delivered_customer_date and order_purchase_timestamp as this is what we would want to predict whenever we get a new order i.e.
-given this order, how many number of hours are required to deliver the product to the customer.
-
-The wrangled dataset after incorporating various datasets available to us looked like below:
-```{r Final wrangled data,echo=FALSE}
-wrangled_data %>% head(2) %>% as.matrix()  %>% t() %>% 
-  knitr::kable(col.names=c('Example Value1','Example Value2'), caption='Final wrangled data set')
-```
-
-## Performance Metric
-
-Before we dived into building our delivery prediction system, we identified our metric of choice i.e.
-RMSE for evaluating current delivery estimates being made by Olist and for comparing our delivery prediction system against that, in addition to improving our system as we built it.
-
-*  RMSE is the root mean square error and that's what was used for this project to evaluate models and to optimize them for fitment.
-* As we know, $RMSE = sqrt( 1/N * sum((\hat{y}-y)^2) )$
-    + Here, $\hat{y}$ is the prediction. In context of this project, predicted number of hours for delivering the order item,
-    + and y is the actual outcome. In context of this project, actual number of hours taken for delivering the order item.
-
-### Current RMSE
-
-Before we dived into building our delivery prediction system, we looked into current RMSE.
-To calculate this, we leveraged order_estimated_delivery_date i.e.
-estimated delivery date and order_delivered_customer_date i.e.
-actual delivery timestamp, which was available in orders data set.
-
-While order_estimated_delivery_date did not have time details, order_delivered_customer_date did.
-Since order_delivered_customer_date could be rounded off to same day or next day, we computed range of RMSE which can be calculated using either of the approach.
-
-This RMSE i.e. our performance metric was later used as benchmark for modelling. The delivery prediction system being built attempted to improve on this metric.
-
-Below were the current RMSE calculations based on how we interpret the actual delivery timestamp.
 ```{r Range of current rmse, echo=FALSE}
@@
 ```{r Method to derive distance data, include = FALSE}
-# we will use distGeo from geosphere package to compute distance in km between seller and customer.
-considered to be highly accurate.
-if it takes more time, we can consider using distVincentyEllipsoid or distHaversine as we are fine with accuracy of a km.
+# we will use distGeo from geosphere package to compute distance in km between seller and customer. considered to be highly accurate.
 extract_distance_between_seller_customer <- function(df) { 
@@
   semi_join(training_dataset,by="customer_zip_code_prefix")
-## TODO review if we want to change this join
-# validation_dataset <- validation_dataset %>% semi_join(training_dataset,by='seller_id')
 
@@
   weight_to_use_per_partition <- sapply(1:3, function(index) {
-  #lambda_data_set <- prepare_lambda_data_set(index)
-  #train_lambda_set <- lambda_data_set[[1]]
-  #test_lambda_set <- lambda_data_set[[2]]  
   train_weight_set <- train_subset_for_encoding[-test_weight_indices[,index],]
@@
   dataset_list <- enhance_datasets_with_derived_features(train_weight_set,test_weight_set,TRUE)
-  train_weight_set <- dataset_list[[1]] %>% select(-seller_id)#,-customer_zip_code_prefix)
-  test_weight_set <- dataset_list[[2]] %>% select(-seller_id)#,-customer_zip_code_prefix)
-
+  train_weight_set <- dataset_list[[1]] %>% select(-seller_id)
+  test_weight_set <- dataset_list[[2]] %>% select(-seller_id)
+
+  # calculating rmse for each weight
   weight_rmses <- sapply(weights, function(weight) {
@@
       scaled_test_subset <- predict(preProcessScalingValues, test_weight_set)
-
-      # TODO remove below scaling was erroneous. train scaling needs to be used for test data
-      # below is just demo of how to use scale
-      # scaled_train_subset <- train_subset %>% mutate(across(c(mean_delay_in_handover,distance_between_two_points,volume,product_weight_g,freight_value,price),scale)) 
-      # scaled_test_subset <- test_subset %>% mutate(across(c(mean_delay_in_handover,distance_between_two_points,volume,product_weight_g,freight_value,price),scale))
-
-      # TODO can we skip modelling while evaluating weight to use or for target encoding?
 
@@
 
+  # calculating rmse for each weight
   weight_rmses <- sapply(weights, function(weight) {
@@
 
-      # TODO remove below scaling was erroneous. train scaling needs to be used for test data
-      # below is just demo of how to use scale
-      # scaled_train_subset <- train_subset %>% mutate(across(c(mean_delay_in_handover,distance_between_two_points,volume,product_weight_g,freight_value,price),scale)) 
-      # scaled_test_subset <- test_subset %>% mutate(across(c(mean_delay_in_handover,distance_between_two_points,volume,product_weight_g,freight_value,price),scale))
-
-      # TODO can we skip modelling while evaluating weight to use or for target encoding?
-
       tuningControl=trainControl(method = "cv", number = 3)
@@
 scaled_test_subset <- predict(preProcessScalingValues, test_subset)
-
-# TODO remove below scaling was erroneous. train scaling needs to be used for test data
-# below is just demo of how to use scale
-# scaled_train_subset <- train_subset %>% mutate(across(c(mean_delay_in_handover,distance_between_two_points,volume,product_weight_g,freight_value,price),scale)) 
-# scaled_test_subset <- test_subset %>% mutate(across(c(mean_delay_in_handover,distance_between_two_points,volume,product_weight_g,freight_value,price),scale))
 
@@
 rf_fit <- train(number_of_hours~., train_subset, method='rf',tuneGrid = data.frame(mtry = c(1,2,3,4,5)),trControl=tuningControl,ntree=200)
-#rf_fit <- train(number_of_hours~., train_subset %>% select(-seller_state), method='rf',tuneGrid = data.frame(mtry = seq(5, 50, 10)),trControl=tuningControl,ntree=250) # nodesize = 50,maxnodes=2000,ntree=200
 
@@
 performance_metric
-# 173.3481 for mtry 2. Technically there are total of 11 predictors. The seller_state was not of much help. it was actually increasing RMSE. So have ignored it.
+# 173.3481 for mtry 2. There are total of 11 predictors.
 ```
@@
 # mean approach
-
 performance_metric <- performance_metric %>% 
@@
 #175.7183
-
-# max approach
-RMSE(predictions_df %>% mutate(max=pmax(predictions_knn,predictions_rf)) %>% pull(max),(test_subset %>% pull(number_of_hours)))
-#177.7854
-
-# min approach
-RMSE(predictions_df %>% mutate(min=pmin(predictions_knn,predictions_rf)) %>% pull(min),(test_subset %>% pull(number_of_hours)))
-#178.7218
 ```

### 2021-07-20 analysis.R

@@ -1,3 +1,1519 @@
-# TODO re
-library(dplyr)
-products_dataset <- read.csv(filte.path('data','olist_products_dataset.csv'),header=TRUE)
+#######################################################################################
+################################# Prepare Environment #################################
+#######################################################################################
+#Install or load packages
+if (!require('dplyr')) install.packages('dplyr'); library('dplyr')
+if (!require('lubridate')) install.packages('lubridate'); library('lubridate')
+if (!require('tidyverse')) install.packages('tidyverse'); library('tidyverse')
+if (!require('geosphere')) install.packages('geosphere'); library('geosphere')
+if (!require('caret')) install.packages('caret'); library('caret')
+if (!require('randomForest')) install.packages('randomForest'); library('randomForest')
+if (!require('knitr')) install.packages('knitr'); library('knitr')
+
+
+#######################################################################################
+#################################### Data wrangling ###################################
+#######################################################################################
+
+####################### Data wrangling for products data set
+
+# read the data from source and remove irrelevant columns
+products_dataset <- read.csv(file.path('data','olist_products_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) %>%
+  select(-product_name_lenght,-product_description_lenght,-product_photos_qty)
+
+# check how they are structured
+products_dataset %>% head()
+products_dataset %>% str()
+
+# load the product category name translations
+product_category_name_translation <- read.csv(file.path('data','product_category_name_translation.csv'),header=TRUE,stringsAsFactors = FALSE)
+
+product_category_name_translation %>% head()
+
+# convert product_name to english form wherever possible. If not available, existing product category name is retained
+products_dataset <- products_dataset %>% 
+  left_join(product_category_name_translation,by='product_category_name') %>% 
+  mutate(product_category_name=if_else(is.na(product_category_name_english),
+                                       product_category_name,
+                                       product_category_name_english)) %>% 
+  select(-product_category_name_english) 
+
+# further remove products for which data is not available i.e is NA
[…]

### 2021-07-20 finalReportPrep.Rmd
fine tuned files

@@
 ```
-
-== TODO additive smoothing/Bayes estimator with target encoding
 
@@
 
-However, to convert categorical variables to numeric values and still be able to limit features and make model training and evaluation less expensive, we applied **target encoding** on the product categories attribute.
-
-But to do that, first, we checked number of observations per product category and noticed that the distribution could require **regularization** to prevent overfitting.
-Secondly, target encoding could be erroneous if done against the random dataset.
-
-To solve these, we identified the regularization weight to use by carrying out target encoding for different datasets set of weights, selecting single weight per dataset based on performance metric i.e.
+However, to convert categorical variables to numeric values and still be able to limit features and make model training and evaluation less expensive, we applied additive smoothing based **target encoding** on the product categories attribute.
+
+Additive smoothing based target encoding for a categorical variable = $(global\_mean * weight + N * u_t,c)/(weight+N)$
+
+where $u_{t,c}$ is mean of target/outcome for a given category and weight is the smoothing factor or amount of regularization
+
+But to do that, first we checked number of observations per product category, to understand the range of weights to consider.
+The target encoding could be erroneous if done with larger weight or against the random dataset.
+
+To solve these, we identified the smoothing factor(or weight) to use by carrying out target encoding for different datasets for a set of weights, selecting single weight per dataset based on performance metric i.e.
 RMSE and then we took a mean of weights identified across the datasets.
 
-Once weight was identified, to avoid any randomness due to choice of the dataset, we computed target encoding for product category attribute across 10 datasets and then took a mean for them to settle at the final target encoded values to use for the concerned feature.
+Once weight was identified, to avoid any randomness due to choice of the dataset, we computed additive smoothing based target encoding for product category attribute across 10 datasets and then took a mean for them to settle at the final target encoded values to use for the concerned feature.
 
@@
 Incorporating them as it is would not have worked for many machine learning models which require numerical values.
-As a result, for seller ids we applied target encoding and followed strategy used for product categories.
+As a result, for seller ids we applied additive smoothing based target encoding and followed strategy used for product categories.
 
@@
 Incorporating them as it is would not have worked for many machine learning models which require numerical values.
-As a result, for customer states we applied target encoding and followed strategy used for product categories.
+As a result, for customer states we applied additive smoothing based target encoding and followed strategy used for product categories.
 
@@
 
-In addition, as earlier shared, categorical features were converted to numeric form with the help of target encoding with a weight term computed earlier for purpose of regularization and to avoid overfitting.
+In addition, as earlier shared, categorical features were converted to numeric form with the help of additive smoothing based target encoding with a smoothing factor/weight term computed earlier to avoid overfitting.
 Also, any undesired features were removed, reducing train_subset and test_subset to only the desired features.
@@
 
-In addition, categorical features were converted to numeric form with the help of target encoding with a weight term computed earlier for purpose of regularization and to avoid overfitting.
+In addition, categorical features were converted to numeric form with the help of additive smoothing based target encoding with a weight term computed earlier for purpose of smoothing and to avoid overfitting.
 Also, any undesired features were removed, reducing training_dataset and validation_dataset to only the desired features.
@@
 During the course of project, we carried out data cleaning, identified features and derived features from existing information available.
-In addition, converted categorical features to numeric form and applied various techniques like target encoding, scaling and normalization as required and regularized the required features.
+In addition, converted categorical features to numeric form and applied various techniques like additive smoothing based target encoding, scaling and normalization as required.
 

### 2021-07-21 finalReportPrep.Rmd
updated code with file references

@@
 ```{r trained knn model based on train_subset, echo=FALSE }
-load('rda/knn_model.rda')
-knn_fit
-
+if(file.exists('rda/knn_model.rda')) { 
+  load('rda/knn_model.rda'); 
+  print(knn_fit);
+} else {
+  lines <- readLines('captured-outputs/knn_model_output.txt')
+  cat(lines,sep="\n")
+
+  rm(lines)
+}
+```
+
+```{r print plot for knn model based on train_subset, echo=FALSE, out.width ="100%", fig.align='center'}
+if(file.exists('rda/knn_model.rda')) { 
     knn_fit %>% ggplot() 
-```
-
-Based on model results, k value of `r knn_fit$finalModel$k` was identified to be the one with lowest RMSE.
+} else {
+  knitr::include_graphics('captured-outputs/knn_model_plot.png')
+}
+```
+
+Based on model results, k value of `r if(exists("knn_fit",mode="list")) { knn_fit$finalModel$k } else {77}` was identified to be the one with lowest RMSE.
 
@@
 ```{r trained random forest model based on train_subset, echo=FALSE}
-load('rda/rf_model.rda')
-rf_fit
-
+if(file.exists('rda/rf_model.rda')) { 
+  load('rda/rf_model.rda'); 
+  print(rf_fit);
+} else {
+  lines <- readLines('captured-outputs/rf_model_output.txt')
+  cat(lines,sep="\n")
+
+  rm(lines)
+}
+```
+
+```{r print plot for rf model based on train_subset, echo=FALSE, out.width ="100%", fig.align='center'}
+if(file.exists('rda/rf_model.rda')) { 
     rf_fit %>% ggplot() 
+} else {
+  knitr::include_graphics('captured-outputs/rf_model_plot.png')
+}
 ```
@@
 
-Based on model results, mtry value of `r rf_fit$finalModel$mtry` was identified to be the one with lowest RMSE.
+Based on model results, mtry value of `r if(exists("rf_fit",mode="list")) { rf_fit$finalModel$mtry } else {2}` was identified to be the one with lowest RMSE.
 
@@
 ```{r Tuning parameters to use for final random forest, echo=FALSE}
-load('rda/rf_model.rda')
-
-data.frame(tuning_parameter='mtry', value=rf_fit$finalModel$mtry) %>% add_row(tuning_parameter='number_of_trees', value=500) %>% knitr::kable(caption='Random forest tuning parameters')
+mtry_value <- 2
+if(file.exists('rda/rf_model.rda')) { 
+  load('rda/rf_model.rda');
+  mtry_value = rf_fit$finalModel$mtry
+}
+
+data.frame(tuning_parameter='mtry', value=mtry_value) %>% add_row(tuning_parameter='number_of_trees', value=500) %>% knitr::kable(caption='Random forest tuning parameters')
+
+rm(mtry_value)
 ```
@@
 
+Details of the final trained random forest model can be found below:
+
 ```{r trained final random forest model based on training_dataset, echo=FALSE}
+if(file.exists('rda/final_model.rda')) {
   load('rda/final_model.rda')
-#final_model
-
+  sink('captured-outputs/final_model_output.txt')
+  print(final_model)
+  sink()
+}
+
+lines <- readLines('captured-outputs/final_model_output.txt')
+cat(lines,sep="\n")
+
+if(exists("final_model",mode="list")) {
   plot(final_model,main='Random forest model error against number of trees')
+} else {
+  plot_object <- readRDS('rds/final_model_plot.rds')
+  plot(plot_object,type='l',main='Random forest model error against number of trees', xlab='trees', ylab='Error')
+
+  rm(plot_object)
+}
 ```
@@
 performance_metric %>% knitr::kable(caption='Performance Metric(RMSE) Summary')
-```
-
-```{r Final knn model traing with data scaled, include=FALSE, eval= FALSE}
-set.seed(seed=1,sample.kind = 'Rounding')
-preProcessScalingValues <- preProcess(training_dataset %>% select(-number_of_hours), method = c("center", "scale"))
-
-scaled_train_subset <- predict(preProcessScalingValues, training_dataset) 
-scaled_test_subset <- predict(preProcessScalingValues, validation_dataset)
-
-final_knn_fit <- knnreg(number_of_hours~., scaled_train_subset,k = knn_fit$finalModel$k)
-
-final_knn_fit
-
-predictions <- predict(final_knn_fit,scaled_test_subset %>% select(-number_of_hours))
-
-RMSE(predictions,(scaled_test_subset %>% pull(number_of_hours) ))
-# 202.8552
-
-save(final_knn_fit,file='rda/final_knn_fit.rda')
-```
-
-```{r Final ensemble model, include=FALSE, eval= FALSE}
-load('rda/final_knn_fit.rda')
-load('rda/final_model.rda')
-
-set.seed(seed=1,sample.kind = 'Rounding')
-
-predictions_knn <- predict(final_knn_fit,scaled_test_subset %>% select(-number_of_hours))
-predictions_rf <- predict(final_model,validation_dataset %>% select(-number_of_hours))
-
-predictions_df <- data.frame(predictions_knn,predictions_rf)
-
-RMSE(rowMeans(predictions_df),(validation_dataset %>% pull(number_of_hours) ))
-# 193.3459
 ```

### 2021-07-21 finalReport.Rmd
intermediate state for record keeping

@@
 
+Combine segregated data sets
+- create derived info i.e. distance between two points
+
 Derived data set should be generated later on observations
@@
 
+```{r smooth fit for price, echo=FALSE}
+training_dataset %>% ggplot(aes(x=price,y=number_of_hours)) + geom_smooth() 
+```
+
 ## Explore distance relationship between two points with number_of_hours
@@
 # let's plot errorbar for number of hours vs volume of product group
-training_dataset %>% mutate(volume_group=round(volume/1000)+1) %>% mutate(volume_group=as.factor(volume_group)) %>% group_by(volume_group) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% ggplot(aes(x=volume_group,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))
+training_dataset %>% mutate(volume_group=round(volume/5000)+1) %>% mutate(volume_group=as.factor(volume_group)) %>% group_by(volume_group) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% ggplot(aes(x=volume_group,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))
 ```
@@
 
+== TODO additive smoothing/Bayes estimator with target encoding
+
 ```{r Dataframe to store target encoding weights, include=FALSE}
@@
 
-```{r Apply Target encoding for seller_id, echo = FALSE}
+```{r Apply Target encoding for product_category_name, echo = FALSE}
 #TODO you may need to remove customer_state and seller_state from dataset train_subset_for_encoding before proceeding further.
 
-  train_subset_for_encoding <- train_subset[1:30000,] %>% select(-customer_loc, -customer_zip_code_prefix, -customer_state)
+  train_subset_for_encoding <- train_subset[1:30000,] %>% select(-seller_id,-customer_loc, -customer_zip_code_prefix, -customer_state)
 
@@
 
-  weights <- seq(0,6,3)
+  weights <- seq(0,10,5)
 
@@
   test_weight_set <- test_weight_set %>%
+    semi_join(train_weight_set,by='product_category_name')
+
+  weight_rmses <- sapply(weights, function(weight) {
+      product_category_effects <- train_weight_set %>% 
+        group_by(product_category_name) %>% 
+        summarise(mean=mean(number_of_hours),n=n()) %>%
+        mutate(mean=(n*mean+weight*global_mean)/(n+weight)) %>% 
+        select(-n) %>% rename()
+
+      train_weight_set <- train_weight_set %>% left_join(product_category_effects,by='product_category_name') %>% rename(product_category_effect=mean) %>% select(-product_category_name)
+
+      test_weight_set <- test_weight_set %>% left_join(product_category_effects,by='product_category_name') %>% rename(product_category_effect=mean) %>% select(-product_category_name)
+
+      set.seed(seed=1,sample.kind = 'Rounding')
+      preProcessScalingValues <- preProcess(train_weight_set %>% select(-number_of_hours), method = c("center", "scale"))
+
+      scaled_train_subset <- predict(preProcessScalingValues, train_weight_set)
+      scaled_test_subset <- predict(preProcessScalingValues, test_weight_set)
+
+      # TODO remove below scaling was erroneous. train scaling needs to be used for test data
+      # below is just demo of how to use scale
+      # scaled_train_subset <- train_subset %>% mutate(across(c(mean_delay_in_handover,distance_between_two_points,volume,product_weight_g,freight_value,price),scale)) 
+      # scaled_test_subset <- test_subset %>% mutate(across(c(mean_delay_in_handover,distance_between_two_points,volume,product_weight_g,freight_value,price),scale))
+
+      # TODO can we skip modelling while evaluating weight to use or for target encoding?
+
+      tuningControl=trainControl(method = "cv", number = 3, p = .8)
+      knn_fit_for_weight <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(27,34,41)),trControl=tuningControl)
+
+      knn_fit_for_weight
+      print(knn_fit_for_weight)
+
+      predictions <- predict(knn_fit_for_weight,scaled_test_subset %>% select(-number_of_hours))
+
+      RMSE(scaled_test_subset$number_of_hours,predictions) 
+  })
+
+  # selecting weight leading to lowest RMSE
+  weight_to_use <- weights[which.min(weight_rmses)]
+  paste('Weight to use: ', weight_to_use, ', corresponding RMSE: ', weight_rmses[which.min(weight_rmses)])
+
+  print(weight_rmses)
+  weight_to_use
+})
+
+weight_to_use_per_partition
+
+weight_to_use <- mean(weight_to_use_per_partition)  
+# 3
+
+target_encoding_weights <- target_encoding_weights %>% add_row(effect='product_category',weight=weight_to_use)
+```
+
+```{r Post regularization for product_category effect, echo=FALSE}
+
+# prepare product_category_effect based on regularisation weight we have identified
+  global_mean <- mean(train_subset$number_of_hours)
+
+  set.seed(1,sample.kind='Rounding')
+  test_weight_indices <- createDataPartition(train_subset$number_of_hours,p=0.2,times=10,list=FALSE)
+
+product_category_effects_list <- lapply(1:10, function(index) {
+    train_weight_set <- train_subset[-test_weight_indices[,index],]
+
+    product_category_effects <- train_weight_set %>% 
+      group_by(product_category_name) %>%
+      summarise(mean=mean(number_of_hours),n=n()) %>%
+      mutate(mean=(n*mean+weight_to_use*global_mean)/(n+weight_to_use)) %>% 
+      select(-n) %>% rename()
+
+    product_category_effects
+  }
+)
+
+pce <- product_category_effects_list[[1]]
+for (index in 2:10) {
+   pce <- pce %>% full_join(product_category_effects_list[[index]],by='product_category_name', suffix=c("1",index))
+}
+
+# wherever value is NA, we have used global mean
+pce <- pce %>% mutate(across(-product_category_name,~if_else(is.na(.x),global_mean,.x)))
+
+# as different partitions may not have exactly same data set combined the results
+product_category_effects <- tibble(pce[,1], product_category_effect=rowMeans(pce[,-1],na.rm = TRUE))
+
+train_subset <- train_subset %>% left_join(product_category_effects,by='product_category_name') %>% select(-product_category_name) %>% mutate(product_category_effect=if_else(is.na(product_category_effect),global_mean,product_category_effect))
+
+test_subset <- test_subset %>% left_join(product_category_effects,by='product_category_name') %>% select(-product_category_name)
+```
+
+```{r Apply Target encoding for seller_id, echo = FALSE}
+#TODO you may need to remove customer_state and seller_state from dataset train_subset_for_encoding before proceeding further.
+
+  train_subset_for_encoding <- train_subset[1:30000,] %>% select(-customer_loc, -customer_zip_code_prefix, -customer_state)
+
+  global_mean <- mean(train_subset_for_encoding$number_of_hours)
+
+  set.seed(1,sample.kind='Rounding')
+  test_weight_indices <- createDataPartition(train_subset_for_encoding$number_of_hours,p=0.2,times=3,list=FALSE)
+
+  weights <- seq(0,6,3)
+
+  weight_to_use_per_partition <- sapply(1:3, function(index) {
+  #lambda_data_set <- prepare_lambda_data_set(index)
+  #train_lambda_set <- lambda_data_set[[1]]
+  #test_lambda_set <- lambda_data_set[[2]]  
+  train_weight_set <- train_subset_for_encoding[-test_weight_indices[,index],]
+  test_weight_set <- train_subset_for_encoding[test_weight_indices[,index],]
+
+  test_weight_set <- test_weight_set %>%
     semi_join(train_weight_set,by='product_category_name') %>%
@@
   set.seed(1,sample.kind='Rounding')
-  test_weight_indices <- createDataPartition(train_subset$number_of_hours,p=0.2,times=5,list=FALSE)
-
-seller_id_effects_list <- lapply(1:5, function(index) {
+  test_weight_indices <- createDataPartition(train_subset$number_of_hours,p=0.2,times=10,list=FALSE)
+
+seller_id_effects_list <- lapply(1:10, function(index) {
     train_weight_set <- train_subset[-test_weight_indices[,index],]
@@
 sie <- seller_id_effects_list[[1]]
-for (index in 2:5) {
+for (index in 2:10) {
    sie <- sie %>% full_join(seller_id_effects_list[[index]],by='seller_id', suffix=c("1",index))
@@
 
-# whereever value is NA, we have used global mean
+# wherever value is NA, we have used global mean
 sie <- sie %>% mutate(across(-seller_id,~if_else(is.na(.x),global_mean,.x)))
@@
 
-train_subset <- train_subset %>% left_join(seller_id_effects,by='seller_id') %>% select(-product_category_name,-seller_id) %>% mutate(seller_id_effect=if_else(is.na(seller_id_effect),global_mean,seller_id_effect))
-
-test_subset <- test_subset %>% left_join(seller_id_effects,by='seller_id') %>% select(-product_category_name,-seller_id)
+train_subset <- train_subset %>% left_join(seller_id_effects,by='seller_id') %>% select(-seller_id) %>% mutate(seller_id_effect=if_else(is.na(seller_id_effect),global_mean,seller_id_effect))
+
+test_subset <- test_subset %>% left_join(seller_id_effects,by='seller_id') %>% select(-seller_id)
 ```
@@
   set.seed(1,sample.kind='Rounding')
-  test_weight_indices <- createDataPartition(train_subset$number_of_hours,p=0.2,times=5,list=FALSE)
-
-customer_state_effects_list <- lapply(1:5, function(index) {
+  test_weight_indices <- createDataPartition(train_subset$number_of_hours,p=0.2,times=10,list=FALSE)
+
+customer_state_effects_list <- lapply(1:10, function(index) {
     train_weight_set <- train_subset[-test_weight_indices[,index],]
@@
 cse <- customer_state_effects_list[[1]]
-for (index in 2:5) {
+for (index in 2:10) {
    cse <- cse %>% full_join(customer_state_effects_list[[index]],by='customer_state', suffix=c("1",index))
@@
 
-# whereever value is NA, we have used global mean
+# wherever value is NA, we have used global mean
 cse <- cse %>% mutate(across(-customer_state,~if_else(is.na(.x),global_mean,.x)))
@@
 
-tuningControl=trainControl(method = "cv", number = 3, p = .8)
-knn_fit <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(27,34,41)),trControl=tuningControl)
+tuningControl=trainControl(method = "cv", number = 5, p = .8)
+knn_fit <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(56,63,70,77)),trControl=tuningControl)
 
@@
 RMSE(predictions,(scaled_test_subset %>% pull(number_of_hours) ))
-#184.7906 for k=34.
+#183.2447 for k=70
 
@@
 ```{r correlation, eval=FALSE}
-cor(train_subset %>% select(price,freight_value,customer_loc_effect,customer_state_effect,seller_id_effect,distance_between_two_points,mean_delay_in_handover,volume,number_of_hours,product_weight_g),use="pairwise.complete.obs") %>% knitr::kable()
+cor(train_subset %>% select(price,freight_value,product_category_effect,customer_loc_effect,customer_state_effect,seller_id_effect,distance_between_two_points,mean_delay_in_handover,volume,number_of_hours,product_weight_g),use="pairwise.complete.obs") %>% knitr::kable()
 ```
@@
 tuningControl=trainControl(method = "cv", number = 2, p = .8)
-rf_fit <- train(number_of_hours~., train_subset, method='rf',tuneGrid = data.frame(mtry = seq(5, 40, 10)),trControl=tuningControl,ntree=250)
+rf_fit <- train(number_of_hours~., train_subset, method='rf',tuneGrid = data.frame(mtry = c(1,2,3,4,5,10)),trControl=tuningControl,ntree=250)
 #rf_fit <- train(number_of_hours~., train_subset %>% select(-seller_state), method='rf',tuneGrid = data.frame(mtry = seq(5, 50, 10)),trControl=tuningControl,ntree=250) # nodesize = 50,maxnodes=2000,ntree=200
@@
 RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
-# 175.0988 for mtry 15.
-Technically there are total of 82 variables due to 73 product categories and 9 other predictors.
-The seller_state was not of much help.
-it was actually increasing RMSE by .6.
-So have ignored it.
+# 173.9907 for mtry 2. Technically there are total of 10 predictors. The seller_state was not of much help. it was actually increasing RMSE. So have ignored it.
 
@@
 RMSE(predictions,(scaled_test_subset %>% pull(number_of_hours) ))
-# 185.1074 uses GCV.Cp method and select=FALSE
+# 183.4923 uses GCV.Cp method and select=TRUE
 
 save(gam_fit,file='rda/gam_model.rda')
+```
+
+```{r Prepare product_category effect based on target encoding, include = FALSE}
+# prepare product_category_effect based on regularisation weight we have identified
+prepare_product_category_effect <- function(train_df, weight_to_use) {
+  global_mean <- mean(train_df$number_of_hours)
+
+  set.seed(1,sample.kind='Rounding')
+  test_weight_indices <- createDataPartition(train_df$number_of_hours,p=0.2,times=10,list=FALSE)
+
+  product_category_effects_list <- lapply(1:10, function(index) {
+      train_weight_set <- train_df[-test_weight_indices[,index],]
+
+      product_category_effects <- train_weight_set %>% 
+        group_by(product_category_name) %>%
+        summarise(mean=mean(number_of_hours),n=n()) %>%
+        mutate(mean=(n*mean+weight_to_use*global_mean)/(n+weight_to_use)) %>% 
+        select(-n) %>% rename()
+
+      product_category_effects
+    }
+  )
+
+  pce <- product_category_effects_list[[1]]
+  for (index in 2:10) {
+     pce <- pce %>% full_join(product_category_effects_list[[index]],by='product_category_name', suffix=c("1",index))
+  }
+
+  # wherever value is NA, we have used global mean
+  pce <- pce %>% mutate(across(-product_category_name,~if_else(is.na(.x),global_mean,.x)))
+
+  # as different partitions may not have exactly same data set combined the results
+  product_category_effects <- tibble(pce[,1], product_category_effect=rowMeans(pce[,-1],na.rm = TRUE))
+
+  #product_category effects are ready
+  product_category_effects
+}
 ```
@@
 
+```{r Enhance training and test dataset for product_category effect based on target encoding, include = FALSE}
+enhance_datasets_using_product_category_effect <- function(product_category_effects,train_df,test_df,cleanup) {
+  #we will use seller id effects computed using training data in test data.
+  train_df <- train_df %>% left_join(product_category_effects,by='product_category_name') %>% mutate(product_category_effect=if_else(is.na(product_category_effect),global_mean,product_category_effect))
+
+  test_df <- test_df %>% left_join(product_category_effects,by='product_category_name')
+
+  # cleanup
+  if (cleanup) {
+    train_df <- train_df %>% select(-product_category_name)
+    test_df <- test_df %>% select(-product_category_name)
+  }
+
+  list(train_df,test_df)
+}
+```
+
 ```{r Enhance training and test dataset for seller_id effect based on target encoding, include = FALSE}
@@
   if (cleanup) {
-    train_df <- train_df %>% select(-product_category_name,-seller_id)
-    test_df <- test_df %>% select(-product_category_name,-seller_id)
+    train_df <- train_df %>% select(-seller_id)
+    test_df <- test_df %>% select(-seller_id)
   }
@@
 
+  print('Enhancing data for product category effects')
+  weight_to_use <- target_encoding_weights %>% 
+    filter(effect=='product_category') %>% 
+    pull(weight)
+  product_category_effects <- prepare_product_category_effect(train_ds,weight_to_use)
+  dataset_list <- enhance_datasets_using_product_category_effect(product_category_effects,train_ds,test_ds,cleanup)
+  train_ds <- dataset_list[[1]]
+  test_ds <- dataset_list[[2]]
+
   print('Enhancing data for seller id effects')
@@
 
-validation_dataset <- validation_dataset %>%
-  semi_join(training_dataset,by='product_category_name') %>% 
-  semi_join(training_dataset,by='seller_id') %>% 
-  semi_join(training_dataset,by='customer_state') %>%
-  semi_join(training_dataset,by='customer_loc')
-
 # cleanup
 training_dataset <- training_dataset %>%
-  select(-handed_over_to_carrier_date,-shipping_limit_date,-product_category_name,-seller_id,-customer_state,-customer_loc,-customer_zip_code_prefix)
+  select(-handed_over_to_carrier_date,-shipping_limit_date,-product_category_name,-seller_id,-customer_state,-customer_loc,-customer_zip_code_prefix,-order_purchase_timestamp,-payment_approved_at)
 validation_dataset <- validation_dataset %>%
-  select(-handed_over_to_carrier_date,-shipping_limit_date,-product_category_name,-seller_id,-customer_state,-customer_loc,-customer_zip_code_prefix)
+  select(-handed_over_to_carrier_date,-shipping_limit_date,-product_category_name,-seller_id,-customer_state,-customer_loc,-customer_zip_code_prefix,-order_purchase_timestamp,-payment_approved_at)
 ```
@@
 RMSE((validation_dataset %>% pull(number_of_hours) ),predictions)
-# 190.184 for mtry=15, ntree=500
+# 189.2705 for mtry=2, ntree=500
+```
+
+```{r Final model analysis, echo=FALSE}
+validation_dataset %>% select(number_of_hours) %>% mutate(diff=(number_of_hours-predictions)) %>% ggplot(aes(diff)) + geom_histogram(binwidth = 1) + scale_x_log10()
+
+validation_dataset %>% select(number_of_hours) %>% mutate(diff=number_of_hours-predictions) %>% summary()
+
+```
+
+```{r Final knn model traing with data scaled, echo=FALSE}
+set.seed(seed=1,sample.kind = 'Rounding')
+preProcessScalingValues <- preProcess(training_dataset %>% select(-number_of_hours), method = c("center", "scale"))
+
+scaled_train_subset <- predict(preProcessScalingValues, training_dataset) 
+scaled_test_subset <- predict(preProcessScalingValues, validation_dataset)
+
+final_knn_fit <- knnreg(number_of_hours~., scaled_train_subset,k = knn_fit$finalModel$k)
+
+final_knn_fit
+
+predictions <- predict(final_knn_fit,scaled_test_subset %>% select(-number_of_hours))
+
+RMSE(predictions,(scaled_test_subset %>% pull(number_of_hours) ))
+# 203.0848
+
+save(final_knn_fit,file='rda/final_knn_fit.rda')
 ```

### 2021-07-21 analysis.R
Cleaning files, renaming files for final presentation (#1)

@@ -1,1519 +1,0 @@
-#######################################################################################
-################################# Prepare Environment #################################
-#######################################################################################
-#Install or load packages
-if (!require('dplyr')) install.packages('dplyr'); library('dplyr')
-if (!require('lubridate')) install.packages('lubridate'); library('lubridate')
-if (!require('tidyverse')) install.packages('tidyverse'); library('tidyverse')
-if (!require('geosphere')) install.packages('geosphere'); library('geosphere')
-if (!require('caret')) install.packages('caret'); library('caret')
-if (!require('randomForest')) install.packages('randomForest'); library('randomForest')
-if (!require('knitr')) install.packages('knitr'); library('knitr')
-
-
-#######################################################################################
-#################################### Data wrangling ###################################
-#######################################################################################
-
-####################### Data wrangling for products data set
-
-# read the data from source and remove irrelevant columns
-products_dataset <- read.csv(file.path('data','olist_products_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) %>%
-  select(-product_name_lenght,-product_description_lenght,-product_photos_qty)
-
-# check how they are structured
-products_dataset %>% head()
-products_dataset %>% str()
-
-# load the product category name translations
-product_category_name_translation <- read.csv(file.path('data','product_category_name_translation.csv'),header=TRUE,stringsAsFactors = FALSE)
-
-product_category_name_translation %>% head()
-
-# convert product_name to english form wherever possible. If not available, existing product category name is retained
-products_dataset <- products_dataset %>% 
-  left_join(product_category_name_translation,by='product_category_name') %>% 
-  mutate(product_category_name=if_else(is.na(product_category_name_english),
-                                       product_category_name,
-                                       product_category_name_english)) %>% 
-  select(-product_category_name_english) 
-
-# further remove products for which data is not available i.e is NA
-products_dataset <- products_dataset %>% 
[…]

### 2021-07-21 end-to-end-script.R

@@ -1,0 +1,1519 @@
+#######################################################################################
+################################# Prepare Environment #################################
+#######################################################################################
+#Install or load packages
+if (!require('dplyr')) install.packages('dplyr'); library('dplyr')
+if (!require('lubridate')) install.packages('lubridate'); library('lubridate')
+if (!require('tidyverse')) install.packages('tidyverse'); library('tidyverse')
+if (!require('geosphere')) install.packages('geosphere'); library('geosphere')
+if (!require('caret')) install.packages('caret'); library('caret')
+if (!require('randomForest')) install.packages('randomForest'); library('randomForest')
+if (!require('knitr')) install.packages('knitr'); library('knitr')
+
+
+#######################################################################################
+#################################### Data wrangling ###################################
+#######################################################################################
+
+####################### Data wrangling for products data set
+
+# read the data from source and remove irrelevant columns
+products_dataset <- read.csv(file.path('data','olist_products_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) %>%
+  select(-product_name_lenght,-product_description_lenght,-product_photos_qty)
+
+# check how they are structured
+products_dataset %>% head()
+products_dataset %>% str()
+
+# load the product category name translations
+product_category_name_translation <- read.csv(file.path('data','product_category_name_translation.csv'),header=TRUE,stringsAsFactors = FALSE)
+
+product_category_name_translation %>% head()
+
+# convert product_name to english form wherever possible. If not available, existing product category name is retained
+products_dataset <- products_dataset %>% 
+  left_join(product_category_name_translation,by='product_category_name') %>% 
+  mutate(product_category_name=if_else(is.na(product_category_name_english),
+                                       product_category_name,
+                                       product_category_name_english)) %>% 
+  select(-product_category_name_english) 
+
+# further remove products for which data is not available i.e is NA
+products_dataset <- products_dataset %>% 
[…]

### 2021-07-21 finalProjectReport.Rmd
fine tuned files

--- 2021-07-21 finalReportPrep.Rmd
@@
 author: "Asham Vohra"
-date: "6/21/2021"
+date: "21st June 2021"
 output:
-  pdf_document: default
+  pdf_document:
+    toc: true
+    toc_depth: 2
+    number_sections: true
   html_document:
@@
 ```
+
+\newpage
 
@@
 
-This was a crucial dataset not only for us to build our delivery estimation system but to even identify the current performance metric i.e. RMSE for the approach Olist was using.
+This was a crucial dataset not only for us to build our delivery prediction system but to even identify the current performance metric i.e. RMSE for the approach Olist was using.
 
@@
 # order_item_id --->sequential number identifying number of items included in the same order.
-# so if there are three items they will be nuber 1 , 2, 3 for the given order
+# so if there are three items they will be numbered 1 , 2, 3 for the given order
 # shipping_limit_date ---> Shows the seller shipping limit date for handling the order over to the logistic partner.
-# freight_value --> item freight value item (if an order has more than one item the freight value is splitted between items)
+# freight_value --> item freight value item (if an order has more than one item the freight value is split between items)
 order_items_dataset <- read.csv(file.path('data','olist_order_items_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) %>% mutate(shipping_limit_date=as_datetime(shipping_limit_date))  %>% select(-order_item_id)
@@
 
-For the above analysis, knn model was used, given that it does not make assumptions about the dataset, which was useful in the early phase of analysis.
+For the above analysis, k-nearest neighbors(knn) model was used, given that it does not make assumptions about the dataset, which was useful in the early phase of analysis.
 However, to use knn model, we **scaled and normalized** the features as knn model uses Euclidean distance.
@@
 
-In addition, for knn model training and tuning, we carried our 5-fold cross validation providing a range of values of k to use for training the model.
+In addition, for knn model training and tuning, we carried our 5-fold cross validation providing a range of values of **k** to use for training the model.
 Here, k means number of neighbors to be considered.
@@
 
-Based on model results, k value of `r if(exists("knn_fit",mode="list")) { knn_fit$finalModel$k } else {77}` was identified to be the one with lowest RMSE.
+Based on model results, **k** value of **`r if(exists("knn_fit",mode="list")) { knn_fit$finalModel$k } else {77}`** was identified to be the one with lowest RMSE.
 
@@
 To train and tune random forest model, we carried our 4-fold cross validation providing a range of values of mtry to use for training the model.
-Here, mtry means number of random features used to determine the best split at each node of the decision tree.
-In addition, we configured the number of decision trees to create as 200.
+Here, **mtry** means number of random features used to determine the best split at each node of the decision tree.
+In addition, we configured the number of decision trees to create as **200**.
 Here, the cross validation not only helped to reduce the probability of overfitting but also helped in identifying the tuning parameter i.e.
@@
 
-Based on model results, mtry value of `r if(exists("rf_fit",mode="list")) { rf_fit$finalModel$mtry } else {2}` was identified to be the one with lowest RMSE.
+Based on model results, **mtry** value of **`r if(exists("rf_fit",mode="list")) { rf_fit$finalModel$mtry } else {2}`** was identified to be the one with lowest RMSE.
 
@@
 
-We noted that weighted ensemble model had lowest RMSE of `r min(rmses_weighted_ensemble_model)` at weight=`r weights_for_ensemble_model[which.min(rmses_weighted_ensemble_model)]`
+We noted that weighted ensemble model had lowest RMSE of **`r min(rmses_weighted_ensemble_model)`** at weight=**`r weights_for_ensemble_model[which.min(rmses_weighted_ensemble_model)]`**
 

### 2021-07-21 end-to-end-script.R

@@ -90,7 +90,7 @@ orders_dataset %>% head(2) %>% as.matrix() %>% t() %>%
 # order_item_id --->sequential number identifying number of items included in the same order.
 # so if there are three items they will be numbered 1 , 2, 3 for the given order
 # shipping_limit_date ---> Shows the seller shipping limit date for handling the order over to the logistic partner.
-# freight_value --> item freight value item (if an order has more than one item the freight value is splitted between items)
+# freight_value --> item freight value item (if an order has more than one item the freight value is split between items)
 order_items_dataset <- read.csv(file.path('data','olist_order_items_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) %>% mutate(shipping_limit_date=as_datetime(shipping_limit_date))  %>% select(-order_item_id)
 
 # check how they are structured
@@ -920,7 +920,7 @@ training_dataset %>% group_by(product_category_name) %>%
 
 
 # Apply Target encoding for product_category_name
-# here we will try to identify the regularization weight to use by carrying out target encoding 
+# here we will try to identify the regularization weight/smoothing factor to use by carrying out target encoding 
 # for different datasets set of weights, selecting single weight per dataset based on 
 # performance metric i.e. RMSE and then we will take a mean of weights identified across the datasets.
 train_subset_for_encoding <- train_subset[1:40000,] %>% select(-customer_loc, -customer_state)
@@ -1007,6 +1007,7 @@ training_dataset %>% group_by(seller_id) %>%
   ylab('Avg number of hours')
 
 # Apply Target encoding for seller_id
+# here we will try to identify the regularization weight/smoothing factor to use by carrying out target encoding
 train_subset_for_encoding <- train_subset[1:40000,] %>% select(-customer_loc, -customer_state)
 
 global_mean <- mean(train_subset_for_encoding$number_of_hours)
@@ -1098,6 +1099,7 @@ training_dataset %>% group_by(customer_state) %>%
   xlab('Customer States') + ylab('Avg number of hours')
 
 # Apply Target encoding for customer_state
+# here we will try to identify the regularization weight/smoothing factor to use by carrying out target encoding 
 train_subset_for_encoding <- train_subset[1:40000,] %>% select(-customer_loc)
 global_mean <- mean(train_subset_for_encoding$number_of_hours)
 
@@ -1199,6 +1201,7 @@ training_dataset %>% group_by(customer_loc) %>%
[…]

### 2021-07-21 finalProjectReport.Rmd
resolved typos

@@
 order_delivered_customer_date and order_purchase_timestamp as this is what we would want to predict whenever we get a new order i.e.
-given this order, how many number of hours are required to deliver the product to the customer.
+given this order, how many number of hours are required to deliver the ordered product to the customer.
 
@@
 
-To understand its influence on the delivery timelines, we plotted numbers of hours taken for delivery against the price of the ordered item.
+To understand its influence on the delivery timelines, we plotted numbers of hours taken for delivery against the price of the ordered product.
 In order to see the trend, gam method based smoothing curve was plotted on the graph.
@@
 ```{r Explore price relation with number_of_hours, echo=FALSE, message=FALSE}
-training_dataset  %>% ggplot(aes(price,number_of_hours)) + geom_smooth() + xlab('Price of ordered item') + ylab('Number of hours taken for delivery')
-```
-In above plot, we noted that number of hours taken for delivery vary as the price of the ordered item changes.
-It appeared to increase significantly as the price of ordered item increased and decreased when the price reached towards higher limits.
+training_dataset  %>% ggplot(aes(price,number_of_hours)) + geom_smooth() + xlab('Price of ordered product') + ylab('Number of hours taken for delivery')
+```
+In above plot, we noted that number of hours taken for delivery vary as the price of the ordered product changes.
+It appeared to increase significantly as the price of ordered product increased and decreased when the price reached towards higher limits.
 Though confidence interval indicated higher variability in delivery timelines at higher prices.
@@
 In addition, it does not make assumptions about the dataset.
-In our case limited dataset and that the alogrithm does not make any assumptions, made it a good candidate for model development and performance evaluation.
+In our case limited dataset and that the algorithm does not make any assumptions, made it a good candidate for model development and performance evaluation.
 
@@
 RMSE = **`r performance_metric %>% filter(type=='Final model') %>% pull(value)`** captured against **validation_dataset** was much better than what Olist store had as identified from the dataset.
-The newly built model can be used for predicting number of hours required for delivery of a given ordered item.
+The newly built model can be used for predicting number of hours required for delivery of a given ordered product.
 
 # Conclusion
-As part of this project, the goal was to build a delivery prediction system which could predict the number of hours required to delivered the ordered item rating with minimal RMSE.
+As part of this project, the goal was to build a delivery prediction system which could predict the number of hours required to delivered the ordered product rating with minimal RMSE.
 
@@
 Post this, we explored multiple ensemble models using techniques like averaging, weighted average, stacking, among others.
-Finally we selected random forest based model and the model so trained performs within desired range of RMSE and can be useful for predicting number of hours required for delivery of a given ordered item.
+Finally we selected random forest based model and the model so trained performs within desired range of RMSE and can be useful for predicting number of hours required for delivery of a given ordered product.
 Due care was taken to prevent overfitting from influencing our model.
